### 2017-03-05 eda_report.rnw
Delivered EDA report to Guido.

\documentclass[12pt]{article}

\usepackage{algorithm2e} 
\usepackage{amsmath} \usepackage{amsthm} 
\usepackage{amsfonts} 
\usepackage{bbm} \usepackage{color,soul} 
\usepackage{framed} 
\usepackage[margin=0.5in]{geometry} \usepackage{hyperref} 
\usepackage{mathtools} 
\usepackage[dvipsnames]{xcolor}  
\newtheorem{theorem}{Theorem}[section] 
\newtheorem{lemma}[theorem]{Lemma} \newtheorem{proposition}[theorem]{Proposition} \newtheorem{corollary}[theorem]{Corollary}  
\newcommand{\D}{\mathrm{d}}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil} \DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\SetKwInput{KwInput}{Input} 
\SetKwInput{KwOutput}{Output}

\begin{document}

\title{Exploring the ``Party City'' effect in the NBA}
\author{Andreas Santucci, Eric Lax}
\maketitle

\subsection*{Motivation}
It's not unreasonable to think that game performance can be affected (in part) by what takes place off the basketball court. We are interested to see if teams exhibit a decline in performance the day following a game in a party city. We exploit data on bookmaker spreads: the expected score differential between two teams after conditioning on observable performance. We expect a team to meet the spread half the time; if not, the bookmakers would open up an opportunity for arbitrage over the long-term. We construct a model which attempts to estimate the causal effect of  visiting a ``party-city'' the day before a game on the probability of meeting the spread.

\paragraph{Defining a party city}
We define a \emph{party city} to be one of the following: Atlanta, Brooklyn, Los Angeles, or Miami. We then look for other teams (not from one of these cities)
who travel to a party city and then play a game in another city the day after; each of the thirty NBA teams encounters this situation on average twice per season.

\subsection*{A simple first model}
What if we pick up the effect of a busy play-schedule, or travel from a far-away city? We consider comparing the difference in performance when a team has visited a party city the day before to the difference in performance when a team has visited any other city the day before. We estimate the probability of meeting the spread
using a logistic regression. Model sketch:
$$
\texttt{meet spread} \sim \substack{\texttt{days since} \\ \texttt{last game}} + \substack{\texttt{games played} \\ \texttt{last week}} + \substack{\texttt{team} \\ \texttt{ranking}} + \substack{\texttt{time}\\ \texttt{difference}} + \substack{\texttt{distance} \\ \texttt{traveled}} + \texttt{party}
$$

\subsection*{Challenges} Each of the thirty NBA teams runs into our quasi-experiment only a few times
per season, so our effective sample size relatively small. One possible remedy is to consider a more granular model, analyzing individual player performance. Another challenge is that our model relies
on spreads as input, for which we would like to know which factors are taken into account, especially
in so far as this model may change from season to season. Although
we cannot directly observe how the spreads model is built, we can observe how well different features
predict the probability of meeting the spread, and from this back out or infer what one such model 
may be.

\subsection*{References}
``The Benefits of College Athletic Success: An Application of the Propensity Score Design'', by Michael L. Anderson. March, 2017. \href{http://www.mitpressjournals.org/doi/pdfplus/10.1162/REST_a_00589}{paper}

\newpage

\section*{Exploratory Analysis}
Here, we explore a few aspects of our data, which are collected from
\href{http://www.goldsheet.com/histnba.php}{gold sheet}. We utilize 2008-2012 data. We start with a sanity check: if spreads are set correctly, they should be met on average half the time; otherwise an arbitrage opportunity exists. This
is verified.

<<echo=F,results='hide',include=FALSE>>=
require(data.table)
require(ggplot2)
require(magrittr)

load(file = 'tmp_data/spreads_wdist.RData')
load(file = 'tmp_data/standings.RData')

data[, season := gsub('_20', '_', season)]

standings <- subset(standings, select = c('team', 'season', 'pct'))
standings[, pct := shift(pct, type = 'lag'), by = team]

data <- merge(data, standings, by = c('season', 'team'), all.x = T)
data <- data[data[, head(.I, 82), by = list(season, team)]$V1]

### Define a Party Variable
party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
party.rgx <- paste0('(', party.cities, ')', collapse = "|")

data[, party := 0]
data[, weekend := grepl("(Friday)|(Saturday)|(Sunday)", day) %>% as.integer]

data[!grepl(party.rgx, team) & grepl(party.rgx, last.game.loc) & ndays.lgame == 1, party := 1]
@

<<echo=F,out.height="6in",warning=F>>=

### How often are teams meeting the spread?
hist(data[, mean(outcome == 'W'), by = list(season, team)][, V1],
     breaks = seq(0.3, 0.75, by = 0.025),
     main = "How often are teams meeting the spread?",
     xlab = "Proportion of games for which a team-year meets the spread")
@

We also check how many observations we will have to test our hypothesis. On average, each team
only encounters our quasi-experiment twice per season. This motivates the need for a more granular
model, which perhaps might look at the party effect on individual performance. We could
further break this down into offensive performance (shots scored per minute of playtime)
and defensive performance, perhaps.

<<echo=F,out.height="5.5in">>=
#data[, .N, by = list(season, team)]
data[party==1, .N, by = list(season, team)][, N] %>%
    hist(main = "Number of party days by season-team", xlab = "Number of days", ylab = "Frequency")
@

\subsection*{Arguing which factors are taken into account in the spread model}
We argue that the input spreads data has already taken into account (at least some of)
a team's ability, in so far as ability can be proxied by last season's average win-rate.

<<echo=F,warning=F,fig.pos='H'>>=

### TODO : Look at outliers here!!
### TODO : Remove playoff games.
m <- data[,#grepl(party.rgx, game.loc) & !grepl(party.rgx, team),
          list(meet.spread = mean(outcome == 'W'), pct = unique(pct), .N), by = list(team, season)]
ggplot(m, aes(x = pct, y = meet.spread)) +
    geom_point() +
    facet_wrap(facets = ~season) + 
    labs(title = "In select seasons, lagged performance not predictive of meet the spread",
         x = "Percentage of games won last season",
         y = "Proportion of games for which spread was met in current season")
@

We remark that in seasons starting in 2008-2010, there is almost no correlation between 
performance in the prior season and proportion of games for which spread was met in the current season. In 2011-12 and 2012-13 we have positive correlation between these two variables, which is interesting; we do not expect this. It is indicative
of a change in the way spreads are calculated, and is worthy of further investigation.

We now also argue that clearly the number of days since last game has been factored into spread data.

<<>>=
m[, cor(meet.spread, pct, use = 'complete'), by = season]
@

We see that on average, teams are meeting the spread half the time no matter if they've played
their most recent game one day ago or one week ago. There are a few outliers when teams are allowed
over one week rest; we note that these games are not of interest to our analysis, and may be excluded
without problem.

<<echo=F,out.height="6in",warning=F,fig.pos='H'>>=
m <- data[, list(meet.spread = mean(outcome == 'W'), .N), by = ndays.lgame][order(ndays.lgame)]
ggplot(m, aes(x = ndays.lgame, y = meet.spread, label = N)) +
    geom_point() +
    geom_label(nudge_y = -.04) +
    geom_hline(yintercept = 0.5) + 
    labs(x = "Number of days since last game", y = "Met the Spread",
         title = paste("Proportion of games which meet spread",
                       "as a function of rest-time",
                       "(Counts indicate number of observations used)", sep = "\n"))
@

The same is true for distance traveled as well.

<<echo=F,warning=F>>=
m <- data[, list(meet.spread = mean(outcome == 'W'), .N), by = list(season, travel.dist = round(travel.dist, digits = -2))]
ggplot(m, aes(x = travel.dist, y = meet.spread, label = N)) +
    geom_point() +
    geom_label(nudge_y = -.01) +
    geom_hline(yintercept = 0.5) +
#    facet_wrap(facets = ~season) + 
    labs(x = "Travel Distance", y = "Met the Spread",
         title = paste("Proportion of games which meet spread",
                       "as a function of travel distance",
                       "(Counts indicate number of observations used)", sep = "\n"))

@

<<>>= 
m[, cor(travel.dist, meet.spread, use = 'complete')]
@

When we partition the above plot by season, we see that there is some linear correlation
between distance traveled and proportion of games for which spread was met; however, this correlation
heavily influenced by a few outliers (in particular, these outliers' averages are calculated based on very few observations).

<<>>=
m[, cor(travel.dist, meet.spread, use = 'complete'), by = season]
@

\section*{A tabular look at the effect of partying by team}

Lastly, we take a look at the effect of partying by season-team. Notice that
some teams are more effected than others; we hypothesize that demographics of the 
team influence the effect size. In particular, younger teams with unmarried players may
exhibit larger variability in performance after visiting party cities.

Some notable examples include the Denver Nuggets, and the Philadelphia 76'ers.
Also of importance: the San Antonio Spurs consistently have the oldest team, no matter which
statistic is used, and they also have a coach which is notoriously strict on partying; for this team, unsurprisingly, the party effect moves in the opposite direction. For 2008-2012 historical data, we display below the proportion of games for which
each team meets the spread under treatment as compared to without, along with the number of observations used to calculate the statistic.

<<>>=

data[, list(meet.spread = mean(outcome == 'W'), .N), by = list(team, party)][order(team, party)]
@

We might also consider that the effect sizes may change as a function of how competitive the team is. Perhaps we may consider running a separate analysis for teams in the bottom quantile of last-year's rankings.
We also remark that there is not a lot of data supporting each mean estimate for
the treatment effect by team (especially at the season-team level), and this does pose issues.
\subsection*{A first model}
For the 2008-2012 seasons, we fit the following model.

<<>>=
load(file = 'tmp_data/spreads_with_standings.RData')
model <- glm(outcome == 'W' ~ party + weekend + ndays.lgame + pct +
                 travel.dist + team + season + team:season,
             family = 'binomial', data = data)

summary(model)$coef[c(1:6, sample(7:nrow(coef(summary(model))), size = 6)), ]
@

\section*{Problems, Extensions}

We remark that the magnitude and sign of the coefficient on \texttt{party} are of practical
importance. Although we don't achieve statistical significance, we may be able to come up with
a formulaic way to determine which teams are susceptible to the treatment effect, e.g.
teams with mostly young and unmarried demographic, and run a separate analysis for this group. It's also noteworthy that the model was teetering on weak statistical significance
before adding in 2008-2009 data, and the results could change slightly as we include more historical
data (either more recent or more past).

\end{document}

### 2017-03-05 clean_spread.r

@@ -21,7 +21,7 @@ source('functions.r')
 ### Load Data
 ##############################
 
-files <- list.files('raw_data', pattern = '[[:digit:]]{4}.txt', full.names = T)
+files <- list.files('raw_data', pattern = '[[:digit:]]{4}.txt', full.names = T)[-(1:3)]
 data <- lapply(files, collectData)
 data <- rbindlist(data)
 

### 2017-03-05 explore.r

@@ -1,14 +1,21 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Exploratory Data Analysis
+###
+### Andreas Santucci
+###
+### Date: March 2017
+###
+### Inputs: 'tmp_data/spreads_wdist.RData', 'tmp_data/standings.RData'
+###
+################################################################################
+################################################################################
 
-## require(RSelenium)
-## # https://cran.r-project.org/web/packages/RSelenium/vignettes/RSelenium-basics.html
-## # andreas@asantucci-MBP:~/Downloads$ java -jar selenium-server-standalone-2.53.1.jar 
-## remDr <- remoteDriver(port = 4444)
-## remDr$open()
-## remDr$navigate('http://www.goldsheet.com/histnba.php')
-
-### To do:
-### 1. is the team a playoff contender?
-### 2. Fix the way year is assigned to date variable.
+##################################################
+### Set up Workspace, load data
+##################################################
 
 require(data.table)
 require(ggmap)
@@ -26,9 +33,12 @@ standings <- subset(standings, select = c('team', 'season', 'pct'))
 
 data <- merge(data, standings, by = c('season', 'team'), all.x = T)
 
-##############################
-## What does it mean to party?
-##############################
+### Remove playoff games (sketchy)
+data <- data[data[, head(.I, 82), by = list(season, team)]$V1]
+
+##################################################
+### Define a Party Variable
+##################################################
 
 party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
 party.rgx <- paste0('(', party.cities, ')', collapse = "|")
@@ -39,21 +49,50 @@ data[, weekend := grepl("(Friday)|(Saturday)|(Sunday)", day) %>% as.integer]
 ### We (1) exclude people who already party, (2) check if last game played against party city
 ### (3) check last game was played 'away' i.e. at party city and (4) party last night
 data[!grepl(party.rgx, team) & grepl(party.rgx, last.game.loc) & ndays.lgame == 1, party := 1]
-## data[grepl(party.rgx, last.opp) & last.loc == 'V' & last.game == 1 & bad.team == 1, party := 1]
+
[…]

### 2017-03-05 functions.r

@@ -7,8 +7,7 @@ Fetch <- function(team, beg, end, data, year) {
         strsplit('[[:blank:]]{2,}')
     datum <- Filter(function(x) !any(grepl("CAN\\.", x)), datum)
     datum <- do.call(rbind, args = datum) %>%
-        data.table(.) %>% cbind(team) # %>% cbind(year)
-#    datum <- data.table(datum)
+        data.table(.) %>% cbind(team)
 }
 
 
@@ -37,3 +36,6 @@ collectData <- function(datafile) {
     setcolorder(data, c('season', 'date', 'team', 'opponent', 'outcome', 'spread', 'score', 'location', 'ou'))
     return(data)
 }
+
+getFullTeamName <- function(needle, haystack)
+    grep(needle, haystack, value = T)

### 2017-03-05 model.r

@@ -1,4 +1,17 @@
-
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Model
+###
+### Andreas Santucci
+###
+### Date: March 2017
+###
+### Inputs: 'tmp_data/spreads_with_standings.RData'
+###
+################################################################################
+################################################################################
 
 require(data.table)
 require(magrittr)
@@ -11,8 +24,6 @@ table(data[ndays.lgame == 1, outcome]) # Win 49% of the games they play the day 
 
 data[, pq := cut(pct, breaks = quantile(pct)), by = season]
 data[, hp := ifelse(pq == '(0.244,0.378]' & party == 1, 1, 0)]
-
-standings[order(season, pct)]
 
 party.teams <- c('charlotte bobcats',   'chicago bulls',      'denver nuggets',
                  'new orleans hornets', 'philadelphia 76ers', 'houston rockets')

### 2017-03-05 scrape_standings.r

@@ -23,7 +23,7 @@ scrapeESPNStandings <- function(year) {
     return(data)
 }
 
-standings <- lapply(2010:2013, scrapeESPNStandings)
+standings <- lapply(2005:2013, scrapeESPNStandings)
 standings <- rbindlist(standings)
 standings[, season := gsub('NBA Standings - ', '', season)]
 

### 2017-03-28 scrape_rosters.r
Init commit. Script to scrape rosters for each team. Works.

@@ -1,0 +1,78 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Scrape NBA Rosters
+###
+### Andreas Santucci
+###
+### Date: March 2017
+###
+### Inputs: None
+###
+### Dependencies: data.table, magrittr, parallel, XML
+###
+################################################################################
+################################################################################
+
+##############################
+### Set up Workspace
+##############################
+require(parallel)
+
+cl <- makeCluster(detectCores())
+clusterCall(cl, function()  {
+    require(data.table)
+    require(magrittr)
+    require(XML)
+})
+
+load(file = 'tmp_data/spreads.RData')
+full.names <- data[, unique(team)]
+
+##############################
+### Create a listing of team abbreviations
+##############################
+
+### Create a data.frame of team-names and team-abbreviations
+### (as used on bball-reference.com)
+doc   <- htmlParse('http://www.basketball-reference.com/teams/')
+links <- xpathSApply(doc, '//a/@href')
+abbrs <- grep('/teams/[A-Z]{3}/$', x = links, value = T) %>%
+    gsub('/teams/([A-Z]{3})/', '\\1', .) %>%
+    `[`(1:30) # Grab 30 active teams.
+
+teams <- data.table(team = full.names, abbr = abbrs)
+
+getRoster <- function(team, link = 'http://www.basketball-reference.com/teams/') {
+    pg <- paste0(link, team)
+    dc <- htmlParse(pg)
+    lk <- xpathSApply(dc, '//a/@href')
+    ### For now, just grab first ten years of roster data.
+    lk <- grep('[A-Z]{3}/[0-9]{4}.html', lk, value = T) %>% unique %>% `[`(1:10)
+    roster <- lapply(lk, function(l) {
+        yr <- gsub('/teams/[A-Z]{3}/([0-9]{4}).html$', '\\1', l)
+        dt <- readHTMLTable(paste0('http://www.basketball-reference.com', l), which = 1)
+        setDT(dt)
+        dt[, `:=`(team = team, year = yr)]
+        setnames(dt, gsub(' ', '', tolower(names(dt))))
+        return(dt)
+    })
+    roster <- rbindlist(roster)
+    return(roster)
+}
+
+##############################
+### Scrape data
+##############################
+
+roster <- parLapplyLB(cl, teams$abbr, getRoster)
[…]

### 2017-03-28 eda_report.rnw
Added getAvgAge function, and incorporated log of avg.age into the model.

@@
 
-data[, list(meet.spread = mean(outcome == 'W'), .N), by = list(team, party)][order(team, party)]
+tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), by = list(team, party)][order(team, party)]
+
+load(file = 'tmp_data/roster.RData')
+roster[, season := gsub('-', '_', season)]
+setkey(roster, team, season)
+getAvgAge <- function(t, s, game.date, roster) {
+    roster[.(t, s), mean(game.date - birthdate)]
+}
+data[, avg.age := getAvgAge(team, season, date, roster), by = list(team, season, date)]
+data[, avg.age := as.numeric(avg.age) %>% log]
+
+tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
+            by = list(avg.age = round(avg.age), party, season)][order(avg.age, season, party)]
+tmp[order(avg.age, party)]
+## ages[, year := as.numeric(year)]
+## data[, year := year(date)] # Note quite right.
+## data <- merge(data, ages, by = c('team', 'year'), all.x = T)
+
 @
@@
 load(file = 'tmp_data/spreads_with_standings.RData')
-model <- glm(outcome == 'W' ~ party + weekend + ndays.lgame + pct +
+model <- glm(outcome == 'W' ~ party + avg.age*party + weekend + ndays.lgame + pct +
                  travel.dist + team + season + team:season,

### 2017-03-29 eda_report.rnw
Added citation. Now including Boston and not Miami as Party City. Examining effect for bottom 50% rank from last season.

@@
 \subsection*{References}
+``Balancing, Regression, Difference-In-Differences and Synthetic Control Methods: A Synthesis'', by Doudchenko, Imbens. Oct 2016. \href{http://www.nber.org/papers/w22791}{paper} \\
 ``The Benefits of College Athletic Success: An Application of the Propensity Score Design'', by Michael L.
@@
 ### Define a Party Variable
-party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
+#party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
+party.cities <- c('boston', 'brooklyn', 'los angeles', 'new york')
 party.rgx <- paste0('(', party.cities, ')', collapse = "|")
@@
      xlab = "Proportion of games for which a team-year meets the spread")
+
 @
@@
     hist(main = "Number of party days by season-team", xlab = "Number of days", ylab = "Frequency")
+@
+
+\subsection*{Which teams to include as party cities?}
+<<echo=-1>>=
+
+tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
+     by = list(last.game.loc, ndays.lgame)][order(last.game.loc, ndays.lgame)]
+
+tmp[ndays.lgame == 1][order(meet.spread)]
+
+@
+
+<<results='hide'>>=
+
+### What about by game location
+ggplot(tmp, aes(x = ndays.lgame, y = meet.spread, label = N)) +
+    geom_point() + 
+    geom_label(nudge_y = 0.2) +
+    geom_hline(yintercept = 0.5) + 
+    facet_wrap(facets = ~ last.game.loc)
+
+@
+
+\subsection*{Does last-season's ranking have an implication?}
+
+<<>>=
+tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
+            by = list(pct, party)][order(pct)]
+ggplot(tmp, aes(x = pct, y = meet.spread, color = as.factor(party), label = N)) + 
+    geom_label()
 @
@@
 
-<<>>=
+<<echo=F>>=
 
@@
 tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
-            by = list(avg.age = round(avg.age), party, season)][order(avg.age, season, party)]
-tmp[order(avg.age, party)]
+            by = list(avg.age = round(avg.age, 2), party)][
+    order(avg.age, party)] %>% na.omit
+
+ggplot(tmp, aes(x = avg.age, y = meet.spread, 
+                color = party %>% as.factor, label = N)) + 
+    geom_point() +
+    geom_label() +
+    geom_smooth()
 ## ages[, year := as.numeric(year)]
@@
 <<>>=
-load(file = 'tmp_data/spreads_with_standings.RData')
+#load(file = 'tmp_data/spreads_with_standings.RData')
 model <- glm(outcome == 'W' ~ party + avg.age*party + weekend + ndays.lgame + pct +
-                 travel.dist + team + season + team:season,
-             family = 'binomial', data = data)
+                 travel.dist + season,
+             family = 'binomial', data = data, subset = pct < 0.5)
 

### 2017-04-02 init_model.rnw
p-val 0.02 on party. Sent a model to Charlie for feedback.

\documentclass[12pt]{article}

\usepackage{algorithm2e} 
\usepackage{amsmath} 
\usepackage{amsthm} 
\usepackage{amsfonts} 
\usepackage{bbm} 
\usepackage{color,soul} 
\usepackage{framed} 
\usepackage[margin=0.5in]{geometry} 
\usepackage{hyperref} 
\usepackage{mathtools} 
\usepackage[dvipsnames]{xcolor}

\newtheorem{theorem}{Theorem}[section] 
\newtheorem{lemma}[theorem]{Lemma} 
\newtheorem{proposition}[theorem]{Proposition} 
\newtheorem{corollary}[theorem]{Corollary}  
\newcommand{\D}{\mathrm{d}}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil} 
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\SetKwInput{KwInput}{Input} 
\SetKwInput{KwOutput}{Output}

\begin{document}

\title{\textbf{NBA Party}}
\author{Andreas Santucci, Eric Lax}
\date{April 2017}
\maketitle

<<echo=F,results='hide'>>=
require(data.table)
require(ggplot2)
load(file = 'tmp_data/covers_lines.RData')
lines[, `:=`(pct = NULL, weekend = NULL)]
others <- grep("party", names(lines), value = T, invert = T)
setcolorder(lines, c(others, c('party')))
@

\subsection*{Data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house. To this data, we've added features such as 
the number of days since last game, the last game location, the travel distance,
average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}, and a party indicator.

<<>>= 
head(lines)
@ 
Our data range from 2010-2011 season through present, which includes part of the 2016-2017 season.

\paragraph{What does it mean to party?}

\begin{align*}   
  \texttt{party} = \begin{cases}     
    1 &\textrm{ if } \textrm{ team played in Brooklyn, Los Angeles, or New York yesterday} \\ 
    0 &\textrm{ otherwise}   \end{cases} 
\end{align*}

\subsection*{A significant first model}

<<>>=

model <- glm(outcome == 'W' ~ party + ndays.lgame + I(log(travel.dist+1)) + season,
             data = lines, family = 'binomial', subset = season < 2016)

summary(model)

@

The magnitude of our \texttt{party} variable is both practically and 
statistically significant.

\subsection*{A model worth betting on}
We mention that we don't have very many games to bet on, but that we do make a profit in expectation.

<<>>=
### We're only interested in making predictions when teams party.
yhat <- predict(model, newdata = lines[season >= 2016 & party == 1], type = 'response')
yhat <- ifelse(yhat < 0.41325, 0, 1) # We use a simple threshold based on hist(yhat).
table(yhat, lines[season >= 2016 & party == 1, outcome])
@

\subsection*{Where to go from here?}
There are several approaches.

\paragraph{Refining the definition of party, Estimating travel schedules} Right now, we define our party variable
crudely, that is, we use whether the team played in a party city the day before
as a proxy for whether the team spent the night in a party city the day before the 
game. Realize that we could further identify our effect if we could more accurately
distinguish when treatment is applied.

One idea is to utilize the time each game is played alongside a sample flight schedule
between stadiums. In the simplest example, if we know that a team plays a game on the one coast and a game on the opposing coast the following day, it may be possible to 
infer with reasonable accuracy which day the team \emph{actually} traveled. We can
then use this logical prediction (on where the team spent the night) to define
our \texttt{party} variable accordingly.

The general idea is that we either hope to add a few more treatment observations
to our data, or refine the fidelity of existing observations, so that we may identify
a more concentrated effect.

\paragraph{Focusing on player demographics} We remark that we do not have
player age in the model, even though we did take time to incorporate this into our data. Including an interaction between \texttt{party} and \texttt{avg.age} introduces
colinearity in our model and robs us of statistical significance. However, there
is both intuition and mild evidence that younger teams are affected more by our treatment effect. We also hypothesize that the effect is stronger for unmarried players.
With this in mind, we may consider exploring in more detail the demographics of our teams. We detail a few approaches.

One idea is to create an \texttt{average age} variable for each observation which, instead of simply taking the average across all players on the roster, uses a weighted average based on which players are estimated to be starters and which are on the bench (this can be determined either from the ``depth chart'' or from an estimate based on the current season's historical minutes). The problem here is that further refining our age variable isn't going to help if it's colinear with \texttt{party}.

Another idea is to look at the proportion of players on a team which are unmarried. Our intuition is that younger, unmarried players will tend to exhibit a stronger treatment effect. If adding
carefully constructed demographic features helps our model predict team-performance, this may motivate us to further consider individual bets. That is, we can build a more granular model which
examines the performance of each player each game. This has the added benefit that with 5 players on each team each game, we get a 10x increase in the number of bets we can take.
This is helpful since right now, although our margins are high, our profits limited due to limited number of bets we can take.

\paragraph{Knowing our own data better}
We remark that there are still unanswered questions in our existing data.
For example, we tried scraping the team's performance in the season prior and using this as a feature to predict whether the team will meet the spread. Although the magnitude of the coefficient wasn't practically significant, it was still extremely statistically significant in our model. This is surprising, since we expect the House to take advantage of such an obvious feature when constructing their model.
This opens up the question: can we back out from our black-box how Vegas builds their model?

\end{document}

### 2017-04-06 init_model.rnw
Stratified random sampling to estimate last-location fixed effects. These are then used as features in a second stage model, which achieves very significant results.

@@
 load(file = 'tmp_data/covers_lines.RData')
+load(file = 'tmp_data/ndrink_estabs.RData')
+drinks[, county := NULL]
+
 lines[, `:=`(pct = NULL, weekend = NULL)]
-others <- grep("party", names(lines), value = T, invert = T)
-setcolorder(lines, c(others, c('party')))
+
+### We're careful to merge in ndrinking_estabs according to the PREV game played.
+lines <- merge(lines, drinks, 
+               by.x = c('last.game.loc', 'season'), 
+               by.y = c('team',          'season'))
+setkey(lines, date)
+others <- grep("(party)|(last.game)", names(lines), value = T, invert = T)
+setcolorder(lines, c(others, c('last.game.loc', 'party')))
+
+require(sampling)
+set.seed(04052017)
+setkey(lines, season, last.game.loc)
+
+n = length(unique(lines$last.game.loc))
+m = length(unique(lines$season))
+s = strata(data = lines, stratanames = c('season', 'last.game.loc'),
+           size = rep(5, n*m), method = 'srswor')
+
+var.creation.indices = s$ID_unit
+var.creation = lines[var.creation.indices]
+training     = lines[-var.creation.indices]
+
+m = glm(outcome == 'W' ~ 0 + last.game.loc, data = var.creation, family = 'binomial')
+coef <- m$coef
+names(coef) <- gsub('last.game.loc', '', names(coef))
+coef = data.table(last.game.loc = names(coef), coef)
+
+setkey(training, last.game.loc)
+training = training[coef]
+
+m = glm(outcome == 'W' ~ I(coef*(ndays.lgame==1)), data = training, family = 'binomial', subset = season < 2017) # Significant.
+
+### Not quite worth betting on, yet.
+yhat <- predict(m, newdata = training[season >= 2017 & ndays.lgame == 1], type = 'response')
+yhat <- ifelse(yhat < 0.519, 0, 1)
+table(yhat, training[season >= 2017 & ndays.lgame==1, outcome])
+
 @
@@
 
-model <- glm(outcome == 'W' ~ party + ndays.lgame + I(log(travel.dist+1)) + season,
+model <- glm(outcome == 'W' ~ party + ndays.lgame + I(log(travel.dist+1)) + 
+                 season + nmusicians,
              data = lines, family = 'binomial', subset = season < 2016)
@@
 <<>>=
-### We're only interested in making predictions when teams party.
+### We're only interested in making predictions when teams party. When they do, we bet against them.
 yhat <- predict(model, newdata = lines[season >= 2016 & party == 1], type = 'response')

### 2017-04-07 init_model.rnw
After meeting with Eric at Arillaga.

@@
 <<echo=F,results='hide'>>=
+
+### To Do: estimate coef by last.game.loc but exclude obs if last.game.loc == team
+### i.e. just like in estimating party coef, we don't want to estimate effect of last 
+### playing in boston if the team itself is boston.
+### 
 require(data.table)
@@
 
-### We're careful to merge in ndrinking_estabs according to the PREV game played.
-lines <- merge(lines, drinks, 
+inner.merge = merge(lines, drinks, 
                     by.x = c('last.game.loc', 'season'),
                     by.y = c('team',          'season'))
-setkey(lines, date)
-others <- grep("(party)|(last.game)", names(lines), value = T, invert = T)
-setcolorder(lines, c(others, c('last.game.loc', 'party')))
+setkey(lines, season, team, date)
+setkey(inner.merge, season, team, date)
+outter = lines[!inner.merge]
+
+### We're careful to merge in ndrinking_estabs according to the PREV game played.
+## lines <- merge(lines, drinks, 
+##                by.x = c('last.game.loc', 'season'), 
+##                by.y = c('team',          'season'), all.x = T) ## Toronto raptors?
+## setkey(lines, date)
+## others <- grep("(party)|(last.game)", names(lines), value = T, invert = T)
+## setcolorder(lines, c(others, c('last.game.loc', 'party')))
 
@@
 set.seed(04052017)
-setkey(lines, season, last.game.loc)
+setkey(lines, season, last.game.loc, date)
 
@@
 
-m = glm(outcome == 'W' ~ 0 + last.game.loc, data = var.creation, family = 'binomial')
+m = glm(outcome == 'W' ~ 0 + last.game.loc, data = var.creation, family = 'binomial',
+        subset = last.game.loc != team)
 coef <- m$coef
@@
 
-m = glm(outcome == 'W' ~ I(coef*(ndays.lgame==1)), data = training, family = 'binomial', subset = season < 2017) # Significant.
+m = glm(outcome == 'W' ~ 0 + coef + I(coef*(ndays.lgame==1)) + ndays.lgame + travel.dist,
+        data = training, family = 'binomial', subset = season < 2017) # Significant.
+
+synth  = cbind(coef, ndays.lgame = 1)
+actual = lines[ndays.lgame == 1, list(meet.spread = mean(outcome == 'W')), keyby = last.game.loc]
+preds  = predict(m, newdata = synth, type = 'response')
+actual$preds = preds
+require(ggplot2)
+ggplot(actual, aes(x = meet.spread, y = preds, label = last.game.loc)) + 
+    geom_point()+
+    geom_label()
+plot(actual$meet.spread, preds)
 

### 2017-04-07 init_model.rnw
Using nmusicians interacted with ndays.lgame == 1, we get a significant model (p = 0.0014) and a modest success rate on hold-out data (55%).

--- 2017-04-02 init_model.rnw
@@
 <<echo=F,results='hide'>>=
+
+### Reshape data
+### One model, outcome ~ loast.game.llocation, last.city.size, last.city.region,
+
+### Group data by region.
+
 require(data.table)
 require(ggplot2)
+require(magrittr)
+require(sampling)
+set.seed(4052017)
+
 load(file = 'tmp_data/covers_lines.RData')
+load(file = 'tmp_data/ndrink_estabs.RData')
+
+lines <- merge(lines, drinks, 
+               by.x = c('season', 'last.game.loc'),
+               by.y = c('season', 'team'), all.x = T)
+
+m <- glm(outcome == 'W' ~ I(nmusicians*(ndays.lgame==1)), 
+         data = lines[season < 2017 & last.game.loc != team])
+p <- predict(m, newdata = lines[season == 2017 & ndays.lgame == 1 & last.game.loc != team], type = 'response')
+table(round(p), lines[season == 2017 & ndays.lgame == 1 & last.game.loc != team, outcome])
+
+## lines = lines[team != last.game.loc]
 lines[, `:=`(pct = NULL, weekend = NULL)]
-others <- grep("party", names(lines), value = T, invert = T)
-setcolorder(lines, c(others, c('party')))
+
+melted = melt(lines, id.vars = c('season', 'date', 'location'))
+setkey(melted, season, date, location)
+melted = melted[variable == 'team']
+casted = dcast(melted, formula = season + date + location ~ .)
+
+valid.idx = which(lines$season == 2017) %>% sample(., size = floor(length(.) / 12))
+valid = lines[valid.idx]
+lines = lines[-valid.idx]
+
+n = length(unique(lines$last.game.loc))
+m = length(unique(lines$season))
+setkey(lines, season, last.game.loc, date)
+
+### Stratified random sampling.
+s = strata(data = lines, stratanames = c('season', 'last.game.loc'),
+           size = rep(7, n*m), method = 'srswor')
+
+### Partition data for two-stage least-squares.
+var.creation.indices = s$ID_unit
+var.creation = lines[var.creation.indices]
+training     = lines[-var.creation.indices]
+
+### First stage model.
+coefs <- sapply(1:60, function(x) {
+    indices = sample(1:nrow(var.creation), size = nrow(var.creation), replace = T)
+    glm(outcome == 'W' ~ 0 + last.game.loc, 
+        data = var.creation[indices], family = 'binomial')$coef
+})
+
+rownames(coefs) <- gsub('last.game.loc', '', rownames(coefs))
+coef = data.table(last.game.loc = rownames(coefs), coef = rowMeans(coefs))
+
+## m = glm(outcome == 'W' ~ 0 + last.game.loc, 
+##         data = var.creation[last.game.loc != team], 
+##         family = 'binomial')
+## coef <- m$coef
+## names(coef) <- gsub('last.game.loc', '', names(coef))
+## coef = data.table(last.game.loc = names(coef), coef)
+
+setkey(training, last.game.loc)
+training = training[coef]
+
+### Second stage model.
+m = glm(outcome == 'W' ~ 1 + I(coef*(ndays.lgame==1)),
+        data = training[last.game.loc != team], 
+        family = 'binomial', subset = season < 2017) # Significant.
+
+### Meet the spread on 2017 data vs. prediction.
+setkey(valid, last.game.loc)
+valid = valid[coef]
+preds = predict(m, newdata = valid[ndays.lgame == 1], type = 'response')
+table(round(preds), valid[ndays.lgame == 1, outcome])
+
+## ggplot(actual, aes(x = meet.spread, y = preds, label = last.game.loc)) + 
+##     geom_point()+
+##     geom_label()
+
+### Not quite worth betting on, yet.
+## yhat <- predict(m, newdata = training[season >= 2017 & ndays.lgame == 1], type = 'response') %>% round
+## table(yhat, training[season >= 2017 & ndays.lgame==1, outcome])
+
 @

### 2017-04-09 init_model.rnw
Rough draft of an updated model - outcome ~ (ndays.lgame==1)*nmusicians is highly significant and bets well on holdout data.

--- 2017-04-06 init_model.rnw
@@
 <<echo=F,results='hide'>>=
+
 require(data.table)
 require(ggplot2)
+require(magrittr)
+require(sampling)
+set.seed(4052017)
+
 load(file = 'tmp_data/covers_lines.RData')
-load(file = 'tmp_data/ndrink_estabs.RData')
-drinks[, county := NULL]
-
-lines[, `:=`(pct = NULL, weekend = NULL)]
-
-### We're careful to merge in ndrinking_estabs according to the PREV game played.
-lines <- merge(lines, drinks, 
-               by.x = c('last.game.loc', 'season'), 
-               by.y = c('team',          'season'))
-setkey(lines, date)
-others <- grep("(party)|(last.game)", names(lines), value = T, invert = T)
-setcolorder(lines, c(others, c('last.game.loc', 'party')))
-
-require(sampling)
-set.seed(04052017)
-setkey(lines, season, last.game.loc)
-
-n = length(unique(lines$last.game.loc))
-m = length(unique(lines$season))
-s = strata(data = lines, stratanames = c('season', 'last.game.loc'),
-           size = rep(5, n*m), method = 'srswor')
-
-var.creation.indices = s$ID_unit
-var.creation = lines[var.creation.indices]
-training     = lines[-var.creation.indices]
-
-m = glm(outcome == 'W' ~ 0 + last.game.loc, data = var.creation, family = 'binomial')
-coef <- m$coef
-names(coef) <- gsub('last.game.loc', '', names(coef))
-coef = data.table(last.game.loc = names(coef), coef)
-
-setkey(training, last.game.loc)
-training = training[coef]
-
-m = glm(outcome == 'W' ~ I(coef*(ndays.lgame==1)), data = training, family = 'binomial', subset = season < 2017) # Significant.
-
-### Not quite worth betting on, yet.
-yhat <- predict(m, newdata = training[season >= 2017 & ndays.lgame == 1], type = 'response')
-yhat <- ifelse(yhat < 0.519, 0, 1)
-table(yhat, training[season >= 2017 & ndays.lgame==1, outcome])
+load(file = 'tmp_data/nmusician_estabs.RData')
 
@@
 
-\subsection*{Data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
+\subsection*{Data}
+
+\paragraph{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
 To this data, we've added features such as
-the number of days since last game, the last game location, the travel distance,
-average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}, and a party indicator.
+the number of days since last game, the last game location, the travel distance, and
+average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}.
+
+\paragraph{BLS data} We also obtained data from
+\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of
+establishments by business type at the county level. As a proxy for how much night-life
+there is in a city, we look toward the number of sound recording studios, musical groups,
+and music publishers there are. For each year and each county, we simply calculate
+the average number of establishments across all three music categories listed above.
+
+We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data
+for this team.
 
 <<>>=
-head(lines)
+musicians[, list(team, season, nmusicians)]
 @
+
+We then merge this in with our lines data, taking care to do so according to the
+last game's location. We also lag our BLS data by one year.
+
+<<>>=
+### Merge lines data with BLS data.
+lines <- merge(lines, musicians,
+               by.x = c('season', 'last.game.loc'),
+               by.y = c('season', 'team'), all.x = T)
+
+setkey(lines, season, team, date)
+lines[, list(date, team, opponent, location, 
+             line, score, outcome, ndays.lgame, 
+             travel.dist, avg.age, nmusicians)]
+
+@
+
 Our data range from 2010-2011 season through present, which includes part of the 2016-2017 season.
 
-\paragraph{What does it mean to party?}
+\paragraph{What cities have nightlife?}
 
-\begin{align*}   
-  \texttt{party} = \begin{cases}     
-    1 &\textrm{ if } \textrm{ team played in Brooklyn, Los Angeles, or New York yesterday} \\ 
-    0 &\textrm{ otherwise}   \end{cases} 
-\end{align*}
+We examine which cities have a notable nightlife, as measured by our proxy.
 
-\subsection*{A significant first model}
+<<>>= 
+musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)), by = team]
+@
+
+Los Angeles stands out, as does New York. Some runner ups include Chicago and Miami.
+
+\subsection*{A significant first model} We keep things simple, and try to explain the likelihood
+of meeting the spread as a function of whether a team visited a city with an active night-life
+the day before.
 
@@
 
-model <- glm(outcome == 'W' ~ party + ndays.lgame + I(log(travel.dist+1)) + 
-                 season + nmusicians,
-             data = lines, family = 'binomial', subset = season < 2016)
+m <- glm(outcome == 'W' ~ I(nmusicians*(ndays.lgame==1)) + ndays.lgame + travel.dist + avg.age, 
+         data = lines[season < 2017 & last.game.loc != team])
 
-summary(model)
+summary(m)
 
@@
 
-The magnitude of our \texttt{party} variable is both practically and 
+The magnitude of our ``\texttt{party}'' variable is both practically and 
 statistically significant.
@@
 \subsection*{A model worth betting on}
-We mention that we don't have very many games to bet on, but that we do make a profit in expectation.
+If we re-train an even simpler model, regressing on only the variable of interest, 
+we get great betting performance.
 
 <<>>=
-### We're only interested in making predictions when teams party. When they do, we bet against them.
-yhat <- predict(model, newdata = lines[season >= 2016 & party == 1], type = 'response')
-yhat <- ifelse(yhat < 0.41325, 0, 1) # We use a simple threshold based on hist(yhat).
-table(yhat, lines[season >= 2016 & party == 1, outcome])
+m <- glm(outcome == 'W' ~ I(nmusicians*(ndays.lgame==1)), 
+         data = lines[season < 2017 & last.game.loc != team])
+
+### We're only interested in making predictions when teams party.
+p <- predict(m, newdata = lines[season == 2017 & 
+                                ndays.lgame == 1 & 
+                                last.game.loc != team], type = 'response')
+table(round(p), lines[season == 2017 & ndays.lgame == 1 & last.game.loc != team, outcome])
+
 @
 
-\subsection*{Where to go from here?}
-There are several approaches.
-
-\paragraph{Refining the definition of party, Estimating travel schedules} Right now, we define our party variable
-crudely, that is, we use whether the team played in a party city the day before
-as a proxy for whether the team spent the night in a party city the day before the 
-game. Realize that we could further identify our effect if we could more accurately
-distinguish when treatment is applied.
-
-One idea is to utilize the time each game is played alongside a sample flight schedule
-between stadiums. In the simplest example, if we know that a team plays a game on the one coast and a game on the opposing coast the following day, it may be possible to 
-infer with reasonable accuracy which day the team \emph{actually} traveled. We can
-then use this logical prediction (on where the team spent the night) to define
-our \texttt{party} variable accordingly.
-
-The general idea is that we either hope to add a few more treatment observations
-to our data, or refine the fidelity of existing observations, so that we may identify
-a more concentrated effect.
-
-\paragraph{Focusing on player demographics} We remark that we do not have
-player age in the model, even though we did take time to incorporate this into our data. Including an interaction between \texttt{party} and \texttt{avg.age} introduces
-colinearity in our model and robs us of statistical significance. However, there
-is both intuition and mild evidence that younger teams are affected more by our treatment effect. We also hypothesize that the effect is stronger for unmarried players.
-With this in mind, we may consider exploring in more detail the demographics of our teams. We detail a few approaches.
-
-One idea is to create an \texttt{average age} variable for each observation which, instead of simply taking the average across all players on the roster, uses a weighted average based on which players are estimated to be starters and which are on the bench (this can be determined either from the ``depth chart'' or from an estimate based on the current season's historical minutes).
-The problem here is that further refining our age variable isn't going to help if it's colinear with \texttt{party}.
-
-Another idea is to look at the proportion of players on a team which are unmarried. Our intuition is that younger, unmarried players will tend to exhibit a stronger treatment effect. If adding
-carefully constructed demographic features helps our model predict team-performance, this may motivate us to further consider individual bets. That is, we can build a more granular model which
-examines the performance of each player each game. This has the added benefit that with 5 players on each team each game, we get a 10x increase in the number of bets we can take.
-This is helpful since right now, although our margins are high, our profits limited due to limited number of bets we can take.
-
-\paragraph{Knowing our own data better}
-We remark that there are still unanswered questions in our existing data.
-For example, we tried scraping the team's performance in the season prior and using this as a feature to predict whether the team will meet the spread.
-Although the magnitude of the coefficient wasn't practically significant, it was still extremely statistically significant in our model.
-This is surprising, since we expect the House to take advantage of such an obvious feature when constructing their model.
-This opens up the question: can we back out from our black-box how Vegas builds their model?
-
 \end{document}

### 2017-04-09 scrape_covers.r
Added error handling code.

@@ -29,7 +29,9 @@ clusterCall(cl, function() {
 })
 
 load(file = 'tmp_data/game_days.RData')
-game.days <- Filter(function(x) grepl("^201[^0]", x), game.days) 
+game.days <- Filter(function(x) x < Sys.Date() & x > '2010-11-25', game.days)
+## game.days <- Filter(function(x) grepl("^201[^0]", x), game.days)
+## game.days <- Filter(function(x) x < Sys.Date(), game.days)
 game.days <- gsub('-0([0-9])-', '-\\1-', game.days)
 
 ##############################
@@ -57,7 +59,8 @@ scrapeGame <- function(link) {
     })
 }
 
-scrapeCoverLines <- function(date, rescrape = F, save.path = 'raw_data/covers/') {
+scrapeCoverLines <- function(date, rescrape = F, save.path = 'raw_data/covers/', ntries = 0) {
+    if (ntries > 5) return(NULL)
     if (!rescrape && file.exists(paste0(save.path, date, '.csv')))
         return(NULL)
     cat(paste0("Now trying to download covers for date: ", date, "\n"))
@@ -83,7 +86,7 @@ scrapeCoverLines <- function(date, rescrape = F, save.path = 'raw_data/covers/')
     }, error = function(e) {
         Sys.sleep(1)
         cat(paste0('Error when downloading data for date: ', date, '\n'))
-        scrapeCoverLines(date)
+        scrapeCoverLines(date, ntries = ntries + 1)
     })
 }
 
@@ -95,4 +98,4 @@ scrapeCoverLines <- function(date, rescrape = F, save.path = 'raw_data/covers/')
 ### an individual core will need to re-call this function.
 clusterExport(cl, c('scrapeGame', 'scrapeCoverLines'))
 lines <- parLapplyLB(cl, game.days, scrapeCoverLines)
-
+#lapply(game.days, scrapeCoverLines)

### 2017-04-09 clean_bls.r
Updated variable names to be consistent.

@@ -15,6 +15,9 @@
 ################################################################################
 ################################################################################
 
+### Notes and data-sources.
+  # http://www.nciaa.com/content.aspx?page_id=22&club_id=160641&module_id=29898
+
 ##################################################
 ### Set up Workspace
 ##################################################
@@ -36,8 +39,10 @@ clusterCall(cl, function() {
 ### Subset BLS Data into manageable chunks.
 ##############################
 
-SubsetBLS <- function(year, raw.path = 'raw_data/bls/', save.path = 'tmp_data/bls/',
-                      party = c('wine', 'spirit', 'beer', 'liquor', 'drinking')) {
+party <- c('sound recording', 'music publisher', 'musical group')
+
+SubsetBLS <- function(year, party.regex,
+                      raw.path = 'raw_data/bls/', save.path = 'tmp_data/bls/') {
     ### Create a listing of files for all relevant counties, load data.
     foldr <- list.files(path = raw.path, pattern = paste0(year, '.q1'), full.names = T)
     files <- list.files(path = foldr,
@@ -51,16 +56,16 @@ SubsetBLS <- function(year, raw.path = 'raw_data/bls/', save.path = 'tmp_data/bl
     ### We will take our own total later, so for now we take granular data.
     bls <- bls[grep("County, NAICS 6-digit", agglvl.title)]
     ### Create regex to subset to drinking establshimensts.
-    party <- paste0('(\\<', party, '\\>)', collapse = '|')
-    bls <- bls[grep(party, industry.title, ignore.case = T)]    
-    write.csv(x = bls, file = paste0(save.path, year, '.csv'))
+    party.regex <- paste0('(', party.regex, ')', collapse = '|')
+    bls <- bls[grep(party.regex, industry.title, ignore.case = T)]    
+    write.csv(x = bls, file = paste0(save.path, year, '.csv'), row.names = F)
     rm(bls)
     gc()
     return(NULL)
 }
 
 years <- 2010:2016
-parLapplyLB(cl, years, SubsetBLS)
+parLapplyLB(cl, years, SubsetBLS, party.regex = party)
 
 ##################################################
 ### Merge in BLS data with corresponding NBA teams.
@@ -90,14 +95,19 @@ locs[, county := ifelse(is.na(administrative.area.level.2),
 locs <- data.table(team = teams, county = locs$county)
[…]

### 2017-04-09 init_model.rnw
Using continuous measure of party-city. Tidied up formatting. Sent a draft to Guido.

@@
 
-\subsection*{Data}
+\section*{Data Sources}
 
 \paragraph{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
-To this data, we've added features such as
-the number of days since last game, the last game location, the travel distance, and
-average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}.
+<<echo=F>>=
+lines[, list(date, team, opponent, line, score)]
+@
 
 \paragraph{BLS data} We also obtained data from
-\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of
-establishments by business type at the county level. As a proxy for how much night-life
-there is in a city, we look toward the number of sound recording studios, musical groups,
-and music publishers there are. For each year and each county, we simply calculate
-the average number of establishments across all three music categories listed above.
+\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the county-quarter level. \ul{As a 
+proxy for how much night-life there is in a city, we look toward the 
+number of sound recording studios, musical groups, and music publishers there 
+are for a particular county-year}: we simply calculate the average number of establishments across all three music categories listed above.\footnote{We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.}
 
-We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data
-for this team.
-
-<<>>=
-musicians[, list(team, season, nmusicians)]
+<<echo=F>>=
+musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
+musicians[, list(county, team, season, nmusicians = round(nmusicians, 1))]
 @
@@
 last game's location. We also lag our BLS data by one year.
+To this data, we've created features such as 
+the number of days since last game, the last game location, the travel distance, and
+average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}.
 
-<<>>=
+<<echo=2>>=
 ### Merge lines data with BLS data.
@@
 
-\paragraph{What cities have nightlife?}
+\subsection*{What cities have nightlife?}
 
@@
 <<>>= 
-musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)), by = team]
+musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)), by = team][order(average.nmusicians)]
 @
@@
 
-\subsection*{A significant first model} We keep things simple, and try to explain the likelihood
+\subsection*{A significant model} We keep things simple, and try to explain the likelihood
 of meeting the spread as a function of whether a team visited a city with an active night-life
@@
 
-<<>>=
+<<echo=-4>>= 
+lines[, party := ifelse(ndays.lgame == 1, nmusicians, 0)]
 
-m <- glm(outcome == 'W' ~ I(nmusicians*(ndays.lgame==1)) + ndays.lgame + travel.dist + avg.age, 
+m <- glm(outcome == 'W' ~ party + ndays.lgame + travel.dist + avg.age, 
          data = lines[season < 2017 & last.game.loc != team])
@@
 
-The magnitude of our ``\texttt{party}'' variable is both practically and 
+The sign of our ``\texttt{party}'' variable is intuitively correct, and 
 statistically significant.
@@
 \subsection*{A model worth betting on}
-If we re-train an even simpler model, regressing on only the variable of interest, 
-we get great betting performance.
+If we re-train an even simpler model, we obtain fantastic betting performance.
 
@@
 
+summary(m)$coef %>% round(., 5)
+
 ### We're only interested in making predictions when teams party.
-p <- predict(m, newdata = lines[season == 2017 & 
-                                ndays.lgame == 1 & 
-                                last.game.loc != team], type = 'response')
+p <- predict(m, newdata = lines[season == 2017 & ndays.lgame == 1 & last.game.loc != team], 
+             type = 'response')
 table(round(p), lines[season == 2017 & ndays.lgame == 1 & last.game.loc != team, outcome])

### 2017-04-09 clean_covers.r
Moved exploratory analysis code into explore_covers.r.

@@ -15,6 +15,18 @@
 ################################################################################
 ################################################################################
 
+### 1. Investigate party cities -- Increase quality of observations, be really confident they *stayed* in a party city overnight.
+###    Intuitively: you stayed in a city the night before where there is a lot of partying
+###    E.g. an East Coast team may fly into LA the day before and we may want to define that game to be a party city
+###    they may also fly out after an afternoon game and miss the party-effect completely.
+###    Data Sources: tweets
+###    Figure out: LA -- rest day -- Boston, we know they won't take the red-eye
+### 2. Weighting age by starting minutes
+### 3. Marital status
+
+### Bring game-time into data
+### Try and get a sense of what a travel schedule might be..
+
 ##################################################
 ### Set up Workspace, Read in Raw Data
 ##################################################
@@ -37,7 +49,7 @@ lines[, date := as.Date(date, format = '%A, %B %d %Y')]
 lines[, line := gsub('^(-?[0-9.]+)/.*$', '\\1', line)]
 lines[, line.ts := NULL]
 lines[, ou := NULL]
-lines[, game.time := NULL]
+## lines[, game.time := NULL]
 
 ##################################################
 ### Expand team abbreviations.
@@ -65,7 +77,7 @@ setnames(lines, 'team', 'away.team')
 ### Remove missing observations. Set line as numeric.
 ##################################################
 
-lines <- lines[line != 'OFF', list(date, home.team, away.team, home.pts, away.pts, line)]
+lines <- lines[line != 'OFF', list(date, game.time, home.team, away.team, home.pts, away.pts, line)]
 lines[, line := as.numeric(line)]
 
 ### Season spans the new year, i.e. starts in October ends in April.
@@ -85,7 +97,7 @@ lines[, outcome := ifelse(home.pts + line >= away.pts, 'W', 'L')]  # Ties?
 lines[, score := paste(home.pts, away.pts, sep = '-')]
 lines[, `:=`(home.pts = NULL, away.pts = NULL)]
 lines[, location := team]
-lines <- lines[, list(season, date, team, opponent, location, line, score, outcome)]
+lines <- lines[, list(season, date, game.time, team, opponent, location, line, score, outcome)]
 
 ##################################################
[…]

### 2017-04-11 scrape_population.r
Init commit. Scraping population by county-year.

@@ -1,0 +1,59 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Scraping Population Data
+###
+### Andreas Santucci
+###
+### Date: April 2017
+###
+### Inputs: 
+###
+### Dependencies: 
+###
+################################################################################
+################################################################################
+
+require(data.table)
+require(magrittr)
+require(RCurl)
+require(XML)
+
+### Download population estimates by county-year.
+pop <- fread('https://www2.census.gov/programs-surveys/popest/datasets/2010-2016/counties/totals/co-est2016-alldata.csv',
+             select = c('STNAME', 'CTYNAME', paste0('POPESTIMATE', 2010:2016)))
+setnames(pop, tolower(names(pop)))
+
+### Don't forget Washington D.C., which is technically a metro area.
+dc <- fread('https://www2.census.gov/programs-surveys/popest/datasets/2010-2016/metro/totals/csa-est2016-alldata.csv',
+            select = c('NAME', paste0('POPESTIMATE', 2010:2016)))
+Sys.setlocale('LC_ALL','C') 
+dc <- dc[grep("district", NAME, ignore.case = T)]
+dc[, stname := 'Washington DC']
+setnames(dc, c('ctyname', paste0('popestimate', 2010:2016), 'stname'))
+dc[, ctyname := 'District of Columbia']
+pop <- rbind(pop, dc)
+
+### Reshape our data.
+pop <- melt(pop, id.vars = c('stname', 'ctyname'), variable.name = 'year')
+pop[, year := gsub('popestimate', '', year) %>% as.integer]
+setnames(pop, 'value', 'population')
+
+setnames(pop, c('stname', 'ctyname'), c('state', 'county'))
+
+### We use lagged data.
+setnames(pop, 'year', 'season')
+pop[, season := season + 1]
+pop <- pop[state != county]
+pop[, county := paste(county, state, sep = ', ')]
+pop[, state := NULL]
+pop[county == 'District of Columbia, Washington DC', county := 'District of Columbia']
+
+save(pop, file = 'tmp_data/population_by_county_year.RData')
+
+### Old
+## fips <- readLines('https://www2.census.gov/geo/docs/reference/codes/files/national_county.txt')
+## fips <- strsplit(fips, split = ',', fixed = T) %>% do.call(rbind, .) %>% data.table
+## fips[, ncol(fips) := NULL, with = F]
+## setnames(fips, c('state', 'state.fips', 'county.fips', 'county'))

### 2017-04-11 init_model.rnw
Brought in population data.

@@
 load(file = 'tmp_data/nmusician_estabs.RData')
+load(file = 'tmp_data/population_by_county_year.RData')
+
+musicians <- merge(musicians, pop, by = c('county', 'season'), all.x = T)
 
@@
 <<>>= 
-musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)), by = team][order(average.nmusicians)]
+musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)),
+          by = team][order(average.nmusicians)]
 @

### 2017-04-12 clean_covers.r
Tidied up 'date' variable. Added points admitted and scored by each team.

@@ -15,18 +15,6 @@
 ################################################################################
 ################################################################################
 
-### 1. Investigate party cities -- Increase quality of observations, be really confident they *stayed* in a party city overnight.
-###    Intuitively: you stayed in a city the night before where there is a lot of partying
-###    E.g. an East Coast team may fly into LA the day before and we may want to define that game to be a party city
-###    they may also fly out after an afternoon game and miss the party-effect completely.
-###    Data Sources: tweets
-###    Figure out: LA -- rest day -- Boston, we know they won't take the red-eye
-### 2. Weighting age by starting minutes
-### 3. Marital status
-
-### Bring game-time into data
-### Try and get a sense of what a travel schedule might be..
-
 ##################################################
 ### Set up Workspace, Read in Raw Data
 ##################################################
@@ -43,13 +31,13 @@ lines <- lapply(files, fread) %>% rbindlist
 ### Clean up date and line variables. Remove extraneous vars.
 ##################################################
 
-lines[, game.time := gsub('^.* ([0-9:APM]+)$', '\\1', date)]
-lines[, date := as.Date(date, format = '%A, %B %d %Y')]
+lines[, date := as.POSIXct(date, format = '%A, %B %d %Y %I:%M%p')]
+## lines[, game.time := gsub('^.* ([0-9:APM]+)$', '\\1', date)]
+## lines[, date := as.Date(date, format = '%A, %B %d %Y')]
 
 lines[, line := gsub('^(-?[0-9.]+)/.*$', '\\1', line)]
 lines[, line.ts := NULL]
 lines[, ou := NULL]
-## lines[, game.time := NULL]
 
 ##################################################
 ### Expand team abbreviations.
@@ -77,7 +65,7 @@ setnames(lines, 'team', 'away.team')
 ### Remove missing observations. Set line as numeric.
 ##################################################
 
-lines <- lines[line != 'OFF', list(date, game.time, home.team, away.team, home.pts, away.pts, line)]
+lines <- lines[line != 'OFF', list(date, home.team, away.team, home.pts, away.pts, line)]
 lines[, line := as.numeric(line)]
 
 ### Season spans the new year, i.e. starts in October ends in April.
[…]

### 2017-04-12 init_model.rnw
Now using 'nhours.lgame' to obtain a very significant model.

@@
 lines[, list(date, team, opponent, location, 
-             line, score, outcome, ndays.lgame, 
+             line, score, outcome, nhours.lgame, 
              travel.dist, avg.age, nmusicians)]
@@
 <<echo=-4>>=
-lines[, party := ifelse(ndays.lgame == 1, nmusicians, 0)]
 
-m <- glm(outcome == 'W' ~ party + ndays.lgame + travel.dist + avg.age, 
+HOURS <- 24
+lines[, party := ifelse(nhours.lgame <= HOURS, nmusicians, 0)]
+
+m <- glm(outcome == 'W' ~ party + nhours.lgame + travel.dist + avg.age, 
          data = lines[season < 2017 & last.game.loc != team])
@@
 <<>>=
-m <- glm(outcome == 'W' ~ I(nmusicians*(ndays.lgame==1)), 
+HOURS <- 24
+m <- glm(outcome == 'W' ~ I(nmusicians*(nhours.lgame <= HOURS)), 
          data = lines[season < 2017 & last.game.loc != team])
@@
 ### We're only interested in making predictions when teams party.
-p <- predict(m, newdata = lines[season == 2017 & ndays.lgame == 1 & last.game.loc != team], 
+p <- predict(m, newdata = lines[season == 2017 & nhours.lgame <= HOURS & last.game.loc != team], 
              type = 'response')
-table(round(p), lines[season == 2017 & ndays.lgame == 1 & last.game.loc != team, outcome])
+p <- ifelse(p < .45, 0, 1)
+table(round(p), lines[season == 2017 & nhours.lgame <= HOURS & last.game.loc != team, outcome])
+
+@
+
+<<>>=
+
+require(lubridate)
+
+lines[, tpa := team.pts.admitted - mean(team.pts.admitted), by = list(team, season)]
+lines[, tps := team.pts.scored   - mean(team.pts.scored), by = list(team, season)]
+
+lm(tpa ~ last.game.time:party + I(hour(date))  + nhours.lgame, data = lines[nhours.lgame <= HOURS & last.game.loc != team]) %>% summary
+#lm(tps ~ party, data = lines) %>% summary
+
+lines[nhours.lgame <= 36, mean(tpa), by = last.game.loc][order(V1)]
+lines[nhours.lgame <= 36, mean(tps), by = last.game.loc][order(V1)]
+lines[nhours.lgame <= 36, mean(tps) - mean(tpa), by = last.game.loc][order(V1)]
+
+HOURS <- 24
+lines[, party := ifelse(nhours.lgame <= HOURS, nmusicians / log(population), 0)]
+m <- glm(outcome == 'W' ~ party + nhours.lgame + travel.dist + avg.age, 
+         data = lines[season < 2017 & last.game.loc != team])
+
+summary(m)
 

### 2017-04-12 scrape_espn.r
Added ability to avoid rescraping data.

@@ -48,21 +48,26 @@ scrapeGame <- function(gameID, date, save.path = 'tmp_data/espn/') {
     away <- teams[1]
     home <- teams[2]
     names(data) <- c('linescore', away, home)
-    data <- rbind(cbind(data[[2]], team = away, date),
-                  cbind(data[[3]], team = home, date))
-    write.csv(data, file = paste0(save.path, gameID, '.csv'), row.names = F)
+    tryCatch(expr = {
+        data <- rbind(cbind(data[[2]], team = away, date),
+                      cbind(data[[3]], team = home, date))
+        write.csv(data, file = paste0(save.path, gameID, '.csv'), row.names = F)
+    }, error = function(e) return(NULL))
     return(NULL)
 }
 
-scrapeGames <- function(date, gameIDs)
+scrapeGames <- function(date, gameIDs, rescrape = F, path = 'tmp_data/espn') {
+    if (!rescrape) {
+        scraped <- list.files(path) %>% gsub('\\.csv', '', .)
+        gameIDs <- setdiff(gameIDs, scraped)
+    }
     lapply(gameIDs, scrapeGame, date = date)
+}
 
 ### Scrape.
 clusterExport(cl, varlist = c('scrapeGame', 'readHTMLTable', '%>%'))
-clusterMap(cl, fun = scrapeGames, gameID = gameIDs, date = game.days,
+clusterMap(cl, fun = scrapeGames, gameIDs = gameIDs, date = game.days,
            .scheduling = 'dynamic')
-
-
 
 
 

### 2017-04-12 init_model.rnw
Added ESPN player data and associated models.

@@
 
+##################################################
+### Total Points Allowed
+##################################################
+
 lines[, tpa := team.pts.admitted - mean(team.pts.admitted), by = list(team, season)]
@@
 
+##################################################
+### musicians per log population
+##################################################
+
 HOURS <- 24
@@
 
+##################################################
+### ESPN
+##################################################
+
+espn <- fread('tmp_data/espn_player_data.csv')
+espn[, date := as.Date(date)]
+lines[, simple.date := as.Date(date)]
+data <- merge(lines, espn, by.x = c('team', 'simple.date'), by.y = c('team', 'date')) ### Careful which obs we keep!
+
+data[, dmeaned.free  := free.made/min - mean(free.made/min), by = list(starters, season)]
+data[, dmeaned.three := three.made/min - mean(three.made/min), by = list(starters, season)]
+data[, dmeaned.reb   := reb / min - mean(reb / min), by = list(starters, season)]
+
+setnames(data, '+/-', 'pm')
+
+lm(dmeaned.three ~ party, data = data[season < 2017 & last.game.loc != team]) %>% summary
+lm(dmeaned.reb ~ party, data = data[season < 2017 & last.game.loc != team]) %>% summary
+
+### WOW! Earn 1-2 points less per minute on average. (Further condition on player position?)
+### Next goal: we should be able to recover our grand mean estimate above. That is, test model efficacy 
+### by predicting the number of points a team will score using ex-ante minutes of play-time.
+lm(I(as.numeric(pm) / min) ~ party, data = data[season < 2017 & last.game.loc != team]) %>% summary
+
 @

### 2017-04-13 explore_starters.r
Init commit. Script will be used to explore player-specific stats.

@@ -1,0 +1,117 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Exploring Starters Data (from ESPN)
+###
+### Andreas Santucci
+###
+### Date: April 2017
+###
+### Inputs: 
+###
+### Dependencies: 
+###
+################################################################################
+################################################################################
+
+
+require(data.table)
+require(ggplot2)
+require(lubridate)
+require(magrittr)
+require(sampling)
+set.seed(4052017)
+
+load(file = 'tmp_data/covers_lines.RData')
+load(file = 'tmp_data/nmusician_estabs.RData')
+load(file = 'tmp_data/population_by_county_year.RData')
+
+musicians <- merge(musicians, pop, by = c('county', 'season'), all.x = T)
+musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
+
+lines <- merge(lines, musicians,
+               by.x = c('season', 'last.game.loc'),
+               by.y = c('season', 'team'), all.x = T)
+
+setkey(lines, season, team, date)
+
+lines[, party := ifelse(nhours.lgame <= HOURS, nmusicians, 0)]
+
+
+##################################################
+### Total Points Allowed
+##################################################
+
+lines[, tpa := team.pts.admitted - mean(team.pts.admitted), by = list(team, season)]
+lines[, tps := team.pts.scored   - mean(team.pts.scored), by = list(team, season)]
+
+lm(tpa ~ party + I(hour(date))  + nhours.lgame, data = lines[nhours.lgame <= HOURS & last.game.loc != team]) %>% summary
+
+##################################################
+### ESPN
+##################################################
+
+espn <- fread('tmp_data/espn_player_data.csv')
+espn[, date := as.Date(date)]
+lines[, simple.date := as.character(date) %>% substr(., 1, 10) %>% as.Date]
+data <- merge(lines, espn,
+              by.x = c('team', 'simple.date'),
+              by.y = c('team', 'date'), all.x = T)
+
+### Create player-season demeaned features.
[…]

### 2017-04-13 scrape_espn.r
Now scraping all players, not just starters.

@@ -41,22 +41,29 @@ gameIDs <- parSapplyLB(cl, game.days, extractGameIDs, simplify = F)
 save(gameIDs, file = 'tmp_data/espn_gameIDs.RData')
 
 ### For each game ID, we scrape the player data.
-scrapeGame <- function(gameID, date, save.path = 'tmp_data/espn/') {
+scrapeGame <- function(gameID, date, save.path = 'raw_data/espn/') {
     pg <- paste0('http://www.espn.com/nba/boxscore?gameId=', gameID)
+    doc <- htmlParse(pg)
+    l <- xpathApply(doc, path = '//tbody/tr/td')
+    a <- sapply(l, xmlValue)
+    Extract <- function(xmlvals, starting = 1) {
+        xmlvals <- xmlvals[starting:length(xmlvals)]
+        bidx <- grep("^(.{2,})\\1[A-Z]{1,2}$", xmlvals)[1]
+        lidx <- grep("(DNP)|(TEAM)", xmlvals)[1]
+        offset <- ifelse(xmlvals[lidx] == 'DNP', 2, 1)
+        return(xmlvals[bidx:(lidx-offset)] %>% matrix(., ncol = 15, byrow = T))
+    }
+    away <- Extract(a, 1)
+    home <- Extract(a, starting = grep("%", a)[1])
     data <- readHTMLTable(pg)[1:3]
     teams <- data$linescore[[1]] %>% as.character
-    away <- teams[1]
-    home <- teams[2]
-    names(data) <- c('linescore', away, home)
-    tryCatch(expr = {
-        data <- rbind(cbind(data[[2]], team = away, date),
-                      cbind(data[[3]], team = home, date))
-        write.csv(data, file = paste0(save.path, gameID, '.csv'), row.names = F)
-    }, error = function(e) return(NULL))
+    data <- rbind(cbind(away, teams[1], date %>% as.character),
+                  cbind(home, teams[2], date %>% as.character))
+    write.csv(data, file = paste0(save.path, gameID, '.csv'), row.names = F)
     return(NULL)
 }
 
-scrapeGames <- function(date, gameIDs, rescrape = F, path = 'tmp_data/espn') {
+scrapeGames <- function(date, gameIDs, rescrape = F, path = 'raw_data/espn') {
     if (!rescrape) {
         scraped <- list.files(path) %>% gsub('\\.csv', '', .)
[…]

### 2017-04-13 init_model.rnw
Fixed issue concerning 'simple.date' and merging in ESPN data. Added ideas for models, which have now been moved to explore_starters.r

--- 2017-04-12 init_model.rnw
@@
 <<>>=
-HOURS <- 24
+HOURS <- 36
 m <- glm(outcome == 'W' ~ I(nmusicians*(nhours.lgame <= HOURS)), 
@@
              type = 'response')
-p <- ifelse(p < .45, 0, 1)
 table(round(p), lines[season == 2017 & nhours.lgame <= HOURS & last.game.loc != team, outcome])
@@
 
-lines[nhours.lgame <= 36, mean(tpa), by = last.game.loc][order(V1)]
-lines[nhours.lgame <= 36, mean(tps), by = last.game.loc][order(V1)]
-lines[nhours.lgame <= 36, mean(tps) - mean(tpa), by = last.game.loc][order(V1)]
+## lines[nhours.lgame <= 36, mean(tpa), by = last.game.loc][order(V1)]
+## lines[nhours.lgame <= 36, mean(tps), by = last.game.loc][order(V1)]
+## lines[nhours.lgame <= 36, mean(tps) - mean(tpa), by = last.game.loc][order(V1)]
+
+##################################################
+### musicians per log population
+##################################################
 
@@
 
+##################################################
+### ESPN
+##################################################
+
+espn <- fread('tmp_data/espn_player_data.csv')
+espn[, date := as.Date(date)]
+lines[, simple.date := as.character(date) %>% substr(., 1, 10) %>% as.Date]
+data <- merge(lines, espn, by.x = c('team', 'simple.date'), by.y = c('team', 'date'), all.x = T) ### Careful which obs we keep!
+
+data[, dmeaned.free  := free.made/min - mean(free.made/min), by = list(starters, season)]
+data[, dmeaned.three := three.made/min - mean(three.made/min), by = list(starters, season)]
+data[, dmeaned.reb   := reb / min - mean(reb / min), by = list(starters, season)]
+data[, dmeaned.oreb   := oreb / min - mean(oreb / min), by = list(starters, season)]
+data[, dmeaned.dreb   := dreb / min - mean(dreb / min), by = list(starters, season)]
+data[, dmeaned.pts   := pts / min - mean(pts / min,na.rm=T), by = list(starters, season)]
+
+data[, chg.pos := three.made + fg.made + free.attempted/2 + 
+           to + (fg.attempted-fg.made + three.attempted - three.made - oreb)]
+
+changes <- data[, list(ttl.chg.pos = sum(chg.pos)), 
+            keyby = list(season, team, date, opponent)]
+lines <- merge(lines, changes, by = c('season', 'team', 'date', 'opponent'))
+
+setkey(lines, season, team, date)
+lines[, lag.chg.pos := c(NA, lag(ttl.chg.pos)[1:.N-1]), by = list(season, team)]
+
+glm(outcome == 'W' ~ party + nhours.lgame + I(log(travel.dist+1)) + lag.chg.pos,
+    data = lines[season < 2017 & last.game.loc != team], family = 'binomial') %>% summary
+
+setnames(data, '+/-', 'pm')
+
+lm(dmeaned.three ~ party, data = data[season < 2017 & last.game.loc != team]) %>% summary
+lm(dmeaned.reb ~ party, data = data[season < 2017 & last.game.loc != team]) %>% summary
+lm(dmeaned.dreb ~ party, data = data[season < 2017 & last.game.loc != team]) %>% summary
+
+data[, pm := as.numeric(pm)]
+data[, pf := as.numeric(pf)]
+### WOW! Earn 1-2 points less per minute on average. (Further condition on player position?)
+### Next goal: we should be able to recover our grand mean estimate above. That is, test model efficacy 
+### by predicting the number of points a team will score using ex-ante minutes of play-time.
+lm(I(pm / min) ~ party, data = data[season < 2017 & last.game.loc != team]) %>% summary
+
+lm(dmeaned.pts ~ party, data = data[season < 2017 & last.game.loc != team]%>%na.omit) %>% summary
+
+## data[grep("(los angeles)|(new york)|(brooklyn)", last.game.loc), 
+##      list(average.dmeaned.dreb = mean(dmeaned.dreb)), by = starters][order(average.dmeaned.dreb)] %>%
+##     write.csv(., file = 'dmeaned.dreb.csv')
+
+## data[grep("(los angeles)|(new york)|(brooklyn)", last.game.loc), 
+##      list(average.dmeaned.pts = mean(dmeaned.pts)), by = starters][order(average.dmeaned.pts)] %>%
+##     write.csv(., file = 'dmeaned.pts.csv')
+
 @

### 2017-04-15 init_model.rnw
Updated to include Minutes data and corresponding player specific model.

--- 2017-04-12 init_model.rnw
@@
 
-\section*{Data Sources}
+\section{Data Sources}
 
-\paragraph{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
+\subsection{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
 <<echo=F>>=
-lines[, list(date, team, opponent, line, score)]
+lines[, list(date = as.Date(date), team, opponent, line, score)]
 @
 
-\paragraph{BLS data} We also obtained data from
-\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the county-quarter level. \ul{As a 
+\subsection{BLS data} We also obtained data from
+\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the county-quarter level. As a 
 proxy for how much night-life there is in a city, we look toward the 
 number of sound recording studios, musical groups, and music publishers there 
-are for a particular county-year}: we simply calculate the average number of establishments across all three music categories listed above.\footnote{We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.}
+are for a particular county-year: we simply calculate the average number of establishments across all three music categories listed above.\footnote{We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.}
 
@@
 
-We then merge this in with our lines data, taking care to do so according to the
-last game's location. We also lag our BLS data by one year.
+We then merge this in with our lines data, taking care to do so according to each season-team's last game location; we also lag our BLS data by one year.
 To this data, we've created features such as 
 the number of days since last game, the last game location, the travel distance, and
-average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}.
+average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}. Our data range from 2010-2011 season through present, which includes part of the 2016-2017 season.
 
-<<echo=2>>=
+<<echo=F>>=
 ### Merge lines data with BLS data.
@@
 setkey(lines, season, team, date)
-lines[, list(date, team, opponent, location, 
-             line, score, outcome, nhours.lgame, 
-             travel.dist, avg.age, nmusicians)]
+
+## lines[, list(date, team, opponent, location, 
+##              line, score, outcome, nhours.lgame, 
+##              travel.dist, avg.age, nmusicians)]
 
@@
 
-Our data range from 2010-2011 season through present, which includes part of the 2016-2017 season.
+\subsection{Minutes data}
+We also gathered minutes data from \href{http://www.espn.com/nba/scoreboard/_/date/}{ESPN}.
 
-\subsection*{What cities have nightlife?}
+<<echo=F>>=
+espn <- fread('tmp_data/espn_player_data.csv')
+espn[, date := as.Date(date)]
+espn[1:5, list(player, mins, reb, ast, stl, pm, pts, date, team)]
+@
 
-We examine which cities have a notable nightlife, as measured by our proxy.
+The data also contain information on the number of free throws, field goals, and
+three pointers made and attempted by each player in each game.
+
+\section{Which cities (don't) have nightlife?}
+
+We examine which cities have a notable nightlife, as measured by our proxy;
+bottom and top five.
 
@@
 musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)),
-          by = team][order(average.nmusicians)]
+          by = team][order(average.nmusicians)][c(1:5, 25:29)]
 @
@@
 
-\subsection*{A significant model} We keep things simple, and try to explain the likelihood
+\subsection{Defining a party variable}
+
+We have a free parameter: the time-window which we believe the hangover effect
+to last. Suppose we only look toward games which have been played in a party
+city at most 24 hours ago. We take care to ignore
+teams playing in their own cities.
+
+<<>>=
+HOURS <- 24
+lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, nmusicians, 0)]
+@
+
+\section{Possible confounders}
+What if we pick up effects of a busy play schedule?
+Aside from accounting for how much rest time a team is allowed in between games,
+and travel distance between games, we're also interested to get a measure of
+fatigue incurred on the court.
+We account for this by bringing in minutes data, and counting the number of 
+times the players run back and forth on the court. This is crudely measured
+by the number of changes in posession. We back out this quantity.
+
+<<echo=6>>=
+
+lines[, simple.date := as.character(date) %>% substr(., 1, 10) %>% as.Date]
+data <- merge(lines, espn,
+              by.x = c('team', 'season', 'simple.date'),
+              by.y = c('team', 'season', 'date'), all.x = T)
+
+### Estimate player fatigue, by looking at number of changes in posession.
+data[, chg.pos := three.made + fg.made + free.attempted/2 + 
+           to + (fg.attempted-fg.made + three.attempted - three.made - oreb)]
+setkey(data, season, player, date)
+data[, lag.chg.pos := shift(chg.pos), by = list(season, player)]
+
+# Since we sum by location-date we double count.
+changes <- data[, list(ttl.chg.pos = sum(chg.pos) / 2), 
+            keyby = list(season, location, date)]
+lines <- merge(lines, changes, by = c('season', 'location', 'date'))
+
+setkey(lines, season, team, date)
+lines[, lag.chg.pos := c(NA, lag(ttl.chg.pos)[1:.N-1]), by = list(season, team)]
+
+@
+
+To end up with something like
+
+<<echo=F>>=
+lines[, list(date = simple.date, location = substr(location, 1, 7), 
+             team = substr(team, 1, 7), opponent = substr(opponent, 1, 7), 
+             line, score, outcome, 
+             avg.age, last.game.loc, nhours.lgame, travel.dist, party, lag.chg.pos)]
+@
+
+\section{A significant model} We keep things simple, and try to explain the likelihood
 of meeting the spread as a function of whether a team visited a city with an active night-life
@@
 
-HOURS <- 24
-lines[, party := ifelse(nhours.lgame <= HOURS, nmusicians, 0)]
-
-m <- glm(outcome == 'W' ~ party + nhours.lgame + travel.dist + avg.age, 
-         data = lines[season < 2017 & last.game.loc != team])
+m <- glm(outcome == 'W' ~ party + lag.chg.pos + travel.dist + 
+             nhours.lgame + I(hour(date)),
+         data = lines[season < 2017], family = 'binomial',
+         na.action = 'na.exclude')
 
@@
 The sign of our ``\texttt{party}'' variable is intuitively correct, and 
-statistically significant.
+statistically significant. The magnitude is also practically significant
+once we account for the units of \texttt{party} variable, i.e.
+visiting a party city such as New York or LA increases number of musicians
+by one or two hundred.
 
-\subsection*{A model worth betting on}
+\section{A model worth betting on}
+\paragraph{High precision model}
 If we re-train an even simpler model, we obtain fantastic betting performance.
@@
 
-summary(m)$coef %>% round(., 5)
-
 ### We're only interested in making predictions when teams party.
-p <- predict(m, newdata = lines[season == 2017 & nhours.lgame <= HOURS & last.game.loc != team], 
-             type = 'response')
-p <- ifelse(p < .45, 0, 1)
-table(round(p), lines[season == 2017 & nhours.lgame <= HOURS & last.game.loc != team, outcome])
+p <- predict(m, newdata = lines[season == 2017 & nhours.lgame <= HOURS & 
+                                last.game.loc != team], type = 'response')
+table(round(p), lines[season == 2017 & nhours.lgame <= HOURS & 
+                      last.game.loc != team, outcome])
 
@@
 
-<<>>=
+\paragraph{A model which realizes more games to bet on}
+We can try different variants, for example, supposing the treatment effect of visiting
+a city filled with musicians depends on the teams average age.
 
-require(lubridate)
-
-lines[, tpa := team.pts.admitted - mean(team.pts.admitted), by = list(team, season)]
-lines[, tps := team.pts.scored   - mean(team.pts.scored), by = list(team, season)]
-
-lm(tpa ~ last.game.time:party + I(hour(date))  + nhours.lgame, data = lines[nhours.lgame <= HOURS & last.game.loc != team]) %>% summary
-#lm(tps ~ party, data = lines) %>% summary
-
-lines[nhours.lgame <= 36, mean(tpa), by = last.game.loc][order(V1)]
-lines[nhours.lgame <= 36, mean(tps), by = last.game.loc][order(V1)]
-lines[nhours.lgame <= 36, mean(tps) - mean(tpa), by = last.game.loc][order(V1)]
-
-HOURS <- 24
-lines[, party := ifelse(nhours.lgame <= HOURS, nmusicians / log(population), 0)]
-m <- glm(outcome == 'W' ~ party + nhours.lgame + travel.dist + avg.age, 
+<<echo=1>>=
+m <- glm(outcome == 'W' ~ I(nmusicians*(nhours.lgame <= HOURS)/avg.age), 
          data = lines[season < 2017 & last.game.loc != team])
 
-summary(m)
+p <- predict(m, newdata = lines[season == 2017 & nhours.lgame <= HOURS & 
+                                last.game.loc != team], type = 'response')
+table(round(p), lines[season == 2017 & nhours.lgame <= HOURS & 
+                      last.game.loc != team, outcome])
 
@@
 
+\section{Player specific model}
+
+Can we measure fatigue on a player specific model? Consider something like
+rebounds, which by player anecdote require a lot of energy.
+Does party have an effect on this outcome? We might be curious to disentangle
+what other factors effect player fatigue, as these could be possible confounders
+for our above analysis.
+
+In particular, for each player, we calculate the number of rebounds per minute
+earned in a game, less the players average number of rebounds per minute for that
+season. We explore how this response variable can be explained as a function of
+\texttt{party}, the lagged number of changes in posessions, travel distance,
+and rest time since last game.
+
+<<echo=F>>=
+require(mboost)
+lines[, game.time := hour(date)]
+
+lm(demeaned.reb.per.min ~ party + lag.chg.pos +
+       I(log(travel.dist+1)/nhours.lgame) + nhours.lgame,
+   data = data[season < 2017 & last.game.loc != team]) %>%
+   summary
+@
+
 \end{document}

### 2017-04-15 explore_starters.r
Added gbm model and rpart.

@@ -37,7 +37,7 @@ lines <- merge(lines, musicians,
 setkey(lines, season, team, date)
 
 HOURS <- 24
-lines[, party := ifelse(nhours.lgame <= HOURS, nmusicians, 0)]
+lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, nmusicians, 0)]
 
 ##################################################
 ### Total Points Allowed
@@ -46,7 +46,7 @@ lines[, party := ifelse(nhours.lgame <= HOURS, nmusicians, 0)]
 lines[, tpa := team.pts.admitted - mean(team.pts.admitted), by = list(team, season)]
 lines[, tps := team.pts.scored   - mean(team.pts.scored), by = list(team, season)]
 
-lm(tpa ~ party + I(hour(date))  + nhours.lgame, data = lines[nhours.lgame <= HOURS & last.game.loc != team]) %>% summary
+lm(tpa ~ party + I(hour(date))  + nhours.lgame, data = lines[nhours.lgame <= HOURS]) %>% summary
 
 ##################################################
 ### ESPN
@@ -77,22 +77,26 @@ setkey(lines, season, team, date)
 lines[, lag.chg.pos := c(NA, lag(ttl.chg.pos)[1:.N-1]), by = list(season, team)]
 
 ### Good news: party significant after accounting for fatigue.
-m <- glm(outcome == 'W' ~ party + lag.chg.pos + I(travel.dist*(travel.dist>1800)) + nhours.lgame +
-             I(hour(date)),
-         data = lines[season < 2017 & last.game.loc != team], family = 'binomial')
+require(splines)
+m <- glm(outcome == 'W' ~ party + lag.chg.pos + travel.dist + 
+             nhours.lgame + I(hour(date)),
+         data = lines[season < 2017], family = 'binomial',
+         na.action = 'na.exclude')
+
+#plot(m$data$nhours.lgame, residuals(m))
 
 ### SVM for out of the box performance.
-lines <- lines[last.game.loc != team,
-               list(hour = hour(date), last.game.time, travel.dist,
-                    party, lag.chg.pos, season, outcome)] %>% na.omit
+## lines <- lines[last.game.loc != team,
+##                list(hour = hour(date), last.game.time, travel.dist,
+##                     party, lag.chg.pos, season, outcome)] %>% na.omit
 
-require(e1071)
[…]

### 2017-04-15 init_model.rnw
Added an SVM for its out of the box performance.

@@
 
-The data also contain information on the number of free throws, field goals, and
+The minutes data also contain information on the number of free throws, field goals, and
 three pointers made and attempted by each player in each game.
@@
 
-<<>>= 
+<<echo=F>>= 
 musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)),
@@
 
+\paragraph{An out of the box machine learning model}
+For fun, we try a simple support vector machine. We posit that whether
+teams party also depends on how much time they have between games; even
+when restricting our data to the twenty-four hour window after teams visit a party city, there is quite a bit of heterogeneity in when yesterday's game finished and when 
+today's starts that a non-linear classifier may be able to take advantage of.
+
+<<echo=2:5>>=
+require(e1071)
+svm.data <- lines[last.game.loc != team, list(hour = hour(date), last.game.time,
+                                              party, season, outcome)] %>% na.omit
+
+m <- svm(x = svm.data[season < 2016, -c('outcome', 'season'), with = F],
+         y = svm.data[season < 2016, outcome == 'W'],
+         scale = T, type = 'C-classification', probability = T, kernel = 'radial')
+
+preds <- predict(m, newdata = svm.data[season == 2016, -c('outcome', 'season'), with = F])
+table(preds, svm.data[season == 2016 , outcome])
+@
+
 \section{Player specific model}

### 2017-04-24 init_model.rnw
Added section on replicating analysis for MLB.

@@
 
-<<echo=F,results='hide'>>=
+<<echo=F,results='hide',messages=F>>=
 
@@
 load(file = 'tmp_data/population_by_county_year.RData')
-
 musicians <- merge(musicians, pop, by = c('county', 'season'), all.x = T)
@@
 @
+
+For results, skip to section \ref{seq: significant model}.
 
@@
 
-<<echo=F>>=
+<<echo=F,results='hide'>>=
 musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
-musicians[, list(county, team, season, nmusicians = round(nmusicians, 1))]
 @
@@
 
-To end up with something like
+To end up with a quantity that reflects the total number of changes in possession across both teams within each game.
 
@@
 
-\section{A significant model} We keep things simple, and try to explain the likelihood
+\section{A significant model} 
+\label{seq: significant model}
+We keep things simple, and try to explain the likelihood
 of meeting the spread as a function of whether a team visited a city with an active night-life
@@
 require(e1071)
-svm.data <- lines[last.game.loc != team, list(hour = hour(date), last.game.time,
-                                              party, season, outcome)] %>% na.omit
+svm.data <- lines[last.game.loc != team, 
+                  list(hour = hour(date), last.game.time, party, 
+                       loss.season, season, outcome)] %>%
+    na.omit
 
+### tune.svm(..., gamma = 2^(-5:5))
 m <- svm(x = svm.data[season < 2016, -c('outcome', 'season'), with = F],
               y = svm.data[season < 2016, outcome == 'W'],
-         scale = T, type = 'C-classification', probability = T, kernel = 'radial')
-
+              scale = T, type = 'C-classification', probability = T, 
+              kernel = 'radial')
 preds <- predict(m, newdata = svm.data[season == 2016, -c('outcome', 'season'), with = F])
@@
        I(log(travel.dist+1)/nhours.lgame) + nhours.lgame,
-   data = data[season < 2017 & last.game.loc != team]) %>%
+   data = data[season < 2017 & last.game.loc != team] %>% na.omit) %>%
    summary
@@
 
+\section{Robustness Check: Replicating the analysis for MLB}
+We go ahead and repeat this process for major league baseball.
+The only difference here is that when we define our party variable,
+we implicitly interact it with \texttt{weekend}.
+
+<<echo=-1>>=
+load(file = 'tmp_data/covers_lines_mlb.RData')
+lines[, party := ifelse(last.game.loc != team & weekend == 1, nmusicians, 0)]
+
+m <- glm(I(team.score > opponent.score) ~ party + I(team == location) + 
+             odds + nhours.lgame + weekend + travel.dist,
+         data = lines, family = 'binomial')
+summary(m)
+@
+
+We note that although we don't quite achieve conventional statistical significance,
+we're not that far off the mark.
+
+\paragraph{Betting on MLB}
+We can also earn a profit making predictions, even after accounting for the cut
+taken by the betting house.
+
+<<echo=F>>= 
+source('functions.r')
+
+m <- glm(I(team.score > opponent.score) ~ party,
+         data = lines[year(date) < 2016], family = 'binomial')
+p <- predict(m, newdata = lines[year(date) == 2016], type = 'response')
+
+idx <- which(p < .45)
+p1 <- mapply(Predict, prediction = p[idx], vegas = lines[year(date) == 2016, odds][idx],
+             MoreArgs=list(threshold = 0.01))
+b <- mapply(Bet, our.prediction = p1,
+            actual.outcome = lines[year(date) == 2016, team.score > opponent.score][idx],
+            odds = lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
+
+sum(b, na.rm = T)
+
+@
+
+If we were to bet one hundred dollars on each game in 2016,
+we would earn a profit of just less than \$400 over the course of 45 games.
+
 \end{document}

### 2017-04-26 init_model.rnw
p=.101 on mlb.

@@
 
-\paragraph{An out of the box machine learning model}
-For fun, we try a simple support vector machine. We posit that whether
-teams party also depends on how much time they have between games; even
-when restricting our data to the twenty-four hour window after teams visit a party city, there is quite a bit of heterogeneity in when yesterday's game finished and when 
-today's starts that a non-linear classifier may be able to take advantage of.
-
-<<echo=2:5>>=
-require(e1071)
-svm.data <- lines[last.game.loc != team, 
-                  list(hour = hour(date), last.game.time, party, 
-                       loss.season, season, outcome)] %>%
-    na.omit
-
-### tune.svm(..., gamma = 2^(-5:5))
-m <- svm(x = svm.data[season < 2016, -c('outcome', 'season'), with = F],
-              y = svm.data[season < 2016, outcome == 'W'],
-              scale = T, type = 'C-classification', probability = T, 
-              kernel = 'radial')
-preds <- predict(m, newdata = svm.data[season == 2016, -c('outcome', 'season'), with = F])
-table(preds, svm.data[season == 2016 , outcome])
-@
-
 \section{Player specific model}
@@
 load(file = 'tmp_data/covers_lines_mlb.RData')
-lines[, party := ifelse(last.game.loc != team & weekend == 1, nmusicians, 0)]
+lines[, party := ifelse(team != location & weekend == 1, nmusicians, 0)]
 
-m <- glm(I(team.score > opponent.score) ~ party + I(team == location) + 
-             odds + nhours.lgame + weekend + travel.dist,
+m <- glm(I(team.score > opponent.score) ~ party + odds + I(team == location) + 
+             ndays.lgame + travel.dist + weekend,
          data = lines, family = 'binomial')
@@
 We note that although we don't quite achieve conventional statistical significance,
-we're not that far off the mark.
+we're not that far off the mark. The difference in significance between 
+sports could possibly be explained by the heterogeneity in athleticism 
+required to play basketball versus baseball, wherein the former requires far more 
+non-stop sprinting. We also remark that the p-value on \texttt{party}
+similar in magnitude to the p-value on our \texttt{home-team advantage} indicator,
+which is an important feature as deemed by the literature.\cite{hometeamadv}
 
@@
 
-<<echo=F>>= 
+<<echo=2:4>>= 
 source('functions.r')
-
-m <- glm(I(team.score > opponent.score) ~ party,
+m <- glm(I(team.score > opponent.score) ~ party + odds,
          data = lines[year(date) < 2016], family = 'binomial')
@@
 sum(b, na.rm = T)
-
 @
@@
 If we were to bet one hundred dollars on each game in 2016,
-we would earn a profit of just less than \$400 over the course of 45 games.
+we would earn a profit of about \Sexpr{format(round(sum(b, na.rm=T), -1), big.mark = ',')} dollars
+over the course of \Sexpr{sum(!is.na(b))} games.
+
+\begin{thebibliography}{9}
+  \bibitem{hometeamadv}
+    Jeremy P. Jamieson.
+    \textit{The Home Field Advantage in Athletics: A Meta-Analysis}
+    \href{http://onlinelibrary.wiley.com/doi/10.1111/j.1559-1816.2010.00641.x/full}{paper}
+     \end{thebibliography}
 

### 2017-04-26 clean_covers_mlb.r
Added features for 'first in series' for each matchup.

@@ -19,6 +19,8 @@ require(data.table)
 require(magrittr)
 require(sampling)
 
+source('functions.r')
+
 ### A character vector of full team names.
 fulls <- c('arizona diamondbacks', 'atlanta braves', 'baltimore orioles', 'boston redsox',
            'chicago cubs', 'chicago whitesox', 'cincinati reds', 'cleveland indians',
@@ -131,9 +133,15 @@ lines <- rbind(data, lines.dup)
 lines[, season := year(date)]
 
 setkey(lines, team, season, date)
-lines[, nhours.lgame  := c(NA, diff(date)),            by = list(team, season)]
-lines[, last.game.loc := c(NA, lag(location)[1:.N-1]), by = list(team, season)]
+### Some observations have duplicates. Can't have these.
+lines <- lines[!lines[, .N, by = list(team, season, date)][N > 1]]
+lines[, ndays.lgame  := c(NA, diff(as.Date(date))),   by = list(team, season)]
+lines[, last.game.loc := shift(location), by = list(team, season)]
+lines <- lines[ndays.lgame != 0]
 
+##################################################
+### Team Locations and Distance Traveled.
+##################################################
 ### Here, we geocode team locations to get lat-lon, and also addresses.
 ## require(ggmap)
 ## require(sp)
@@ -160,58 +168,49 @@ lines[, travel.dist := getDist(location, last.game.loc, dmat), by = list(locatio
 lines[, c('team.score', 'opponent.score') := tstrsplit(score, split = '-')]
 lines[, weekend := weekday %in% c('Saturday', 'Sunday')]
 
+##################################################
+### Musicians from BLS
+##################################################
 load(file = 'tmp_data/nmusician_estabs_mlb.RData')
 lines <- merge(lines, musicians,
                by.x = c('season', 'last.game.loc'),
                by.y = c('season', 'team'), all.x = T)
 
 ### What does it mean to party?
-lines[, party := ifelse(last.game.loc != team & weekend == 1, nmusicians, 0)]
+setkey(lines, season, team, date)
+lines[, last.game.loc := shift(location), by = list(team, season)]
+lines[, lag.opponent := shift(opponent),  by = list(team, season)]
+lines[, first.in.series := ifelse(is.na(lag.opponent) | opponent != lag.opponent, 1, 0)]
+lines[, not.first := !first.in.series]
 
[…]

### 2017-04-26 writing/Thesis.tex
Init commit. Included author names and instutition.

%% ----------------------------------------------------------------
%% Thesis.tex -- MAIN FILE (the one that you compile with LaTeX)
%% ----------------------------------------------------------------

% Set up the document
\documentclass[a4paper, 11pt, oneside]{Thesis}  % Use the "Thesis" style, basedon the ECS Thesis style by Steve Gunn
\graphicspath{Figures/}  % Location of the graphics files (set up for graphics to be in PDF format)

% Include any extra LaTeX packages required
\usepackage[square, numbers, comma, sort&compress]{natbib}  % Use the "Natbib" style for the references in the Bibliography
\usepackage{verbatim}  % Needed for the "comment" environment to make LaTeX comments
\usepackage{vector}  % Allows "\bvec{}" and "\buvec{}" for "blackboard" style bold vectors in maths
\hypersetup{urlcolor=blue, colorlinks=true}  % Colours hyperlinks in blue, but this can be distracting if there are many links.

%% ----------------------------------------------------------------
\begin{document}
\frontmatter      % Begin Roman style (i, ii, iii, iv...) page numbering

% Set up the Title Page
\title  {On the effects of partying on next-day sports performance}
\authors  {\texorpdfstring
            {\href{https://stanford.edu/~santucci/}{A. Santucci and E. Lax}}
            {}
            }
\addresses  {\groupname\\\deptname\\\univname}  % Do not change this here, instead these must be set in the "Thesis.cls" file, please look through it instead
\date       {\today}
\subject    {}
\keywords   {}

\maketitle
%% ----------------------------------------------------------------

\setstretch{1.3}  % It is better to have smaller font and larger line spacing than the other way round

% Define the page headers using the FancyHdr package and set up for one-sided printing
\fancyhead{}  % Clears all page headers and footers
\rhead{\thepage}  % Sets the right side header to show the page number
\lhead{}  % Clears the left side page header

\pagestyle{fancy}  % Finally, use the "fancy" page style to implement the FancyHdr headers

%% ----------------------------------------------------------------
% Declaration Page required for the Thesis, your institution may give you a different text to place here
\Declaration{

\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics

I, AUTHOR NAME, declare that this thesis titled, `THESIS TITLE' and the work presented in it are my own. I confirm that:

\begin{itemize} 
\item[\tiny{$\blacksquare$}] This work was done wholly or mainly while in candidature for a research degree at this University.

\item[\tiny{$\blacksquare$}] Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.

\item[\tiny{$\blacksquare$}] Where I have consulted the published work of others, this is always clearly attributed.

\item[\tiny{$\blacksquare$}] Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.

\item[\tiny{$\blacksquare$}] I have acknowledged all main sources of help.

\item[\tiny{$\blacksquare$}] Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.
\\
\end{itemize}

Signed:\\
\rule[1em]{25em}{0.5pt}  % This prints a line for the signature

Date:\\
\rule[1em]{25em}{0.5pt}  % This prints a line to write the date
}
\clearpage  % Declaration ended, now start a new page

%% ----------------------------------------------------------------
% The "Funny Quote Page"
\pagestyle{empty}  % No headers or footers for the following pages

\null\vfill
% Now comes the "Funny Quote", written in italics
\textit{``Write a funny quote here.''}

\begin{flushright}
If the quote is taken from someone, their name goes here
\end{flushright}

\vfill\vfill\vfill\vfill\vfill\vfill\null
\clearpage  % Funny Quote page ended, start a new page
%% ----------------------------------------------------------------

% The Abstract Page
\addtotoc{Abstract}  % Add the "Abstract" page entry to the Contents
\abstract{
\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics

The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots

}

\clearpage  % Abstract ended, start a new page
%% ----------------------------------------------------------------

\setstretch{1.3}  % Reset the line-spacing to 1.3 for body text (if it has changed)

% The Acknowledgements page, for thanking everyone
\acknowledgements{
\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics

The acknowledgements and the people to thank go here, don't forget to include your project advisor\ldots

}
\clearpage  % End of the Acknowledgements
%% ----------------------------------------------------------------

\pagestyle{fancy}  %The page style headers have been "empty" all this time, now use the "fancy" headers as defined before to bring them back

%% ----------------------------------------------------------------
\lhead{\emph{Contents}}  % Set the left side page header to "Contents"
\tableofcontents  % Write out the Table of Contents

%% ----------------------------------------------------------------
\lhead{\emph{List of Figures}}  % Set the left side page header to "List if Figures"
\listoffigures  % Write out the List of Figures

%% ----------------------------------------------------------------
\lhead{\emph{List of Tables}}  % Set the left side page header to "List of Tables"
\listoftables  % Write out the List of Tables

%% ----------------------------------------------------------------
\setstretch{1.5}  % Set the line spacing to 1.5, this makes the following tables easier to read
\clearpage  % Start a new page
\lhead{\emph{Abbreviations}}  % Set the left side page header to "Abbreviations"
\listofsymbols{ll}  % Include a list of Abbreviations (a table of two columns)
{
% \textbf{Acronym} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or \\
\textbf{LAH} & \textbf{L}ist \textbf{A}bbreviations \textbf{H}ere \\

}

%% ----------------------------------------------------------------
\clearpage  % Start a new page
\lhead{\emph{Physical Constants}}  % Set the left side page header to "Physical Constants"
\listofconstants{lrcl}  % Include a list of Physical Constants (a four column table)
{
% Constant Name & Symbol & = & Constant Value (with units) \\
Speed of Light & $c$ & $=$ & $2.997\ 924\ 58\times10^{8}\ \mbox{ms}^{-\mbox{s}}$ (exact)\\

}

%% ----------------------------------------------------------------
\clearpage  %Start a new page
\lhead{\emph{Symbols}}  % Set the left side page header to "Symbols"
\listofnomenclature{lll}  % Include a list of Symbols (a three column table)
{
% symbol & name & unit \\
$a$ & distance & m \\
$P$ & power & W (Js$^{-1}$) \\
& & \\ % Gap to separate the Roman symbols from the Greek
$\omega$ & angular frequency & rads$^{-1}$ \\
}
%% ----------------------------------------------------------------
% End of the pre-able, contents and lists of things
% Begin the Dedication page

\setstretch{1.3}  % Return the line spacing back to 1.3

\pagestyle{empty}  % Page style needs to be empty for this page
\dedicatory{For/Dedicated to/To my\ldots}

\addtocontents{toc}{\vspace{2em}}  % Add a gap in the Contents, for aesthetics

%% ----------------------------------------------------------------
\mainmatter	  % Begin normal, numeric (1,2,3...) page numbering
\pagestyle{fancy}  % Return the page headers back to the "fancy" style

% Include the chapters of the thesis, as separate files
% Just uncomment the lines as you write the chapters

\input{Chapters/Chapter1} % Introduction

%\input{Chapters/Chapter2} % Background Theory

%\input{Chapters/Chapter3} % Experimental Setup

%\input{Chapters/Chapter4} % Experiment 1

%\input{Chapters/Chapter5} % Experiment 2

%\input{Chapters/Chapter6} % Results and Discussion

%\input{Chapters/Chapter7} % Conclusion

%% ----------------------------------------------------------------
% Now begin the Appendices, including them as separate files

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

\appendix % Cue to tell LaTeX that the following 'chapters' are Appendices

\input{Appendices/AppendixA}	% Appendix Title

%\input{Appendices/AppendixB} % Appendix Title

%\input{Appendices/AppendixC} % Appendix Title

\addtocontents{toc}{\vspace{2em}}  % Add a gap in the Contents, for aesthetics
\backmatter

%% ----------------------------------------------------------------
\label{Bibliography}
\lhead{\emph{Bibliography}}  % Change the left side page header to "Bibliography"
\bibliographystyle{unsrtnat}  % Use the "unsrtnat" BibTeX style for formatting the Bibliography
\bibliography{Bibliography}  % The references (bibliography) information are stored in the file named "Bibliography.bib"

\end{document}  % The End
%% ----------------------------------------------------------------

### 2017-05-02 init_model.rnw
Now using logged  for party variable. Only helps significance, and doesn't hurt betting.

@@
 HOURS <- 24
-lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, nmusicians, 0)]
+lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians), 0)]
 @
@@
 load(file = 'tmp_data/covers_lines_mlb.RData')
-lines[, party := ifelse(team != location & weekend == 1, nmusicians, 0)]
+lines[, party := ifelse(team != location & weekend == 1, log(nmusicians), 0)]
 

### 2017-05-05 init_model.rnw
Added a plot of running profits from betting in 2016.

@@
 
-m <- glm(outcome == 'W' ~ party + lag.chg.pos + travel.dist + 
+lines[, ew := sin(travel.dir %% 360 / 360 * 2 * pi)]
+#plot(lines$travel.dir, lines$ew)
+
+m <- glm(outcome == 'W' ~ party + lag.chg.pos + travel.dist*ew + 
              nhours.lgame + I(hour(date)),
@@
 
-lm(demeaned.reb.per.min ~ party + lag.chg.pos +
-       I(log(travel.dist+1)/nhours.lgame) + nhours.lgame,
-   data = data[season < 2017 & last.game.loc != team] %>% na.omit) %>%
+lm(demeaned.reb.per.min ~ party + lag.chg.pos + travel.dist + nhours.lgame,
+   data = data[last.game.loc != team] %>% na.omit) %>%
    summary
@@
 sum(b, na.rm = T)
+
+### 
+betting.days <- lines[year(date) == 2016][idx, date]
+running.prof <- cumsum(ifelse(is.na(b), 0, b))
+plot(x = 1:length(b), y = running.prof, 
+     main = 'Profits over the 2016 season', 
+     xlab = 'Game Number', ylab = 'Profit', type = 'l')
+abline(a = 0, b = 0, lty = 3)
+points(x = which.min(running.prof), y = min(running.prof), type = 'p')
+text(x = which.min(running.prof), y = min(running.prof), labels = min(running.prof), pos = 2)
+
 @

### 2017-05-05 init_model.rnw
Now using just party variable for prediction in NBA, instead of scaling by avg team age.

@@
                                 last.game.loc != team], type = 'response')
+
 table(round(p), lines[season == 2017 & nhours.lgame <= HOURS & 
@@
 <<echo=1>>=
-m <- glm(outcome == 'W' ~ I(nmusicians*(nhours.lgame <= HOURS)/avg.age), 
-         data = lines[season < 2017 & last.game.loc != team])
+source('functions.r')
+m <- glm(outcome == 'W' ~ party,
+         data = lines[season < 2016 & last.game.loc != team])
 
-p <- predict(m, newdata = lines[season == 2017 & nhours.lgame <= HOURS & 
+p <- predict(m, newdata = lines[season == 2016 & nhours.lgame <= HOURS & 
                                 last.game.loc != team], type = 'response')
-table(round(p), lines[season == 2017 & nhours.lgame <= HOURS & 
+
+table(round(p), lines[season == 2016 & nhours.lgame <= HOURS & 
                       last.game.loc != team, outcome])
@@
 <<echo=2:4>>= 
-source('functions.r')
 m <- glm(I(team.score > opponent.score) ~ party + odds,
@@
 idx <- which(p < .45)
-p1 <- mapply(Predict, prediction = p[idx], vegas = lines[year(date) == 2016, odds][idx],
+p1 <- mapply(Predict, prediction = p[idx], 
+             vegas = lines[year(date) == 2016, odds][idx],
              MoreArgs=list(threshold = 0.01))

### 2017-05-06 poster/poster.tex
Init commit. Nice templated outline.

\documentclass[24pt, a0paper, portrait]{tikzposter} 
 \usepackage{algorithm2e} \usepackage{amsmath} \usepackage{amsthm} \usepackage{amsfonts} \usepackage{bbm} \usepackage{color,soul} \usepackage{framed} \usepackage{hyperref} \usepackage{mathtools} \usepackage[utf8]{inputenc} 
 \newtheorem{theorem}{Theorem}[section] \newtheorem{lemma}[theorem]{Lemma} \newtheorem{proposition}[theorem]{Proposition} \newtheorem{corollary}[theorem]{Corollary}  \DeclareMathOperator*{\argmin}{arg\,min} \DeclareMathOperator*{\argmax}{arg\,max} \newcommand{\D}{\mathrm{d}} \SetKwInput{KwInput}{Input} \SetKwInput{KwOutput}{Output}   \title{The effects of night-light on subsequent day sports performance} \author{Eric Lax, Andreas Santucci} \date{May 19th, 2017} \institute{Institute for Computational and Mathematical Engineering}   \usetheme{Board}   \begin{document}   \maketitle

\block{~} {     It's not unreasonable to think that game performance can be affected (in part) by what takes place off the court. We are interested to see if teams exhibit a decline in performance the day following a game in a party city. We exploit data on bookmaker spreads, the expected score differential between two teams after conditioning on observable performance. We expect a team to meet the spread half the time, since this is how bookmakers minimize risk on their end. We construct a model which attempts to estimate the causal effect of  visiting a "party-city" the day before a game on the probability of meeting the spread. }

\begin{columns}     \column{0.4}     \block{Data Sources}{
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
      Text and more text
    }       \column{0.6}     \block{NBA model}{Here, \vspace{4cm}} \end{columns}

\begin{columns}     \column{0.5}     \block{Profits on MLB Betting in 2016}     {         \begin{tikzfigure}             \includegraphics[width=0.4\textwidth]{figures/mlb_bets}         \end{tikzfigure}     }     \column{0.5}     \block{Description of the figure}{
      On the left, we plot profits as a function of time over the 2016 season. For each relevant
      game, we place a bet if our model instructs us of odds sufficiently different from what
      the betting house suggests. We account for the bookmaker's cut. In the early season,
      at its worst we are out of pocket \$750 dollars. By season end, we have recouped all of
      our initial investment plus an additional \$13.5k.
    }
    \note[         
      targetoffsetx=-8cm,          
      targetoffsety=-6.5cm,          
      width=0.5\linewidth         
    ]{e-mail: \texttt{santucci@stanford.edu}}   \end{columns}   \end{document}

### 2017-05-06 poster/poster.tex
Working version with data-information. Not yet working: model tables.

@@
  \usepackage{algorithm2e} \usepackage{amsmath} \usepackage{amsthm} \usepackage{amsfonts} \usepackage{bbm} \usepackage{color,soul} \usepackage{framed} \usepackage{hyperref} \usepackage{mathtools} \usepackage[utf8]{inputenc}
+\usepackage{subcaption}
  \newtheorem{theorem}{Theorem}[section] \newtheorem{lemma}[theorem]{Lemma} \newtheorem{proposition}[theorem]{Proposition} \newtheorem{corollary}[theorem]{Corollary}  \DeclareMathOperator*{\argmin}{arg\,min} \DeclareMathOperator*{\argmax}{arg\,max} \newcommand{\D}{\mathrm{d}} \SetKwInput{KwInput}{Input} \SetKwInput{KwOutput}{Output}   \title{The effects of night-light on subsequent day sports performance} \author{Eric Lax, Andreas Santucci} \date{May 19th, 2017} \institute{Institute for Computational and Mathematical Engineering}   \usetheme{Board}   \begin{document}   \maketitle
 
-\block{~} {     It's not unreasonable to think that game performance can be affected (in part) by what takes place off the court.
+\block{Idea} {     It's not unreasonable to think that game performance can be affected (in part) by what takes place off the court.
 We are interested to see if teams exhibit a decline in performance the day following a game in a party city.
@@
 
+\block{Mathematical Model}{
+  \[
+    \texttt{meet spread} \sim \texttt{party} + \substack{\texttt{days since} \\ \texttt{last game}} + \substack{\texttt{time}\\ \texttt{difference}} + \substack{\texttt{distance} \\ \texttt{traveled}}
+  \]
+}
+
 \begin{columns}     \column{0.4}     \block{Data Sources}{
-      Text and more text
-      Text and more text
-      Text and more text
-      Text and more text
-      Text and more text
-      Text and more text
-      Text and more text
-      Text and more text
-      Text and more text
-    }       \column{0.6}     \block{NBA model}{Here, \vspace{4cm}} \end{columns}
+    \paragraph{Betting house data and game outcomes}
+    We obtained data from 
+    \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, 
+    which records the outcome of each game alongside the line set by a betting house.
+
+    \paragraph{Data which determines the night-life index of a city}
+    We also obtained data from
+    \href{https://www.bls.gov/data/}{Bureau of Labor Statistics} 
+    which records the number of establishments by business type at the county-quarter level. As a 
+    proxy for how much night-life there is in a city, we look toward the 
+    number of sound recording studios, musical groups, and music publishers there 
+    are for a particular county-year: we simply calculate the average number of establishments 
+    across all three music categories listed above.
+
+We then merge this in with our lines data, taking care to do so according to each season-team's last game location; we also lag our BLS data by one year.
+To this data, we've created features such as 
+the number of days since last game, the last game location, the travel distance, and
+average age. Our data range from 2010-2011 season through present, which includes part of the 2016-2017 season.
+
+\paragraph{Minutes data on player specific performance}
+We also gathered minutes data from \href{http://www.espn.com/nba/scoreboard/_/date/}{ESPN}.
+The minutes data also contain information on the number of free throws, field goals, and
+three pointers made and attempted by each player in each game.
+
+    }
+    \column{0.6}     
+    \block{NBA model}{
+      Here,
+      % Table created by stargazer v.5.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu % Date and time: Sat, May 06, 2017 - 09:58:05 AM
+
+} 
+\end{columns}
+
+\begin{columns}   \column{0.4}
+  \block{Player Specific Model}{
+    Here
+  }
+  \column{0.6}
+  \block{MLB Model}{
+    Here
+  } \end{columns}
 
@@
       width=0.5\linewidth         
-    ]{e-mail: \texttt{santucci@stanford.edu}}   \end{columns}   \end{document}
+    ]{e-mail: \texttt{santucci@stanford.edu}}   \end{columns}
+
+\block{Meta Analysis}{
+  Over 6k lines of code written, with ~120 commits.
+
+  \begin{tikzfigure}
+    \includegraphics[width=0.3\textwidth]{figures/gh_commits}
+  \end{tikzfigure}
+  \begin{tikzfigure}     \includegraphics[width=0.7\textwidth]{figures/punch_card}    \end{tikzfigure}
+}
+
+\end{document}

### 2017-05-06 poster/poster.tex
Added stargazer tables and adjusted sticker-note containing email addresses.

@@
 
-\block{Mathematical Model}{
+\block{Model}{
   \[
@@
   \]
+  Our setup relies on a quasi-experiment: we look for instances where a team plays in a city
+  with active nightlife, then plays a game in a different location the next day. Since next-day
+  opponent is uncorrelated with exposure to treatment, we have identification in our variable of 
+  interest.
 }
 
-\begin{columns}     \column{0.4}     \block{Data Sources}{
+\begin{columns}     \column{0.6}     \block{Data Sources}{
     \paragraph{Betting house data and game outcomes}
@@
     }
-    \column{0.6}     
+    \column{0.4}     
     \block{NBA model}{
-      Here,
       % Table created by stargazer v.5.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu % Date and time: Sat, May 06, 2017 - 09:58:05 AM
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & outcome == "W" \\  \hline \\[-1.8ex]   party & $-$0.052$^{**}$ (0.022) \\    lag.chg.pos & 0.0001 (0.002) \\    travel.dist & $-$0.00002 (0.00004) \\    ew & $-$0.035 (0.047) \\    nhours.lgame & 0.001 (0.001) \\    I(hour(date)) & 0.006 (0.012) \\    travel.dist:ew & 0.00004 (0.00005) \\    Constant & $-$0.144 (0.348) \\   \textit{N} & 9,979 \\  Log Likelihood & $-$6,912.426 \\  Akaike Inf.
+Crit.
+& 13,840.850 \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
 
@@
   \block{Player Specific Model}{
-    Here
+
+    \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & demeaned.reb.per.min \\  \hline \\[-1.8ex]   party & $-$0.001$^{*}$ (0.0005) \\    lag.chg.pos & 0.00003 (0.0001) \\    travel.dist & 0.00000$^{**}$ (0.00000) \\    nhours.lgame & 0.00002 (0.00002) \\    Constant & $-$0.003$^{**}$ (0.002) \\   \textit{N} & 44,190 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0003 \\  Residual Std.
+Error & 0.105 (df = 44185) \\  F Statistic & 3.857$^{***}$ (df = 4; 44185) \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
+
   }
@@
   \block{MLB Model}{
-    Here
+
+    \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & I(team.score \textgreater  opponent.score) \\  \hline \\[-1.8ex]   party & $-$0.034$^{*}$ (0.019) \\    odds & 2.602$^{***}$ (0.162) \\    I(team == location) & 0.028 (0.033) \\    ndays.lgame & 0.012 (0.019) \\    travel.dist & 0.00000 (0.00003) \\    weekend & 0.042 (0.036) \\    Constant & $-$1.333$^{***}$ (0.080) \\   \textit{N} & 25,215 \\  Log Likelihood & $-$17,274.460 \\  Akaike Inf.
+Crit.
+& 34,562.920 \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
+
   } \end{columns}
@@
     }
-    \note[         
-      targetoffsetx=-8cm,          
-      targetoffsety=-6.5cm,          
-      width=0.5\linewidth         
-    ]{e-mail: \texttt{santucci@stanford.edu}}   \end{columns}
 
@@
   \begin{tikzfigure}
-    \includegraphics[width=0.3\textwidth]{figures/gh_commits}
+    \includegraphics[width=0.6\linewidth]{figures/gh_commits}
   \end{tikzfigure}
-  \begin{tikzfigure}     \includegraphics[width=0.7\textwidth]{figures/punch_card}    \end{tikzfigure}
 }
+\note[         
+  targetoffsetx=-8cm,          
+  targetoffsety=-8.5cm,          
+  width=0.5\linewidth         
+]{\texttt{elax13@stanford.edu}, \, \hspace{3pt} \texttt{santucci@stanford.edu}}
+
+\end{columns}
 

### 2017-05-08 init_model.rnw
Added in plot of average return as a function of threshold for MLB.

@@
 <<>>=
+
 HOURS <- 24
 lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians), 0)]
+
 @
@@
 
+# stargazer(m, style = 'qje', single.row = T)
 @
@@
 <<>>=
+
 HOURS <- 36
-m <- glm(outcome == 'W' ~ I(nmusicians*(nhours.lgame <= HOURS)), 
-         data = lines[season < 2017 & last.game.loc != team])
+m <- glm(outcome == 'W' ~ party, 
+         data = lines[season < 2016 & last.game.loc != team])
 
 ### We're only interested in making predictions when teams party.
-p <- predict(m, newdata = lines[season == 2017 & nhours.lgame <= HOURS & 
+p <- predict(m, newdata = lines[season == 2016 & nhours.lgame <= HOURS & 
                                 last.game.loc != team], type = 'response')
 
-table(round(p), lines[season == 2017 & nhours.lgame <= HOURS & 
+idx <- which(p < 0.476)
+out <- table(lines[season == 2016 & nhours.lgame <= HOURS & last.game.loc != team][idx, outcome])
+
+(out['L'] * 100 - 110*out['W']) %>% as.numeric
+
+table(round(p), lines[season == 2016 & nhours.lgame <= HOURS & 
                       last.game.loc != team, outcome])
@@
 <<echo=F>>=
-require(mboost)
 lines[, game.time := hour(date)]
 
-lm(demeaned.reb.per.min ~ party + lag.chg.pos + travel.dist + nhours.lgame,
-   data = data[last.game.loc != team] %>% na.omit) %>%
-   summary
+m <- lm(demeaned.reb.per.min ~ party + lag.chg.pos + travel.dist + nhours.lgame,
+   data = data[last.game.loc != team] %>% na.omit)
+
+#stargazer(m, style = 'qje', single.row = T)
 @
@@
 summary(m)
+# stargazer(m, style = 'qje', single.row = T)
 @
@@
 points(x = which.min(running.prof), y = min(running.prof), type = 'p')
-text(x = which.min(running.prof), y = min(running.prof), labels = min(running.prof), pos = 2)
+text(x = which.min(running.prof), y = min(running.prof), labels = min(running.prof), pos = 4)
+
+### the following plot shows sufficient deviation from house odds to motivate
+### an optimized betting strategy.
+hist(p- lines[year(date) == 2016, odds])
+
+tholds <- seq(.01, .1, by = .005)
+tmp <- sapply(tholds, function(t) {
+    p1 <- mapply(Predict, prediction = p[idx], 
+                 vegas = lines[year(date) == 2016, odds][idx],
+                 MoreArgs=list(threshold = t))
+    b <- mapply(Bet, our.prediction = p1,
+                actual.outcome = lines[year(date) == 2016, team.score > opponent.score][idx],
+                odds = lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
+    return(c(sum(b, na.rm = T) / sum(!is.na(b)), sum(!is.na(b)))) # Avg profit per game.
+})
+
+plot(x = tholds, y = tmp[1, ], main = 'Return per game\nas a function of threshold', xlab = 'threshold', ylab = 'average profit')
+text(x = tholds, y = tmp[1, ], labels = tmp[2, ], pos = 1)
+
+### plot return per game as a function of difference in odds between us and vegas.
 

### 2017-05-08 poster/poster.tex
Formatting. Crucial typo fix in title.

@@
 \usepackage{subcaption}
- \newtheorem{theorem}{Theorem}[section] \newtheorem{lemma}[theorem]{Lemma} \newtheorem{proposition}[theorem]{Proposition} \newtheorem{corollary}[theorem]{Corollary}  \DeclareMathOperator*{\argmin}{arg\,min} \DeclareMathOperator*{\argmax}{arg\,max} \newcommand{\D}{\mathrm{d}} \SetKwInput{KwInput}{Input} \SetKwInput{KwOutput}{Output}   \title{The effects of night-light on subsequent day sports performance} \author{Eric Lax, Andreas Santucci} \date{May 19th, 2017} \institute{Institute for Computational and Mathematical Engineering}   \usetheme{Board}   \begin{document}   \maketitle
+ \newtheorem{theorem}{Theorem}[section] \newtheorem{lemma}[theorem]{Lemma} \newtheorem{proposition}[theorem]{Proposition} \newtheorem{corollary}[theorem]{Corollary}  \DeclareMathOperator*{\argmin}{arg\,min} \DeclareMathOperator*{\argmax}{arg\,max} \newcommand{\D}{\mathrm{d}} \SetKwInput{KwInput}{Input} \SetKwInput{KwOutput}{Output}   \title{The effects of night-life on subsequent day sporting performance} \author{Eric Lax, Andreas Santucci} \date{May 19th, 2017} \institute{Institute for Computational and Mathematical Engineering}   \usetheme{Board}   \begin{document}   \maketitle
 
@@
       % Table created by stargazer v.5.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu % Date and time: Sat, May 06, 2017 - 09:58:05 AM
+\begin{center}
 \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & outcome == "W" \\  \hline \\[-1.8ex]   party & $-$0.052$^{**}$ (0.022) \\    lag.chg.pos & 0.0001 (0.002) \\    travel.dist & $-$0.00002 (0.00004) \\    ew & $-$0.035 (0.047) \\    nhours.lgame & 0.001 (0.001) \\    I(hour(date)) & 0.006 (0.012) \\    travel.dist:ew & 0.00004 (0.00005) \\    Constant & $-$0.144 (0.348) \\   \textit{N} & 9,979 \\  Log Likelihood & $-$6,912.426 \\  Akaike Inf.
@@
 & 13,840.850 \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
+\end{center}
 
@@
 
-\begin{columns}   \column{0.4}
+\begin{columns}   \column{0.5}
   \block{Player Specific Model}{
 
+\begin{center}
     \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & demeaned.reb.per.min \\  \hline \\[-1.8ex]   party & $-$0.001$^{*}$ (0.0005) \\    lag.chg.pos & 0.00003 (0.0001) \\    travel.dist & 0.00000$^{**}$ (0.00000) \\    nhours.lgame & 0.00002 (0.00002) \\    Constant & $-$0.003$^{**}$ (0.002) \\   \textit{N} & 44,190 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0003 \\  Residual Std.
 Error & 0.105 (df = 44185) \\  F Statistic & 3.857$^{***}$ (df = 4; 44185) \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
+\end{center}
 
   }
-  \column{0.6}
+  \column{0.5}
   \block{MLB Model}{
 
+\begin{center}
     \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & I(team.score \textgreater  opponent.score) \\  \hline \\[-1.8ex]   party & $-$0.034$^{*}$ (0.019) \\    odds & 2.602$^{***}$ (0.162) \\    I(team == location) & 0.028 (0.033) \\    ndays.lgame & 0.012 (0.019) \\    travel.dist & 0.00000 (0.00003) \\    weekend & 0.042 (0.036) \\    Constant & $-$1.333$^{***}$ (0.080) \\   \textit{N} & 25,215 \\  Log Likelihood & $-$17,274.460 \\  Akaike Inf.
@@
 & 34,562.920 \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
-
+\end{center}
   } \end{columns}
@@
   targetoffsetx=-8cm,          
-  targetoffsety=-8.5cm,          
+  targetoffsety=-7.8cm,          
   width=0.5\linewidth         

### 2017-05-08 functions.r
Now accounting for Bovada's 10$ cut.

@@ -40,9 +40,9 @@ Bet <- function(our.prediction, actual.outcome, odds) {
     if (is.na(our.prediction)) return(NA_integer_)
     if (our.prediction == actual.outcome) {
         if (our.prediction > 0)
-            return(100 / odds - 100 - 5)
+            return(100 / odds - 100 - 10)
         else if (our.prediction == 0)
-            return(100 / (1-odds) - 100 - 5)
+            return(100 / (1-odds) - 100 - 10)
     } else if (our.prediction != actual.outcome)
         return(-100)
 }

### 2017-05-09 init_model.rnw
Now saving mlb_bets.pdf

@@
 ### 
+pdf('mlb_bets.pdf')
 betting.days <- lines[year(date) == 2016][idx, date]
@@
 text(x = which.min(running.prof), y = min(running.prof), labels = min(running.prof), pos = 4)
+dev.off()
 

### 2017-05-09 poster/poster.tex
Formatting to ensure everything fits on one page.

@@
 
-\block{Idea} {     It's not unreasonable to think that game performance can be affected (in part) by what takes place off the court.
-We are interested to see if teams exhibit a decline in performance the day following a game in a party city.
+\block{~} {     It's not unreasonable to think that game performance can be affected (in part) by what takes place off the court.
+We are interested to see if teams exhibit a decline in performance the day following a game in a city with active nightlife.
 We exploit data on bookmaker spreads, the expected score differential between two teams after conditioning on observable performance.
 We expect a team to meet the spread half the time, since this is how bookmakers minimize risk on their end.
-We construct a model which attempts to estimate the causal effect of  visiting a "party-city" the day before a game on the probability of meeting the spread.
+We construct a model which attempts to estimate the causal effect of  visiting a "party-city" on subsequent day performance as measured by the spread.
+Since next-day
+  opponent is uncorrelated with exposure to treatment, we have identification in our variable of 
+  interest.
 }
@@
   \]
-  Our setup relies on a quasi-experiment: we look for instances where a team plays in a city
-  with active nightlife, then plays a game in a different location the next day. Since next-day
-  opponent is uncorrelated with exposure to treatment, we have identification in our variable of 
-  interest.
 }
@@
       at its worst we are out of pocket \$750 dollars. By season end, we have recouped all of
-      our initial investment plus an additional \$13.5k.
+      our initial investment plus an additional \$11.5k.
     }

### 2017-05-20 init_model.rnw
Now creating a betting model for each year in 2012-2016.

--- 2017-05-08 init_model.rnw
@@
 <<echo=2:4>>=
+
+pdf('mlb_bets_by_season.pdf')
+par(mfrow = c(3,2))
+lapply(2012:2016, function(yr) {
+    m <- glm(I(team.score > opponent.score) ~ party + odds,
+             data = lines[year(date) < yr], family = 'binomial')
+    p <- predict(m, newdata = lines[year(date) == yr], type = 'response')
+    idx <- which(p < .45)
+    p1 <- mapply(Predict, prediction = p[idx], 
+                 vegas = lines[year(date) == yr, odds][idx],
+                 MoreArgs=list(threshold = 0.01))
+    b <- mapply(Bet, our.prediction = p1,
+                actual.outcome = lines[year(date) == yr, team.score > opponent.score][idx],
+                odds = lines[year(date) == yr, odds][idx], SIMPLIFY = T)
+    cat(paste0(sum(b, na.rm = T), '\n'))
+    ### 
+#    pdf(paste0('mlb_bets_', yr, '.pdf'))
+    betting.days <- lines[year(date) == yr][idx, date]
+    running.prof <- cumsum(ifelse(is.na(b), 0, b))
+    plot(x = 1:length(b), y = running.prof, 
+         main = paste0('Profits over the ', yr, ' season\n',
+                       'Training on 2011-', yr-1, ' data'),
+         xlab = 'Game Number', ylab = 'Profit', type = 'l')
+    abline(a = 0, b = 0, lty = 3)
+    points(x = which.min(running.prof), y = min(running.prof), type = 'p')
+    text(x = which.min(running.prof), y = min(running.prof), 
+         labels = round(min(running.prof)), pos = 4)
+#    dev.off()
+})
+dev.off()
+
+@
+
+\includegraphics{mlb_bets_by_season}
+
+If we were to bet one hundred dollars on each game in 2016,
+we would earn a profit of about \Sexpr{format(round(sum(b, na.rm=T), -1), big.mark = ',')} dollars
+over the course of \Sexpr{sum(!is.na(b))} games.
+
+<<>>=
+
 m <- glm(I(team.score > opponent.score) ~ party + odds,
@@
 p <- predict(m, newdata = lines[year(date) == 2016], type = 'response')
-
 idx <- which(p < .45)
@@
             odds = lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
-
-sum(b, na.rm = T)
-
-### 
-betting.days <- lines[year(date) == 2016][idx, date]
-running.prof <- cumsum(ifelse(is.na(b), 0, b))
-plot(x = 1:length(b), y = running.prof, 
-     main = 'Profits over the 2016 season', 
-     xlab = 'Game Number', ylab = 'Profit', type = 'l')
-abline(a = 0, b = 0, lty = 3)
-points(x = which.min(running.prof), y = min(running.prof), type = 'p')
-text(x = which.min(running.prof), y = min(running.prof), labels = min(running.prof), pos = 4)
 
@@
 
-If we were to bet one hundred dollars on each game in 2016,
-we would earn a profit of about \Sexpr{format(round(sum(b, na.rm=T), -1), big.mark = ',')} dollars
-over the course of \Sexpr{sum(!is.na(b))} games.
-
 \begin{thebibliography}{9}

### 2017-05-22 writing/draft.tex
Init commit.

%% This file is to be used as a template for your submission. 
%% Rename this file and replace the text with the text of 
%% your manuscript.
%%
%% The standard LaTeX document class "article" is recommended. 
%% Use options letterpaper and 12pt.
\documentclass[letterpaper,12pt]{article}

%% This is the recommended preamble for your document.

%% Load De Gruyter specific settings 
\usepackage{dgjournal}

%% The mathptmx package is recommended for Times compatible math symbols.
%% Use mtpro2 or mathtime instead of mathptmx if you have the commercially
%% available MathTime fonts.
%% Other options are txfonts (free) or belleek (free) or TM-Math (commercial)
\usepackage{mathptmx}

%% Use the graphics package to include figures
\usepackage{graphics}

%% Use natbib with these recommended options
\usepackage[authoryear,comma,longnamesfirst,sectionbib]{natbib}

%% Start your document body here
\begin{document}

%% Do NOT include any fronmatter information; including the title, author names,
%% institutes, acknowledgments and title footnotes (author information, funding
%% sources, etc.). Start the document with the first section or paragraph of
%% the article.

\section{Introduction}
\section{Previous Research}
\section{Materials and Methods}
\subsection{Data}
\section{Results}
\section{Conclusions}
\section{References}

\section{This should be the first section of your document}

Your text goes here. Start with the first section or paragraph of your
article. Do not set use either headers or footers and do not set any running
heads or change any page numbers.

The title page along with headers and footers will be inserted by the EdiKit
system. Use the ``revise manuscript'' link to enter this information in the
EdiKit system.

Use the standard LaTeX commands to set the text of the article. The dgjournal
package works with any of the standard LaTeX document classes - article, report
or book. Other document classes should work but compliance with De Gruyter
requirements is not assured. For more information on LaTeX, see
\cite{lamport,mittelbach,oetiker}. \cite{mittelbach} is highly recommended.

Use the standard LaTeX sectioning commands for your headings.

\subsection{A second order heading}

Some text under the subheading. Paragraphs that follow heads are not
indented.

Math should also be set in Times. Use the mathptmx package if you do not have
any of the commercially available fonts that are compatible with Times.
\begin{equation}
    y^{(n)} = \sum_{i=0}^{n-1} a_i(x) y^{(i)} + r(x) 
\end{equation}

All environments provided by the standard LaTeX document classes are
unchanged. Vertical spaces within lists have been altered to comply with De Gruyter
requirements.
\begin{enumerate}
\item This is the first item within the list. Some more text here in order to
  display the alignment.
\item Another item in the list.
\item Yet another item in the list.
\end{enumerate}

Here is an example of a Figure. It's the same as in standard LaTeX.

\begin{figure}[!h]
%% Use the graphics package to insert figures
%% \includegraphics{figure.eps}
% Use \centering to center the table
\centering
%% A small box in place of a figure
\framebox{%
  \begin{minipage}{10pc}
    \begin{center}
      \vspace{1cm}\par
      A figure\par
      \vspace{1cm}
    \end{center}
  \end{minipage}}
\caption{Insert your caption here. If you wish to label your figure for
  cross-referencing, use a label either within the caption or after it.}
\label{fig1}
\end{figure}

An example of a table follows. This is also the same as in standard LaTeX.

\begin{table}[!h]
% Use \centering to center the table
\centering
\caption{Insert your table caption here. If you wish to label the table for
  cross-referencing, use a label either within the caption or after it.}
\begin{tabular}{llll}
\hline
Symbol        & LaTeX Command      & Symbol      & LaTeX Command \\
\hline
$\alpha$      & \verb+\alpha+      & $\zeta$     & \verb+\zeta+ \\
$\beta$       & \verb+\beta+       & $\eta$      & \verb+\eta+ \\
$\gamma$      & \verb+\gamma+      & $\theta$    & \verb+\theta+ \\
$\delta$      & \verb+\delta+      & $\vartheta$ & \verb+\vartheta+ \\
$\epsilon$    & \verb+\epsilon+    & $\iota$     & \verb+\iota+ \\
$\varepsilon$ & \verb+\varepsilon+ & $\kappa$    & \verb+\kappa+ \\
\hline
\end{tabular}
\end{table}

Use the \verb+thebibliography+ environment for the references.  BibTeX users may
use the provided BibTeX style file DeGruyter.bst.

%% BibTeX support
\bibliographystyle{DeGruyter}
\bibliography{sample}

\end{document}

### 2017-05-24 writing/draft.tex
Added a skeleton outline and some rough thoughts to Intro, and Data sections.

@@
 \usepackage{dgjournal}
+
+\usepackage{hyperref}
 
@@
 \section{Introduction}
+It's not unreasonable to think that game performance can be affected 
+(in part) by what takes place off the court. We are interested to see if 
+teams exhibit a decline in performance the day following a game in a city 
+with active nightlife. We exploit data on bookmaker spreads, the expected score 
+differential between two teams after conditioning on observable performance. 
+We expect a team to meet the spread half the time, since this is how bookmakers 
+minimize risk on their end. We construct a model which attempts to 
+estimate the causal effect of  visiting a "party-city" on subsequent day 
+performance as measured by the spread. Since next-day
+opponent is uncorrelated with exposure to treatment, we have identification in our variable of 
+interest.
+
 \section{Previous Research}
+
 \section{Materials and Methods}
 \subsection{Data}
+\subsubsection{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
+
+\subsubsection{BLS data} We also obtained data from
+\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the county-quarter level. As a 
+proxy for how much night-life there is in a city, we look toward the 
+number of sound recording studios, musical groups, and music publishers there 
+are for a particular county-year: we simply calculate the average number of establishments across all three music categories listed above.\footnote{We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.}
+
+We then merge this in with our lines data, taking care to do so according to each season-team's last game location; we also lag our BLS data by one year.
+To this data, we've created features such as 
+the number of days since last game, the last game location, the travel distance, and
+average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}. Our data range from 2010-2011 season through present, which includes part of the 2016-2017 season.
+
+\subsubsection{Minutes data}
+We also gathered minutes data from \href{http://www.espn.com/nba/scoreboard/_/date/}{ESPN}.
+
+The minutes data also contain information on the number of free throws, field goals, and
+three pointers made and attempted by each player in each game.
+
+\subsubsection{Which cities (don't) have nightlife?}
+
+We examine which cities have a notable nightlife, as measured by our proxy;
+bottom and top five.
+
+Los Angeles stands out, as does New York. Some runner ups include Chicago and Miami.
+
+\subsection{Method: Defining a continuous measure of night life}
+
+We have a free parameter: the time-window which we believe the hangover effect
+to last. Suppose we only look toward games which have been played in a party
+city at most 24 hours ago. We take care to ignore
+teams playing in their own cities.
+
+\subsection{Possible confounders}
+What if we pick up effects of a busy play schedule?
+Aside from accounting for how much rest time a team is allowed in between games,
+and travel distance between games, we're also interested to get a measure of
+fatigue incurred on the court.
+We account for this by bringing in minutes data, and counting the number of 
+times the players run back and forth on the court. This is crudely measured
+by the number of changes in posession. We back out this quantity.
+
 \section{Results}
+
 \section{Conclusions}
+
 \section{References}

### 2017-05-27 writing/draft.tex
Init commit.

@@
 
+Citations toward anecdotal. % The ``JR Smith'' effect.
+
 \section{Previous Research}
+Studies on effects of alcohol on sports performance in a controlled setting.
 
 \section{Materials and Methods}
+Determining when players partied. One might consider to look at Twitter data
+and or night clubs per city.
+
+Insights: major party cities are LA and NY, and athletes party is well correlated with musicians. This led to two separate definitions of party. Our most basic analysis was to consider performance 
+the day after visiting LA or NY.
+We also considered the number of musical establishments per city.
+
+\paragraph{Logic for using spread data}
+The logic for using spread data: it controls for a lot, but we don't know exactly
+what it controls for.
+
 \subsection{Data}
@@
 \section{Results}
+
+\subsection{Extending the research}
+We can also validate where points are lost in MLB in a similar way.
+
+\subsection{Betting performance}
 

### 2017-05-29 writing/draft.tex
Worked with Eric on first draft.

--- 2017-05-24 writing/draft.tex
@@
 It's not unreasonable to think that game performance can be affected 
-(in part) by what takes place off the court. We are interested to see if 
+(in part) by what takes place off the court. 
+Anecdotally, it has been reported that professional sports teams party on 
+the road, sometimes to the detriment of the game.\cite{hoch}
+J.R. Smith is infamous for citing his improvement in performance due to a
+lack of nightlife when moving to Cleveland from New York.\cite{price,ley}
+A few interviews with former professional athletes revealed that location
+matters more than day of the week, i.e. parties can happen any day of the week,
+but they are more likely to occur in select cities.
+
+We are interested to see if 
 teams exhibit a decline in performance the day following a game in a city 
@@
 \section{Previous Research}
+In 2014, Ezekowitz tried to answer the question we're interested in.\cite{ezekowitz} 
+However, the study only
+looks at performance in party cities, as opposed to subsequent day performance (i.e.
+a hangover effect). Moreover, their sample size was much smaller. Finally, there was no
+methodology used behind the selection of party cities. Between these factors, no
+conclusion was reached with statistical significance. The data do weakly suggest
+that teams are less likely to win on the road in party cities.
+
+In 2015, Chase also looked at a statistical ranking of the biggest partiers in
+the NBA.\cite{chase}
+
+Studies on effects of alcohol on sports performance in a controlled setting.
+Same goes for sleep deprivation and possibly narcotics. 
+Both of those well correlated with active night life.
 
 \section{Materials and Methods}
+One of the hardest parts of this analysis is that partying is a latent variable.
+In order to Determine when players partied, one might consider to look at Twitter data
+for direct mentions of partying, either by athletes or witnesses. However, due to
+intense public scrutiny that athletes face, there is a strong reluctance to post
+such mentions on social media. There are notable exceptions, such
+as Oddel Beckham Jr.'s trip to Miami in 2017 season.\cite{bleler}
+The reports are few and far enough between,
+and could be biased based on whether a particular city has nightclubs with VIP rooms,
+that picking up statistical evidence would be a fruitless.
+
+\paragraph{Two approaches: a discrete indicator and a continuous measure of nightlife}
+Because famous athletes often enjoy a celebrity lifestyle, using conventional statistics
+such as number of drinking establishments in each city does not work; athletes instead
+tend to visit high end nightclubs. We again turned to player interviews for insight,
+and found two major insights. We obtained a listing of cities which were anecdotally
+mentioned as having active night-life. Between players, the rank ordering of these
+was consistent to the extent that LA and NY were top
+offenders. This motivates using a binary indicator for party in these two cities. 
+We also found that athletes tend to party with other A-list celebrities,
+particularly musicians. This motivated a continuous measure of nightlife, based on a
+simple count of the number of musicians. We pursue both approaches.
+
+Insights: major party cities are LA and NY, and athletes party is well correlated with musicians. This led to two separate definitions of party. Our most basic analysis was to consider performance 
+the day after visiting LA or NY.
+We also considered the number of musical establishments per city.
+
+\paragraph{Logic for using spread data}
+In order to assert a causal relation between cities with active nightlife 
+and subsequent day sports performance, we need to control for salient
+features such as the ability of players on either team in the match-up,
+whether any players are injured, the momentum of success each team is carrying
+in the season, etc. To that end, we utilize spread data. The benefit of
+using spread data is that it controls for a lot, but the downside is that
+we don't know exactly what it controls for.
+
+For example, if book-keepers
+are already aware of the effects of night-life on performance, it could
+already be factored into the spread; if this were the case, then our variable
+of interest would not appear to have significant explanatory power
+on the margin, when also conditioning on spread, in spite of there being
+a true effect. Alternatively,
+if the house book-keeper systematically over-valued
+teams which you are more likely to play after a party city,
+we could see a spurious effect.
+That said, the amount of money that is at stake and the effort put into creating these spreads 
+suggests that they are reliable. So, if we do find an effect in our variable of interest,
+we can be confident in our conclusion that there is a causal effect and not some
+other systematic bias.
+
+\subsection{Possible confounders}
+
+\paragraph{Home team effect}
+Home-team advantage is well-studied.\cite{jones07, jones08}  
+Because teams play a lot of their games on the road,
+most games after playing in a party city will be an away game.
+This should be covered by spread data, but we also include
+a home-team indicator to be sure.
+
+\paragraph{Fatigue from previous game}
+What if we pick up effects of a busy play schedule?
+Aside from accounting for how much rest time a team is allowed in between games,
+we're also interested to get a measure of
+fatigue incurred on the court.
+We account for this by bringing in minutes data, and counting the number of 
+times the players run back and forth on the court. This is crudely measured
+by the number of changes in posession. We back out this quantity.
+
+\paragraph{Fatigue from Jet lag}
+What about jet-lag? Prolonged air travel has in general been shown
+to have a negative effect on sports performance.\cite{leeandgalvez}
+Previous studies have shown that jet lag is a strong enough effect
+to erase home field advantage in MLB.\cite{songetal}
+
+\paragraph{The problem of an unknown travel schedule} In NBA, we have the difficulty of
+not observing when teams travel. 
+In MLB, each match-up is actually a series of games.
+
 \subsection{Data}
@@
 
-\subsubsection{Which cities (don't) have nightlife?}
-
-We examine which cities have a notable nightlife, as measured by our proxy;
-bottom and top five.
-
-Los Angeles stands out, as does New York. Some runner ups include Chicago and Miami.
+\subsubsection{Loctaion and Travel Distance Data} is obtained
+by API.
 
@@
 
-\subsection{Possible confounders}
-What if we pick up effects of a busy play schedule?
-Aside from accounting for how much rest time a team is allowed in between games,
-and travel distance between games, we're also interested to get a measure of
-fatigue incurred on the court.
-We account for this by bringing in minutes data, and counting the number of 
-times the players run back and forth on the court. This is crudely measured
-by the number of changes in posession. We back out this quantity.
+\paragraph{Discrete indicator}
+If a team plays in Los Angeles or New York within 24 hours of their current game,
+we indicate one, else zero.
+
+\paragraph{Continuous measure}
+If the last game was played at most 36 hours ago, use log average number
+of musicians in last game location, else zero.
+
+\subsection{Model} We use a logistic regression to estimate the likelihood of meeting
+the spread as a function of whether the team partied as well as controlling for other confounders
+listed above.
 
 \section{Results}
+\subsection{Using spread data}
+\begin{center}
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & outcome == "W" \\  \hline \\[-1.8ex]   party & $-$0.052$^{**}$ (0.022) \\    lag.chg.pos & 0.0001 (0.002) \\    travel.dist & $-$0.00002 (0.00004) \\    ew & $-$0.035 (0.047) \\    nhours.lgame & 0.001 (0.001) \\    I(hour(date)) & 0.006 (0.012) \\    travel.dist:ew & 0.00004 (0.00005) \\    Constant & $-$0.144 (0.348) \\   \textit{N} & 9,979 \\  Log Likelihood & $-$6,912.426 \\  Akaike Inf.
+Crit.
+& 13,840.850 \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
+\end{center}
+
+\subsection{Using performance statistics}
+We additionally looked at actual performance instead of relying on spread data. 
+This allows us to avoid using the black-box of spread data, and also pin-point
+more exactly what elements of the game are in fact effected by partying.
+This motivates looking at number of points allowed, number of points scored.
+
+We also considered rebounds, fouls, and even injuries. 
+\begin{center}
+    \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & demeaned.reb.per.min \\  \hline \\[-1.8ex]   party & $-$0.001$^{*}$ (0.0005) \\    lag.chg.pos & 0.00003 (0.0001) \\    travel.dist & 0.00000$^{**}$ (0.00000) \\    nhours.lgame & 0.00002 (0.00002) \\    Constant & $-$0.003$^{**}$ (0.002) \\   \textit{N} & 44,190 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0003 \\  Residual Std.
+Error & 0.105 (df = 44185) \\  F Statistic & 3.857$^{***}$ (df = 4; 44185) \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
+\end{center}
+
+\subsection{Robustness check}
+We verify that this phenomena persists in the MLB. In MLB, money-lines
+are used instead of point-spreads. So as opposed to predicting
+the likelihood of meeting the spread, we predict the probability
+a team would win against their opponent as both a function of the book-makers'
+odds of winning and party. Additionally, we interacted weekend.
+One of the major differences between NBA and MLB games is that MLB games
+are played in series, so the same match-up repeated multiple times. This remedies
+the problem of an unknown travel schedule we faced with NBA because we are sure 
+teams stay overnight in the same location within a series.
+
+\begin{center}
+    \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & I(team.score \textgreater  opponent.score) \\  \hline \\[-1.8ex]   party & $-$0.034$^{*}$ (0.019) \\    odds & 2.602$^{***}$ (0.162) \\    I(team == location) & 0.028 (0.033) \\    ndays.lgame & 0.012 (0.019) \\    travel.dist & 0.00000 (0.00003) \\    weekend & 0.042 (0.036) \\    Constant & $-$1.333$^{***}$ (0.080) \\   \textit{N} & 25,215 \\  Log Likelihood & $-$17,274.460 \\  Akaike Inf.
+Crit.
+& 34,562.920 \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\ \end{tabular}
+\end{center}
+
+\subsubsection{Betting performance}
+\begin{figure}[!h]
+  \centering
+  \includegraphics{mlb_bets_by_season}
+  \caption{Betting performance by season. In each season, The 
+    circled point represents the worst out of pocket expense incurred during the season.
+    We first train a model on historical data then use said model to predict on ``current year''
+    data, ploting profits as a function of time over the ``current'' season. For each relevant
+    game, we place a bet if our model instructs us of odds sufficiently different from what
+    the betting house suggests. We account for the bookmaker's cut. E.g. in the early 2016 season,
+    at its worst we are out of pocket \$750 dollars. By season end, we have recouped all of
+    our initial investment plus an additional \$11.5k.}
+  \label{bettingperf}
+\end{figure}
+
+\subsection{Extending the research}
+We can also validate where points are lost in MLB in a similar way.
+It would be interesting to look at number of shots taken, and number of shots taken by stars.
+It's possible there are prolonged effects of partying, and applying some sort of kernel
+on the number of hours played since the last game could be interesting. It could also be
+that different (types of) teams or players are more or less susceptible to the effects
+of night-life; this could be demographic characteristics such as age or marital status.
 

### 2017-05-29 writing/draft.tex
Added explanation on previous research on alcohol and sports performance.

@@
 the road, sometimes to the detriment of the game.\cite{hoch}
-J.R. Smith is infamous for citing his improvement in performance due to a
+Within the NBA, J.R. Smith is infamous for citing his improvement in performance due to a
 lack of nightlife when moving to Cleveland from New York.\cite{price,ley}
@@
 minimize risk on their end. We construct a model which attempts to 
-estimate the causal effect of  visiting a "party-city" on subsequent day 
+estimate the causal effect of  visiting a ``party-city'' on subsequent day 
 performance as measured by the spread. Since next-day
@@
 \section{Previous Research}
-In 2014, Ezekowitz tried to answer the question we're interested in.\cite{ezekowitz} 
+There has been an attempt to answer this question before.\cite{ezekowitz} 
 However, the study only
-looks at performance in party cities, as opposed to subsequent day performance (i.e.
-a hangover effect). Moreover, their sample size was much smaller. Finally, there was no
-methodology used behind the selection of party cities. Between these factors, no
-conclusion was reached with statistical significance. The data do weakly suggest
-that teams are less likely to win on the road in party cities.
+looks at performance in party cities, as opposed to subsequent day performance (i.e. the study does not consider a delayed hangover effect). Moreover, their sample size was much smaller. 
+Finally, there was no methodology used behind the selection of party cities. The author does conclude
+that teams systematically play worse on Sunday afternoons when on the road,
+regardless of cities.
 
-In 2015, Chase also looked at a statistical ranking of the biggest partiers in
-the NBA.\cite{chase}
+Studies have also been carried out on statistical ranking of the players
+most notorious for pursuing active nightlife in the NBA.\cite{chase} The study
+does conclude that players perform worse on Sundays when compared with other days
+as measured by the ratio of shots scored to shots taken.
 
-Studies on effects of alcohol on sports performance in a controlled setting.
+Studies on effects of alcohol on sports performance in a controlled setting
+show that alcohol impairs the nervous system, resulting in decrements in both
+cognitive function and motor skill.\cite{shirreffs} In a study on 
+the effects of alcohol on recovery of male athletes, it was found that
+consumption is correlated with reduced production of testosterone
+which ultimately leads to inhibited ability to recover and adapt to exercise.\cite{mjbarnes}
 Same goes for sleep deprivation and possibly narcotics. 
-Both of those well correlated with active night life.
+These are all well correlated with active night life.
 
@@
 One of the hardest parts of this analysis is that partying is a latent variable.
-In order to Determine when players partied, one might consider to look at Twitter data
+In order to determine when players partied, one might consider to look at social media data
 for direct mentions of partying, either by athletes or witnesses. However, due to
@@
 and could be biased based on whether a particular city has nightclubs with VIP rooms,
-that picking up statistical evidence would be a fruitless.
+that picking up statistical evidence would be fruitless.
 

### 2017-05-29 writing/draft.tex
Added citations toward alcohol and sleep-deprivation. Cleaned up intro to section three on materials and methods.

@@
 \usepackage{hyperref}
+\usepackage{color,soul}
 
@@
 which ultimately leads to inhibited ability to recover and adapt to exercise.\cite{mjbarnes}
-Same goes for sleep deprivation and possibly narcotics. 
-These are all well correlated with active night life.
+Sleep deprivation has also been shown to strongly impair cognitive and motor performance.\cite{pilcher} Both alcohol and sleep deprivation are well correlated with lifestyles associated
+with an active night life.
 
@@
 The reports are few and far enough between,
-and could be biased based on whether a particular city has nightclubs with VIP rooms,
-that picking up statistical evidence would be fruitless.
+and whether they are even reported at all is strongly biased on whether a particular city 
+has nightclubs with VIP rooms, etc; picking up statistical evidence would be fruitless in this case.
 
@@
 such as number of drinking establishments in each city does not work; athletes instead
-tend to visit high end nightclubs. We again turned to player interviews for insight,
-and found two major insights. We obtained a listing of cities which were anecdotally
+tend to visit high end nightclubs. We again turned to player interviews for ideas,
+and found two major insights. From our conversations, 
+we obtained a listing of cities which were anecdotally
 mentioned as having active night-life. Between players, the rank ordering of these
@@
 particularly musicians. This motivated a continuous measure of nightlife, based on a
-simple count of the number of musicians. We pursue both approaches.
-
-Insights: major party cities are LA and NY, and athletes party is well correlated with musicians. This led to two separate definitions of party. Our most basic analysis was to consider performance 
-the day after visiting LA or NY.
-We also considered the number of musical establishments per city.
+simple count of the number of musicians in the county in which the stadium resides. 
+We pursue both approaches.
 
@@
 of interest would not appear to have significant explanatory power
-on the margin, when also conditioning on spread, in spite of there being
+after conditioning on spread, in spite of there being
 a true effect. Alternatively,
@@
 we could see a spurious effect.
-That said, the amount of money that is at stake and the effort put into creating these spreads 
+
+\hl{Careful} That said, the amount of money that is at stake and the effort put into creating these spreads 
 suggests that they are reliable. So, if we do find an effect in our variable of interest,

### 2017-05-30 writing/draft.tex
Major updates through results section. Added party_discrete results.

@@
 \usepackage{color,soul}
+\usepackage{amsmath}
 
@@
 whether any players are injured, the momentum of success each team is carrying
-in the season, etc. To that end, we utilize spread data. The benefit of
+in the season, etc.\cite{anderson} To that end, we utilize spread data. The benefit of
 using spread data is that it controls for a lot, but the downside is that
@@
 we could see a spurious effect. 
-
-\hl{Careful} That said, the amount of money that is at stake and the effort put into creating these spreads 
+That said, the amount of money that is at stake and the effort put into creating these spreads 
 suggests that they are reliable. So, if we do find an effect in our variable of interest,
@@
 
-\paragraph{The problem of an unknown travel schedule} In NBA, we have the difficulty of
-not observing when teams travel. 
-In MLB, each match-up is actually a series of games.
+\paragraph{The problem of an unknown travel schedule} In NBA, we have the difficulty of not observing when teams travel. E.g. if a particular team  plays
+in Los Angeles on Friday and another game in New York the following day, we don't
+know whether the team traveled on Friday evening or Saturday morning. Heuristics for
+travel schedules can be made, but it is hard to obtain reliable team-travel data.
+In MLB, on the other hand, each match-up is actually a series of games. E.g.
+if the Dodgers play the Giants in San Francisco, they will play several games spread
+across several days, each in the same city; in these cases, we can infer that the team
+is in fact spending the night in said city.
 
 \subsection{Data}
-\subsubsection{Lines data} We obtained data from \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{Covers.com}, which records the outcome of each game alongside the line set by a betting house.
+\paragraph{Lines data} We first look toward 
+\href{http://www.basketball-reference.com/leagues/}{basketball-reference.com}
+for a listing of game-days by season. With this listing, we queried \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{covers.com},
+for all games occurring on a particular game, and for each we observe
+outcome of each game alongside the line set by a betting house. We obtain
+data for seasons 2010-11 through 2016-17.
 
-\subsubsection{BLS data} We also obtained data from
+\paragraph{BLS data} We also obtained data from
 \href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the county-quarter level. As a 
@@
 number of sound recording studios, musical groups, and music publishers there 
-are for a particular county-year: we simply calculate the average number of establishments across all three music categories listed above.\footnote{We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.}
+are for a particular county-year: we simply calculate the average number of establishments across all three music categories listed above.
+We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.
 
-We then merge this in with our lines data, taking care to do so according to each season-team's last game location; we also lag our BLS data by one year.
+We then merge this in with our lines data, taking care to do so according to each season-team's last game location such that the feature captures the proper
+party effect. In addition, we lag our BLS data by one year for two reasons:
+realistically, player expectations of which cities have the most active night-life may not
+update in real-time and also doing so allows us to create a betting model, i.e.
+we don't rely on any features we can't obtain before making a game-time prediction. 
 To this data, we've created features such as 
-the number of days since last game, the last game location, the travel distance, and
-average age\footnote{We use $\log \textrm{\# days old}$, at the time the game is played.}. Our data range from 2010-2011 season through present, which includes part of the 2016-2017 season.
+the number of days since last game, the last game location, and the travel distance. 
+Our data range from 2010-2011 season through present, which includes part of the 2016-2017 season.
 
-\subsubsection{Minutes data}
+\paragraph{Minutes data}
 We also gathered minutes data from \href{http://www.espn.com/nba/scoreboard/_/date/}{ESPN}.
+The minutes data also contain information on the number of free throws, field goals, and three pointers made and attempted by each player in each game.
 
-The minutes data also contain information on the number of free throws, field goals, and
-three pointers made and attempted by each player in each game.
+\paragraph{Location and Travel Distance Data} We are also interested
+to account for the effects of jet-lag. For each team, we queried
+a maps search engine to obtain a longitude and latitude listing.\cite{ggmap}
+For each pairwise set of coordinates, we calculated distance between the two
+points.\cite{sp} To account for jet-lag, we first computed the final direction 
+(bearing) when arriving at the 
+second point after starting from the first and following the shortest path on a sphere 
+(following a great circle).\cite{lallensack}
 
-\subsubsection{Loctaion and Travel Distance Data} is obtained
-by API.
-
-\subsection{Method: Defining a continuous measure of night life}
-
+\subsection{Method: Defining what it means for a city to have nightlife}
 We have a free parameter: the time-window which we believe the hangover effect
@@
 city at most 24 hours ago. We take care to ignore
-teams playing in their own cities.
+teams playing in their own cities, i.e. if a team's last game was played in at home 
+we say that they were not exposed to any nightlife.
 
 \paragraph{Discrete indicator}
-If a team plays in Los Angeles or New York within 24 hours of their current game,
-we indicate one, else zero.
+Our first sanity check was to define an indicator variable for when a team plays in 
+Los Angeles or New York within 24 hours of their current game, we indicate one, else zero. Anecdotally, these are mentioned as hot-spots for
+nightlife.
 
 \paragraph{Continuous measure}
-If the last game was played at most 36 hours ago, use log average number
-of musicians in last game location, else zero.
+We also defined a continuous measure of nightlife:
+if the last game was played at most 24 hours ago, use the log of the average number
+of musicians in last game location, else emit zero. Intuitively, this captures
+how much nightlife there is in a city since musical performances are correlated
+with parties, if not an essential feature.
 
@@
 the spread as a function of whether the team partied as well as controlling for other confounders
-listed above.
+listed above. Our basic model is as follows.
+
+\[
+\texttt{meet spread} \,  \sim \, \substack{\texttt{days since} \\ \texttt{last game}} \, + \, \texttt{jetlag} \, + \, \substack{\texttt{home team} \\ \texttt{effect}} \, + \, \texttt{party}
+\]
 
@@
 \subsection{Using spread data}
-\begin{center}
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & outcome == "W" \\  \hline \\[-1.8ex]   party & $-$0.052$^{**}$ (0.022) \\    lag.chg.pos & 0.0001 (0.002) \\    travel.dist & $-$0.00002 (0.00004) \\    ew & $-$0.035 (0.047) \\    nhours.lgame & 0.001 (0.001) \\    I(hour(date)) & 0.006 (0.012) \\    travel.dist:ew & 0.00004 (0.00005) \\    Constant & $-$0.144 (0.348) \\   \textit{N} & 9,979 \\  Log Likelihood & $-$6,912.426 \\  Akaike Inf.
-Crit.
-& 13,840.850 \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
-\end{center}
+We find that after accounting for player fatigue incurred during the last game, jetlag and travel fatigue, and recovery time, and home-team effect, that both our discrete indicator and continuous measure of party
+have a statistically significant effect on likelihood of meeting the spread.
 
@@
 This motivates looking at number of points allowed, number of points scored.
+We also considered rebounds, fouls, and even injuries.
 
-We also considered rebounds, fouls, and even injuries. 
-\begin{center}
-    \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & demeaned.reb.per.min \\  \hline \\[-1.8ex]   party & $-$0.001$^{*}$ (0.0005) \\    lag.chg.pos & 0.00003 (0.0001) \\    travel.dist & 0.00000$^{**}$ (0.00000) \\    nhours.lgame & 0.00002 (0.00002) \\    Constant & $-$0.003$^{**}$ (0.002) \\   \textit{N} & 44,190 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0003 \\  Residual Std.
-Error & 0.105 (df = 44185) \\  F Statistic & 3.857$^{***}$ (df = 4; 44185) \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\  \end{tabular}
-\end{center}
+In particular, we see that the average number of rebounds obtained per minute
+of game-play does decrease as a function of being exposed to a city with active night-life.
+The method was to first calculate for each player and each game the number of rebounds obtained
+per minute of game-play. Then, for each player and each season we subtract from each observation
+the player's average number of rebounds obtained per minute of game-play using the entire season of data. This is an ad-hoc approach that cannot be used for betting.
 
 \subsection{Robustness check}
-We verify that this phenomena persists in the MLB. In MLB, money-lines
+We verify that this phenomena persists in Major League Baseball. In the MLB, money-lines
 are used instead of point-spreads. So as opposed to predicting
@@
 a team would win against their opponent as both a function of the book-makers'
-odds of winning and party. Additionally, we interacted weekend.
+odds of winning and our party variable. Additionally, we interacted our party variable with weekend.
 One of the major differences between NBA and MLB games is that MLB games
@@
 
-\begin{center}
-    \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & I(team.score \textgreater  opponent.score) \\  \hline \\[-1.8ex]   party & $-$0.034$^{*}$ (0.019) \\    odds & 2.602$^{***}$ (0.162) \\    I(team == location) & 0.028 (0.033) \\    ndays.lgame & 0.012 (0.019) \\    travel.dist & 0.00000 (0.00003) \\    weekend & 0.042 (0.036) \\    Constant & $-$1.333$^{***}$ (0.080) \\   \textit{N} & 25,215 \\  Log Likelihood & $-$17,274.460 \\  Akaike Inf.
+% Table created by stargazer v.5.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu % Date and time: Mon, May 29, 2017 - 11:56:17 PM
+\hspace{-105pt}
+\begin{tabular}{@{\extracolsep{5pt}}lcccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{4}{c}{\textit{Dependent variable:}} \\  \cline{2-5}  \\[-1.8ex] & \multicolumn{2}{c}{NBA: Meet Spread} & demeaned reb/min & MLB: Win Game \\  \\[-1.8ex] & \multicolumn{2}{c}{\textit{logistic}} & \textit{OLS} & \textit{logistic} \\  \\[-1.8ex] & (1) & (2) & (3) & (4)\\  \hline \\[-1.8ex]  party.discrete & $-$0.406$^{***}$ (0.133) &  &  &  \\    party &  & $-$0.053$^{**}$ (0.022) & $-$0.001$^{*}$ (0.0005) & $-$0.035$^{*}$ (0.020) \\    lag.chg.pos & 0.0001 (0.002) & 0.0001 (0.002) & 0.00003 (0.0001) &  \\    odds &  &  &  & 2.602$^{***}$ (0.162) \\    log(travel.dist + 1) & $-$0.017 (0.025) & $-$0.015 (0.025) & 0.001 (0.001) & $-$0.001 (0.005) \\    ew & $-$0.114 (0.230) & $-$0.101 (0.230) &  &  \\    nhours.lgame & 0.001 (0.001) & 0.001 (0.001) & 0.00003 (0.00002) &  \\    I(hour(date)) & 0.007 (0.012) & 0.005 (0.012) &  &  \\    weekend &  &  &  & 0.040 (0.036) \\    I(team == location) & $-$0.011 (0.043) & $-$0.001 (0.043) &  & 0.025 (0.034) \\    log(travel.dist + 1):ew & 0.018 (0.035) & 0.016 (0.035) &  &  \\    ndays.lgame &  &  &  & 0.014 (0.019) \\    Constant & $-$0.077 (0.374) & $-$0.041 (0.374) & $-$0.007$^{*}$ (0.004) & $-$1.331$^{***}$ (0.080) \\   \hline \\[-1.8ex]  Observations & 10,000 & 9,979 & 44,190 & 25,215 \\  R$^{2}$ &  &  & 0.0003 &  \\  Adjusted R$^{2}$ &  &  & 0.0002 &  \\  Log Likelihood & $-$6,925.180 & $-$6,912.606 &  & $-$17,274.450 \\  Akaike Inf.
 Crit.
-& 34,562.920 \\  \hline  \hline \\[-1.8ex]  \textit{Notes:} & \multicolumn{1}{r}{$^{***}$Significant at the 1 percent level.} \\   & \multicolumn{1}{r}{$^{**}$Significant at the 5 percent level.} \\   & \multicolumn{1}{r}{$^{*}$Significant at the 10 percent level.} \\ \end{tabular}
-\end{center}
+& 13,868.360 & 13,843.210 &  & 34,562.900 \\  Residual Std.
+Error &  &  & 0.105 (df = 44185) &  \\  F Statistic &  &  & 3.064$^{**}$ (df = 4; 44185) &  \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
 \subsubsection{Betting performance}
+We also validate our model by seeing how well we can perform against the
+market.
 \begin{figure}[!h]
@@
 
+\clearpage
+
 \subsection{Extending the research}
@@
 
-\section{This should be the first section of your document}
-
-Your text goes here. Start with the first section or paragraph of your
-article. Do not set use either headers or footers and do not set any running
-heads or change any page numbers.
-
-The title page along with headers and footers will be inserted by the EdiKit
-system. Use the ``revise manuscript'' link to enter this information in the
-EdiKit system.
-
-Use the standard LaTeX commands to set the text of the article. The dgjournal
-package works with any of the standard LaTeX document classes - article, report
-or book. Other document classes should work but compliance with De Gruyter
-requirements is not assured. For more information on LaTeX, see
-\cite{lamport,mittelbach,oetiker}. \cite{mittelbach} is highly recommended.
-
-Use the standard LaTeX sectioning commands for your headings.
-
-\subsection{A second order heading}
-
-Some text under the subheading. Paragraphs that follow heads are not
-indented.
-
-Math should also be set in Times. Use the mathptmx package if you do not have
-any of the commercially available fonts that are compatible with Times.
-\begin{equation}
-    y^{(n)} = \sum_{i=0}^{n-1} a_i(x) y^{(i)} + r(x) 
-\end{equation}
-
-All environments provided by the standard LaTeX document classes are
-unchanged. Vertical spaces within lists have been altered to comply with De Gruyter
-requirements.
-\begin{enumerate}
-\item This is the first item within the list. Some more text here in order to
-  display the alignment.
-\item Another item in the list.
-\item Yet another item in the list.
-\end{enumerate}
-
-Here is an example of a Figure. It's the same as in standard LaTeX.
-
-\begin{figure}[!h]
-%% Use the graphics package to insert figures
-%% \includegraphics{figure.eps}
-% Use \centering to center the table
-\centering
-%% A small box in place of a figure
-\framebox{%
-  \begin{minipage}{10pc}
-    \begin{center}
-      \vspace{1cm}\par
-      A figure\par
-      \vspace{1cm}
-    \end{center}
-  \end{minipage}}
-\caption{Insert your caption here. If you wish to label your figure for
-  cross-referencing, use a label either within the caption or after it.}
-\label{fig1}
-\end{figure}
-
-An example of a table follows. This is also the same as in standard LaTeX.
-
-\begin{table}[!h]
-% Use \centering to center the table
-\centering
-\caption{Insert your table caption here. If you wish to label the table for
-  cross-referencing, use a label either within the caption or after it.}
-\begin{tabular}{llll}
-\hline
-Symbol        & LaTeX Command      & Symbol      & LaTeX Command \\
-\hline
-$\alpha$      & \verb+\alpha+      & $\zeta$     & \verb+\zeta+ \\
-$\beta$       & \verb+\beta+       & $\eta$      & \verb+\eta+ \\
-$\gamma$      & \verb+\gamma+      & $\theta$    & \verb+\theta+ \\
-$\delta$      & \verb+\delta+      & $\vartheta$ & \verb+\vartheta+ \\
-$\epsilon$    & \verb+\epsilon+    & $\iota$     & \verb+\iota+ \\
-$\varepsilon$ & \verb+\varepsilon+ & $\kappa$    & \verb+\kappa+ \\
-\hline
-\end{tabular}
-\end{table}
-
-Use the \verb+thebibliography+ environment for the references.  BibTeX users may
-use the provided BibTeX style file DeGruyter.bst.
-
 %% BibTeX support

### 2017-05-31 writing/draft.tex
Solid rough draft.

@@
 the number of days since last game, the last game location, and the travel distance.
-Our data range from 2010-2011 season through present, which includes part of the 2016-2017 season.
 
@@
 second point after starting from the first and following the shortest path on a sphere 
-(following a great circle).\cite{lallensack}
+(following a great circle), which yields a direction in degrees. 
+We then obtain a measure for east-west travel by looking at the trigonometric-$\sin$ of this angle.
+Interacting this with travel-distance gets us a measure of jetlag.\cite{lallensack}
 
@@
 teams playing in their own cities, i.e. if a team's last game was played in at home 
-we say that they were not exposed to any nightlife.
+we say that they were not exposed to an exogenous shock in nightlife.
 
@@
 \subsection{Using performance statistics}
-We additionally looked at actual performance instead of relying on spread data. 
-This allows us to avoid using the black-box of spread data, and also pin-point
-more exactly what elements of the game are in fact effected by partying.
-This motivates looking at number of points allowed, number of points scored.
-We also considered rebounds, fouls, and even injuries.
+We additionally looked at actual performance metrics instead of relying on spread data. 
+This allows us to avoid using a black-box feature, and also pin-point
+more exactly what elements of the game are in fact effected after visiting a city with active 
+nightlife. This motivates looking at number of points allowed, and number of points scored;
+we also considered how rebounds, fouls, and even injuries might be explained (in part) by
+a hangover effect.
 
@@
 The method was to first calculate for each player and each game the number of rebounds obtained
-per minute of game-play. Then, for each player and each season we subtract from each observation
-the player's average number of rebounds obtained per minute of game-play using the entire season of data. This is an ad-hoc approach that cannot be used for betting.
+per minute of game-play. Then, we subtract from each observation the player's average number of rebounds per minute taken across the entire season.
+This is an ad-hoc approach that cannot be used for betting, since at game-time we are relying on future data for use in our average statistic which we must calculate and sweep out from each observation.
 
 \subsection{Robustness check}
-We verify that this phenomena persists in Major League Baseball. In the MLB, money-lines
+As a robustness check, we verify that this phenomena persists in an entirely separate
+sport with entirely different players. To that end, we look toward
+Major League Baseball, which is also well set up for our quasi-experiment since
+there are many games played back-to-back in the season (more in fact, than NBA).
+
+In general, we can replicate our analysis, but there are also some notable differences.
+In the MLB, money-lines
 are used instead of point-spreads. So as opposed to predicting
@@
 a team would win against their opponent as both a function of the book-makers'
-odds of winning and our party variable. Additionally, we interacted our party variable with weekend.
-One of the major differences between NBA and MLB games is that MLB games
+odds of winning and our party variable. 
+Additionally, our methodology changes ever so slightly: when creating our
+continuous measure of nightlife, we interact the variable with weekend. The logic
+is that because there are so many more games played in MLB, the weekend effect is more salient.
+One of the last major differences between NBA and MLB games is that MLB games
 are played in series, so the same match-up repeated multiple times. This remedies
@@
 \hspace{-105pt}
-\begin{tabular}{@{\extracolsep{5pt}}lcccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{4}{c}{\textit{Dependent variable:}} \\  \cline{2-5}  \\[-1.8ex] & \multicolumn{2}{c}{NBA: Meet Spread} & demeaned reb/min & MLB: Win Game \\  \\[-1.8ex] & \multicolumn{2}{c}{\textit{logistic}} & \textit{OLS} & \textit{logistic} \\  \\[-1.8ex] & (1) & (2) & (3) & (4)\\  \hline \\[-1.8ex]  party.discrete & $-$0.406$^{***}$ (0.133) &  &  &  \\    party &  & $-$0.053$^{**}$ (0.022) & $-$0.001$^{*}$ (0.0005) & $-$0.035$^{*}$ (0.020) \\    lag.chg.pos & 0.0001 (0.002) & 0.0001 (0.002) & 0.00003 (0.0001) &  \\    odds &  &  &  & 2.602$^{***}$ (0.162) \\    log(travel.dist + 1) & $-$0.017 (0.025) & $-$0.015 (0.025) & 0.001 (0.001) & $-$0.001 (0.005) \\    ew & $-$0.114 (0.230) & $-$0.101 (0.230) &  &  \\    nhours.lgame & 0.001 (0.001) & 0.001 (0.001) & 0.00003 (0.00002) &  \\    I(hour(date)) & 0.007 (0.012) & 0.005 (0.012) &  &  \\    weekend &  &  &  & 0.040 (0.036) \\    I(team == location) & $-$0.011 (0.043) & $-$0.001 (0.043) &  & 0.025 (0.034) \\    log(travel.dist + 1):ew & 0.018 (0.035) & 0.016 (0.035) &  &  \\    ndays.lgame &  &  &  & 0.014 (0.019) \\    Constant & $-$0.077 (0.374) & $-$0.041 (0.374) & $-$0.007$^{*}$ (0.004) & $-$1.331$^{***}$ (0.080) \\   \hline \\[-1.8ex]  Observations & 10,000 & 9,979 & 44,190 & 25,215 \\  R$^{2}$ &  &  & 0.0003 &  \\  Adjusted R$^{2}$ &  &  & 0.0002 &  \\  Log Likelihood & $-$6,925.180 & $-$6,912.606 &  & $-$17,274.450 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{4}{c}{\textit{Dependent variable:}} \\  \cline{2-5}  \\[-1.8ex] & \multicolumn{2}{c}{NBA: Meet Spread} & Demeaned Reb/Min & MLB: Win Game \\  \\[-1.8ex] & \multicolumn{2}{c}{\textit{logistic}} & \textit{OLS} & \textit{logistic} \\  \\[-1.8ex] & (1) & (2) & (3) & (4)\\  \hline \\[-1.8ex]  party.discrete & $-$0.406$^{***}$ (0.133) &  &  &  \\    party &  & $-$0.053$^{**}$ (0.022) & $-$0.001$^{*}$ (0.0005) & $-$0.035$^{*}$ (0.020) \\    lag.chg.pos & 0.0001 (0.002) & 0.0001 (0.002) & 0.00003 (0.0001) &  \\    odds &  &  &  & 2.602$^{***}$ (0.162) \\    log(travel.dist + 1) & $-$0.017 (0.025) & $-$0.015 (0.025) & 0.001 (0.001) & $-$0.001 (0.005) \\    ew & $-$0.114 (0.230) & $-$0.101 (0.230) &  &  \\    nhours.lgame & 0.001 (0.001) & 0.001 (0.001) & 0.00003 (0.00002) &  \\    I(hour(date)) & 0.007 (0.012) & 0.005 (0.012) &  &  \\    weekend &  &  &  & 0.040 (0.036) \\    I(team == location) & $-$0.011 (0.043) & $-$0.001 (0.043) &  & 0.025 (0.034) \\    log(travel.dist + 1):ew & 0.018 (0.035) & 0.016 (0.035) &  &  \\    ndays.lgame &  &  &  & 0.014 (0.019) \\    Constant & $-$0.077 (0.374) & $-$0.041 (0.374) & $-$0.007$^{*}$ (0.004) & $-$1.331$^{***}$ (0.080) \\   \hline \\[-1.8ex]  Observations & 10,000 & 9,979 & 44,190 & 25,215 \\  R$^{2}$ &  &  & 0.0003 &  \\  Adjusted R$^{2}$ &  &  & 0.0002 &  \\  Log Likelihood & $-$6,925.180 & $-$6,912.606 &  & $-$17,274.450 \\  Akaike Inf.
 Crit.
@@
 of night-life; this could be demographic characteristics such as age or marital status.
+Future work may consider similar effects in European soccer leagues.
 
 \section{Conclusions}
-
-\section{References}
+We do find that visiting a city with active nightlife the day before a game does have 
+a statistically significant negative impact on a team's likelihood of meeting the spread in NBA.
+The decline in likelihood to meet the spread is possibly caused by a lack of control over the ball,
+since we also find that players capture less rebounds per minute of gameplay the day after
+visiting a city with active nightlife. We verify that this phenomena persists across sport, and 
+to that end look toward major league baseball money-lines. Using these, we back out gambling
+house's estimate for the team's probability of winning the game, and after conditioning on this
+we find that visiting a city with active nightlife the day before a game does negatively 
+impact the team's likelihood of winning the game.
+With a causal model established, we realize a profitable betting strategy in each year of testing.
 

### 2017-05-31 writing/draft.tex
Formatting and polishing the first rough draft. Updated results table-names. Added a count of # of games in dataset.

@@
 lack of nightlife when moving to Cleveland from New York.\cite{price,ley}
-A few interviews with former professional athletes revealed that location
+Matt Barnes recently complained that ``there's no nightlife in Utah'' when 
+asked about playing Jazz in the second round of NBA playoffs in 2017.\cite{cestone}
+A few interviews with former professional NBA athletes revealed that location
 matters more than day of the week, i.e. parties can happen any day of the week,
@@
 estimate the causal effect of  visiting a ``party-city'' on subsequent day 
-performance as measured by the spread. Since next-day
-opponent is uncorrelated with exposure to treatment, we have identification in our variable of 
-interest.
+performance as measured by the spread. To the extent that next-day
+opponent is uncorrelated with exposure to treatment, we have identification in our variable of interest.
 
@@
 for direct mentions of partying, either by athletes or witnesses. However, due to
-intense public scrutiny that athletes face, there is a strong reluctance to post
-such mentions on social media. There are notable exceptions, such
-as Oddel Beckham Jr.'s trip to Miami in 2017 season.\cite{bleler}
-The reports are few and far enough between,
+intense public scrutiny that athletes face, there is a strong reluctance to disclose
+their personal lives on social media. There are notable exceptions, such
+as Oddel Beckham Jr.'s recent trip to Miami.\cite{bleler}
+However, these reports are few and far enough between,
 and whether they are even reported at all is strongly biased on whether a particular city 
@@
 mentioned as having active night-life. Between players, the rank ordering of these
-was consistent to the extent that LA and NY were top
+was consistent to the extent that Los Angeles and New York city were top
 offenders. This motivates using a binary indicator for party in these two cities. 
@@
 a true effect. Alternatively,
-if the house book-keeper systematically over-valued
-teams which you are more likely to play after a party city,
+if the house book-keeper systematically over-values
+teams which are more likely to be encountered after playing a city with active nightlife,
 we could see a spurious effect. 
 That said, the amount of money that is at stake and the effort put into creating these spreads 
-suggests that they are reliable. So, if we do find an effect in our variable of interest,
+suggests that they are reasonably reliable. 
+So, if we do find an effect in our variable of interest,
 we can be confident in our conclusion that there is a causal effect and not some 
@@
 
-\paragraph{Fatigue from Jet lag}
-What about jet-lag? Prolonged air travel has in general been shown
+\paragraph{Fatigue from jetlag}
+What about jetlag? Prolonged air travel has in general been shown
 to have a negative effect on sports performance.\cite{leeandgalvez}
-Previous studies have shown that jet lag is a strong enough effect
-to erase home field advantage in MLB.\cite{songetal}
+Previous studies have shown that jetlag is a strong enough effect
+to erase home field advantage in major league baseball, for example.\cite{songetal}
 
@@
 for a listing of game-days by season. With this listing, we queried \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{covers.com},
-for all games occurring on a particular game, and for each we observe
-outcome of each game alongside the line set by a betting house. We obtain
-data for seasons 2010-11 through 2016-17.
+for all games occurring on a particular day, and for each we observe the
+outcome of each game alongside the line set by a betting house.\cite{hrisports} We obtain
+data for seasons 2010-11 through 2016-17. We are missing the first couple months of games 
+in the 2010-11 season, and our 2011-12 season has some 229 missing games. In each remaining
+season, we have at most two games missing of the $82 \times 30 / 2 = 1,230$ played.
+We have a total of 7,829 games in our data-set.
 
@@
 (following a great circle), which yields a direction in degrees. 
-We then obtain a measure for east-west travel by looking at the trigonometric-$\sin$ of this angle.
+We then obtain a measure for east-west travel by looking at the trigonometric-$\sin$ of this angle after converting to radians.
 Interacting this with travel-distance gets us a measure of jetlag.\cite{lallensack}
@@
 there are many games played back-to-back in the season (more in fact, than NBA).
+There are also anecdotes of players celebrating with alcohol in public.\cite{miller,campbell}
 
-In general, we can replicate our analysis, but there are also some notable differences.
+In general, we can replicate our analysis and use the same resources to obtain our data,
+ but there are also some notable differences.
 In the MLB, money-lines
@@
 \hspace{-105pt}
-\begin{tabular}{@{\extracolsep{5pt}}lcccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{4}{c}{\textit{Dependent variable:}} \\  \cline{2-5}  \\[-1.8ex] & \multicolumn{2}{c}{NBA: Meet Spread} & Demeaned Reb/Min & MLB: Win Game \\  \\[-1.8ex] & \multicolumn{2}{c}{\textit{logistic}} & \textit{OLS} & \textit{logistic} \\  \\[-1.8ex] & (1) & (2) & (3) & (4)\\  \hline \\[-1.8ex]  party.discrete & $-$0.406$^{***}$ (0.133) &  &  &  \\    party &  & $-$0.053$^{**}$ (0.022) & $-$0.001$^{*}$ (0.0005) & $-$0.035$^{*}$ (0.020) \\    lag.chg.pos & 0.0001 (0.002) & 0.0001 (0.002) & 0.00003 (0.0001) &  \\    odds &  &  &  & 2.602$^{***}$ (0.162) \\    log(travel.dist + 1) & $-$0.017 (0.025) & $-$0.015 (0.025) & 0.001 (0.001) & $-$0.001 (0.005) \\    ew & $-$0.114 (0.230) & $-$0.101 (0.230) &  &  \\    nhours.lgame & 0.001 (0.001) & 0.001 (0.001) & 0.00003 (0.00002) &  \\    I(hour(date)) & 0.007 (0.012) & 0.005 (0.012) &  &  \\    weekend &  &  &  & 0.040 (0.036) \\    I(team == location) & $-$0.011 (0.043) & $-$0.001 (0.043) &  & 0.025 (0.034) \\    log(travel.dist + 1):ew & 0.018 (0.035) & 0.016 (0.035) &  &  \\    ndays.lgame &  &  &  & 0.014 (0.019) \\    Constant & $-$0.077 (0.374) & $-$0.041 (0.374) & $-$0.007$^{*}$ (0.004) & $-$1.331$^{***}$ (0.080) \\   \hline \\[-1.8ex]  Observations & 10,000 & 9,979 & 44,190 & 25,215 \\  R$^{2}$ &  &  & 0.0003 &  \\  Adjusted R$^{2}$ &  &  & 0.0002 &  \\  Log Likelihood & $-$6,925.180 & $-$6,912.606 &  & $-$17,274.450 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{4}{c}{\textit{Dependent variable:}} \\  \cline{2-5}  \\[-1.8ex] & \multicolumn{2}{c}{NBA: Meet Spread} & Demeaned Reb/Min & MLB: Win Game \\  \\[-1.8ex] & \multicolumn{2}{c}{\textit{logistic}} & \textit{OLS} & \textit{logistic} \\  \\[-1.8ex] & (1) & (2) & (3) & (4)\\  \hline \\[-1.8ex]  party.discrete & $-$0.406$^{***}$ (0.133) &  &  &  \\    party.continuous &  & $-$0.053$^{**}$ (0.022) & $-$0.001$^{*}$ (0.0005) & $-$0.035$^{*}$ (0.020) \\    lag.change.position & 0.0001 (0.002) & 0.0001 (0.002) & 0.00003 (0.0001) &  \\    odds &  &  &  & 2.602$^{***}$ (0.162) \\    log(travel.dist+1) & $-$0.017 (0.025) & $-$0.015 (0.025) & 0.001 (0.001) & $-$0.001 (0.005) \\    east-west.travel.dir& $-$0.114 (0.230) & $-$0.101 (0.230) &  &  \\    nhours.last.game & 0.001 (0.001) & 0.001 (0.001) & 0.00003 (0.00002) &  \\    hour(date) & 0.007 (0.012) & 0.005 (0.012) &  &  \\    weekend &  &  &  & 0.040 (0.036) \\    (team == location) & $-$0.011 (0.043) & $-$0.001 (0.043) &  & 0.025 (0.034) \\    log(travel.dist+1):ew & 0.018 (0.035) & 0.016 (0.035) &  &  \\    ndays.last.game &  &  &  & 0.014 (0.019) \\    Constant & $-$0.077 (0.374) & $-$0.041 (0.374) & $-$0.007$^{*}$ (0.004) & $-$1.331$^{***}$ (0.080) \\   \hline \\[-1.8ex]  Observations & 10,000 & 9,979 & 44,190 & 25,215 \\  R$^{2}$ &  &  & 0.0003 &  \\  Adjusted R$^{2}$ &  &  & 0.0002 &  \\  Log Likelihood & $-$6,925.180 & $-$6,912.606 &  & $-$17,274.450 \\  Akaike Inf.
 Crit.

### 2017-06-01 init_model.rnw
Added table and plot showing effect of points admitted as a function of party.

--- 2017-05-08 init_model.rnw
@@
 <<echo=F>>=
-lines[, list(date = as.Date(date), team, opponent, line, score)]
+nba.lines <- lines
+nba.lines[, list(date = as.Date(date), team, opponent, line, score)]
 @
@@
 ### Merge lines data with BLS data.
-lines <- merge(lines, musicians,
+nba.lines <- merge(nba.lines, musicians,
                by.x = c('season', 'last.game.loc'),
@@
 
-setkey(lines, season, team, date)
+setkey(nba.lines, season, team, date)
 
-## lines[, list(date, team, opponent, location, 
+### Exclude playoff games.
+nba.lines <- nba.lines[, head(.SD, n=82), by=list(season,team)]
+
+## nba.lines[, list(date, team, opponent, location, 
 ##              line, score, outcome, nhours.lgame, 
@@
 HOURS <- 24
-lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians), 0)]
+nba.lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians), 0)]
 
@@
 
-lines[, simple.date := as.character(date) %>% substr(., 1, 10) %>% as.Date]
-data <- merge(lines, espn,
+nba.lines[, simple.date := as.character(date) %>% substr(., 1, 10) %>% as.Date]
+data <- merge(nba.lines, espn,
               by.x = c('team', 'season', 'simple.date'),
@@
             keyby = list(season, location, date)]
-lines <- merge(lines, changes, by = c('season', 'location', 'date'))
+nba.lines <- merge(nba.lines, changes, by = c('season', 'location', 'date'))
 
-setkey(lines, season, team, date)
-lines[, lag.chg.pos := c(NA, lag(ttl.chg.pos)[1:.N-1]), by = list(season, team)]
+setkey(nba.lines, season, team, date)
+nba.lines[, lag.chg.pos := c(NA, lag(ttl.chg.pos)[1:.N-1]), by = list(season, team)]
 
@@
 <<echo=F>>=
-lines[, list(date = simple.date, location = substr(location, 1, 7), 
+nba.lines[, list(date = simple.date, location = substr(location, 1, 7), 
              team = substr(team, 1, 7), opponent = substr(opponent, 1, 7), 
@@
 
-lines[, ew := sin(travel.dir %% 360 / 360 * 2 * pi)]
-#plot(lines$travel.dir, lines$ew)
+nba.lines[, ew := sin(travel.dir %% 360 / 360 * 2 * pi)]
+#plot(nba.lines$travel.dir, nba.lines$ew)
 
-m <- glm(outcome == 'W' ~ party + lag.chg.pos + travel.dist*ew + 
-             nhours.lgame + I(hour(date)),
-         data = lines[season < 2017], family = 'binomial',
+nba.lines[, party.discrete := grepl("(los angeles)|(new york)|(brooklyn)", last.game.loc) & nhours.lgame <= 24]
+
+m0 <- glm(outcome == 'W' ~ party.discrete + lag.chg.pos + log(travel.dist+1)*ew + 
+             nhours.lgame + I(hour(date)) + I(team==location),
+         data = nba.lines[season < 2017], family = 'binomial',
          na.action = 'na.exclude')
 
-summary(m)
+m1 <- glm(outcome == 'W' ~ party + lag.chg.pos + log(travel.dist+1)*ew + 
+             nhours.lgame + I(hour(date)) + I(team==location),
+         data = nba.lines[season < 2017], family = 'binomial',
+         na.action = 'na.exclude')
 
-# stargazer(m, style = 'qje', single.row = T)
+summary(m1)
+
+## nba.lines[,.N/2,by=list(season)]
+## stargazer(m)
+
 @
@@
 m <- glm(outcome == 'W' ~ party, 
-         data = lines[season < 2016 & last.game.loc != team])
+         data = nba.lines[season < 2016 & last.game.loc != team])
 
 ### We're only interested in making predictions when teams party.
-p <- predict(m, newdata = lines[season == 2016 & nhours.lgame <= HOURS & 
+p <- predict(m, newdata = nba.lines[season == 2016 & nhours.lgame <= HOURS & 
                                 last.game.loc != team], type = 'response')
@@
 idx <- which(p < 0.476)
-out <- table(lines[season == 2016 & nhours.lgame <= HOURS & last.game.loc != team][idx, outcome])
+out <- table(nba.lines[season == 2016 & nhours.lgame <= HOURS & last.game.loc != team][idx, outcome])
 
@@
 
-table(round(p), lines[season == 2016 & nhours.lgame <= HOURS & 
+table(round(p), nba.lines[season == 2016 & nhours.lgame <= HOURS & 
                       last.game.loc != team, outcome])
@@
 m <- glm(outcome == 'W' ~ party,
-         data = lines[season < 2016 & last.game.loc != team])
+         data = nba.lines[season < 2016 & last.game.loc != team])
 
-p <- predict(m, newdata = lines[season == 2016 & nhours.lgame <= HOURS & 
+p <- predict(m, newdata = nba.lines[season == 2016 & nhours.lgame <= HOURS & 
                                 last.game.loc != team], type = 'response')
 
-table(round(p), lines[season == 2016 & nhours.lgame <= HOURS & 
+table(round(p), nba.lines[season == 2016 & nhours.lgame <= HOURS & 
                       last.game.loc != team, outcome])
@@
 <<echo=F>>=
-lines[, game.time := hour(date)]
+nba.lines[, game.time := hour(date)]
 
-m <- lm(demeaned.reb.per.min ~ party + lag.chg.pos + travel.dist + nhours.lgame,
+m2 <- lm(demeaned.reb.per.min ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame,
    data = data[last.game.loc != team] %>% na.omit)
+summary(m2)
 
+lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame, data=data[last.game.loc!=team] %>% na.omit) %>% summary
+
+### Points  allowed as afunction of last game location (by party)
+nba.lines[, demeaned.pts.admitted := team.pts.admitted - mean(team.pts.admitted), by = list(season, team)]
+nba.lines[, mean(demeaned.pts.admitted), by = list(last.game.loc, party.discrete)][order(V1)]
+tmp <- nba.lines[, list(demeaned.pts.admitted = mean(demeaned.pts.admitted)), 
+                 by = list(last.game.loc, party)][order(demeaned.pts.admitted)]
+plot(x=jitter(tmp$party),y=tmp$demeaned.pts.admitted)
+abline(lm(demeaned.pts.admitted ~ party, data=tmp))
 #stargazer(m, style = 'qje', single.row = T)
+## stargazer(m)
+
 @
@@
 load(file = 'tmp_data/covers_lines_mlb.RData')
-lines[, party := ifelse(team != location & weekend == 1, log(nmusicians), 0)]
+mlb.lines <- lines
+mlb.lines[, party := ifelse(team != location & weekend == 1, log(nmusicians), 0)]
 
-m <- glm(I(team.score > opponent.score) ~ party + odds + I(team == location) + 
-             ndays.lgame + travel.dist + weekend,
-         data = lines, family = 'binomial')
-summary(m)
-# stargazer(m, style = 'qje', single.row = T)
+setkey(mlb.lines, season, team, date)
+mlb.lines <- mlb.lines[, head(.SD, n=162),by=list(season,team,date)] #Exclude playoff games
+
+m3 <- glm(I(team.score > opponent.score) ~ party + odds + I(team == location) + 
+             ndays.lgame + log(travel.dist+1) + weekend,
+         data = mlb.lines, family = 'binomial')
+summary(m3)
+#stargazer(m0, m1, m2, m3, single.row = T)
 @
@@
 <<echo=2:4>>=
+
+pdf('mlb_bets_by_season.pdf')
+par(mfrow = c(3,2))
+b <- lapply(2012:2016, function(yr) {
     m <- glm(I(team.score > opponent.score) ~ party + odds,
-         data = lines[year(date) < 2016], family = 'binomial')
-p <- predict(m, newdata = lines[year(date) == 2016], type = 'response')
-
+             data = mlb.lines[year(date) < yr], family = 'binomial')
+    p <- predict(m, newdata = mlb.lines[year(date) == yr], type = 'response')
     idx <- which(p < .45)
     p1 <- mapply(Predict, prediction = p[idx], 
-             vegas = lines[year(date) == 2016, odds][idx],
+                 vegas = mlb.lines[year(date) == yr, odds][idx],
                  MoreArgs=list(threshold = 0.01))
     b <- mapply(Bet, our.prediction = p1,
-            actual.outcome = lines[year(date) == 2016, team.score > opponent.score][idx],
-            odds = lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
-
-sum(b, na.rm = T)
-
+                actual.outcome = mlb.lines[year(date) == yr, team.score > opponent.score][idx],
+                odds = mlb.lines[year(date) == yr, odds][idx], SIMPLIFY = T)
+    cat(paste0(sum(b, na.rm = T), '\n'))
     ### 
-betting.days <- lines[year(date) == 2016][idx, date]
+#    pdf(paste0('mlb_bets_', yr, '.pdf'))
+    betting.days <- mlb.lines[year(date) == yr][idx, date]
     running.prof <- cumsum(ifelse(is.na(b), 0, b))
     plot(x = 1:length(b), y = running.prof, 
-     main = 'Profits over the 2016 season', 
+         main = paste0('Profits over the ', yr, ' season\n',
+                       'Training on 2011-', yr-1, ' data'),
          xlab = 'Game Number', ylab = 'Profit', type = 'l')
@@
     points(x = which.min(running.prof), y = min(running.prof), type = 'p')
-text(x = which.min(running.prof), y = min(running.prof), labels = min(running.prof), pos = 4)
+    text(x = which.min(running.prof), y = min(running.prof), 
+         labels = round(min(running.prof)), pos = 4)
+    return(b)
+#    dev.off()
+})
+dev.off()
+
+@
+
+\includegraphics{mlb_bets_by_season}
+
+If we were to bet one hundred dollars on each game in 2016,
+we would earn a profit of about \Sexpr{format(round(sum(b[[length(b)]], na.rm=T), -1), big.mark = ',')} dollars
+over the course of \Sexpr{sum(!is.na(b))} games.
+
+<<>>=
+
+m <- glm(I(team.score > opponent.score) ~ party + odds,
+         data = mlb.lines[year(date) < 2016], family = 'binomial')
+p <- predict(m, newdata = mlb.lines[year(date) == 2016], type = 'response')
+idx <- which(p < .45)
+p1 <- mapply(Predict, prediction = p[idx], 
+             vegas = mlb.lines[year(date) == 2016, odds][idx],
+             MoreArgs=list(threshold = 0.01))
+b <- mapply(Bet, our.prediction = p1,
+                actual.outcome = mlb.lines[year(date) == 2016, team.score > opponent.score][idx],
+            odds = mlb.lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
 
@@
 ### an optimized betting strategy.
-hist(p- lines[year(date) == 2016, odds])
+hist(p- mlb.lines[year(date) == 2016, odds])
 
@@
     p1 <- mapply(Predict, prediction = p[idx], 
-                 vegas = lines[year(date) == 2016, odds][idx],
+                 vegas = mlb.lines[year(date) == 2016, odds][idx],
                  MoreArgs=list(threshold = t))
     b <- mapply(Bet, our.prediction = p1,
-                actual.outcome = lines[year(date) == 2016, team.score > opponent.score][idx],
-                odds = lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
+                actual.outcome = mlb.lines[year(date) == 2016, team.score > opponent.score][idx],
+                odds = mlb.lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
     return(c(sum(b, na.rm = T) / sum(!is.na(b)), sum(!is.na(b)))) # Avg profit per game.

### 2017-06-01 writing/draft.tex
Changed how tables are broken out.

@@
 Anecdotally, it has been reported that professional sports teams party on 
-the road, sometimes to the detriment of the game.\cite{hoch}
+the road, sometimes to the detriment of the game.
 Within the NBA, J.R. Smith is infamous for citing his improvement in performance due to a
@@
 which ultimately leads to inhibited ability to recover and adapt to exercise.\cite{mjbarnes}
-Sleep deprivation has also been shown to strongly impair cognitive and motor performance.\cite{pilcher} Both alcohol and sleep deprivation are well correlated with lifestyles associated
-with an active night life.
+Sleep deprivation has also been shown to strongly impair cognitive and motor performance.\cite{pilcher} Both alcohol and sleep deprivation are well correlated with nightlife.
 
@@
 whether any players are injured, the momentum of success each team is carrying
-in the season, etc.\cite{anderson} To that end, we utilize spread data. The benefit of
+in the season, etc. To that end, we utilize spread data.\cite{anderson} The benefit of
 using spread data is that it controls for a lot, but the downside is that
@@
 a true effect. Alternatively,
-if the house book-keeper systematically over-values
+if the house book-keeper systematically mis-values
 teams which are more likely to be encountered after playing a city with active nightlife,
@@
 So, if we do find an effect in our variable of interest,
-we can be confident in our conclusion that there is a causal effect and not some 
+we can be reasonably confident in our conclusion that there is a causal effect and not some 
 other systematic bias.
@@
 in Los Angeles on Friday and another game in New York the following day, we don't
-know whether the team traveled on Friday evening or Saturday morning. Heuristics for
+know whether the team traveled on Friday evening or Saturday morning and hence don't know
+which city they ``visited'' on Friday night. Heuristics for
 travel schedules can be made, but it is hard to obtain reliable team-travel data.
@@
 for all games occurring on a particular day, and for each we observe the
-outcome of each game alongside the line set by a betting house.\cite{hrisports} We obtain
-data for seasons 2010-11 through 2016-17. We are missing the first couple months of games 
-in the 2010-11 season, and our 2011-12 season has some 229 missing games. In each remaining
+outcome of each game alongside the line set by a betting house. We interpret the lines
+according to a standard guide.\cite{hrisports} We obtain
+data for seasons 2010-11 through 2016-17. There are exactly 30 teams in the NBA,
+each plays 82 games per season.
+We are missing the first couple months of games 
+in the 2010-11 season, and our 2011-12 season has 229 missing games. In each remaining
 season, we have at most two games missing of the $82 \times 30 / 2 = 1,230$ played.
@@
 to account for the effects of jet-lag. For each team, we queried
-a maps search engine to obtain a longitude and latitude listing.\cite{ggmap}
+a maps search engine to obtain a longitude and latitude describing their
+stadium location.\cite{ggmap}
 For each pairwise set of coordinates, we calculated distance between the two
@@
 \subsection{Using spread data}
-We find that after accounting for player fatigue incurred during the last game, jetlag and travel fatigue, and recovery time, and home-team effect, that both our discrete indicator and continuous measure of party
+We find that after accounting for player fatigue incurred during the last game, jetlag and travel fatigue, recovery time, and home-team effect that both our discrete indicator and continuous measure of party
 have a statistically significant effect on likelihood of meeting the spread.
+In laymens terms, teams under perform book-maker's expectations after playing in cities 
+with active nightlife.
+This supports the hypothesis that players enjoy nightlife systematically in certain
+cities and that a ``hangover'' effect is felt on the court the next day.
+
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]    party.discrete & $-$0.433$^{***}$ (0.118) &  \\    party.continuous &  & $-$0.053$^{**}$ (0.023) \\    lag.change.position & $-$0.0001 (0.002) & 0.00005 (0.002) \\    log(travel.dist + 1) & $-$0.019 (0.026) & $-$0.015 (0.026) \\    east-west.travel.dir & $-$0.151 (0.231) & $-$0.116 (0.231) \\    nhours.last.game & 0.001 (0.001) & 0.001 (0.001) \\    hour(date) & 0.006 (0.013) & 0.005 (0.013) \\    (team == location) & $-$0.007 (0.043) & 0.004 (0.043) \\    log(travel.dist + 1):ew & 0.022 (0.036) & 0.018 (0.036) \\    Constant & $-$0.019 (0.378) & $-$0.037 (0.378) \\   \hline \\[-1.8ex]  Observations & 9,844 & 9,823 \\  Log Likelihood & $-$6,814.705 & $-$6,804.350 \\  Akaike Inf.
+Crit.
+& 13,647.410 & 13,626.700 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+
+We mention that it's not surprising that features describing jet-lag and recovery time
+are not significant: such salient information has already been taken into account in
+setting the spread.
+In the next section, we explore which specific areas of gameplay are affected.
 
@@
 This allows us to avoid using a black-box feature, and also pin-point
-more exactly what elements of the game are in fact effected after visiting a city with active 
-nightlife. This motivates looking at number of points allowed, and number of points scored;
-we also considered how rebounds, fouls, and even injuries might be explained (in part) by
-a hangover effect.
+more precisely what elements of the game effected by visiting a nightlife city.
+This motivates looking at number of points allowed, and number of points scored; we also considered how rebounds, fouls, and even injuries might be
+explained (in part) by a hangover effect.
 
+\paragraph{Rebounds: the ``hustle'' statistic}
 In particular, we see that the average number of rebounds obtained per minute
-of game-play does decrease as a function of being exposed to a city with active night-life.
+of game-play does decrease as a function of being exposed to a city with active night-life. Rebounds are often considered a ``hustle-statistic'' in the sense 
+that anecdotally, players who work harder on the court and expend energy boxing out
+other players will have higher rebound rates. If this common knowledge is accepted,
+our analysis suggests that players are working less hard following travel to a 
+city with active nightlife.
+
 The method was to first calculate for each player and each game the number of rebounds obtained
@@
 This is an ad-hoc approach that cannot be used for betting, since at game-time we are relying on future data for use in our average statistic which we must calculate and sweep out from each observation.
+
+\paragraph{The hangover correlates with allowing more points}
+Similarly, common knowledge also suggests that the overall defense of a team
+is effected strongly by individual player effort. It seems likely that a hangover effect
+would be more prevalent on the defensive aspects of gameplay. To that end, we examine points allowed and points scored as a function of exposure to nightlife.
+It's worth noting that travel to party cities is correlated with allowing
+more points on the subsequent day of gameplay, but the null hypothesis
+can't be rejected for points scored.
+
+It's also interesting to see that after visiting LA or NY, for any given
+team their next day performance is strongly correlated with allowing the other team to
+score two and a half points more. 
+It's even more interesting to note that points spreads are on a similar order of magnitude,  
+i.e. this could often be the difference between meeting the spread or not.
+
+\hspace{-80pt}
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & demeaned.pts.admitted & team.pts.admitted & team.pts.scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   party.discrete & 2.603$^{***}$ (0.542) &  &  \\    party.continuous &  & 0.486$^{***}$ (0.141) & $-$0.003 (0.138) \\    lag.change.position &  & 0.267$^{***}$ (0.016) & 0.285$^{***}$ (0.016) \\    log(travel.dist+1) &  & $-$0.429$^{**}$ (0.191) & 0.140 (0.187) \\    nhours.last.game &  & 0.012$^{*}$ (0.007) & 0.016$^{**}$ (0.007) \\    Constant & $-$0.072 (0.090) & 70.598$^{***}$ (2.266) & 64.914$^{***}$ (2.218) \\   \hline \\[-1.8ex]  Observations & 15,658 & 6,426 & 6,426 \\  R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Adjusted R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Residual Std.
+Error & 11.157 (df = 15656) & 11.937 (df = 6421) & 11.688 (df = 6421) \\  F Statistic & 23.077$^{***}$ (df = 1; 15656) & 75.897$^{***}$ (df = 4; 6421) & 86.787$^{***}$ (df = 4; 6421) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 One of the last major differences between NBA and MLB games is that MLB games
-are played in series, so the same match-up repeated multiple times. This remedies
+are played in series, so the same match-up is repeated multiple times. This remedies
 the problem of an unknown travel schedule we faced with NBA because we are sure 
@@
 
-% Table created by stargazer v.5.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu % Date and time: Mon, May 29, 2017 - 11:56:17 PM
-\hspace{-105pt}
-\begin{tabular}{@{\extracolsep{5pt}}lcccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{4}{c}{\textit{Dependent variable:}} \\  \cline{2-5}  \\[-1.8ex] & \multicolumn{2}{c}{NBA: Meet Spread} & Demeaned Reb/Min & MLB: Win Game \\  \\[-1.8ex] & \multicolumn{2}{c}{\textit{logistic}} & \textit{OLS} & \textit{logistic} \\  \\[-1.8ex] & (1) & (2) & (3) & (4)\\  \hline \\[-1.8ex]  party.discrete & $-$0.406$^{***}$ (0.133) &  &  &  \\    party.continuous &  & $-$0.053$^{**}$ (0.022) & $-$0.001$^{*}$ (0.0005) & $-$0.035$^{*}$ (0.020) \\    lag.change.position & 0.0001 (0.002) & 0.0001 (0.002) & 0.00003 (0.0001) &  \\    odds &  &  &  & 2.602$^{***}$ (0.162) \\    log(travel.dist+1) & $-$0.017 (0.025) & $-$0.015 (0.025) & 0.001 (0.001) & $-$0.001 (0.005) \\    east-west.travel.dir& $-$0.114 (0.230) & $-$0.101 (0.230) &  &  \\    nhours.last.game & 0.001 (0.001) & 0.001 (0.001) & 0.00003 (0.00002) &  \\    hour(date) & 0.007 (0.012) & 0.005 (0.012) &  &  \\    weekend &  &  &  & 0.040 (0.036) \\    (team == location) & $-$0.011 (0.043) & $-$0.001 (0.043) &  & 0.025 (0.034) \\    log(travel.dist+1):ew & 0.018 (0.035) & 0.016 (0.035) &  &  \\    ndays.last.game &  &  &  & 0.014 (0.019) \\    Constant & $-$0.077 (0.374) & $-$0.041 (0.374) & $-$0.007$^{*}$ (0.004) & $-$1.331$^{***}$ (0.080) \\   \hline \\[-1.8ex]  Observations & 10,000 & 9,979 & 44,190 & 25,215 \\  R$^{2}$ &  &  & 0.0003 &  \\  Adjusted R$^{2}$ &  &  & 0.0002 &  \\  Log Likelihood & $-$6,925.180 & $-$6,912.606 &  & $-$17,274.450 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & (team.score $ > $ opponent.score) \\  \hline \\[-1.8ex]   party & $-$0.035$^{*}$ (0.020) \\    odds & 2.602$^{***}$ (0.162) \\    I(team == location) & 0.025 (0.034) \\    ndays.lgame & 0.014 (0.019) \\    log(travel.dist + 1) & $-$0.001 (0.005) \\    weekend & 0.040 (0.036) \\    Constant & $-$1.331$^{***}$ (0.080) \\   \hline \\[-1.8ex]  Observations & 25,215 \\  Log Likelihood & $-$17,274.450 \\  Akaike Inf.
 Crit.
-& 13,868.360 & 13,843.210 &  & 34,562.900 \\  Residual Std.
-Error &  &  & 0.105 (df = 44185) &  \\  F Statistic &  &  & 3.064$^{**}$ (df = 4; 44185) &  \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{4}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 34,562.900 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
 \subsubsection{Betting performance}
-We also validate our model by seeing how well we can perform against the
+We also validate our MLB model by seeing how well we can perform against the
 market.
@@
   \caption{Betting performance by season. In each season, The 
-    circled point represents the worst out of pocket expense incurred during the season.
+    circled point represents the worst out of pocket expense incurred.
     We first train a model on historical data then use said model to predict on ``current year''
-    data, ploting profits as a function of time over the ``current'' season. For each relevant
-    game, we place a bet if our model instructs us of odds sufficiently different from what
-    the betting house suggests. We account for the bookmaker's cut. E.g. in the early 2016 season,
+    data, plotting profits as a function of time over the ``current'' season. For each relevant
+    game, we place a bet of \$100 if the expected value of the bet is positive after accounting f
+    or the bookmaker's cut. E.g. in the     early 2016 season,
     at its worst we are out of pocket \$750 dollars. By season end, we have     recouped all of

### 2017-06-01 writing/draft.tex
Tightened up first rough draft with Eric.

@@
 the road, sometimes to the detriment of the game.
-Within the NBA, J.R. Smith is infamous for citing his improvement in performance due to a
-lack of nightlife when moving to Cleveland from New York.\cite{price,ley}
+Within the NBA, the player J.R. Smith underwent a substantial improvement
+in performance metrics after being traded from New York to Cleveland; when an interviewer
+asked what contributed to this bump in statistics, he infamously replied
+that in New York he would have been out partying, but because there's no nightlife
+in Cleveland, he ends up hitting the gym.\cite{price,ley}
 Matt Barnes recently complained that ``there's no nightlife in Utah'' when 
@@
 One of the hardest parts of this analysis is that partying is a latent variable.
-In order to determine when players partied, one might consider to look at social media data
-for direct mentions of partying, either by athletes or witnesses. However, due to
+In order to determine when players partied, one may consider looking at social media data
+for direct mentions of partying, either by athletes or witnesses. However, due to the
 intense public scrutiny that athletes face, there is a strong reluctance to disclose
@@
 as Oddel Beckham Jr.'s recent trip to Miami.\cite{bleler}
-However, these reports are few and far enough between,
-and whether they are even reported at all is strongly biased on whether a particular city 
-has nightclubs with VIP rooms, etc; picking up statistical evidence would be fruitless in this case.
+However, these reports are few and far enough between that picking up statistical 
+evidence would be fruitless.
 
@@
 we don't rely on any features we can't obtain before making a game-time prediction. 
-To this data, we've created features such as 
+To this data, we've added features such as 
 the number of days since last game, the last game location, and the travel distance.
@@
 We have a free parameter: the time-window which we believe the hangover effect
-to last. Suppose we only look toward games which have been played in a party
-city at most 24 hours ago. We take care to ignore
-teams playing in their own cities, i.e. if a team's last game was played in at home 
-we say that they were not exposed to an exogenous shock in nightlife.
+to last. We consider games which have been played at most 24 hours ago, and take care to ignore
+teams playing in their own city, i.e. if a team's last game was played at home  we say that they were not exposed to an exogenous shock in nightlife.
 
@@
 the likelihood of meeting the spread, we predict the probability
-a team would win against their opponent as both a function of the book-makers'
+a team would win as both a function of the book-makers'
 odds of winning and our party variable. 
 Additionally, our methodology changes ever so slightly: when creating our
-continuous measure of nightlife, we interact the variable with weekend. The logic
-is that because there are so many more games played in MLB, the weekend effect is more salient.
+continuous measure of nightlife, we interact the variable with weekend. Interviews with athletes suggest that MLB players are more likely to
+restrict their partying to weekends.
 One of the last major differences between NBA and MLB games is that MLB games
@@
 & 34,562.900 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+
+As shown above, our continuous measure of nightlife remains statistically significant.
 
@@
     data, plotting profits as a function of time over the ``current'' season. For each relevant
-    game, we place a bet of \$100 if the expected value of the bet is positive after accounting f
-    or the bookmaker's cut. E.g. in the     early 2016 season,
+    game, we place a bet of \$100 if the expected value of the bet is positive after accounting for the bookmaker's cut. E.g. in the     early 2016 season,
     at its worst we are out of pocket \$750 dollars. By season end, we have     recouped all of
@@
 \subsection{Extending the research}
-We can also validate where points are lost in MLB in a similar way.
-It would be interesting to look at number of shots taken, and number of shots taken by stars.
-It's possible there are prolonged effects of partying, and applying some sort of kernel
-on the number of hours played since the last game could be interesting. It could also be
-that different (types of) teams or players are more or less susceptible to the effects
-of night-life; this could be demographic characteristics such as age or marital status.
+\paragraph{Within the NBA} 
+We're interested to explore what other factors may affect a team's propensity to suffer hangover effects beyond simply visiting a party city; e.g. are teams more likely to go out after
+a win or a loss? Do teams who aren't going to make the playoffs party more? Are there demographic
+factors which influence partying, e.g. are younger, unmarried players more susceptible to treatment effect? Within a game, it would be interesting to look at 
+number of shots taken, and number of shots taken by 
+stars as a function of previous day visitation to party cities. We could also investigate 
+effects on injury risk and flagrant foul rates.
+
+\paragraph{MLB}
+It would also be interesting to examine the particular ways that partying
+effects performance statistics similar to how we did with NBA.
+We might consider whether batting performance (averages, slugging percentage)
+pitching (ERA, WHIP) or general fielding stats are effected by party.
+
+\paragraph{General}
+In both sports, it's possible there are prolonged effects of partying, and applying some sort of kernel on the number of hours played since the last game could be interesting. 
 Future work may consider similar effects in European soccer leagues.
@@
 We do find that visiting a city with active nightlife the day before a game does have 
-a statistically significant negative impact on a team's likelihood of meeting the spread in NBA.
-The decline in likelihood to meet the spread is possibly caused by a lack of control over the ball,
-since we also find that players capture less rebounds per minute of gameplay the day after
-visiting a city with active nightlife. We verify that this phenomena persists across sport, and 
+a statistically significant negative impact on a team's likelihood of meeting the spread in NBA. This decreased likelihood to meet the spread may be explained 
+by the team's defense letting up more points. In turn, we believe that the drop in
+defensive performance may be caused by a drop in physical exertion.
+We verify that this phenomena persists across sport, and 
 to that end look toward major league baseball money-lines. Using these, we back out gambling
@@
 we find that visiting a city with active nightlife the day before a game does negatively 
-impact the team's likelihood of winning the game.
+impact the team's likelihood of winning.
 With a causal model established, we realize a profitable betting strategy in each year of testing.

### 2017-06-03 writing/draft.tex
Read-through with Eric in Kennedy.

@@
 (in part) by what takes place off the court. 
-Anecdotally, it has been reported that professional sports teams party on 
+There are numerous reports of professional sports teams partying on 
 the road, sometimes to the detriment of the game.
-Within the NBA, the player J.R. Smith underwent a substantial improvement
-in performance metrics after being traded from New York to Cleveland; when an interviewer
+Within the NBA, J.R. Smith's performance metrics substantially improved 
+ after being traded from New York to Cleveland; when an interviewer
 asked what contributed to this bump in statistics, he infamously replied
@@
 that teams systematically play worse on Sunday afternoons when on the road,
-regardless of cities.
+regardless of the city.
 
@@
 most notorious for pursuing active nightlife in the NBA.\cite{chase} The study
-does conclude that players perform worse on Sundays when compared with other days
-as measured by the ratio of shots scored to shots taken.
+ concludes that players perform worse on Sundays when compared with other days, as measured by the ratio of shots scored to shots taken.
 
@@
 simple count of the number of musicians in the county in which the stadium resides. 
+This captures
+how much nightlife there is in a city since musical performances are correlated
+with parties, if not an essential feature.
 We pursue both approaches.
@@
 Previous studies have shown that jetlag is a strong enough effect
-to erase home field advantage in major league baseball, for example.\cite{songetal}
+to erase home field advantage in major league baseball.\cite{songetal}
 
 \paragraph{The problem of an unknown travel schedule} In NBA, we have the difficulty of not observing when teams travel. E.g. if a particular team  plays
-in Los Angeles on Friday and another game in New York the following day, we don't
+in Los Angeles on Friday and another game in Boston the following day, we don't
 know whether the team traveled on Friday evening or Saturday morning and hence don't know
@@
 
-We then merge this in with our lines data, taking care to do so according to each season-team's last game location such that the feature captures the proper
+We then merge this in with our lines data, taking care to do so according to each team's last game location such that the feature captures the proper
 party effect. In addition, we lag our BLS data by one year for two reasons:
@@
 We also gathered minutes data from \href{http://www.espn.com/nba/scoreboard/_/date/}{ESPN}.
-The minutes data also contain information on the number of free throws, field goals, and three pointers made and attempted by each player in each game.
+The minutes data contains information on the number of free throws, field goals, and three pointers made and attempted by each player in each game.
 
@@
 \subsection{Method: Defining what it means for a city to have nightlife}
-We have a free parameter: the time-window which we believe the hangover effect
-to last. We consider games which have been played at most 24 hours ago, and take care to ignore
+We consider games which have been played at most 24 hours ago, and take care to ignore
 teams playing in their own city, i.e. if a team's last game was played at home  we say that they were not exposed to an exogenous shock in nightlife.
@@
 Our first sanity check was to define an indicator variable for when a team plays in 
-Los Angeles or New York within 24 hours of their current game, we indicate one, else zero. Anecdotally, these are mentioned as hot-spots for
-nightlife.
+Los Angeles or New York within 24 hours of their current game, we indicate one, else zero.
 
@@
 if the last game was played at most 24 hours ago, use the log of the average number
-of musicians in last game location, else emit zero. Intuitively, this captures
-how much nightlife there is in a city since musical performances are correlated
-with parties, if not an essential feature.
+of musicians in last game location, else emit zero.
 
@@
 
-We mention that it's not surprising that features describing jet-lag and recovery time
-are not significant: such salient information has already been taken into account in
-setting the spread.
+The features describing jet-lag and recovery time are not significant, but this is to be
+expected: such salient information has already been taken into account in setting the spread.
 In the next section, we explore which specific areas of gameplay are affected.
@@
 As a robustness check, we verify that this phenomena persists in an entirely separate
-sport with entirely different players. To that end, we look toward
+sport. To that end, we look toward
 Major League Baseball, which is also well set up for our quasi-experiment since
@@
   \includegraphics{mlb_bets_by_season}
-  \caption{Betting performance by season. In each season, The 
+  \caption{Betting performance by season. In each season, the 
     circled point represents the worst out of pocket expense incurred.
@@
     at its worst we are out of pocket \$750 dollars. By season end, we have     recouped all of
-    our initial investment plus an additional \$11.5k.}
+    our initial investment plus an additional \$11.5k. We are profitable in each season, and mention that 2015 is the ``worst'', but we still come out ahead.}
   \label{bettingperf}

### 2017-06-10 init_model.rnw
Added dot-plot of continuous-party by team-location.

@@
 HOURS <- 24
-nba.lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians), 0)]
+nba.lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians+1), 0)]
+nba.lines[, party.discrete := grepl("(los angeles)|(new york)|(brooklyn)", last.game.loc) & 
+                nhours.lgame <= HOURS & 
+                last.game.loc != team]
 
@@
 
-nba.lines[, party.discrete := grepl("(los angeles)|(new york)|(brooklyn)", last.game.loc) & nhours.lgame <= 24]
-
 m0 <- glm(outcome == 'W' ~ party.discrete + lag.chg.pos + log(travel.dist+1)*ew + 
              nhours.lgame + I(hour(date)) + I(team==location),
-         data = nba.lines[season < 2017], family = 'binomial',
+         data = nba.lines[season < 2017 & !is.na(party)], family = 'binomial',
          na.action = 'na.exclude')
 
-m1 <- glm(outcome == 'W' ~ party + lag.chg.pos + log(travel.dist+1)*ew + 
-             nhours.lgame + I(hour(date)) + I(team==location),
-         data = nba.lines[season < 2017], family = 'binomial',
-         na.action = 'na.exclude')
+m1 <- update(m0, . ~ party + . - party.discrete)
 
@@
 ## nba.lines[,.N/2,by=list(season)]
-## stargazer(m)
+stargazer(m0, m1, covariate.labels = c('party discrete', 'party continuous', 'lag changes in posession', 'logged travel distance', 'east-west travel direction', 'number hours rest time', 'time of game during day', 'home team effect', 'logged travel distance * east-west', 'Constant'),
+          dep.var.labels = 'Meet the Spread')
+
+tmp <- musicians[, list(party.continuous = mean(log(nmusicians+1))), 
+                 by = team][order(party.continuous)][!is.na(party.continuous)]
+
+##################################################
+### Dot plot of Party-Continuous by Team Location
+##################################################
+pdf('writing/Party_by_Team_Location.pdf', width = 14, height = 8.5)
+ticks = 1:nrow(tmp)
+plot(x = ticks, 
+     y = tmp$party.continuous, 
+     main = 'Continuous measure of party by Team Location',
+     ylab = 'Log Avg of Musicians and Sound Recording Studios',
+     xlab = '',
+     axes = F)
+axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>% 
+                               gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
+## text(x = ticks, y = 1.25, labels = tmp$team, srt = 90)
+axis(side = 2)
+dev.off()
 
@@
 
-lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame, data=data[last.game.loc!=team] %>% na.omit) %>% summary
+mpa <- lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame, 
+   data=nba.lines[last.game.loc!=team] %>% na.omit)
+mpo <- update(mpa, team.pts.scored ~ .)
 
@@
 nba.lines[, mean(demeaned.pts.admitted), by = list(last.game.loc, party.discrete)][order(V1)]
+mpa1 <- lm(demeaned.pts.admitted ~ party.discrete, data = nba.lines)
+#stargazer(mpa1, mpa, mpo, single.row=T)
+
 tmp <- nba.lines[, list(demeaned.pts.admitted = mean(demeaned.pts.admitted)), 
@@
 summary(m3)
+
+mlb.lines[, party.discrete := grepl("(los angeles)|(new york)", last.game.loc) & ndays.lgame == 1 & weekend == T]
 #stargazer(m0, m1, m2, m3, single.row = T)

### 2017-06-11 writing/draft.tex
Restructured according to Charlie's comments.

@@
 \usepackage{amsmath}
+\usepackage{rotating}
 
@@
 (in part) by what takes place off the court. 
-There are numerous reports of professional sports teams partying on 
-the road, sometimes to the detriment of the game.
+We are interested to see if 
+teams exhibit a decline in performance the day following a game in a city 
+with active nightlife; part of the question is determining which cities are notorious
+for nightlife.
+
+\paragraph{Why do we care} 
+Inquiring about this causal relationship is worthwhile for several reasons. The
+NBA Commissioner Adam Silver has reported that there is \$400 billion per year spent 
+on sports betting in the U.S. alone.\citep{asilver,cnnsilver} Apart from arbitrage opportunities,
+knowing how player performance is affected by external factors could inform coaches of a more
+optimal play strategy. E.g. if visitation to a city with active nightlife is correlated with
+next-day injury, perhaps coaches might bench their star players if the game isn't crucial.
+
+\paragraph{What do we do}
+We exploit data on bookmaker spreads, the expected score 
+differential between two teams after conditioning on observable performance. 
+We expect a team to meet the spread half the time, since this is one of the easiest ways for 
+bookmakers to guarantee a profit.\citep{dubnerlevitt} We construct a model which attempts to 
+estimate the causal effect of  visiting a ``party-city'' on subsequent day 
+performance as measured by the spread. In particular, we are only looking at
+cases where games are played back-to-back within 24 hours of each other.
+To the extent that next-day opponent is uncorrelated with exposure to treatment, 
+we have identification in our variable of interest.
+
+\paragraph{What do we find}
+We replicate this analysis for both the National Basketball Association
+(NBA) and Major League Baseball (MLB). In particular, we find that visiting
+a city with active nightlife the day prior to a game does have a statistically
+significant negative effect on a team's likelihood of meeting the point-spread set
+by a betting house, for both NBA and MLB. Within the NBA, we analyze specific performance
+metrics such as number of points allowed and number of rebounds obtained, and find that
+both are negatively effected by previous day visitation to a city with active nightlife.
+Lastly, having established a causal model, we realize a profitable betting scheme for MLB.
+
+\section{Effects of Partying on Athletic Performance}
+
+\subsection{Anecdotal evidence of the effects on sporting performance}
+Players openly admit they are interested in nightlife: Matt Barnes recently complained that 
+``there's no nightlife in Utah'' when asked about playing Jazz in the second round of NBA playoffs 
+in 2017.\citep{cestone}
+There are also numerous reports of teams partying to the detriment of the game.
 Within the NBA, J.R. Smith's performance metrics substantially improved 
@@
 asked what contributed to this bump in statistics, he infamously replied
-that in New York he would have been out partying, but because there's no nightlife
-in Cleveland, he ends up hitting the gym.\cite{price,ley}
-Matt Barnes recently complained that ``there's no nightlife in Utah'' when 
-asked about playing Jazz in the second round of NBA playoffs in 2017.\cite{cestone}
-A few interviews with former professional NBA athletes revealed that location
-matters more than day of the week, i.e. parties can happen any day of the week,
-but they are more likely to occur in select cities.
+that in New York he would have been out partying, but because there's no nightlife in Cleveland, he ends up hitting the gym.\citep{price,ley} Before being traded,
+he would score 42 percent of shots taken over the course of his career, but only 34 percent
+on Sunday afternoon games in the 2012-13 season, presumably because such games followed a long
+weekend of partying.\citep{princeofthecity}
 
-We are interested to see if 
-teams exhibit a decline in performance the day following a game in a city 
-with active nightlife. We exploit data on bookmaker spreads, the expected score 
-differential between two teams after conditioning on observable performance. 
-We expect a team to meet the spread half the time, since this is how bookmakers 
-minimize risk on their end. We construct a model which attempts to 
-estimate the causal effect of  visiting a ``party-city'' on subsequent day 
-performance as measured by the spread. To the extent that next-day
-opponent is uncorrelated with exposure to treatment, we have identification in our variable of interest.
+\subsection{Individual Players}
+Studies have also been carried out on statistical ranking of the players
+most notorious for pursuing active nightlife in the NBA.\citep{chase} In particular, 
+the author rejects the null hypothesis that J.R. Smith shoots just as well on Sundays as he does
+every other day of the week at the $p=0.01$ level of significance; however, the author also
+cautions that with so many players in the NBA, it is expected that we find a split as extreme
+as J.R.'s simply due to chance.
+The study goes on to look
+at historical performance for all NBA players who began their career after 1974, and compares
+Sunday afternoon performance to overall career performance. A binomial test is carried out, and the 
+study concludes with strong statistical significance that across the NBA, 
+individual players perform worse on Sundays when compared with other days, 
+as measured by the ratio of shots scored to shots taken.
 
-\section{Previous Research}
-There has been an attempt to answer this question before.\cite{ezekowitz} 
-However, the study only
-looks at performance in party cities, as opposed to subsequent day performance (i.e. the study does not consider a delayed hangover effect). Moreover, their sample size was much smaller. 
-Finally, there was no methodology used behind the selection of party cities. The author does conclude
-that teams systematically play worse on Sunday afternoons when on the road,
-regardless of the city.
+\subsection{Teams}
+There has been an attempt to answer the question at the team-level before: i.e. do teams perform
+worse after a night on the town, and if so does it depend on which city they 
+visited.\citep{ezekowitz} However, the study only
+looks at performance in party cities, as opposed to subsequent day performance (i.e.
+the study does not consider a delayed hangover effect).
+Moreover, their sample size was much smaller, because the author only considered Sunday afternoon games
+from 2006-14. Finally, there was no methodology used behind the selection of party cities;
+the authors first tried to define ad-hoc five cities as ``party cities'', and when that did not yield
+a significant result they tried estimating fixed-effects for each city. The author does conclude
+with weak statistical evidence that teams systematically play worse on Sunday 
+afternoons when on the road, regardless of the city.
 
-Studies have also been carried out on statistical ranking of the players
-most notorious for pursuing active nightlife in the NBA.\cite{chase} The study
- concludes that players perform worse on Sundays when compared with other days, as measured by the ratio of shots scored to shots taken.
-
+\subsection{Laboratory Studies}
 Studies on effects of alcohol on sports performance in a controlled setting
-show that alcohol impairs the nervous system, resulting in decrements in both
-cognitive function and motor skill.\cite{shirreffs} In a study on 
+show that alcohol impairs the nervous system, resulting in a decline in both
+cognitive function and motor skill.\citep{shirreffs} In a study on 
 the effects of alcohol on recovery of male athletes, it was found that
 consumption is correlated with reduced production of testosterone
-which ultimately leads to inhibited ability to recover and adapt to exercise.\cite{mjbarnes}
-Sleep deprivation has also been shown to strongly impair cognitive and motor performance.\cite{pilcher} Both alcohol and sleep deprivation are well correlated with nightlife.
+which ultimately leads to inhibited ability to recover and adapt to exercise.\citep{mjbarnes}
+Sleep deprivation has also been shown to strongly impair cognitive and motor performance.\citep{pilcher} Both alcohol and sleep deprivation are well correlated with nightlife.
 
-\section{Materials and Methods}
+\section{Defining a Measure for Partying}
 One of the hardest parts of this analysis is that partying is a latent variable.
-In order to determine when players partied, one may consider looking at social media data
-for direct mentions of partying, either by athletes or witnesses. However, due to the
-intense public scrutiny that athletes face, there is a strong reluctance to disclose
-their personal lives on social media. There are notable exceptions, such
-as Oddel Beckham Jr.'s recent trip to Miami.\cite{bleler}
-However, these reports are few and far enough between that picking up statistical 
-evidence would be fruitless.
+In order to determine when players partied, one may consider looking at social media data for direct mentions of partying, either by athletes or witnesses.
+However, due to the intense public scrutiny that athletes face, there is a strong reluctance to disclose their personal lives on social media.
+There are notable exceptions, such
+as Oddel Beckham Jr.'s recent trip to Miami.\citep{bleler}
+However, these reports are few and far enough between that picking up statistical evidence would be fruitless.
+We develop two approaches: a discrete indicator and a continuous measure of nightlife.
 
-\paragraph{Two approaches: a discrete indicator and a continuous measure of nightlife}
+\paragraph{A discrete indicator}
 Because famous athletes often enjoy a celebrity lifestyle, using conventional statistics
@@
 mentioned as having active night-life. Between players, the rank ordering of these
-was consistent to the extent that Los Angeles and New York city were top
-offenders. This motivates using a binary indicator for party in these two cities. 
+was consistent to the extent that Los Angeles and New York city were top. 
+This motivates using a binary indicator for party when either of these two cities have been visited within 24-hours of another game.
+
+\begin{figure}
+%\includegraphics[scale=0.5]{travel_dist_by_party}
+  \centering
+  \includegraphics[scale=0.55]{travel_dist_density_by_party}
+  \label{fig: density by party}
+  \caption{We examine travel distance to current game location as a function of
+    whether the last game was in a (non) party city. We see that in general,
+  travel distance follows a similar distribution regardless of last game location, i.e. 
+  it's not as though players encounter significantly
+  more travel fatigue after playing in a party city.}
+\end{figure}
+
+\paragraph{A continuous measure}
 We also found that athletes tend to party with other A-list celebrities,
-particularly musicians. This motivated a continuous measure of nightlife, based on a
-simple count of the number of musicians in the county in which the stadium resides. 
+particularly musicians.
+This motivated a continuous measure of nightlife, based on a simple count of the number of musicians in the county in which the stadium resides: if the last game was played at most 24-hours ago, use the log of the average number
+of musicians in last game location, else emit zero.  
 This captures
@@
 with parties, if not an essential feature.
-We pursue both approaches.
 
-\paragraph{Logic for using spread data}
+% A *density* plot with overlay instead of two histograms
+
+\begin{sidewaysfigure}
+  \hspace{-55pt}
+  \includegraphics[scale=0.7]{Party_by_Team_Location}
+\end{sidewaysfigure}
+
+\section{Data}
+\subsection{Points-spread and money-lines data} We first look toward 
+\href{http://www.basketball-reference.com/leagues/}{basketball-reference.com} and
+\href{http://www.baseball-reference.com/leagues/}{baseball-reference.com}
+for a listing of game-days by season. With these listings, we queried \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{covers.com},
+for all games occurring on a particular day, and for each we observe the
+outcome of each game alongside the either the point-spread (for NBA) or money-line (for MLB) 
+set by a betting house. 
+We interpret the point-spread according to a standard guide.\citep{hrisports}
+
+\paragraph{NBA}
+We obtain
+data for seasons 2010-11 through 2016-17; in basketball, the season runs from October through April. 
+There are exactly 30 teams in the NBA,
+each plays 82 games per season.
+We are missing the first couple months of games 
+in the 2010-11 season, and our 2011-12 season has 229 missing games. In each remaining
+season, we have at most two games missing of the $82 \times 30 / 2 = 1,230$ played.
+We have a total of 7,829 games in our data-set.
+To this data, we've added features such as 
+the number of days since last game, the last game location, and the travel 
+distance between the last game location and the current game location.
+
+\paragraph{MLB} We also obtain data for seasons 2010 through 2017; each season runs from 
+April through October. There are also exactly 30 teams in the MLB, each plays 162 games per season.
+We don't have all $162 \times 30 / 2 = 2,430$ games played each season, but instead only have
+on average almost 2,100 games per season. We have a total of 12,709 games in our data-set. Again,
+we add features such as the number of days since the last game, the last game location, and travel
+distance between game locations.
+
+\subsubsection{Logic for using points-spread and money-line data}
 In order to assert a causal relation between cities with active nightlife 
@@
 whether any players are injured, the momentum of success each team is carrying
-in the season, etc. To that end, we utilize spread data.\cite{anderson} The benefit of
-using spread data is that it controls for a lot, but the downside is that
+in the season, etc. To that end, we utilize spread data.\citep{anderson} The benefit of
+using spread data is that it controls for a lot, but the downside is
 we don't know exactly what it controls for.
 
+\paragraph{How an unknown spread variable may (not) affect our inference}
 For example, if book-keepers
@@
 So, if we do find an effect in our variable of interest,
-we can be reasonably confident in our conclusion that there is a causal effect and not some 
-other systematic bias.
+we can be reasonably confident in our conclusion that there is a causal effect and not some other systematic bias.
 
-\subsection{Possible confounders}
+\subsection{BLS data} We also obtained data from
+\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the county-quarter level. We obtain this data for each county corresponding
+to an NBA or MLB team from years 2010-2016. 
+As a proxy for how much night-life there is in a city, we look toward the 
+number of sound recording studios, musical groups, and music publishers there 
+are for a particular county-year: we simply calculate the average number of establishments across all three music categories listed above.
+We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.\footnote{We don't expect this to be a party city.}
+
+We then merge this in with our lines data, taking care to do so according to each team's last game location such that the feature captures the proper
+party effect. In addition, we lag our BLS data by one year for two reasons:
+realistically, player expectations of which cities have the most active night-life may not
+update in real-time and also doing so allows us to create a betting model, i.e.
+we don't rely on any features we can't obtain before making a game-time prediction.
+
+\subsection{Minutes data for NBA}
+We also gathered minutes data from \href{http://www.espn.com/nba/scoreboard/_/date/}{ESPN}.
+The minutes data contains information on the number of free throws, field goals, and three pointers made and attempted by each player in each game.
+For each season and each team, we have nearly complete 
+minutes data for all games played: 85\% of our season-teams' have no more than 3 missing games.
+
+\subsection{Location data} We are also interested
+to account for the effects of jet-lag. For each team, we queried
+a maps search engine to obtain a longitude and latitude describing their
+stadium location.\citep{ggmap}
+
+\section{Analysis of Team Performance}
+
+\subsection{Modeling Approach}
 
 \paragraph{Home team effect}
-Home-team advantage is well-studied.\cite{jones07, jones08}  
-Because teams play a lot of their games on the road,
-most games after playing in a party city will be an away game.
+Home-team advantage is well-studied.\citep{jones07, jones08}  
+Empirically, teams play about half their games on the road, and half of their games at home. However,
+it's possible that when a team plays on the road, their next game is more likely to be
+an away game as well. I.e. most games after playing in a party city will be an away game.
 This should be covered by spread data, but we also include
@@
 What about jetlag? Prolonged air travel has in general been shown
-to have a negative effect on sports performance.\cite{leeandgalvez}
+to have a negative effect on sports performance.\citep{leeandgalvez}
 Previous studies have shown that jetlag is a strong enough effect
-to erase home field advantage in major league baseball.\cite{songetal}
+to erase home field advantage in major league baseball.\citep{songetal}
+To account for this, we calculate the travel distance between stadiums,
+consider a measure for east-west travel (which induces a change in time-zone).
+
+Specifically, for each pairwise set of coordinates, we calculated distance between the two
+points.\citep{sp} To account for jet-lag, we need to not only consider travel distance
+but the \emph{direction} of travel: if it's west to east the players' must adjust their internal
+clock forward and hence they lose sleep, if it's east to west then the players gain rest time.
+Hence we are interested to interact some continuous measure of east-west travel with
+travel-distance in order to come up with a measure for jet-lag.
+To that end, we first computed the final direction 
+(bearing) when arriving at the 
+second point after starting from the first and following the shortest path on a sphere 
+(following a great circle), which yields a direction in degrees. 
+We then obtain a measure for east-west travel by looking at the trigonometric-$\sin$ of this angle after converting to radians.
+Interacting this with travel-distance gets us a measure of jetlag.\citep{lallensack}
+
+\paragraph{Model} We use a logistic regression to estimate the likelihood of meeting
+the spread as a function of whether the team partied as well as controlling for other confounders
+listed above. Our basic model is as follows accounts for the rest-time since
+last game, jet-lag, home-team effect, player fatigue incurred during the last game,
+and the party effect.
+
+\subsection{Results}
+\subsubsection{NBA: using points-spread data}
+We find that after accounting for player fatigue incurred during the last game, jetlag and travel fatigue, recovery time, and home-team effect that both our discrete indicator and continuous measure of party
+have a statistically significant effect on likelihood of meeting the spread.
+In laymens terms, teams under perform book-maker's expectations after playing in cities 
+with active nightlife.
+This supports the hypothesis that players enjoy nightlife systematically in certain
+cities and that a ``hangover'' effect is felt on the court the next day.
+
+\begin{center}
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   party continuous &  & $-$0.050$^{**}$ \\    &  & (0.022) \\    & & \\   lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   logged travel distance & $-$0.017 & $-$0.015 \\    & (0.026) & (0.026) \\    & & \\   east-west travel direction & $-$0.146 & $-$0.115 \\    & (0.231) & (0.231) \\    & & \\   number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   time of game during day & 0.005 & 0.005 \\    & (0.013) & (0.013) \\    & & \\   home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.036 \\    & (0.378) & (0.378) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,804.511 \\  Akaike Inf.
+Crit.
+& 13,619.350 & 13,627.020 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+\end{center} 
+The features describing jet-lag and recovery time are not significant, but this is to be
+expected: such salient information has already been taken into account in setting the spread.
+In the next section, we explore which specific areas of gameplay are affected.
+
+\subsubsection{MLB: using money-lines data}
+We replicate the same analysis with
+Major League Baseball, which is also well set up for our quasi-experiment since
+there are many games played back-to-back in the season (more in fact, than NBA).
+There are also anecdotes of players celebrating with alcohol in public.\citep{miller,campbell}
+
+\paragraph{Differences between NBA and MLB as it pertains to our analysis}
+In the MLB, money-lines
+are used instead of point-spreads. So as opposed to predicting
+the likelihood of meeting the spread, we predict the probability
+a team would win as both a function of the book-makers'
+odds of winning and our party variable. 
+Additionally, our methodology changes ever so slightly: when creating our
+continuous measure of nightlife, we interact the variable with weekend. Interviews with athletes suggest that MLB players are more likely to
+restrict their partying to weekends.
+One of the last major differences between NBA and MLB games is that MLB games
+are played in series, so the same match-up is repeated multiple times. This remedies
+the problem of an unknown travel schedule we faced with NBA because we are sure 
+teams stay overnight in the same location within a series.
+
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Probability of Winning \\  \hline \\[-1.8ex]   Continuous measure of Nightlife & $-$0.035$^{*}$ \\    & (0.020) \\    & \\   Bookmaker's odds & 2.602$^{***}$ \\    & (0.162) \\    & \\   Home-team effect & 0.025 \\    & (0.034) \\    & \\   Number of rest days & 0.014 \\    & (0.019) \\    & \\   Logged travel distance & $-$0.001 \\    & (0.005) \\    & \\   Weekend & 0.040 \\    & (0.036) \\    & \\   Constant & $-$1.331$^{***}$ \\    & (0.080) \\    & \\  \hline \\[-1.8ex]  Observations & 25,215 \\  Log Likelihood & $-$17,274.450 \\  Akaike Inf.
+Crit.
+& 34,562.900 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+
+As shown above, our continuous measure of nightlife has a statistically significant negative effect for MLB teams likelihood of winning.
+
+\subsubsection{Betting performance}
+We also validate our MLB model by seeing how well we can perform against the
+market.
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.82]{mlb_bets_by_season}
+  \caption{Betting performance by season. In each season, the 
+    circled point represents the worst out of pocket expense incurred.
+    We first train a model on historical data then use said model to predict on ``current year''
+    data, plotting profits as a function of time over the ``current'' season. For each relevant
+    game, we place a bet of \$100 if the expected value of the bet is positive after accounting for the bookmaker's cut. E.g. in the     early 2016 season,
+    at its worst we are out of pocket \$750 dollars. By season end, we have     recouped all of
+    our initial investment plus an additional \$11.5k. We are profitable in each season, and mention that 2015 is the ``worst'', but we still come out ahead.}
+  \label{bettingperf}
+\end{figure}
+
+\subsubsection{Placebo Tests} As a sanity check, we interact our
+discrete-indicator for party city with an indicator for more than 24 hours of rest
+between games, and call this our placebo effect. 
+The idea is that with sufficient rest time, the hangover effect dissipates.
+
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread \\  \hline \\[-1.8ex]   party placebo & 0.066 \\    & (0.088) \\    & \\   lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   logged travel distance & $-$0.006 \\    & (0.026) \\    & \\   east-west travel direction & $-$0.102 \\    & (0.232) \\    & \\   number hours rest time & 0.001 \\    & (0.001) \\    & \\   time of game during day & 0.002 \\    & (0.013) \\    & \\   home team effect & $-$0.008 \\    & (0.043) \\    & \\   logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.067 \\    & (0.378) \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.801 \\  Akaike Inf.
+Crit.
+& 13,631.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+
+Indeed, we see that if players have enough recovery time after visiting cities such as Los
+Angeles or New York, we cannot reject the null hypothesis that player performance is unaffected
+by visitation to cities with active nightlife.
+
+\subsection{Possible Confounders or potential issues}
 
@@
 
-\subsection{Data}
-\paragraph{Lines data} We first look toward 
-\href{http://www.basketball-reference.com/leagues/}{basketball-reference.com}
-for a listing of game-days by season. With this listing, we queried \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{covers.com},
-for all games occurring on a particular day, and for each we observe the
-outcome of each game alongside the line set by a betting house. We interpret the lines
-according to a standard guide.\cite{hrisports} We obtain
-data for seasons 2010-11 through 2016-17. There are exactly 30 teams in the NBA,
-each plays 82 games per season.
-We are missing the first couple months of games 
-in the 2010-11 season, and our 2011-12 season has 229 missing games. In each remaining
-season, we have at most two games missing of the $82 \times 30 / 2 = 1,230$ played.
-We have a total of 7,829 games in our data-set.
-
-\paragraph{BLS data} We also obtained data from
-\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the county-quarter level. As a 
-proxy for how much night-life there is in a city, we look toward the 
-number of sound recording studios, musical groups, and music publishers there 
-are for a particular county-year: we simply calculate the average number of establishments across all three music categories listed above.
-We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.
-
-We then merge this in with our lines data, taking care to do so according to each team's last game location such that the feature captures the proper
-party effect. In addition, we lag our BLS data by one year for two reasons:
-realistically, player expectations of which cities have the most active night-life may not
-update in real-time and also doing so allows us to create a betting model, i.e.
-we don't rely on any features we can't obtain before making a game-time prediction. 
-To this data, we've added features such as 
-the number of days since last game, the last game location, and the travel distance.
-
-\paragraph{Minutes data}
-We also gathered minutes data from \href{http://www.espn.com/nba/scoreboard/_/date/}{ESPN}.
-The minutes data contains information on the number of free throws, field goals, and three pointers made and attempted by each player in each game.
-
-\paragraph{Location and Travel Distance Data} We are also interested
-to account for the effects of jet-lag. For each team, we queried
-a maps search engine to obtain a longitude and latitude describing their
-stadium location.\cite{ggmap}
-For each pairwise set of coordinates, we calculated distance between the two
-points.\cite{sp} To account for jet-lag, we first computed the final direction 
-(bearing) when arriving at the 
-second point after starting from the first and following the shortest path on a sphere 
-(following a great circle), which yields a direction in degrees. 
-We then obtain a measure for east-west travel by looking at the trigonometric-$\sin$ of this angle after converting to radians.
-Interacting this with travel-distance gets us a measure of jetlag.\cite{lallensack}
-
-\subsection{Method: Defining what it means for a city to have nightlife}
-We consider games which have been played at most 24 hours ago, and take care to ignore
-teams playing in their own city, i.e. if a team's last game was played at home  we say that they were not exposed to an exogenous shock in nightlife.
-
-\paragraph{Discrete indicator}
-Our first sanity check was to define an indicator variable for when a team plays in 
-Los Angeles or New York within 24 hours of their current game, we indicate one, else zero.
-
-\paragraph{Continuous measure}
-We also defined a continuous measure of nightlife:
-if the last game was played at most 24 hours ago, use the log of the average number
-of musicians in last game location, else emit zero.
-
-\subsection{Model} We use a logistic regression to estimate the likelihood of meeting
-the spread as a function of whether the team partied as well as controlling for other confounders
-listed above. Our basic model is as follows.
-
-\[
-\texttt{meet spread} \,  \sim \, \substack{\texttt{days since} \\ \texttt{last game}} \, + \, \texttt{jetlag} \, + \, \substack{\texttt{home team} \\ \texttt{effect}} \, + \, \texttt{party}
-\]
-
-\section{Results}
-\subsection{Using spread data}
-We find that after accounting for player fatigue incurred during the last game, jetlag and travel fatigue, recovery time, and home-team effect that both our discrete indicator and continuous measure of party
-have a statistically significant effect on likelihood of meeting the spread.
-In laymens terms, teams under perform book-maker's expectations after playing in cities 
-with active nightlife.
-This supports the hypothesis that players enjoy nightlife systematically in certain
-cities and that a ``hangover'' effect is felt on the court the next day.
-
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]    party.discrete & $-$0.433$^{***}$ (0.118) &  \\    party.continuous &  & $-$0.053$^{**}$ (0.023) \\    lag.change.position & $-$0.0001 (0.002) & 0.00005 (0.002) \\    log(travel.dist + 1) & $-$0.019 (0.026) & $-$0.015 (0.026) \\    east-west.travel.dir & $-$0.151 (0.231) & $-$0.116 (0.231) \\    nhours.last.game & 0.001 (0.001) & 0.001 (0.001) \\    hour(date) & 0.006 (0.013) & 0.005 (0.013) \\    (team == location) & $-$0.007 (0.043) & 0.004 (0.043) \\    log(travel.dist + 1):ew & 0.022 (0.036) & 0.018 (0.036) \\    Constant & $-$0.019 (0.378) & $-$0.037 (0.378) \\   \hline \\[-1.8ex]  Observations & 9,844 & 9,823 \\  Log Likelihood & $-$6,814.705 & $-$6,804.350 \\  Akaike Inf.
-Crit.
-& 13,647.410 & 13,626.700 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
-The features describing jet-lag and recovery time are not significant, but this is to be
-expected: such salient information has already been taken into account in setting the spread.
-In the next section, we explore which specific areas of gameplay are affected.
-
-\subsection{Using performance statistics}
+\section{Analysis of Player Performance}
 We additionally looked at actual performance metrics instead of relying on spread data. 
@@
 This is an ad-hoc approach that cannot be used for betting, since at game-time we are relying on future data for use in our average statistic which we must calculate and sweep out from each observation.
+We then run a linear regression
+explaining demeaned rebounds per minute as a function of our nightlife variable of interest
+as well as our usual features accounting for player-fatigue.
+
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.001$^{*}$ \\    & (0.0005) \\    & \\   Lag change in possessions & 0.00003 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00003 \\    & (0.00002) \\    & \\   Constant & $-$0.008$^{*}$ \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 43,547 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
+Error & 0.104 (df = 43542) \\  F Statistic & 2.981$^{**}$ (df = 4; 43542) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 
-\hspace{-80pt}
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & demeaned.pts.admitted & team.pts.admitted & team.pts.scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   party.discrete & 2.603$^{***}$ (0.542) &  &  \\    party.continuous &  & 0.486$^{***}$ (0.141) & $-$0.003 (0.138) \\    lag.change.position &  & 0.267$^{***}$ (0.016) & 0.285$^{***}$ (0.016) \\    log(travel.dist+1) &  & $-$0.429$^{**}$ (0.191) & 0.140 (0.187) \\    nhours.last.game &  & 0.012$^{*}$ (0.007) & 0.016$^{**}$ (0.007) \\    Constant & $-$0.072 (0.090) & 70.598$^{***}$ (2.266) & 64.914$^{***}$ (2.218) \\   \hline \\[-1.8ex]  Observations & 15,658 & 6,426 & 6,426 \\  R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Adjusted R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Residual Std.
-Error & 11.157 (df = 15656) & 11.937 (df = 6421) & 11.688 (df = 6421) \\  F Statistic & 23.077$^{***}$ (df = 1; 15656) & 75.897$^{***}$ (df = 4; 6421) & 86.787$^{***}$ (df = 4; 6421) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+\hspace{-100pt}
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.828$^{***}$ &  &  \\    & (0.699) &  &  \\    & & & \\   Continuous party measure &  & 0.475$^{***}$ & 0.009 \\    &  & (0.139) & (0.136) \\    & & & \\   Lag change in possessions &  & 0.267$^{***}$ & 0.285$^{***}$ \\    &  & (0.016) & (0.016) \\    & & & \\   Logged travel distance &  & $-$0.430$^{**}$ & 0.142 \\    &  & (0.191) & (0.187) \\    & & & \\   Number hours since last game &  & 0.012$^{*}$ & 0.017$^{**}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Constant & $-$0.047 & 70.581$^{***}$ & 64.886$^{***}$ \\    & (0.090) & (2.267) & (2.220) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,658 & 6,426 & 6,426 \\  R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Adjusted R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Residual Std.
+Error & 11.159 (df = 15656) & 11.937 (df = 6421) & 11.688 (df = 6421) \\  F Statistic & 16.357$^{***}$ (df = 1; 15656) & 75.841$^{***}$ (df = 4; 6421) & 86.788$^{***}$ (df = 4; 6421) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
-\subsection{Robustness check}
-As a robustness check, we verify that this phenomena persists in an entirely separate
-sport. To that end, we look toward
-Major League Baseball, which is also well set up for our quasi-experiment since
-there are many games played back-to-back in the season (more in fact, than NBA).
-There are also anecdotes of players celebrating with alcohol in public.\cite{miller,campbell}
-
-In general, we can replicate our analysis and use the same resources to obtain our data,
- but there are also some notable differences.
-In the MLB, money-lines
-are used instead of point-spreads. So as opposed to predicting
-the likelihood of meeting the spread, we predict the probability
-a team would win as both a function of the book-makers'
-odds of winning and our party variable. 
-Additionally, our methodology changes ever so slightly: when creating our
-continuous measure of nightlife, we interact the variable with weekend. Interviews with athletes suggest that MLB players are more likely to
-restrict their partying to weekends.
-One of the last major differences between NBA and MLB games is that MLB games
-are played in series, so the same match-up is repeated multiple times. This remedies
-the problem of an unknown travel schedule we faced with NBA because we are sure 
-teams stay overnight in the same location within a series.
-
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & (team.score $ > $ opponent.score) \\  \hline \\[-1.8ex]   party & $-$0.035$^{*}$ (0.020) \\    odds & 2.602$^{***}$ (0.162) \\    I(team == location) & 0.025 (0.034) \\    ndays.lgame & 0.014 (0.019) \\    log(travel.dist + 1) & $-$0.001 (0.005) \\    weekend & 0.040 (0.036) \\    Constant & $-$1.331$^{***}$ (0.080) \\   \hline \\[-1.8ex]  Observations & 25,215 \\  Log Likelihood & $-$17,274.450 \\  Akaike Inf.
-Crit.
-& 34,562.900 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
-As shown above, our continuous measure of nightlife remains statistically significant.
-
-\subsubsection{Betting performance}
-We also validate our MLB model by seeing how well we can perform against the
-market.
-\begin{figure}[!h]
-  \centering
-  \includegraphics{mlb_bets_by_season}
-  \caption{Betting performance by season. In each season, the 
-    circled point represents the worst out of pocket expense incurred.
-    We first train a model on historical data then use said model to predict on ``current year''
-    data, plotting profits as a function of time over the ``current'' season. For each relevant
-    game, we place a bet of \$100 if the expected value of the bet is positive after accounting for the bookmaker's cut. E.g. in the     early 2016 season,
-    at its worst we are out of pocket \$750 dollars. By season end, we have     recouped all of
-    our initial investment plus an additional \$11.5k. We are profitable in each season, and mention that 2015 is the ``worst'', but we still come out ahead.}
-  \label{bettingperf}
-\end{figure}
-
-\clearpage
-
+\section{Conclusions and Avenues for Future Research}
 \subsection{Extending the research}
@@
 
-\section{Conclusions}
+\subsection{Conclusion}
 We do find that visiting a city with active nightlife the day before a game does have 

### 2017-06-11 writing/draft.tex
Added heat map for last-game location vs. next day opponent.

@@
 \section{Defining a Measure for Partying}
+A key assumption in our analysis is that next-day opponent is reasonably uncorrelated with
+last game location. We plot a heat-map displaying the relationship between last game location
+and next game location, and further break this down according to whether the game(s)
+were played at home or away.
+
+\begin{figure}
+  \hspace{-44pt}
+  \includegraphics[scale=0.45]{next_day_opponent}
+  \caption{We examine the relationship between last game location and next day opponent in the NBA.
+Starting from the top left, notice that the most likely scenario across all seasons is that a game will be played at home, and the last game was also at home.
+This makes sense; 1/2 of all games are played at home, and 1/2 are played away, so clearly each team is most likely to be found playing a home game for any given game.
+When exactly one of the last game or current game was played away,
+we observe that some combinations cease to exist in our data; e.g.
+we never observe Charlotte Bobcats play Utah Jazz on the road and then fly directly back home for a game.
+Lastly, notice the patterns exhibited when last game is played on the road, and current game also played on the road.
+We see that when last game location was in Los Angeles, next day opponent likely to be a west coast team.
+There appears to be less of a correlation between last game location in New York or Brooklyn and next day opponent, perhaps because players ``gain'' time when flying east-to-west.} \end{figure}
+
+Our takeaway from the figure above is that indeed there is a correlation between last game location
+and next day opponent, in particular this is true for west coast teams. We hypothesize that
+this is due to practical limitations on flight schedules.
+The fact that there is a correlation between last game location and next day opponent increases the motivation for conditioning on spread, which accounts may help account for systematic biases to the extent that some of these next-day opponents are particularly (not) skilled.
+
+\paragraph{Partying as a latent variable}
 One of the hardest parts of this analysis is that partying is a latent variable.
@@
   \label{fig: density by party}
-  \caption{We examine travel distance to current game location as a function of
+  \caption{Within the NBA, we examine travel distance to current game location as a function of
     whether the last game was in a (non) party city. We see that in general,
@@
   it's not as though players encounter significantly
-  more travel fatigue after playing in a party city.}
+  more travel fatigue after playing in a party city.
+We mention that the same plot is not interesting to look at within the MLB, since games are played in series, and therefore most travel-distances between games are zero miles.}
 \end{figure}
@@
   \hspace{-55pt}
-  \includegraphics[scale=0.7]{Party_by_Team_Location}
+  \includegraphics[scale=0.7,angle=180]{Party_by_Team_Location}
 \end{sidewaysfigure}
@@
 
+\section{Appendix}
+
+\begin{figure}   
+  \centering
+  \hspace{-55pt}
+  \includegraphics[scale=0.5]{Party_by_Team_Location_mlb}
+  \caption{Continuous measure of nightlife by team location for MLB.
+What's interesting here is that because New York Yankees are in the Bronx, they show up as having few musicians and sound recording studios.
+We might wonder if this is a bad thing, since with NBA New York city was deemed a major party city.
+It turns out that when we define a discrete indicator of party based on New York or Los Angeles, it is not at all significant, and in fact the sign is incorrect.
+When we use our continuous
+ measure of nightlife, its significant, and it helps inform our profitable betting model.
+This is good news for our proxy, although begs the question why New York should be treated so differently by MLB players when compared with NBA players.
+One point worth noting is that we haven't yet made the important distinction that in the MLB, the Los Angeles Angels are in Orange County, for example, which is much less bustling than the NBA's  Los Angeles Lakers who are located in downtown Los Angeles.}
+\end{figure}
+
+\begin{figure}   
+  \centering
+  \hspace{-55pt}
+  \includegraphics[scale=0.5]{next_day_opponent_mlb} 
+\end{figure}
+
 \end{document}

### 2017-06-11 init_model.rnw
Added plots for density-of-travel-dist-by-party and also heat-map-of-next-day-opponenty-by-last-game-location.

@@
                 last.game.loc != team]
+nba.lines[, party.placebo := grepl("(los angeles)|(new york)|(brooklyn)", last.game.loc) & 
+                nhours.lgame > HOURS & 
+                last.game.loc != team]
 
@@
 
+##################################################
+### Dot plot of Party-Continuous by Team Location
+##################################################
 tmp <- musicians[, list(party.continuous = mean(log(nmusicians+1))), 
@@
 
-##################################################
-### Dot plot of Party-Continuous by Team Location
-##################################################
 pdf('writing/Party_by_Team_Location.pdf', width = 14, height = 8.5)
@@
      y = tmp$party.continuous, 
-     main = 'Continuous measure of party by Team Location',
+     main = 'Continuous measure of party by Team Location: NBA',
      ylab = 'Log Avg of Musicians and Sound Recording Studios',
@@
 axis(side = 2)
+dev.off()
+
+##################################################
+### Histogram of travel dist by party vs not
+##################################################
+pdf('writing/travel_dist_by_party.pdf', width = 11, height = 8.5)
+par(mfrow = c(2,1))
+hist(x = nba.lines[party.discrete == F, travel.dist], 
+     main = 'Histogram of Travel Distance by Last Game Location', xlab = '')
+legend(x = 2250, y = 3500, legend = c('Non Party Cities', 'LA and NY'), 
+       col = c('white', 'black'), fill = c(F, T))
+hist(x = nba.lines[party.discrete == T, travel.dist], 
+     main = '', col = 'black', xlab = 'Travel Distance')
+dev.off()
+
+pdf('writing/travel_dist_density_by_party.pdf', width = 12, height = 11)
+plot(density(nba.lines[(party.discrete) & !is.na(travel.dist), travel.dist]), 
+     xlim = c(-100, 18e2), lty = 1,
+     main = 'Density of travel distance\nfor (non) party cities: NBA',
+     xlab = 'Travel Distance (Miles)')
+lines(density(nba.lines[!(party.discrete) & !is.na(travel.dist), travel.dist]), lty = 2)
+legend(x = 15e2, y = 0.0015, legend = c('LA/NY', 'non-party cities'), lty = 1:2)
+dev.off()
+
+##################################################
+### Placebo model
+##################################################
+mp <- update(m0, . ~ party.placebo + . - party.discrete)
+
+stargazer(mp, covariate.labels = c('party placebo', 'lag changes in posession', 'logged travel distance', 'east-west travel direction', 'number hours rest time', 'time of game during day', 'home team effect', 'logged travel distance * east-west', 'Constant'),
+          dep.var.labels = 'Meet the Spread')
+
+##################################################
+### Table for uncorrelated next day opponent
+##################################################
+
+tmp <- nba.lines[, .N, 
+                 keyby = list(location, last.game.loc, 
+                              cur.game = paste('Current game', 
+                                               ifelse(location == team, 'at home', 'away')),
+                              lst.game = paste('Last game',
+                                               ifelse(last.game.loc == team, 'at home', 'away')))] %>% 
+    na.omit
+
+pdf('writing/next_day_opponent.pdf', width = 14, height = 14)
+ggplot(tmp, aes(x = location, y = last.game.loc, fill = log(N))) + 
+    geom_tile() +
+    facet_wrap(~lst.game + cur.game) + 
+    labs(x = 'Current Game Location', y = 'Last Game Location', 
+         title = 'Distribution of Next Day Opponent - NBA') +
+    theme_bw() +
+    theme(axis.text.x = element_text(angle = 315, vjust = 0.5),
+          plot.title = element_text(hjust=.5))
 dev.off()
@@
 
-mpa <- lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame, 
+tpa <- lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame, 
    data=nba.lines[last.game.loc!=team] %>% na.omit)
-mpo <- update(mpa, team.pts.scored ~ .)
+tpo <- update(tpa, team.pts.scored ~ .)
 
@@
 mpa1 <- lm(demeaned.pts.admitted ~ party.discrete, data = nba.lines)
-#stargazer(mpa1, mpa, mpo, single.row=T)
 
-tmp <- nba.lines[, list(demeaned.pts.admitted = mean(demeaned.pts.admitted)), 
-                 by = list(last.game.loc, party)][order(demeaned.pts.admitted)]
-plot(x=jitter(tmp$party),y=tmp$demeaned.pts.admitted)
-abline(lm(demeaned.pts.admitted ~ party, data=tmp))
-#stargazer(m, style = 'qje', single.row = T)
-## stargazer(m)
+stargazer(mpa1, tpa, tpo, dep.var.labels = c('Points Admitted by Team', 'Team Points Admitted', 'Team Points Scored'),
+          covariate.labels = c('Discrete party indicator',
+                               'Continuous party measure',
+                               'Lag change in possessions',
+                               'Logged travel distance',
+                               'Number hours since last game',
+                               'Constant'))
+
+stargazer(m2, dep.var.labels = 'Demeanead Rebounds Per Minute',
+          covariate.labels = c('Continuous party measure',
+                               'Lag change in possessions',
+                               'Logged travel distance',
+                               'Number hours since last game',
+                               'Constant'))
+
+## tmp <- nba.lines[, list(demeaned.pts.admitted = mean(demeaned.pts.admitted)), 
+##                  by = list(last.game.loc, party)][order(demeaned.pts.admitted)]
+## plot(x=jitter(tmp$party),y=tmp$demeaned.pts.admitted)
+## abline(lm(demeaned.pts.admitted ~ party, data=tmp))
+## #stargazer(m, style = 'qje', single.row = T)
+## ## stargazer(m)
 
@@
 mlb.lines[, party.discrete := grepl("(los angeles)|(new york)", last.game.loc) & ndays.lgame == 1 & weekend == T]
-#stargazer(m0, m1, m2, m3, single.row = T)
+
+## stargazer(m3, covariate.labels = c('Continuous measure of nightlife',
+##                                    'Bookmaker\'s odds', 'Home-team effect',
+##                                    'Number of rest days', 'Logged travel distance',
+##                                    'Weekend', 'Constant'), 
+##           dep.var.capation = 'Probability of Winning')
+
+##################################################
+### Heat Map for next day opponent
+##################################################
+tmp <- mlb.lines[, .N, 
+                 keyby = list(location, last.game.loc, 
+                              cur.game = paste('Current game', 
+                                               ifelse(location == team, 'at home', 'away')),
+                              lst.game = paste('Last game',
+                                               ifelse(last.game.loc == team, 'at home', 'away')))] %>% 
+    na.omit
+
+pdf('writing/next_day_opponent_mlb.pdf', width = 14, height = 14)
+ggplot(tmp, aes(x = location, y = last.game.loc, fill = log(N))) + 
+    geom_tile() +
+    facet_wrap(~lst.game + cur.game) + 
+    labs(x = 'Current Game Location', y = 'Last Game Location', 
+         title = 'Distribution of Next Day Opponent - MLB') +
+    theme_bw() +
+    theme(axis.text.x = element_text(angle = 315, vjust = 0.5),
+          plot.title = element_text(hjust=.5))
+dev.off()
+
+##################################################
+### Party by team location
+##################################################
+load(file = 'tmp_data/nmusician_estabs_mlb.RData')
+tmp <- musicians[, list(party.continuous = mean(log(nmusicians+1))), 
+                 by = team][order(party.continuous)][!is.na(party.continuous)]
+
+pdf('writing/Party_by_Team_Location_mlb.pdf', width = 14, height = 8.5)
+ticks = 1:nrow(tmp)
+plot(x = ticks, 
+     y = tmp$party.continuous, 
+     main = 'Continuous measure of party by Team Location: MLB',
+     ylab = 'Log Avg of Musicians and Sound Recording Studios',
+     xlab = '',
+     axes = F)
+axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>% 
+                               gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
+axis(side = 2)
+dev.off()
+
 @
@@
 
-pdf('mlb_bets_by_season.pdf')
+pdf('writing/mlb_bets_by_season.pdf')
 par(mfrow = c(3,2))

### 2017-06-12 writing/draft.tex
Polished up the first couple pages a bit more. Added a plot explaining why drinking establishments and population doesn't work.

@@
 \section{Introduction}
-It's not unreasonable to think that game performance can be affected 
-(in part) by what takes place off the court. 
+It's not unreasonable to think that in-game sporting performance can be affected 
+partly by what takes place off the court. 
 We are interested to see if 
 teams exhibit a decline in performance the day following a game in a city 
-with active nightlife; part of the question is determining which cities are notorious
-for nightlife.
+with active nightlife; we call this a ``hangover effect''.
+Part of the question is determining a reasonable way to measure
+levels of nightlife, and correspondingly which cities are notorious for it; we colloquially
+refer to such cities as ``party cities''.
 
@@
 optimal play strategy. E.g. if visitation to a city with active nightlife is correlated with
-next-day injury, perhaps coaches might bench their star players if the game isn't crucial.
+next-day injury, perhaps coaches might bench their star players if the latter game isn't crucial.
 
@@
 by a betting house, for both NBA and MLB. Within the NBA, we analyze specific performance
-metrics such as number of points allowed and number of rebounds obtained, and find that
-both are negatively effected by previous day visitation to a city with active nightlife.
-Lastly, having established a causal model, we realize a profitable betting scheme for MLB.
+metrics such as number of points allowed and number of rebounds obtained, and find that both are negatively effected by previous day visitation to a city with active nightlife.
+Having established a causal model, we realize a profitable betting scheme for MLB.
+We also perform a robustness check and see that if players rest more than
+ 24 hours after visiting a ``party-city'', the hangover effect dissipates.
 
@@
 \subsection{Anecdotal evidence of the effects on sporting performance}
-Players openly admit they are interested in nightlife: Matt Barnes recently complained that 
+Professional players openly admit they are interested in nightlife: Matt Barnes recently complained that 
 ``there's no nightlife in Utah'' when asked about playing Jazz in the second round of NBA playoffs 
@@
  after being traded from New York to Cleveland; when an interviewer
-asked what contributed to this bump in statistics, he infamously replied
-that in New York he would have been out partying, but because there's no nightlife in Cleveland, he ends up hitting the gym.\citep{price,ley} Before being traded,
-he would score 42 percent of shots taken over the course of his career, but only 34 percent
-on Sunday afternoon games in the 2012-13 season, presumably because such games followed a long
-weekend of partying.\citep{princeofthecity}
+asked what contributed to this bump in performance, he infamously replied
+that in New York he would have been out partying, but because there's no nightlife in Cleveland he ends up hitting the gym.\citep{price,ley} Before being traded, J.R.
+Smith's career-to-date success rate of shots taken was 42 percent, but somehow he only managed
+34 percent on Sunday afternoon games in the 2012-13 season.
+Presumably, this was because such games followed a long weekend of partying.\citep{princeofthecity}
 
@@
 Moreover, their sample size was much smaller, because the author only considered Sunday afternoon games
-from 2006-14. Finally, there was no methodology used behind the selection of party cities;
-the authors first tried to define ad-hoc five cities as ``party cities'', and when that did not yield
+from 2006-14. Finally, there was no methodology used behind the selection of party cities:
+the authors first tried to define ad-hoc five ``party cities'', and when that did not yield
 a significant result they tried estimating fixed-effects for each city. The author does conclude
@@
 \section{Defining a Measure for Partying}
+
+\paragraph{On the correlation between next day opponent and last game location}
 A key assumption in our analysis is that next-day opponent is reasonably uncorrelated with
@@
 \begin{figure}
-  \hspace{-44pt}
+  \centering
+  \label{fig: next day opponent}
   \includegraphics[scale=0.45]{next_day_opponent}
@@
 we never observe Charlotte Bobcats play Utah Jazz on the road and then fly directly back home for a game.
-Lastly, notice the patterns exhibited when last game is played on the road, and current game also played on the road.
-We see that when last game location was in Los Angeles, next day opponent likely to be a west coast team.
-There appears to be less of a correlation between last game location in New York or Brooklyn and next day opponent, perhaps because players ``gain'' time when flying east-to-west.} \end{figure}
+Lastly, notice the patterns that emerge when both the last game and current game are played on the road.
+E.g.
+we see that when last game location was in Los Angeles, next day opponent likely to be a west coast team.
+There appears to be less of a correlation between last game location in New York or Brooklyn and next day opponent, perhaps because flying from east-to-west
+is less burdensome as it relates to timezone changes.} 
+\end{figure}
 
-Our takeaway from the figure above is that indeed there is a correlation between last game location
-and next day opponent, in particular this is true for west coast teams. We hypothesize that
-this is due to practical limitations on flight schedules.
-The fact that there is a correlation between last game location and next day opponent increases the motivation for conditioning on spread, which accounts may help account for systematic biases to the extent that some of these next-day opponents are particularly (not) skilled.
+Our takeaway from the figure on page \ref{fig: next day opponent} is that indeed there is some correlation between last game location
+and next day opponent, and in particular this is true for west coast teams. We hypothesize that
+this is due to practical limitations on flight schedules. E.g. if a team is unwilling to take a red-eye
+flight, then there is little hope to play an evening west coast game and make it to the east coast for a game the very next day;
+the same is \emph{not true} when flying from east to west, in which case a morning flight is reasonable.
+The fact that there is some correlation between last game location and next day opponent, even if only in west coast cities, increases the motivation for conditioning on spread, which accounts may help account for systematic biases to the extent that some of these next-day opponents are particularly (not) skilled.
 
@@
 However, these reports are few and far enough between that picking up statistical evidence would be fruitless.
-We develop two approaches: a discrete indicator and a continuous measure of nightlife.
+Instead, we develop two approaches which do not rely on this data: a discrete indicator 
+based on player interviews and also a continuous measure of nightlife based on local community patterns.
 
@@
 
-% A *density* plot with overlay instead of two histograms
-
-\begin{sidewaysfigure}
-  \hspace{-55pt}
-  \includegraphics[scale=0.7,angle=180]{Party_by_Team_Location}
-\end{sidewaysfigure}
+\begin{figure}
+  \centering
+  \label{fig: party by team loc}
+  \includegraphics[scale=0.6,angle=270]{Party_by_Team_Location}
+  \caption{We plot our continuous measure of nightlife for each city within the NBA. We present the analagous plot for MLB in the appendix.\ref{fig: party by team location mlb}}
+\end{figure}
 
@@
 with active nightlife.
-This supports the hypothesis that players enjoy nightlife systematically in certain
-cities and that a ``hangover'' effect is felt on the court the next day.
 
 \begin{center}
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   party continuous &  & $-$0.050$^{**}$ \\    &  & (0.022) \\    & & \\   lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   logged travel distance & $-$0.017 & $-$0.015 \\    & (0.026) & (0.026) \\    & & \\   east-west travel direction & $-$0.146 & $-$0.115 \\    & (0.231) & (0.231) \\    & & \\   number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   time of game during day & 0.005 & 0.005 \\    & (0.013) & (0.013) \\    & & \\   home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.036 \\    & (0.378) & (0.378) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,804.511 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\ \hline \\[-1.8ex]   party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   party continuous &  & $-$0.050$^{**}$ \\    &  & (0.022) \\    & & \\   lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   logged travel distance & $-$0.017 & $-$0.015 \\    & (0.026) & (0.026) \\    & & \\   east-west travel direction & $-$0.146 & $-$0.115 \\    & (0.231) & (0.231) \\    & & \\   number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   time of game during day & 0.005 & 0.005 \\    & (0.013) & (0.013) \\    & & \\   home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.036 \\    & (0.378) & (0.378) \\ \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,804.511 \\  Akaike Inf.
 Crit.
-& 13,619.350 & 13,627.020 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 13,619.350 & 13,627.020 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01}
+\end{tabular} 
 \end{center}
+
+This model supports the hypothesis that players enjoy nightlife systematically in certain
+cities and that a ``hangover'' effect is felt on the court the next day.
 The features describing jet-lag and recovery time are not significant, but this is to be
@@
   \centering
+  \label{mlb bets by season}
   \includegraphics[scale=0.82]{mlb_bets_by_season}
@@
   \hspace{-55pt}
+  \label{fig: party by team location mlb}
   \includegraphics[scale=0.5]{Party_by_Team_Location_mlb}
@@
   \hspace{-55pt}
+  \label{fig: next day opponent mlb}
   \includegraphics[scale=0.5]{next_day_opponent_mlb} 
@@
 
+\begin{sidewaysfigure}   
+  \centering
+  \label{fig: why drinks or pop dont work}
+  \includegraphics[scale=0.5]{why_drinks_or_population_dont_work} 
+  \caption{We remark why using a statistic such as number of drinking establishments or 
+    population size doesn't work. Although both of these are linearly correlated with musicians, 
+    and even in spite of the extremal observations being ranked consistently across each statistic, 
+    i.e. Charlotte Bobcats at the bottom and Los Angeles at the top,  
+    the problem is that population and number of drinking establishments have sinusoidal 
+    relationships with respect to the rank ordering imposed by the musicians index 
+    that do not properly reflect variation in nightlife between cities. For example, both 
+    incorrectly rank Detroit Michigan as having a high nightlife index.
+Another example, population incorrectly places New Orleans, Louisiana as having a low nightlife index, when in reality it is famous for its vibrant city life.}
+\end{sidewaysfigure}
+
 \end{document}

### 2017-06-12 init_model.rnw
Added a plot showing why drinking estabs and population don't work.

@@
 load(file = 'tmp_data/covers_lines.RData')
-load(file = 'tmp_data/nmusician_estabs.RData')
+## load(file = 'tmp_data/nmusician_estabs.RData')
 load(file = 'tmp_data/population_by_county_year.RData')
+load(file = 'tmp_data/nba_musicians.RData')
+
+musicians <- copy(dt)
+rm(dt)
+setnames(musicians, 'variable', 'nmusicians')
 musicians <- merge(musicians, pop, by = c('county', 'season'), all.x = T)
@@
           plot.title = element_text(hjust=.5))
+dev.off()
+
+##################################################
+### Why drinking estabs doesn't work.
+##################################################
+load(file = 'tmp_data/nba_drinking.RData')
+drinking <- copy(dt)
+setnames(drinking, 'variable', 'ndrinks')
+rm(dt)
+
+tmpm <- musicians[, list(log.nmusicians = mean(log(nmusicians+1)),
+                         log.population = mean(log(population))), 
+                  by = team][order(log.nmusicians)] %>% na.omit
+tmpd <- drinking[, list(log.ndrinks = mean(log(ndrinks+1))), 
+                  by = team][order(log.ndrinks)] %>% na.omit
+tmp <- merge(tmpm, tmpd, by = 'team')
+setorder(tmp, log.nmusicians)
+
+pdf('writing/why_drinks_or_population_dont_work.pdf', width = 16, height = 8.5)
+ticks = 1:nrow(tmp)
+plot(x = ticks, 
+     y = tmp$log.nmusicians,
+     main = 'Why drinking establishments or population don\'t work as a proxy for nightlife',
+     ylab = 'Log Avg of Variable',
+     xlab = '', axes = F, pch = 1, ylim = c(0, max(tmp$log.population)))
+points(x = ticks, y = tmp$log.ndrinks, pch = 16, col = 'purple')
+points(x = ticks, y = tmp$log.population, pch = 18, col = 'orange')
+axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>% 
+                               gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
+axis(side = 2)
+legend(x = 22.5, y = 10.25, legend = c('Musicians and sound recording studios',
+                                 'Drinking establishments and liquor stores',
+                                 'Population'), pch = c(1, 16, 18), col = c('black', 'purple', 'orange'))
+text(x = 24.6, y = 10.5, labels = paste("Correlation (alcohol):       ", 
+                                       cor(tmp$log.nmusicians, tmp$log.ndrinks) %>% round(., 3)))
+text(x = 24.6, y = 11, labels = paste("Correlation (population): ", 
+                                       cor(tmp$log.nmusicians, tmp$log.population) %>% round(., 3)))
 dev.off()

### 2017-06-13 writing/draft.tex
Moved figures around, added a figure caption, added more detailed analyses.

@@
 the authors first tried to define ad-hoc five ``party cities'', and when that did not yield
-a significant result they tried estimating fixed-effects for each city. The author does conclude
-with weak statistical evidence that teams systematically play worse on Sunday 
-afternoons when on the road, regardless of the city.
+a significant result they tried estimating fixed-effects for each city. The author does carry out a binomial test of how many home-game covers vs. away-game covers were met, 
+and concludes
+with weak statistical evidence that teams systematically play worse on Sunday afternoons 
+when on the road, regardless of city.
 
@@
 
-Our takeaway from the figure on page \ref{fig: next day opponent} is that indeed there is some correlation between last game location
-and next day opponent, and in particular this is true for west coast teams. We hypothesize that
+We plot a heatmap which displays the logged count of co-occurences between last game
+location and current game location, split by home-and-away games.
+Our takeaway from this figure is that indeed there is some correlation between last game location
+and next day opponent (more so when both are away games,  
+and in particular this is true for west coast teams). We hypothesize that
 this is due to practical limitations on flight schedules. E.g. if a team is unwilling to take a red-eye
@@
 the same is \emph{not true} when flying from east to west, in which case a morning flight is reasonable.
-The fact that there is some correlation between last game location and next day opponent, even if only in west coast cities, increases the motivation for conditioning on spread, which accounts may help account for systematic biases to the extent that some of these next-day opponents are particularly (not) skilled.
+The fact that there is some correlation between last game location and next day opponent, even if only in west coast cities, increases the motivation for conditioning on spread, which may help account for systematic biases to the extent that some of these next-day opponents are particularly (not) skilled.
 
@@
 Because famous athletes often enjoy a celebrity lifestyle, using conventional statistics
-such as number of drinking establishments in each city does not work; athletes instead
+such as number of drinking establishments in each city or population does not work (see appendix \ref{fig: why drinks or pop dont work}), athletes instead
 tend to visit high end nightclubs. We again turned to player interviews for ideas,
@@
 
-\begin{figure}
-%\includegraphics[scale=0.5]{travel_dist_by_party}
-  \centering
-  \includegraphics[scale=0.55]{travel_dist_density_by_party}
-  \label{fig: density by party}
-  \caption{Within the NBA, we examine travel distance to current game location as a function of
-    whether the last game was in a (non) party city. We see that in general,
-  travel distance follows a similar distribution regardless of last game location, i.e. 
-  it's not as though players encounter significantly
-  more travel fatigue after playing in a party city.
-We mention that the same plot is not interesting to look at within the MLB, since games are played in series, and therefore most travel-distances between games are zero miles.}
-\end{figure}
-
 \paragraph{A continuous measure}
@@
 This motivated a continuous measure of nightlife, based on a simple count of the number of musicians in the county in which the stadium resides: if the last game was played at most 24-hours ago, use the log of the average number
-of musicians in last game location, else emit zero.  
+of musical groups, sound recording studios, and musical publishers in county of the 
+last game location, else emit zero.  
 This captures
@@
 with parties, if not an essential feature.
-
-\begin{figure}
-  \centering
-  \label{fig: party by team loc}
-  \includegraphics[scale=0.6,angle=270]{Party_by_Team_Location}
-  \caption{We plot our continuous measure of nightlife for each city within the NBA. We present the analagous plot for MLB in the appendix.\ref{fig: party by team location mlb}}
-\end{figure}
 
@@
 stadium location.\citep{ggmap}
+
+\begin{figure}
+%\includegraphics[scale=0.5]{travel_dist_by_party}
+  \centering
+  \includegraphics[scale=0.55]{travel_dist_density_by_party}
+  \label{fig: density by party}
+  \caption{Within the NBA, we examine travel distance to current game location as a function of
+    whether the last game was in a (non) party city. We see that in general,
+  travel distance follows a similar distribution regardless of last game location, i.e. 
+  it's not as though players encounter significantly
+  more travel fatigue after playing in a party city.
+We mention that the same plot is not interesting to look at within the MLB, since games are played in series, and therefore most travel-distances between games are zero miles.}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \label{fig: party by team loc}
+  \includegraphics[scale=0.6,angle=270]{Party_by_Team_Location}
+  \caption{We plot our continuous measure of nightlife for each city within the NBA. We present the analagous plot for MLB in the appendix.\ref{fig: party by team location mlb}}
+\end{figure}
 
@@
 
-\subsubsection{Placebo Tests} As a sanity check, we interact our
-discrete-indicator for party city with an indicator for more than 24 hours of rest
-between games, and call this our placebo effect. 
+\subsubsection{Placebo Tests} As a sanity check with our NBA data, we interact a
+discrete-indicator for Los Angeles and New York with an 
+indicator for \emph{more} than 24 hours of rest
+between games, and call this our placebo effect. The only difference between this feature
+and our discrete indicator for nightlife is the time-window of interest.
 The idea is that with sufficient rest time, the hangover effect dissipates.
@@
 \section{Analysis of Player Performance}
-We additionally looked at actual performance metrics instead of relying on spread data. 
+Within the NBA, we additionally looked at actual performance metrics instead of relying on point-spread data. 
 This allows us to avoid using a black-box feature, and also pin-point
-more precisely what elements of the game effected by visiting a nightlife city.
+more precisely what elements of the game are affected by visiting a nightlife city.
 This motivates looking at number of points allowed, and number of points scored; we also considered how rebounds, fouls, and even injuries might be
@@
   \includegraphics[scale=0.5]{next_day_opponent_mlb} 
+  \caption{For the MLB, there are some differences when compared with the analogous
+  plot for NBA. In particular, notice that a common scenario is for a team to play two away games 
+  back to back at the same location. This is an artifact of baseball match-ups being played 
+  in series as opposed to one-off games like in the NBA. We also notice in back-to-back
+  away games that the sparse rows 
+  (i.e. locations for which next-day opponent is restricted to only a strict subset of the 
+  29 available teams to play) all are attributed by West coast teams. In particular, 
+  Washington Nationals, San Francisco Giants, Oakland Athletics, Los Angeles Dodgers, and 
+  Arizona Diamondbacks are locations from which a next-day game on the East coast are not possible.}
 \end{figure}
@@
     and even in spite of the extremal observations being ranked consistently across each statistic, 
-    i.e. Charlotte Bobcats at the bottom and Los Angeles at the top,  
+    i.e. Charlotte Bobcats toward the bottom and Los Angeles more or less at the top,  
     the problem is that population and number of drinking establishments have sinusoidal 
@@
     that do not properly reflect variation in nightlife between cities. For example, both 
-    incorrectly rank Detroit Michigan as having a high nightlife index.
-Another example, population incorrectly places New Orleans, Louisiana as having a low nightlife index, when in reality it is famous for its vibrant city life.}
+    incorrectly rank Detroit Michigan as having a high nightlife index. Another example, 
+    population incorrectly places New Orleans, Louisiana as having a low nightlife index, 
+    when in reality it is famous for its vibrant city life.}
 \end{sidewaysfigure}

### 2017-06-13 clean_bls.r
Wrote a wrapper function to facilitate creation of multiple data sets, each their own feature/category from BLS.

@@ -36,14 +36,57 @@ clusterCall(cl, function() {
     require(magrittr)
 })
 
+MyGeoCode <- function(teams) {
+    locs  <- lapply(teams, geocode, output = 'more') %>% rbindlist(., fill = T)
+    setnames(locs, gsub('_', '.', names(locs)))
+    locs <- locs[, lapply(.SD, as.character), .SDcols = 1:ncol(locs)]
+    ### Be careful, since Washington Wizards are in Washington DC (so not a county)
+    locs[, county := ifelse(is.na(administrative.area.level.2),
+                            administrative.area.level.1,
+                            paste(administrative.area.level.2,
+                                  administrative.area.level.1, sep = ', '))]
+    locs <- data.table(team = teams, county = locs$county)
+    save(locs, file = 'tmp_data/team_locations.RData')
+}
+
+CleanBLS <- function(party.regex, suffix, years = 2010:2016, RESCRAPE = F) {
+    parLapplyLB(cl, years, SubsetBLS, party.regex = party.regex, sport = suffix)
+    files <- list.files(path = 'tmp_data/bls', pattern = paste0(suffix, '\\.csv$'), full.names = T)
+    bls <- lapply(files, fread) %>% rbindlist
+    load(file = 'tmp_data/covers_lines.RData')
+    lines[, pct := NULL]
+    if (RESCRAPE) MyGeoCode(unique(lines$team))
+    load(file = 'tmp_data/team_locations.RData')
+    dt <- merge(locs, bls, by.x = 'county', by.y = 'area.title', all.x = T)
+    dt <- dt[, list(variable = mean(qtrly.estabs.count %>% as.numeric)),
+             by = list(county, team, year)][order(year, variable)]
+    setnames(dt, 'year', 'season')
+    dt[, season := as.numeric(season) %>% `+`(1)]
+    save(dt, file = paste0('tmp_data/', suffix, '.RData'))
+}
+
+CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
+         suffix = 'nba_musicians')
+CleanBLS(party.regex = c('wine', 'spirit', 'beer', 'liquor', 'drinking'),
+         suffix = 'nba_drinking')
+
+################################################################################
+### All below is now old code, contained in functions above.
+### Leaving it to give a sense of workflow
+################################################################################
+
[…]

### 2017-06-14 writing/draft.tex
Fixed camel-casing in regression tables.

@@
 \begin{center}
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\ \hline \\[-1.8ex]   party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   party continuous &  & $-$0.050$^{**}$ \\    &  & (0.022) \\    & & \\   lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   logged travel distance & $-$0.017 & $-$0.015 \\    & (0.026) & (0.026) \\    & & \\   east-west travel direction & $-$0.146 & $-$0.115 \\    & (0.231) & (0.231) \\    & & \\   number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   time of game during day & 0.005 & 0.005 \\    & (0.013) & (0.013) \\    & & \\   home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.036 \\    & (0.378) & (0.378) \\ \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,804.511 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\ \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   Party continuous &  & $-$0.050$^{**}$ \\    &  & (0.022) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Log of travel distance & $-$0.017 & $-$0.015 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.146 & $-$0.115 \\    & (0.231) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.005 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   Log of travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.036 \\    & (0.378) & (0.378) \\ \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,804.511 \\  Akaike Inf.
 Crit.
@@
 
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Probability of Winning \\  \hline \\[-1.8ex]   Continuous measure of Nightlife & $-$0.035$^{*}$ \\    & (0.020) \\    & \\   Bookmaker's odds & 2.602$^{***}$ \\    & (0.162) \\    & \\   Home-team effect & 0.025 \\    & (0.034) \\    & \\   Number of rest days & 0.014 \\    & (0.019) \\    & \\   Logged travel distance & $-$0.001 \\    & (0.005) \\    & \\   Weekend & 0.040 \\    & (0.036) \\    & \\   Constant & $-$1.331$^{***}$ \\    & (0.080) \\    & \\  \hline \\[-1.8ex]  Observations & 25,215 \\  Log Likelihood & $-$17,274.450 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Probability of Winning \\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.035$^{*}$ \\    & (0.020) \\    & \\   Bookmaker's odds & 2.602$^{***}$ \\    & (0.162) \\    & \\   Home-team effect & 0.025 \\    & (0.034) \\    & \\   Number of rest days & 0.014 \\    & (0.019) \\    & \\   Log of travel distance & $-$0.001 \\    & (0.005) \\    & \\   Weekend & 0.040 \\    & (0.036) \\    & \\   Constant & $-$1.331$^{***}$ \\    & (0.080) \\    & \\  \hline \\[-1.8ex]  Observations & 25,215 \\  Log Likelihood & $-$17,274.450 \\  Akaike Inf.
 Crit.
@@
 
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread \\  \hline \\[-1.8ex]   party placebo & 0.066 \\    & (0.088) \\    & \\   lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   logged travel distance & $-$0.006 \\    & (0.026) \\    & \\   east-west travel direction & $-$0.102 \\    & (0.232) \\    & \\   number hours rest time & 0.001 \\    & (0.001) \\    & \\   time of game during day & 0.002 \\    & (0.013) \\    & \\   home team effect & $-$0.008 \\    & (0.043) \\    & \\   logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.067 \\    & (0.378) \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.801 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread \\  \hline \\[-1.8ex]   Party placebo & 0.066 \\    & (0.088) \\    & \\   Lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   Log of travel distance & $-$0.006 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.102 \\    & (0.232) \\    & \\   Number hours rest time & 0.001 \\    & (0.001) \\    & \\   Time of game during day & 0.002 \\    & (0.013) \\    & \\   Home team effect & $-$0.008 \\    & (0.043) \\    & \\   Log travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.067 \\    & (0.378) \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.801 \\  Akaike Inf.
 Crit.
@@
 \paragraph{General}
-In both sports, it's possible there are prolonged effects of partying, and applying some sort of kernel on the number of hours played since the last game could be interesting. 
+In both sports, it's possible there are prolonged effects of partying, and applying some sort of kernel on the number of hours played since the last game could be interesting; recall that our decision to only look at back-to-back games
+within 24 hours of each other was only motivated by intuition and not guided by data. 
 Future work may consider similar effects in European soccer leagues.
 
-\subsection{Conclusion}
+\subsection{Conclusions}
 We do find that visiting a city with active nightlife the day before a game does have 

### 2017-06-14 writing/draft.tex
Fixed camel-casing in (sub)section titles.

@@
 
-\subsection{Individual Players}
+\subsection{Individual players}
 Studies have also been carried out on statistical ranking of the players
@@
 
-\subsection{Laboratory Studies}
+\subsection{Laboratory studies}
 Studies on effects of alcohol on sports performance in a controlled setting
@@
 
-\subsection{Modeling Approach}
+\subsection{Modeling approach}
 
@@
 
-\subsubsection{Placebo Tests} As a sanity check with our NBA data, we interact a
+\subsubsection{Placebo tests} As a sanity check with our NBA data, we interact a
 discrete-indicator for Los Angeles and New York with an 
@@
 
-\subsection{Possible Confounders or potential issues}
+\subsection{Possible confounders or potential issues}
 

### 2017-07-01 writing/draft.tex
Tidied introduction page.

@@
 partly by what takes place off the court. 
+We can't observe what happens between games directly. Instead, we proxy for the 
+possibility of athletes partying by looking at play following games in party cities.
 We are interested to see if 
@@
 Part of the question is determining a reasonable way to measure
-levels of nightlife, and correspondingly which cities are notorious for it; we colloquially
-refer to such cities as ``party cities''.
+levels of nightlife, and correspondingly which cities are notorious for it; we colloquially refer to such cities as ``party cities''.
+At first glance, the topic might seem callow, but in fact there are many interesting
+nuances to work through.
 
-\paragraph{Why do we care} 
-Inquiring about this causal relationship is worthwhile for several reasons. The
+Inquiring about this causal relationship is worthwhile for several reasons. 
+Knowing how player performance is affected by external factors could inform coaches of a more
+optimal play strategy: if visitation to a city with active nightlife is correlated with
+next-day injury, perhaps coaches might bench their star players if the latter game isn't crucial.
+It could also inform team managers how travel schedules should be set, or even team rules.
+Further, there are potential opportunities for arbitrage.  The
 NBA Commissioner Adam Silver has reported that there is \$400 billion per year spent 
-on sports betting in the U.S. alone.\citep{asilver,cnnsilver} Apart from arbitrage opportunities,
-knowing how player performance is affected by external factors could inform coaches of a more
-optimal play strategy. E.g. if visitation to a city with active nightlife is correlated with
-next-day injury, perhaps coaches might bench their star players if the latter game isn't crucial.
+on sports betting in the U.S. alone.\citep{asilver,cnnsilver}
 
-\paragraph{What do we do}
-We exploit data on bookmaker spreads, the expected score 
-differential between two teams after conditioning on observable performance. 
+% \paragraph{What do we do}
+How do we carry out this study?
+We exploit data on bookmaker spreads: the expected score 
+differential between two teams after conditioning on observable performance in past games
+and expectations about the upcoming game. 
 We expect a team to meet the spread half the time, since this is one of the easiest ways for 
 bookmakers to guarantee a profit.\citep{dubnerlevitt} We construct a model which attempts to 
-estimate the causal effect of  visiting a ``party-city'' on subsequent day 
-performance as measured by the spread. In particular, we are only looking at
-cases where games are played back-to-back within 24 hours of each other.
-To the extent that next-day opponent is uncorrelated with exposure to treatment, 
-we have identification in our variable of interest.
+estimate the causal effect of  visiting a ``party city'' on subsequent day 
+performance as measured by the odds of beating the spread. In particular, we only consider the hangover effect 
+on games played back-to-back within 24 hours of each other.
+To the extent that odds of beating the spread against next day opponent 
+is uncorrelated with playing in a party city the day before, which should be the case 
+under an efficient betting market, we have identification in our variable of interest.
 
-\paragraph{What do we find}
-We replicate this analysis for both the National Basketball Association
-(NBA) and Major League Baseball (MLB). In particular, we find that visiting
-a city with active nightlife the day prior to a game does have a statistically
-significant negative effect on a team's likelihood of meeting the point-spread set
-by a betting house, for both NBA and MLB. Within the NBA, we analyze specific performance
-metrics such as number of points allowed and number of rebounds obtained, and find that both are negatively effected by previous day visitation to a city with active nightlife.
+% \paragraph{What do we find}
+We perform this analysis for both the National Basketball Association
+(NBA) and Major League Baseball (MLB). 
+We find that visiting a city with active nightlife the day prior to a game does have a 
+statistically significant negative effect on a team's likelihood of meeting bookmakers' expectations
+for both NBA and MLB.
+Within the NBA, we analyze specific team performance metrics such as number of points allowed and number of rebounds obtained in a game, and find that both are negatively effected by previous day visitation to a city with active nightlife.
 Having established a causal model, we realize a profitable betting scheme for MLB.
 We also perform a robustness check and see that if players rest more than
- 24 hours after visiting a ``party-city'', the hangover effect dissipates.
+ 24 hours after visiting a ``party city'', the hangover effect dissipates.
 
@@
 Professional players openly admit they are interested in nightlife: Matt Barnes recently complained that 
-``there's no nightlife in Utah'' when asked about playing Jazz in the second round of NBA playoffs 
+``there's no nightlife in Utah'' when asked about playing the Jazz in the second round of NBA playoffs 
 in 2017.\citep{cestone}
@@
 at historical performance for all NBA players who began their career after 1974, and compares
-Sunday afternoon performance to overall career performance. A binomial test is carried out, and the 
-study concludes with strong statistical significance that across the NBA, 
-individual players perform worse on Sundays when compared with other days, 
-as measured by the ratio of shots scored to shots taken.
+Sunday afternoon performance to overall career performance. 
+In particular, the study examines
+2,035 players with at least one shot on Sunday games and finds that 1,092 of them shot worse
+on Sundays than they did on all other days, as measured by the ratio of shots scored to shots
+taken. A binomial test concludes that across the NBA, 
+individual players perform worse on Sundays. Effect sizes are not considered, 
+i.e. although players perform worse, we don't know by how much.
 
@@
 the study does not consider a delayed hangover effect).
-Moreover, their sample size was much smaller, because the author only considered Sunday afternoon games
-from 2006-14. Finally, there was no methodology used behind the selection of party cities:
-the authors first tried to define ad-hoc five ``party cities'', and when that did not yield
-a significant result they tried estimating fixed-effects for each city. The author does carry out a binomial test of how many home-game covers vs. away-game covers were met, 
-and concludes
-with weak statistical evidence that teams systematically play worse on Sunday afternoons 
-when on the road, regardless of city.
+Moreover, the sample size was much smaller than ours, because the author only considered Sunday afternoon games from 2006-14.
+Finally, there was no methodology used behind the selection of party cities: the authors first define ad-hoc five ``party cities'', and when that did not yield a significant result they estimate fixed-effects for each city.
+The author does carry out a binomial test of how many home-game covers vs.
+away-game covers were met, and concludes with weak statistical evidence that teams
+systematically play worse on Sunday afternoons when on the road, regardless of city, as measured by
+their ability to meet bookmakers' expectations.
+
+Allow us to elaborate on why we think that a delayed hangover effect is a more plausible model.
+Suppose for a moment that teams do in fact systematically play worse when playing with cities in active nightlife.
+Notice that a simple fixed effect for the team match-up and an interaction with a home team indicator would be sufficient to model this.
+Given the size of the betting market, it's not hard to believe that such a simple arbitrage opportunity has already been considered
+and factored into bookmakers' expectations. There is also the issue of an unknown travel schedule: if a team plays in a party city,
+we can't infer whether they even had the option of being exposed to nightlife without looking at where they were the day prior.
+This motivates us to examine a feature describing lagged game location.
 

### 2017-07-01 clean_bls.r
Removed old code in favor of a re-usable function.

@@ -22,7 +22,7 @@
 ### Set up Workspace
 ##################################################
 
-source('functions.r') # for subsetBLS
+source('functions.r') # for subsetBLS, CleanBLS.
 require(data.table)
 require(ggmap)
 require(magrittr)
@@ -36,105 +36,5 @@ clusterCall(cl, function() {
     require(magrittr)
 })
 
-MyGeoCode <- function(teams) {
-    locs  <- lapply(teams, geocode, output = 'more') %>% rbindlist(., fill = T)
-    setnames(locs, gsub('_', '.', names(locs)))
-    locs <- locs[, lapply(.SD, as.character), .SDcols = 1:ncol(locs)]
-    ### Be careful, since Washington Wizards are in Washington DC (so not a county)
-    locs[, county := ifelse(is.na(administrative.area.level.2),
-                            administrative.area.level.1,
-                            paste(administrative.area.level.2,
-                                  administrative.area.level.1, sep = ', '))]
-    locs <- data.table(team = teams, county = locs$county)
-    save(locs, file = 'tmp_data/team_locations.RData')
-}
-
-CleanBLS <- function(party.regex, suffix, years = 2010:2016, RESCRAPE = F) {
-    ### Subset BLS data into manageable chunks.
-    parLapplyLB(cl, years, SubsetBLS, party.regex = party.regex, sport = suffix)
-    files <- list.files(path = 'tmp_data/bls', pattern = paste0(suffix, '\\.csv$'), full.names = T)
-    bls <- lapply(files, fread) %>% rbindlist    # Load in chunks of data.
-    load(file = 'tmp_data/covers_lines.RData')   # Load lines data to get listing of team-names (overkill)
-    lines[, pct := NULL]
-    if (RESCRAPE) MyGeoCode(unique(lines$team))  # Geocode the locations for each of our NBA teams.
-    load(file = 'tmp_data/team_locations.RData')
-    dt <- merge(locs, bls, by.x = 'county', by.y = 'area.title', all.x = T)  # Merge BLS with location data.
-    dt <- dt[, list(variable = mean(qtrly.estabs.count %>% as.numeric)),     # Average across variables.
-             by = list(county, team, year)][order(year, variable)]
-    setnames(dt, 'year', 'season')
-    dt[, season := as.numeric(season) %>% `+`(1)]
-    save(dt, file = paste0('tmp_data/', suffix, '.RData'))
-}
-
[…]

### 2017-07-01 clean_bls_mlb.r
Removed old code in favor of a re-useable function.

@@ -1,3 +1,16 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Clean BLS Data for MLB
+###
+### Andreas Santucci
+###
+### Date: July 2017
+###
+################################################################################
+################################################################################
+
 require(data.table)
 require(ggmap)
 require(magrittr)
@@ -11,48 +24,8 @@ clusterCall(cl, function() {
     require(magrittr)
 })
 
+### All the heavy lifting is done in our CleanBLS function...
 source('functions.r')
 
-# 'Audio and video media reproduction', 'Arts, entertainment, and recreation'
-party <- c('sound recording', 'music publisher', 'musical group')
-
-years <- 2010:2016
-parLapplyLB(cl, years, SubsetBLS, party.regex = party, sport = 'mlb')
-
-files <- list.files(path = 'tmp_data/bls', pattern = 'mlb\\.csv$', full.names = T)
-bls <- lapply(files, fread) %>% rbindlist
-
-### Load covers data to get a listing of teams.
-abbrs <- fread(file = 'mlb_abbrs.csv')
-setnames(abbrs, c('abbr', 'full', 'nick'))
-
-### Geocode the locations for each of our MLB teams.
-locs  <- lapply(abbrs$full, geocode, output = 'more') %>% rbindlist(., fill = T)
-setnames(locs, gsub('_', '.', names(locs)))
-locs <- locs[, lapply(.SD, as.character), .SDcols = 1:ncol(locs)]
-
-### Be careful, since Washington Wizards are in Washington DC (so not a county)
-locs[, county := ifelse(is.na(administrative.area.level.2),
-                        administrative.area.level.1,
-                        paste(administrative.area.level.2,
-                              administrative.area.level.1, sep = ', '))]
-locs <- data.table(team = abbrs$full, county = locs$county)
-locs[county == 'Maryland', county := 'Baltimore County, Maryland']
-locs[county == 'Missouri', county := 'Clay County, Missouri']
-locs[team == 'atlanta braves', county := 'Fulton County, Georgia'] # Changes in 2017.
-save(locs, file = 'tmp_data/mlb_team_locations.RData')
-
-musicians <- merge(locs, bls, by.x = 'county', by.y = 'area.title', all.x = T)
-musicians <- musicians[, list(nmusicians = mean(qtrly.estabs.count %>% as.numeric)),
[…]

### 2017-07-01 init_model.rnw
Now sorting teams based on distance to NYC for heatmap, re: next day opponent uncorrelated.

@@
 load(file = 'tmp_data/population_by_county_year.RData')
-load(file = 'tmp_data/nba_musicians.RData')
+## load(file = 'tmp_data/nba_musicians.RData')
+load(file = 'tmp_data/nba_studios.RData')
 
@@
 HOURS <- 24
+## nba.lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team & weekdays(date) %in% c('Saturday', 'Sunday'), log(nmusicians+1), 0)]
 nba.lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians+1), 0)]
@@
 
+### We sort our table according to last.game.loc distance to NYC to show travel restrictions.
+load(file = 'tmp_data/distance_matrix.RData')
+dmat <- melt(dmat)
+setDT(dmat)
+setnames(dmat, c('last.game.loc', 'location', 'distance'))
+dmat <- dmat[location == 'brooklyn nets', list(last.game.loc, distance)] 
+setorder(dmat, distance)
+
+tmp <- merge(tmp, dmat, by = 'last.game.loc')
+setnames(tmp, 'distance', 'distance to NYC')
+setorder(tmp, 'distance to NYC')
+tmp[, last.game.loc := factor(last.game.loc, levels = dmat[order(distance, decreasing = T), last.game.loc])]
+tmp[, location      := factor(location,      levels = dmat[order(distance, decreasing = T), last.game.loc])]
+
 pdf('writing/next_day_opponent.pdf', width = 14, height = 14)
@@
     labs(x = 'Current Game Location', y = 'Last Game Location', 
-         title = 'Distribution of Next Day Opponent - NBA') +
+         title = 'Distribution of Next Day Opponent - NBA\n(sorted by distance to NYC)') +
     theme_bw() +
@@
 
+### We sort our table according to last.game.loc distance to NYC to show travel restrictions.
+load(file = 'tmp_data/distance_matrix_mlb.RData')
+dmat <- melt(dmat)
+setDT(dmat)
+setnames(dmat, c('last.game.loc', 'location', 'distance'))
+dmat <- dmat[location == 'new york yankees', list(last.game.loc, distance)] 
+setorder(dmat, distance)
+
+tmp <- merge(tmp, dmat, by = 'last.game.loc')
+setnames(tmp, 'distance', 'distance to NYC')
+setorder(tmp, 'distance to NYC')
+tmp[, last.game.loc := factor(last.game.loc, levels = dmat[order(distance, decreasing = T), last.game.loc])]
+tmp[, location      := factor(location,      levels = dmat[order(distance, decreasing = T), last.game.loc])]
+
 pdf('writing/next_day_opponent_mlb.pdf', width = 14, height = 14)

### 2017-07-01 functions.r
Added functions to replace one-off code in  files.

@@ -1,5 +1,65 @@
 
-SubsetBLS <- function(year, party.regex, sport,
+### Function: MyGeoCode
+### -------------------
+### Given a listing of teams for a particular sport, this function
+###   uses `ggmap::geocode` to determine a longitude/latitude for each location.
+MyGeoCode <- function(teams, sport) {
+    locs <- lapply(teams, geocode, output = 'more') %>% rbindlist(., fill = T)
+    setnames(locs, gsub('_', '.', names(locs)))
+    locs <- locs[, lapply(.SD, as.character), .SDcols = 1:ncol(locs)]
+    ### Here, we use an if-else since Washington D.C. not a county...
+    locs[, county := ifelse(is.na(administrative.area.level.2),
+                            administrative.area.level.1,
+                            paste(administrative.area.level.2,
+                                  administrative.area.level.1, sep = ', '))]
+    locs <- data.table(team = teams, county = locs$county)
+    if (sport == 'mlb') {
+        locs[county == 'Maryland', county := 'Baltimore County, Maryland']
+        locs[county == 'Missouri', county := 'Clay County, Missouri']
+        locs[team == 'atlanta braves', county := 'Fulton County, Georgia'] # Changes in 2017.
+    }
+    save(locs, file = paste0('tmp_data/', sport, '_team_locations.RData'))
+}
+
+### Function: Clean BLS
+### -------------------
+### Given a string describing what kinds of business patterns to look for, a labeling suffix,
+###   a sport, and an aggregation function, this functional subsets each year of BLS
+###   data such that only observations matching the party.regex are retained.
+###   Results are saved to disk according to a labeling which uses suffix and sport.
+CleanBLS <- function(party.regex, suffix, sport, FUN, years = 2010:2016, RESCRAPE = F) {
+    ### Subset each year of BLS data to only observations matching regular expression.
+    parLapplyLB(cl, years, SubsetBLS, party.regex = party.regex, sport = sport, suffix = suffix)
+    files <- list.files(path = 'tmp_data/bls',
+                        pattern = paste0(sport, '_', suffix, '\\.csv$'), full.names = T)
+    bls <- lapply(files, fread) %>% rbindlist
+    ### Fetch a listing of team-names from our points-spread/money-lines data.
+    if (sport == 'nba') {
[…]

### 2017-07-01 init_model.rnw
Incorporated (some of) Charlie V2 changes.

@@
 load(file = 'tmp_data/covers_lines.RData')
-## load(file = 'tmp_data/nmusician_estabs.RData')
 load(file = 'tmp_data/population_by_county_year.RData')
-## load(file = 'tmp_data/nba_musicians.RData')
-load(file = 'tmp_data/nba_studios.RData')
+load(file = 'tmp_data/studios_and_artists_nba.RData')
 
@@
 
-## nba.lines[, list(date, team, opponent, location, 
-##              line, score, outcome, nhours.lgame, 
-##              travel.dist, avg.age, nmusicians)]
-
 @
@@
 ## nba.lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team & weekdays(date) %in% c('Saturday', 'Sunday'), log(nmusicians+1), 0)]
-nba.lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians+1), 0)]
+
+### We create a discrete indicator for party based on LA or NY. As well as a placebo (longer rest)
 nba.lines[, party.discrete := grepl("(los angeles)|(new york)|(brooklyn)", last.game.loc) &
@@
                 last.game.loc != team]
+
+### We then create a continuous measure of nightlife. We rescale to 0-1 so it's comparable.
+nba.lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, log(nmusicians+1), 0)]
+nba.lines[, party := party / max(party, na.rm = T)]
 
@@
 times the players run back and forth on the court. This is crudely measured
-by the number of changes in posession. We back out this quantity.
+by the number of changes in posession. We back out this quantity by looking
+at the number of three pointers and field goals made, the number of missed shots,
+as well as the number of turnovers and rebounds.
 
@@
 ## nba.lines[,.N/2,by=list(season)]
-stargazer(m0, m1, covariate.labels = c('party discrete', 'party continuous', 'lag changes in posession', 'logged travel distance', 'east-west travel direction', 'number hours rest time', 'time of game during day', 'home team effect', 'logged travel distance * east-west', 'Constant'),
+stargazer(m0, m1, covariate.labels = c('Party discrete', 'Party continuous', 
+                                       'Lag changes in posession', 
+                                       'Logged travel distance', 
+                                       'East-west travel direction', 
+                                       'Number hours rest time', 
+                                       'Time of game during day', 
+                                       'Home team effect', 
+                                       'Logged travel distance * east-west', 
+                                       'Constant'),
           dep.var.labels = 'Meet the Spread')
@@
 stargazer(mp, covariate.labels = c('party placebo', 'lag changes in posession', 'logged travel distance', 'east-west travel direction', 'number hours rest time', 'time of game during day', 'home team effect', 'logged travel distance * east-west', 'Constant'),
-          dep.var.labels = 'Meet the Spread')
+          dep.var.labels = 'Meet the Spread (NBA)')
 
@@
 ##################################################
-### Why drinking estabs doesn't work.
+### Why drinking estabs doesn't work. (NBA)
 ##################################################
-load(file = 'tmp_data/nba_drinking.RData')
+load(file = 'tmp_data/alcohol_and_drinking_nba.RData')
 drinking <- copy(dt)
@@
 
-tmpm <- musicians[, list(log.nmusicians = mean(log(nmusicians+1)),
-                         log.population = mean(log(population))), 
+tmpm <- nba.lines[, list(log.nmusicians = mean(log(nmusicians+1), na.rm = T),
+                         log.population = mean(log(population),   na.rm = T)), 
                   by = team][order(log.nmusicians)] %>% na.omit
@@
 mlb.lines <- lines
-mlb.lines[, party := ifelse(team != location & weekend == 1, log(nmusicians), 0)]
 
@@
 mlb.lines <- mlb.lines[, head(.SD, n=162),by=list(season,team,date)] #Exclude playoff games
+
+mlb.lines[, party := ifelse(team != location & weekend == 1, log(nmusicians), 0)]
+mlb.lines[, party := party / max(party, na.rm = T)]
+
+### See what happens if we don't interact with weekend, i.e. no weekend effect.
+mlb.lines[, party.now := ifelse(team != location, log(nmusicians), 0)]
+mlb.lines[, party.now := party.now / max(party.now, na.rm = T)]
 
@@
 summary(m3)
+m4 <- update(m3, . ~ party.now + . - party)
 
@@
 
-## stargazer(m3, covariate.labels = c('Continuous measure of nightlife',
-##                                    'Bookmaker\'s odds', 'Home-team effect',
-##                                    'Number of rest days', 'Logged travel distance',
-##                                    'Weekend', 'Constant'), 
-##           dep.var.capation = 'Probability of Winning')
+stargazer(m3, m4, covariate.labels = c('Continuous measure of nightlife',
+                                       'Nightlife (no weekend interaction)', 
+                                       'Bookmaker\'s odds', 'Home-team effect',
+                                       'Number of rest days', 'Logged travel distance',
+                                       'Weekend', 'Constant'), 
+          dep.var.labels = 'Probability of Winning')
+
+stargazer(m4, covariate.labels = c('Party placebo',
+                                   'Bookmaker\'s odds', 'Home-team effect',
+                                   'Number of rest days', 'Logged travel distance',
+                                   'Weekend', 'Constant'), 
+          dep.var.labels = 'Probability of Winning (MLB)')
 
@@
 ##################################################
-load(file = 'tmp_data/nmusician_estabs_mlb.RData')
-tmp <- musicians[, list(party.continuous = mean(log(nmusicians+1))), 
+load(file = 'tmp_data/studios_and_artists_mlb.RData')
+setnames(dt, 'variable', 'nmusicians')
+
+tmp <- dt[, list(party.continuous = mean(log(nmusicians+1))),
           by = team][order(party.continuous)][!is.na(party.continuous)]

### 2017-07-01 writing/draft.tex
Incorporated (some of) Charlie V2 changes.

@@
 \usepackage{rotating}
+\usepackage{multicol}
 
@@
   \includegraphics[scale=0.45]{next_day_opponent}
-  \caption{We examine the relationship between last game location and next day opponent in the NBA.
+  \caption{We examine the relationship between last game location and next day opponent in the NBA, sorted by distance from New York City. 
 Starting from the top left, notice that the most likely scenario across all seasons is that a game will be played at home, and the last game was also at home.
@@
 There are notable exceptions, such
-as Oddel Beckham Jr.'s recent trip to Miami.\citep{bleler}
+as Oddel Beckham Jr.'s trip to Miami in the 2016-17 season.\citep{bieler, bleler}
 However, these reports are few and far enough between that picking up statistical evidence would be fruitless.
 Instead, we develop two approaches which do not rely on this data: a discrete indicator 
-based on player interviews and also a continuous measure of nightlife based on local community patterns.
+based on (a recently retired) player interview and also a continuous measure of nightlife based on local community patterns.
 
@@
 particularly musicians.
-This motivated a continuous measure of nightlife, based on a simple count of the number of musicians in the county in which the stadium resides: if the last game was played at most 24-hours ago, use the log of the average number
+This motivated a continuous measure of nightlife, based on a simple count of the number of musicians in the county in which the stadium resides: if the last game was played at most 24-hours ago, use the log of the total number
 of musical groups, sound recording studios, and musical publishers in county of the 
-last game location, else emit zero.  
-This captures
-how much nightlife there is in a city since musical performances are correlated
+last game location, else emit zero.  We also re-scale this feature to lie within the
+interval $[0,1]$ such that the magnitude of the coefficient is comparable with our
+discrete indicator of nightlife.
+This captures how much nightlife there is in a city since musical performances are correlated
 with parties, if not an essential feature.
@@
 outcome of each game alongside the either the point-spread (for NBA) or money-line (for MLB) 
-set by a betting house. 
+set by a betting house. A point-spread describes the expected score differential
+between two teams, and a money-line describes the odds that a
+team will win a match-up.
+
 We interpret the point-spread according to a standard guide.\citep{hrisports} 
+The reason we utilize bookmaker spreads is because this measure already incorporates
+up-to-date information on how well the team as been performing in said season and also
+their expected (dis)-advantage relative to a particular opponent. Without a feature
+describing bookmaker expectations, we would need to construct a separate model (from scratch)
+which estimates the probability that one team will be victorious over another, for all
+$\binom{n}{2}$ possible match-ups.
 
@@
 season, we have at most two games missing of the $82 \times 30 / 2 = 1,230$ played.
-We have a total of 7,829 games in our data-set.
-To this data, we've added features such as 
-the number of days since last game, the last game location, and the travel 
-distance between the last game location and the current game location.
+We have a total of 7,829 games in our data-set. We add features
+to our data such as number of days since last game, lagged game location, and 
+travel distance.
 
@@
 We don't have all $162 \times 30 / 2 = 2,430$ games played each season, but instead only have
-on average almost 2,100 games per season. We have a total of 12,709 games in our data-set. Again,
-we add features such as the number of days since the last game, the last game location, and travel
-distance between game locations.
+on average almost 2,100 games per season. We posit that the missing observations
+are missing in a way that is exogenous to our model; a hand-inspection of several
+reveals that the website URL was defunct.
+We have a total of 12,709 games in our data-set. We again add features describing rest time and travel distance.
 
-\subsubsection{Logic for using points-spread and money-line data}
+% \paragraph{Feature creation} To each of these data sets, we've added features such as 
+% the number of days since last game, the last game location, and the travel 
+% distance between the last game location and the current game location.
+
+\subsubsection{Logic for using betting metrics}
 In order to assert a causal relation between cities with active nightlife 
@@
 number of sound recording studios, musical groups, and music publishers there 
-are for a particular county-year: we simply calculate the average number of establishments across all three music categories listed above.
+are for a particular county-year: we simply calculate the total number of establishments across all three music categories listed above.
 We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.\footnote{We don't expect this to be a party city.}
@@
 an away game as well. I.e. most games after playing in a party city will be an away game.
-This should be covered by spread data, but we also include
+This should be accounted for in the points-spread, but we also include
 a home-team indicator to be sure.
@@
 
-\begin{center}
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\ \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   Party continuous &  & $-$0.050$^{**}$ \\    &  & (0.022) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Log of travel distance & $-$0.017 & $-$0.015 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.146 & $-$0.115 \\    & (0.231) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.005 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   Log of travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.036 \\    & (0.378) & (0.378) \\ \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,804.511 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   Party continuous &  & $-$0.199$^{**}$ \\    &  & (0.100) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Logged travel distance & $-$0.017 & $-$0.014 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.146 & $-$0.112 \\    & (0.231) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.003 \\    & (0.043) & (0.043) \\    & & \\   Logged travel distance * east-west & 0.022 & 0.017 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.034 \\    & (0.378) & (0.378) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,805.106 \\  Akaike Inf.
 Crit.
-& 13,619.350 & 13,627.020 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01}
-\end{tabular} 
-\end{center}
+& 13,619.350 & 13,628.210 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Probability of Winning \\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.035$^{*}$ \\    & (0.020) \\    & \\   Bookmaker's odds & 2.602$^{***}$ \\    & (0.162) \\    & \\   Home-team effect & 0.025 \\    & (0.034) \\    & \\   Number of rest days & 0.014 \\    & (0.019) \\    & \\   Log of travel distance & $-$0.001 \\    & (0.005) \\    & \\   Weekend & 0.040 \\    & (0.036) \\    & \\   Constant & $-$1.331$^{***}$ \\    & (0.080) \\    & \\  \hline \\[-1.8ex]  Observations & 25,215 \\  Log Likelihood & $-$17,274.450 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.142$^{*}$ &  \\    & (0.083) &  \\    & & \\   Nightlife (no weekend interaction) &  & 0.004 \\    &  & (0.107) \\    & & \\   Bookmaker's odds & 2.600$^{***}$ & 2.604$^{***}$ \\    & (0.162) & (0.163) \\    & & \\   Home-team effect & 0.024 & 0.052 \\    & (0.035) & (0.072) \\    & & \\   Number of rest days & 0.015 & 0.015 \\    & (0.019) & (0.020) \\    & & \\   Logged travel distance & $-$0.001 & $-$0.001 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.042 & 0.0004 \\    & (0.037) & (0.030) \\    & & \\   Constant & $-$1.329$^{***}$ & $-$1.348$^{***}$ \\    & (0.080) & (0.102) \\    & & \\  \hline \\[-1.8ex]  Observations & 25,215 & 24,781 \\  Log Likelihood & $-$17,274.540 & $-$16,977.790 \\  Akaike Inf.
 Crit.
-& 34,562.900 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 34,563.090 & 33,969.580 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread \\  \hline \\[-1.8ex]   Party placebo & 0.066 \\    & (0.088) \\    & \\   Lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   Log of travel distance & $-$0.006 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.102 \\    & (0.232) \\    & \\   Number hours rest time & 0.001 \\    & (0.001) \\    & \\   Time of game during day & 0.002 \\    & (0.013) \\    & \\   Home team effect & $-$0.008 \\    & (0.043) \\    & \\   Log travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.067 \\    & (0.378) \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.801 \\  Akaike Inf.
-Crit.
-& 13,631.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
 Indeed, we see that if players have enough recovery time after visiting cities such as Los
 Angeles or New York, we cannot reject the null hypothesis that player performance is unaffected
-by visitation to cities with active nightlife.
+by visitation to cities with active nightlife. The same is true in the MLB.
+
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   party placebo & 0.072 \\    & (0.082) \\    & \\   lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   east-west travel direction & $-$0.100 \\    & (0.232) \\    & \\   number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   time of game during day & 0.001 \\    & (0.013) \\    & \\   home team effect & $-$0.009 \\    & (0.043) \\    & \\   logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.070 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.701 \\  Akaike Inf.
+Crit.
+& 13,631.400 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Probability of Winning (MLB) \\  \hline \\[-1.8ex]   Party placebo & 0.004 \\    & (0.107) \\    & \\   Bookmaker's odds & 2.604$^{***}$ \\    & (0.163) \\    & \\   Home-team effect & 0.052 \\    & (0.072) \\    & \\   Number of rest days & 0.015 \\    & (0.020) \\    & \\   Logged travel distance & $-$0.001 \\    & (0.005) \\    & \\   Weekend & 0.0004 \\    & (0.030) \\    & \\   Constant & $-$1.348$^{***}$ \\    & (0.102) \\    & \\  \hline \\[-1.8ex]  Observations & 24,781 \\  Log Likelihood & $-$16,977.790 \\  Akaike Inf.
+Crit.
+& 33,969.580 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 

### 2017-07-01 init_model.rnw
Formatted plot for why drinks or population don't work.

@@
 axis(side = 2)
-legend(x = 22.5, y = 10.25, legend = c('Musicians and sound recording studios',
+legend(x = 21.5, y = 12.5, legend = c('Musicians and sound recording studios',
                                  'Drinking establishments and liquor stores',
                                  'Population'), pch = c(1, 16, 18), col = c('black', 'purple', 'orange'))
-text(x = 24.6, y = 10.5, labels = paste("Correlation (alcohol):       ", 
+text(x = 24.6, y = 9.5, labels = paste("Correlation (alcohol):       ", 
                                        cor(tmp$log.nmusicians, tmp$log.ndrinks) %>% round(., 3)))
-text(x = 24.6, y = 11, labels = paste("Correlation (population): ", 
+text(x = 24.6, y = 10, labels = paste("Correlation (population): ", 
                                        cor(tmp$log.nmusicians, tmp$log.population) %>% round(., 3)))

### 2017-07-01 clean_bls.r
Added alcohol and drinking file.

@@ -38,3 +38,7 @@ clusterCall(cl, function() {
 
 CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
          suffix = 'studios_and_artists', sport = 'nba', FUN = sum)
+
+
+CleanBLS(party.regex = c('alcohol', 'drinking'),
+         suffix = 'alcohol_and_drinking', sport = 'nba', FUN = sum)

### 2017-07-02 cardinal_analytx_presentation/nightlife_and_sports.tex
Init commit. Working title page template.

\documentclass{beamer} \usepackage{tikz} \usepackage{pdfrender}  \usetikzlibrary{shapes,arrows} \setbeamerfont{author}{size=\Huge} \setbeamerfont{institute}{size=\normalsize\itshape} \setbeamerfont{title}{size=\fontsize{30}{36}\bfseries} \setbeamerfont{subtitle}{size=\Large\normalfont\slshape}  \setbeamertemplate{title page} { \begin{tikzpicture}[remember picture,overlay] \fill[orange]   ([yshift=15pt]current page.west) rectangle (current page.south east); \node[anchor=east]    at ([yshift=-50pt]current page.north east) (author)   {\parbox[t]{.6\paperwidth} {\raggedleft      \usebeamerfont{author}\textcolor{orange} {     \textpdfrender{     TextRenderingMode=FillStroke,     FillColor=orange,     LineWidth=.1ex,     }{\insertauthor}}}};
\node[anchor=north east]
  at ([yshift=-70pt]current page.north east) (institute) {
    \parbox[t]{.78\paperwidth} { 
      \raggedleft \usebeamerfont{institute}\textcolor{gray} {\insertinstitute}}}; 
\node[anchor=south west] at ([yshift=20pt]current page.west) (logo) {
  \parbox[t]{.19\paperwidth} { 
    \raggedleft \usebeamercolor[fg]{titlegraphic} \inserttitlegraphic}}; 
\node[anchor=east] at ([yshift=-10pt,xshift=-20pt]current page.east) (title) {
  \parbox[t]{\textwidth} {
    \raggedleft \usebeamerfont{author}\textcolor{white} {  
      \textpdfrender { 
        TextRenderingMode=FillStroke, FillColor=white, LineWidth=.1ex}{\inserttitle}}}}; \node[anchor=east] at ([yshift=-60pt,xshift=-20pt]current page.east) (subtitle) {
  \parbox[t]{.6\paperwidth}{\raggedleft\usebeamerfont{subtitle}\textcolor{black}{\insertsubtitle}}};  
\end{tikzpicture} }  \author{John Doe} \institute{Harvard College Professor and \\ Johnstone Family Professor in Psychology, Harvard University} \title{Say What?} \subtitle{Linguistics as a Window for Understanding the Brain} % \titlegraphic{\includegraphics[width=2cm]{ctanlion}}

 \begin{document}  \begin{frame} \maketitle \end{frame}  \end{document}

### 2017-07-03 cardinal_analytx_presentation/nightlife_and_sports.tex
First draft.

@@
-\documentclass{beamer} \usepackage{tikz} \usepackage{pdfrender}  \usetikzlibrary{shapes,arrows} \setbeamerfont{author}{size=\Huge} \setbeamerfont{institute}{size=\normalsize\itshape} \setbeamerfont{title}{size=\fontsize{30}{36}\bfseries} \setbeamerfont{subtitle}{size=\Large\normalfont\slshape}  \setbeamertemplate{title page} { \begin{tikzpicture}[remember picture,overlay] \fill[orange]   ([yshift=15pt]current page.west) rectangle (current page.south east); \node[anchor=east]    at ([yshift=-50pt]current page.north east) (author)   {\parbox[t]{.6\paperwidth} {\raggedleft      \usebeamerfont{author}\textcolor{orange} {     \textpdfrender{     TextRenderingMode=FillStroke,     FillColor=orange,     LineWidth=.1ex,     }{\insertauthor}}}};
+\documentclass{beamer} 
+\usepackage{tikz} 
+\usepackage{pdfrender} 
+\usepackage{mathtools}
+\usepackage{bbm}
+\usepackage{hyperref}
+\usepackage{changepage}
+
+\setbeamertemplate{navigation symbols}{}
+
+\usetikzlibrary{shapes,arrows} \setbeamerfont{author}{size=\Huge} \setbeamerfont{institute}{size=\normalsize\itshape} \setbeamerfont{title}{size=\fontsize{30}{36}\bfseries} \setbeamerfont{subtitle}{size=\Large\normalfont\slshape}  \setbeamertemplate{title page} { \begin{tikzpicture}[remember picture,overlay] \fill[orange]   ([yshift=15pt]current page.west) rectangle (current page.south east); \node[anchor=east]    at ([yshift=-50pt]current page.north east) (author)   {\parbox[t]{.6\paperwidth} {\raggedleft      \usebeamerfont{author}\textcolor{orange} {     \textpdfrender{     TextRenderingMode=FillStroke,     FillColor=orange,     LineWidth=.1ex,     }{\insertauthor}}}};
 \node[anchor=north east]
@@
   \parbox[t]{.6\paperwidth}{\raggedleft\usebeamerfont{subtitle}\textcolor{black}{\insertsubtitle}}};  
-\end{tikzpicture} }  \author{John Doe} \institute{Harvard College Professor and \\ Johnstone Family Professor in Psychology, Harvard University} \title{Say What?} \subtitle{Linguistics as a Window for Understanding the Brain} % \titlegraphic{\includegraphics[width=2cm]{ctanlion}}
+\end{tikzpicture} }  \author{Andreas Santucci} \institute{Stanford Teaching Fellow \\ Computational Math, Data Science} \title{Effects of nightlife on sports} \subtitle{Cardinal Analytx Interview} % \titlegraphic{\includegraphics[width=2cm]{ctanlion}}
+ \begin{document}  \begin{frame} \maketitle \end{frame}
 
- \begin{document}  \begin{frame} \maketitle \end{frame}  \end{document}
+\begin{frame}   \frametitle{What is the problem?}
+  \begin{block}{Do pro athletes party to the detriment of their performance?}
+    It's not unreasonable to think that what happens off the court affects outcome, but!
+    \begin{itemize}       \item Can we even measure this?
+      \item Are there systematic tendencies?     \end{itemize}
+  \end{block} \end{frame}
+
+\begin{frame}   
+  \frametitle{Who cares?}
+  \begin{block}{\$400 billion spent on sports betting per year in the US.}
+    Apparently, a lot of people.
+
+    Knowing the causal relation between partying and next-day performance might inform
+    \begin{itemize}       \item How coaches set team-rules, 
+      \item How travel schedules are determined.     \end{itemize}
+  \end{block}
+\end{frame}
+
+\begin{frame}   
+  \frametitle{What is our solution?}
+  \begin{block}{We can't observe whether players party.}
+    Instead, we consider a quasi-experiment.
+    \begin{itemize}       \item Do teams play worse the day after visiting a ``party city''?
+      \item How do we even determine whether a city has nightlife?     \end{itemize}
+  \end{block}
+\end{frame}
+
+\begin{frame}
+  \frametitle{How can we measure unexpected changes in performance?}   \begin{block}{Bookmaker spreads}     The expected score differential between two teams.
+
+    Conditions on observed performance, and future expectations.
+
+    A team is expected to meet the spread $\frac{1}{2}$ the time.   \end{block} \end{frame}
+
+\begin{frame}   \frametitle{Quasi Experiment}
+  \begin{block}{Consider games played back-to-back within 24 hours.}     If odds of beating the spread uncorrelated with next day opponent 
+    $\longrightarrow$ identification.
+
+  (Under an efficient market, this should be the case)    \end{block}  
+\end{frame}
+
+\begin{frame}   \frametitle{Sports: NBA and MLB}
+  \begin{block}{We perform the analysis for both NBA and MLB}     If we yield a causal model, we can realize a betting schema.   \end{block} \end{frame}
+
+\begin{frame}   \frametitle{NBA Data: 2010-11 through 2016-17 season}
+  \begin{itemize}     \item Each team plays 82 games per season.
+    \item 30 teams in the NBA.
+    \item Total of $82 \times 30 / 2 = 1,230$ games per season.  
+  \end{itemize}
+
+  We query \href{http://www.basketball-reference.com/leagues/}{basketball-reference.com}
+  for a listing of game-days, and then 
+  \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{covers.com} for 
+  game outcomes alongside a point-spread.
+
+  \begin{itemize}     \item Missing the first couple months of 2010-11 season.
+    \item At most two missing games in each of remaining seasons.
+    \item Total of 7,829 games in our dataset.   \end{itemize}
+\end{frame}
+
+\begin{frame}   \frametitle{MLB Data: 2011 through 2017 season}
+  \begin{itemize}     \item Each team plays 162 games per season.
+    \item 30 teams in the MLB.
+    \item Total of $162 \times 30 / 2 = 2,430$ games per season.   \end{itemize}
+
+  We query \href{http://www.baseball-reference.com/leagues/}{baseball-reference.com}
+  for a listing of game days, and again look to Covers.com for game outcomes and money-lines.
+
+  \begin{itemize}     \item We have on average 2,100 games per season.
+    \item (Missing observations due to broken links?)
+    \item Total of 12,709 games in our dataset.    \end{itemize} \end{frame}
+
+\begin{frame}   \frametitle{Bureau of Labor Statistics Data: 2010-2016}
+  We look toward BLS data which describes the number of sound recording
+  studios, musical groups, and musical publishers in each metropolitan statistical area
+  for each quarter of each year.
+
+  All teams are in the US with the exception of 
+  \begin{itemize}     \item Toronto Raptors (NBA)
+    \item Toronto Blue Jays (MLB)   \end{itemize}
+
+  These two locations are excluded from our analysis. \end{frame}
+
+\begin{frame}   \frametitle{Minutes data for NBA}
+  For NBA, we also collect minutes data, which describes the number of 
+  \begin{itemize}     \item Free-throws
+    \item Field goals
+    \item Three-pointers   \end{itemize}
+  attempted and made by each player in each game. For each season and each team,
+  we have nearly complete minutes data; 85\% of our season-teams' have no more than
+  3 missing games. \end{frame}
+
+\begin{frame}   
+  \frametitle{Possible Confounders}
+  What might influence game-outcomes apart from player-ability?
+
+  \begin{itemize}     
+    \item Home team effect
+    \item Fatigue from previous game
+    \item Jet lag   
+    \end{itemize}
+
+  The benefit of using spread data is that it controls for salient factors which affect game outcomes.
+  The downside is that we don't know exactly what it controls for. \end{frame}
+
+\begin{frame}   \frametitle{Feature Creation}
+  We add features which describe
+  \begin{itemize}     \item Number of days since the last game (i.e. rest-time),
+    \item Lagged game location, and
+    \item Travel Distance.   \end{itemize}
+  We seek to construct a model which accounts for travel-fatigue and jet-lag.
+  For this, we need a little bit more. \end{frame}
+
+\begin{frame}   \frametitle{Accounting for Jetlag}
+  In order to account for jet-lag, we need to be mindful of the direction of travel. We calculate
+  \begin{itemize}     \item distance traveled between games,
+    \item final direction (bearing) traveling from $a \to b$, and
+    \item direction of travel (in east-west degrees) by taking sine.
+  \end{itemize}
+  Finally, to account for jetlag we interact this east-west measure with travel distance. \end{frame}
+
+\begin{frame}   \frametitle{Next day opponent: NBA}
+  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent} \end{frame}
+
+\begin{frame}   \frametitle{A closer look at NBA games on the road}
+  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent_on_tour} \end{frame}
+
+\begin{frame}   \frametitle{Further motivation for conditioning on the spread}
+  \begin{block}{With some correlation between next-day opponent, this further motivates conditioning
+  on bookmakers' expectations.}
+  \begin{itemize}     \item If bookkeepers aware of hangover effects, our variable of interest may not appear significant, in spite of there being a true effect.
+
+    \item If bookkeepers systematically mis-value teams which are more likely to be encountered after visiting a city with active nightlife, we could see a spurious effect.
+
+      (Given size of betting market, this is unlikely)
+    \end{itemize}   \end{block}
+\end{frame}
+
+\begin{frame}   \frametitle{Partying as a latent variable}
+  \begin{block}{We can't observe when players go for a night on the town.}
+    Instead, we develop two approaches to proxy for partying.
+
+    A discrete indicator and a continuous measure.
+  \end{block} \end{frame}
+
+\begin{frame}   \frametitle{Discrete Indicator}
+  \begin{block}{Anecdotally, LA and NY appear in news articles frequently.}     This motivates a binary indicator for party when either of these two cities have been visited within 24-hours of another game.
+    \[
+      \texttt{party} \coloneqq \mathbbm 1 \left \{\substack{\textrm{last game played in LA or NY} \\ \textrm{at most 24-hours ago}}\right \}
+    \]   \end{block} \end{frame}
+
+\begin{frame}   
+  \frametitle{Travel Distance Comparison}
+  \centering \includegraphics[scale=0.25]{../writing/travel_dist_density_by_party} 
+\end{frame}
+
+\begin{frame}   \frametitle{Continuous Measure}
+  \begin{block}{A-list celebrities tend to party with other celebrities.}     Idea: proxy nightlife using number of musicians in previous location
+
+    \[
+      \texttt{party} \coloneqq \begin{cases}       \log \left(\# \textrm{musical groups}\right) &\substack{\textrm{ if last game played} \\ \textrm{at most 24-hours ago,}} \\
+      0 &\textrm{ otherwise.}
+    \end{cases}
+    \]
+    \newline
+    Specifically, we consider the sum total of
+    \[
+      \textrm{Sound recording studios } + \textrm{ musical groups } + \textrm{ musical publishers}.
+    \]   \end{block} \end{frame}
+
+\begin{frame}   
+  \frametitle{Continuous Measure of Party (NBA)}
+  \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location} 
+\end{frame}
+
+\begin{frame}   \frametitle{Results: NBA}
+  \vspace{-2pt}
+  \centering
+  \tiny{
+  \begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline   \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\ & & \\   [-1.8ex] Party continuous &  & $-$0.181$^{**}$ \\    &  & (0.091) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Logged travel distance & $-$0.017 & $-$0.014 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.145 & $-$0.113 \\    & (0.232) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   Logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.028 \\    & (0.378) & (0.378) \\ [-1.8ex] & & \\ \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.677 & $-$6,805.100 \\  Akaike Inf.
+Crit.
+& 13,619.350 & 13,628.200 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+}   \end{frame}
+
+\begin{frame}   \frametitle{Placebo Test: what if we let players rest?}
+  \centering
+  \tiny{
+  \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   party placebo & 0.072 \\    & (0.082) \\    & \\   lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   east-west travel direction & $-$0.099 \\    & (0.232) \\    & \\   number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   time of game during day & 0.001 \\    & (0.013) \\    & \\   home team effect & $-$0.009 \\    & (0.043) \\    & \\   logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.070 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.702 \\  Akaike Inf.
+Crit.
+& 13,631.400 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+} \end{frame}
+
+\begin{frame}   \frametitle{Differences between NBA and MLB}
+  The same analysis can be replicated for MLB. Notable differences:
+  \begin{itemize}     \item Money-lines are used instead of point-spreads. 
+      So instead of predicting likelihood of meeting the spread,
+      we predict likelihood of winning conditional on bookmakers odds.
+    \item We interact our measure of nightlife with a weekend indicator.
+    \item Games are played in series, so we no longer have the problem of an unknown travel schedule, i.e. players stay overnight in the same location within a series.   \end{itemize} \end{frame}
+
+\begin{frame}   
+  \frametitle{Continuous Measure of Party (MLB)}
+  \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location_mlb}   
+\end{frame}
+
+\begin{frame}   \frametitle{Results: MLB}
+  \centering
+  \tiny{
+  \begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.126$^{*}$ &  \\    & (0.074) &  \\    & & \\   Nightlife (no weekend interaction) &  & $-$0.016 \\    &  & (0.110) \\    & & \\   Bookmaker's odds & 2.638$^{***}$ & 2.583$^{***}$ \\    & (0.163) & (0.165) \\    & & \\   Home-team effect & 0.008 & 0.034 \\    & (0.035) & (0.083) \\    & & \\   Number of rest days & 0.019 & 0.018 \\    & (0.020) & (0.020) \\    & & \\   Logged travel distance & $-$0.002 & $-$0.003 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.044 & 0.003 \\    & (0.038) & (0.030) \\    & & \\   Constant & $-$1.344$^{***}$ & $-$1.324$^{***}$ \\    & (0.081) & (0.113) \\    & & \\  \hline \\[-1.8ex]  Observations & 24,734 & 24,033 \\  Log Likelihood & $-$16,946.080 & $-$16,470.850 \\  Akaike Inf.
+Crit.
+& 33,906.160 & 32,955.700 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+} \end{frame}
+
+\begin{frame}   \frametitle{Validating our model in the market}
+  Our results are statistically significant, but so what? We validate our model by seeing
+  how well we fare against the market.
+
+  \begin{itemize}     \item We hold out data for validation.
+    \item For each unseen relevant game, we place a bet of \$100 if the expected value of the bet
+      is positive after accounting for bookmakers cut (\$10)
+    \item Are we profitable?
+  \end{itemize} \end{frame}
+
+\begin{frame}   \frametitle{Betting in MLB}
+  \centering
+  \includegraphics[scale=0.5]{../writing/mlb_bets_by_season} \end{frame}
+
+\begin{frame}   \frametitle{Returning to NBA: Player Performance}
+  We additionally look toward actual performance metrics, instead of relying on point-spread data.
+
+  \begin{itemize}     \item Avoid using black-box feature
+    \item Pinpoint more precisely what elements of game-play are affected by nightlife visitation.   \end{itemize}
+
+  Anecdotally, rebounds are considered a ``hustle'' statistic, in the sense that players
+  who work harder on the court and expend energy boxing out other players will have higher
+  rebound rates. \end{frame}
+
+\begin{frame}   \frametitle{Analysis: rebounds}
+  \begin{itemize}     \item For each player and each game, we calculate the number of rebounds obtained per-minute of gamep-lay
+    \item From each observation, we subtract the player's average number of rebounds per minute taken across the entire season.
+    \item We then run a linear regression explaining demeaned rebounds as a function of 
+      nightlife visitation and other features describing player fatigue.   \end{itemize} \end{frame}
+
+\begin{frame}   \frametitle{Results: rebounds}
+  \centering
+  \tiny{
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.003$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00003 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00003 \\    & (0.00002) \\    & \\   Constant & $-$0.008$^{*}$ \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 43,547 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
+Error & 0.104 (df = 43542) \\  F Statistic & 2.945$^{**}$ (df = 4; 43542) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+} \end{frame}
+
+\begin{frame}   \frametitle{Hangover correlates with allowing more points}
+  We examine points allowed and points scored (on the team level).
+
+  Common knowledge suggests that overall defense of a team is strongly influenced by individual player effort.
+
+  Is it the case the hangover effects are more prevalent on defensive aspects of gameplay? \end{frame}
+
+\begin{frame}   \frametitle{Defense in the NBA}
+  \begin{adjustwidth}{-.75cm}{}
+  \tiny{
+  \begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.828$^{***}$ &  &  \\    & (0.699) &  &  \\    & & & \\   Continuous party measure &  & 1.667$^{***}$ & 0.080 \\    &  & (0.587) & (0.575) \\    & & & \\   Lag change in possessions &  & 0.267$^{***}$ & 0.285$^{***}$ \\    &  & (0.016) & (0.016) \\    & & & \\   Logged travel distance &  & $-$0.447$^{**}$ & 0.143 \\    &  & (0.191) & (0.187) \\    & & & \\   Number hours since last game &  & 0.011 & 0.017$^{**}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Constant & $-$0.047 & 70.710$^{***}$ & 64.859$^{***}$ \\    & (0.090) & (2.271) & (2.223) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,658 & 6,426 & 6,426 \\  R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Adjusted R$^{2}$ & 0.001 & 0.044 & 0.051 \\  Residual Std.
+Error & 11.159 (df = 15656) & 11.941 (df = 6421) & 11.688 (df = 6421) \\  F Statistic & 16.357$^{***}$ (df = 1; 15656) & 74.877$^{***}$ (df = 4; 6421) & 86.792$^{***}$ (df = 4; 6421) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+}
+\end{adjustwidth} \end{frame}
+
+\begin{frame}   \frametitle{Future Research}
+  Within the NBA
+  \begin{itemize}     \item Are teams more likely to party after a win or a loss?
+    \item Do teams who aren't going to make the playoffs party more?
+    \item Are there demographic differences which influence propensity to party? E.g. are younger, unmarried players more susceptible to our treatment effect?
+    \item We could also examine how shots taken vs. shots scored are affected for star players.
+  \end{itemize}
+  A preliminary and ad-hoc analysis shows that injury risk increases as a function of party.
+
+  This might inform coaches which players to bench and when.
+
+  Within the MLB, we can look at statistics such as batting averages or general fielding statistics. \end{frame}
+
+\begin{frame}   \frametitle{Future Research (General)}
+  \begin{itemize}     \item It's possible there are prolonged effects; recall our magic 24 hour number.
+    \item European soccer leagues.   \end{itemize} \end{frame}
+
+\begin{frame}   \frametitle{Conclusions}
+  \begin{itemize}     \item NBA and MLB players under-perform bookmakers expectations after visiting LA and NY, and in general we see that performance decreases as a function of (our proxy for) nightlife.
+    \item The decrease in likelihood can (in part) be explained by a drop in defensive performance.
+    \item With a causal model established, we realize a profitable betting scheme.   \end{itemize} \end{frame}
+
+\end{document}

### 2017-07-03 clean_bls.r
Now avoiding to re-scrape data when possible.

@@ -39,6 +39,5 @@ clusterCall(cl, function() {
 CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
          suffix = 'studios_and_artists', sport = 'nba', FUN = sum)
 
-
 CleanBLS(party.regex = c('alcohol', 'drinking'),
-         suffix = 'alcohol_and_drinking', sport = 'nba', FUN = sum)
+         suffix = 'alcohol_and_drinking', sport = 'nba', FUN = sum, RESCRAPE = F)

### 2017-07-03 clean_bls.r
Noted dependencies. No longer re-scraping.

@@ -8,9 +8,11 @@
 ###
 ### Date: April 2017
 ###
-### Inputs: 
+### Inputs: 'raw_data/bls/20[xx].q1-q4.by_area'
 ###
-### Dependencies: 
+### Output: 'tmp_data/bls/20[xx]_[sport]_[suffix].csv'
+###
+### Dependencies: data.table, ggmap, magrittr, parallel
 ###
 ################################################################################
 ################################################################################
@@ -36,8 +38,10 @@ clusterCall(cl, function() {
     require(magrittr)
 })
 
+### Loads raw data. Subsets based on regex. Aggregate using FUN. Save using suffix and sport.
 CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
          suffix = 'studios_and_artists', sport = 'nba', FUN = sum)
 
+### We add option 'rescrape = F' to avoid scraping location data a second time.
 CleanBLS(party.regex = c('alcohol', 'drinking'),
          suffix = 'alcohol_and_drinking', sport = 'nba', FUN = sum, RESCRAPE = F)

### 2017-07-03 clean_bls_mlb.r
Outputting data for comparison plot.

@@ -8,6 +8,13 @@
 ###
 ### Date: July 2017
 ###
+### Inputs: 'raw_data/bls/20[xx].q1-q4.by_area'
+###
+### Output: 'tmp_data/bls/[suffix]_[sport].RData'
+###         e.g. 'studios_and_artists_mlb.RData'
+###
+### Dependencies: data.table, ggmap, magrittr, parallel
+###
 ################################################################################
 ################################################################################
 
@@ -27,5 +34,10 @@ clusterCall(cl, function() {
 ### All the heavy lifting is done in our CleanBLS function...
 source('functions.r')
 
+### Load raw data, subset based on party.regex, aggregate using FUN, and save using suffix/sport.
 CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
          suffix = 'studios_and_artists', sport = 'mlb', FUN = sum)
+
+### Avoids rescraping data. This output used for a comparison plot in the paper.
+CleanBLS(party.regex = c('alcohol', 'drinking'),
+         suffix = 'alcohol_and_drinking', sport = 'mlb', FUN = sum, RESCRAPE = F)

### 2017-07-03 clean_covers.r
Listed file dependencies and outputs.

@@ -10,7 +10,9 @@
 ###
 ### Inputs: 'raw_data/covers/[year]-[month]-[day].csv'
 ###
-### Dependencies: data.table, magrittr
+### Output: 'tmp_data/covers_lines.RData'
+###
+### Dependencies: data.table, ggmap, geosphere, magrittr, sp
 ###
 ################################################################################
 ################################################################################
@@ -20,8 +22,9 @@
 ##################################################
 
 require(data.table)
-require(magrittr)
+require(geosphere)
 require(ggmap)
+require(magrittr)
 require(sp)
 
 files <- list.files('raw_data/covers', full.names = T)
@@ -32,8 +35,6 @@ lines <- lapply(files, fread) %>% rbindlist
 ##################################################
 
 lines[, date := as.POSIXct(date, format = '%A, %B %d %Y %I:%M%p')]
-## lines[, game.time := gsub('^.* ([0-9:APM]+)$', '\\1', date)]
-## lines[, date := as.Date(date, format = '%A, %B %d %Y')]
 
 lines[, line := gsub('^(-?[0-9.]+)/.*$', '\\1', line)]
 lines[, line.ts := NULL]
@@ -87,7 +88,6 @@ lines[, score := paste(home.pts, away.pts, sep = '-')]
 lines[, nominal.outcome := home.pts > away.pts %>% as.integer]
 lines[, `:=`(home.pts = NULL, away.pts = NULL)]
 lines[, location := team]
-#lines <- lines[, list(season, date, team, opponent, location, line, score, outcome)]
 
 ##################################################
 ### Creating a panel-data set.
@@ -143,7 +143,8 @@ lines[, travel.dist := getDist(location, last.game.loc, dmat),
 ##################################################
 ### Direction Traveled.
 ##################################################
-
+load(file = 'tmp_data/nba_team_locations.RData')
+locs[, `:=`(lon = as.numeric(lon), lat = as.numeric(lat))]
 ### Get directions.
 dirs <- matrix(nrow = nrow(dmat), ncol = ncol(dmat), dimnames = dimnames(dmat))
 for (i in 1:(nrow(dirs)))
@@ -209,4 +210,3 @@ lines[, `:=`(team.pts.scored = as.numeric(team.pts.scored),
              team.pts.admitted = as.numeric(team.pts.admitted))]
 
 save(lines, file = 'tmp_data/covers_lines.RData')
-

### 2017-07-03 clean_covers_mlb.r
Added file dependencies.

@@ -8,9 +8,11 @@
 ###
 ### Date: April 2017
 ###
-### Inputs:
+### Inputs: 'tmp_data/covers_mlb/[date].csv'
 ###
-### Dependencies: 
+### Output: 'tmp_data/covers_lines_mlb.RData'
+###
+### Dependencies: data.table, magrittr, sampling
 ###
 ################################################################################
 ################################################################################
@@ -99,7 +101,7 @@ data <- data[!is.na(line) & !is.na(matchup)]
 
 ### Set the location of the game.
 ##  Ex: http://www.covers.com/sports/MLB/matchups?selectedDate=2011-4-01
-data[, location := opponent] ### .
+data[, location := opponent]
 
 ### Determine the odds of winning, backed out from the money-line.
 data[, odds := ifelse(line < 0, 100 / (-line + 100), line / (line + 100))]
@@ -171,8 +173,9 @@ lines[, weekend := weekday %in% c('Saturday', 'Sunday')]
 ##################################################
 ### Musicians from BLS
 ##################################################
-load(file = 'tmp_data/nmusician_estabs_mlb.RData')
-lines <- merge(lines, musicians,
+load(file = 'tmp_data/studios_and_artists_mlb.RData')
+setnames(dt, 'variable', 'nmusicians')
+lines <- merge(lines, dt,
                by.x = c('season', 'last.game.loc'),
                by.y = c('season', 'team'), all.x = T)
 
@@ -184,7 +187,7 @@ lines[, first.in.series := ifelse(is.na(lag.opponent) | opponent != lag.opponent
 lines[, not.first := !first.in.series]
 
 ## t <- lines[, mean(team.score > opponent.score), by = list(location, first.in.series)][, diff(V1), by = location][order(V1)]
-lines[, party := ifelse(team != location & weekend == 1, nmusicians, 0)]
+lines[, party := ifelse(team != location & weekend == 1, log(nmusicians), 0)]
 
 glm(I(team.score > opponent.score) ~ party + odds,
          data = lines, family = 'binomial') %>% summary

### 2017-07-03 clean_espn.r
Noted file dependencies

@@ -8,9 +8,12 @@
 ###
 ### Date: April 2017
 ###
-### Inputs: 
+### Inputs: 'tmp_data/team_abbreviations.csv'
+###         ''raw_data/espn[game-id].csv'
 ###
-### Dependencies: 
+### Output: 'tmp_data/espn_player_data.csv'
+###
+### Dependencies: data.table, magrittr, zoo
 ###
 ################################################################################
 ################################################################################
@@ -39,7 +42,6 @@ setnames(data, c('player', 'min', 'fg', 'three.pt', 'ft',
                  'oreb', 'dreb', 'reb', 'ast', 'stl', 'blk',
                  'to', 'pf', 'pm', 'pts', 'team', 'date'))
 data[, team := tolower(team)]
-#data <- na.omit(data)
 
 data <- merge(data, abbrs, by.x = 'team', by.y = 'abbr', all.x = T)
 data[, 1 := NULL, with = F]
@@ -100,3 +102,4 @@ data[, (paste0('lag.', cols)) := shift(.SD),
      by = list(season, player), .SDcols = cols]
 
 write.csv(data, file = 'tmp_data/espn_player_data.csv', row.names = F)
+

### 2017-07-03 init_model.rnw
Mainly formatting changes to plots, and stargazer tables.

@@
 load(file = 'tmp_data/covers_lines.RData')
-load(file = 'tmp_data/population_by_county_year.RData')
+load(file = 'tmp_data/population_by_MSA_year.RData')
 load(file = 'tmp_data/studios_and_artists_nba.RData')
@@
 setnames(musicians, 'variable', 'nmusicians')
-musicians <- merge(musicians, pop, by = c('county', 'season'), all.x = T)
+
+### This is simply ad-hoc.
+musicians[, population := 0.0]
+for (i in 1:nrow(musicians))
+    set(musicians, i, j = 'population', 
+        value = pop[grepl(musicians[i, locality], locality) &
+                    grepl(musicians[i, state],    locality) &
+                    musicians[i, season] == season, mean(population)])
 
@@
 <<echo=F,results='hide'>>=
-musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
+#musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
 @
@@
 
+### Exclude playoff games.
 setkey(nba.lines, season, team, date)
-
-### Exclude playoff games.
 nba.lines <- nba.lines[, head(.SD, n=82), by=list(season,team)]
@@
 musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)),
-          by = team][order(average.nmusicians)][c(1:5, 25:29)]
+          by = team][order(average.nmusicians)][c(1:5, 25:27)]
 @
@@
      main = 'Continuous measure of party by Team Location: NBA',
-     ylab = 'Log Avg of Musicians and Sound Recording Studios',
+     ylab = 'Log sum of Musicians and Sound Recording Studios',
      xlab = '',
@@
 
+pdf('writing/next_day_opponent_on_tour.pdf', width = 14, height = 14)
+ggplot(tmp[cur.game == 'Current game away' &
+           lst.game == 'Last game away'], aes(x = location, y = last.game.loc, fill = log(N))) +
+    geom_tile() +
+    labs(x = 'Current Game Location', y = 'Last Game Location', 
+         title = 'Distribution of Next Day Opponent - NBA\n(sorted by distance to NYC)') +
+    theme_bw() +
+    theme(axis.text.x = element_text(angle = 315, vjust = 0.5),
+          plot.title = element_text(hjust=.5))
+dev.off()
+
 ##################################################
@@
 tmpm <- nba.lines[, list(log.nmusicians = mean(log(nmusicians+1), na.rm = T),
-                         log.population = mean(log(population),   na.rm = T)), 
+                         log.population = mean(log(population+1), na.rm = T)), 
                   by = team][order(log.nmusicians)] %>% na.omit
@@
      main = 'Why drinking establishments or population don\'t work as a proxy for nightlife',
-     ylab = 'Log Avg of Variable',
-     xlab = '', axes = F, pch = 1, ylim = c(0, max(tmp$log.population)))
-points(x = ticks, y = tmp$log.ndrinks, pch = 16, col = 'purple')
-points(x = ticks, y = tmp$log.population, pch = 18, col = 'orange')
+     ylab = 'Log sum of Variable',
+     xlab = '', axes = F, pch = 1, ylim = c(0, max(tmp$log.population)), type = 'l')
+points(x = ticks, y = tmp$log.ndrinks, pch = 16, col = 'purple', type = 'l')
+points(x = ticks, y = tmp$log.population, pch = 18, col = 'orange', type = 'l')
 axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>% 
@@
 axis(side = 2)
-legend(x = 21.5, y = 12.5, legend = c('Musicians and sound recording studios',
+legend(x = 1.5, y = 12.5, legend = c('Musicians and sound recording studios',
                                  'Drinking establishments and liquor stores',
                                  'Population'), pch = c(1, 16, 18), col = c('black', 'purple', 'orange'))
-text(x = 24.6, y = 9.5, labels = paste("Correlation (alcohol):       ", 
+text(x = 3.6, y = 9.5, labels = paste("Correlation (alcohol):       ", 
                                        cor(tmp$log.nmusicians, tmp$log.ndrinks) %>% round(., 3)))
-text(x = 24.6, y = 10, labels = paste("Correlation (population): ", 
+text(x = 3.6, y = 10, labels = paste("Correlation (population): ", 
                                        cor(tmp$log.nmusicians, tmp$log.population) %>% round(., 3)))
@@
 
-stargazer(mpa1, tpa, tpo, dep.var.labels = c('Points Admitted by Team', 'Team Points Admitted', 'Team Points Scored'),
-          covariate.labels = c('Discrete party indicator',
-                               'Continuous party measure',
+stargazer(m2, dep.var.labels = 'Demeanead Rebounds Per Minute',
+          covariate.labels = c('Continuous party measure',
                                'Lag change in possessions',
@@
 
-stargazer(m2, dep.var.labels = 'Demeanead Rebounds Per Minute',
-          covariate.labels = c('Continuous party measure',
+stargazer(mpa1, tpa, tpo, dep.var.labels = c('Points Admitted by Team', 'Team Points Admitted', 'Team Points Scored'),
+          covariate.labels = c('Discrete party indicator',
+                               'Continuous party measure',
                                'Lag change in possessions',
@@
              ndays.lgame + log(travel.dist+1) + weekend,
-         data = mlb.lines, family = 'binomial')
+         data = mlb.lines[season < 2017], family = 'binomial')
 summary(m3)
@@
           dep.var.labels = 'Probability of Winning')
-
-stargazer(m4, covariate.labels = c('Party placebo',
-                                   'Bookmaker\'s odds', 'Home-team effect',
-                                   'Number of rest days', 'Logged travel distance',
-                                   'Weekend', 'Constant'), 
-          dep.var.labels = 'Probability of Winning (MLB)')
 
@@
      main = 'Continuous measure of party by Team Location: MLB',
-     ylab = 'Log Avg of Musicians and Sound Recording Studios',
+     ylab = 'Log sum of Musicians and Sound Recording Studios',
      xlab = '',

### 2017-07-03 functions.r
Works with MSA's, and significant. Fixed several locations which weren't geo-coded correctly.

@@ -9,7 +9,10 @@ MyGeoCode <- function(teams, sport) {
     locs <- locs[, lapply(.SD, as.character), .SDcols = 1:ncol(locs)]
     ### We take care to fetch the MSA for each area (i.e. locality)
     ### Sometimes, we need to make manual adjustments. We do this for NBA and MLB.
-    if (sport == 'nba')  locs[is.na(locality), locality := 'New York']
+    if (sport == 'nba')  {
+        locs[is.na(locality), locality := 'New York']
+        locs[locality == 'El Segundo', locality := 'Los Angeles']
+    }
     locs <- data.table(team = teams, locality = locs$locality,
                        state = locs$administrative.area.level.1, lon = locs$lon, lat = locs$lat)
     if (sport == 'mlb') {
@@ -22,8 +25,9 @@ MyGeoCode <- function(teams, sport) {
 ### Function: Clean BLS
 ### -------------------
 ### Given a string describing what kinds of business patterns to look for, a labeling suffix,
-###   a sport, and an aggregation function, this functional subsets each year of BLS
+###   a sport, and an aggregation function, this functional subsets each year of raw BLS
 ###   data such that only observations matching the party.regex are retained.
+###   The data are then aggregated using the aggregation function FUN.
 ###   Results are saved to disk according to a labeling which uses suffix and sport.
 CleanBLS <- function(party.regex, suffix, sport, FUN, years = 2010:2016, RESCRAPE = T) {
     ### Subset each year of BLS data to only observations matching regular expression.
@@ -46,6 +50,7 @@ CleanBLS <- function(party.regex, suffix, sport, FUN, years = 2010:2016, RESCRAP
     load(file = paste0('tmp_data/', sport, '_team_locations.RData'))
     bls[, area.title := gsub(' msa$', '', area.title, ignore.case = T)]
     states <- data.table(state = state.name, state.abb = state.abb)
+    states <- rbind(states, data.frame(state = 'District of Columbia', state.abb = 'DC'))
     locs <- locs[states, on = 'state', nomatch = 0]
     ### We can't simply merge. We have to look for the county within the MSA.
     dt <- lapply(1:nrow(locs), function(i) { 
@@ -55,7 +60,7 @@ CleanBLS <- function(party.regex, suffix, sport, FUN, years = 2010:2016, RESCRAP
     }) %>% rbindlist
     ## dt <- merge(locs, bls, by.x = 'msa', by.y = 'area.title', all.x = T)
[…]

### 2017-07-03 functions.r
Bookmarked legacy code.

@@ -129,7 +129,11 @@ Bet <- function(our.prediction, actual.outcome, odds) {
         return(-100)
 }
 
+##################################################
+### Legacy code.
+##################################################
 
+### Fetch is used as a helper function within CollectData (below).
 Fetch <- function(team, beg, end, data, year) {
     datum <- data[beg:end]
     ### We have to be careful for data entry errors. Ex: missing delimiter.
@@ -141,34 +145,25 @@ Fetch <- function(team, beg, end, data, year) {
         data.table(.) %>% cbind(team)
 }
 
-
+### CollectData was used in our preliminary analysis when we had data from gold-lines.com
 collectData <- function(datafile) {
     data <- readLines(datafile) # 'raw_data/2011.txt')
-
     ### Remove blank lines.
     blanks <- grep("^[[:blank:]]*$", data)
     data <- data[-blanks]
-
     ### Find team names, aka starting indices. 
     beg <- grep("^\\(SUR", data) - 1     # Look one row before team-summary for team-name.
     end <- c(beg[-1] - 1, length(data))  # Find limits of data for each team.
     teams <- data[beg]
-
     ### This data frame describes where data start-and-stop for each team.
     teams <- data.frame(team = teams, beg = beg+2, end = end)
-
     data <- mapply(Fetch, team = teams$team, beg = teams$beg, end = teams$end, year = '2011',
                    MoreArgs = list(data=data), SIMPLIFY = F)
-
     data <- rbindlist(data)
     data[['season']] = gsub('raw_data/([0-9_]{9})\\.txt', '\\1', datafile)
-
     setnames(data, c('date', 'opponent', 'outcome', 'spread', 'score',
                      'location', 'ou', 'team', 'season'))
     setcolorder(data, c('season', 'date', 'team', 'opponent', 'outcome', 'spread', 'score',
                         'location', 'ou'))
     return(data)
 }
-
-## getFullTeamName <- function(needle, haystack)
-##     grep(needle, haystack, value = T)

### 2017-07-03 scrape_bls.r
Noted file dependencies.

@@ -8,17 +8,19 @@
 ###
 ### Date: April 2017
 ###
-### Inputs: 
+### Inputs: 'https://data.bls.gov/cew/data/files/[year]/csv/[year]_qtrly_by_area.zip'
 ###
-### Dependencies: 
+### Output: 'raw_data/bls/'
+###
+### Dependencies: magrittr, parallel, RCurl, XML 
 ###
 ################################################################################
 ################################################################################
 
-require(XML)
 require(magrittr)
-require(RCurl)
 require(parallel)
+require(RCurl)
+require(XML)
 
 cl <- makeCluster(detectCores())
 

### 2017-07-03 scrape_covers.r
Noted file dependencies.

@@ -8,9 +8,11 @@
 ###
 ### Date: March 2017
 ###
-### Inputs: NA
+### Inputs: 'tmp_data/game_days.RData'
 ###
-### Dependencies: data.table, XML
+### Output: 'raw_data/covers/[date].csv'
+###
+### Dependencies: data.table, magrittr, parallel, XML
 ###
 ################################################################################
 ################################################################################

### 2017-07-03 scrape_covers_mlb.r
Noted file dependencies.

@@ -8,9 +8,11 @@
 ###
 ### Date: April 2017
 ###
-### Inputs: 
+### Inputs: 'tmp_data/game_days_mlb.RData'
 ###
-### Dependencies: 
+### Output: 'tmp_data/covers_mlb/[date].csv'
+###
+### Dependencies: data.table, magrittr, parallel, XML
 ###
 ################################################################################
 ################################################################################
@@ -25,7 +27,7 @@ require(parallel)
 require(XML)
 load(file = 'tmp_data/game_days_mlb.RData')
 
-game.days <- Filter(function(x) x < Sys.Date() & x > as.Date('2011-01-01'),
+game.days <- Filter(function(x) x > as.Date('2011-01-01'),
                     x = game.days) %>% as.character
 
 ##############################

### 2017-07-03 scrape_espn.r
Noted file dependencies.

@@ -8,9 +8,11 @@
 ###
 ### Date: April 2017
 ###
-### Inputs: 
+### Inputs: 'tmp_data/game_days.RData'
 ###
-### Dependencies: 
+### Output: 'raw_data/espn/[gameID].csv'
+###
+### Dependencies: data.table, magrittr, parallel, RCurl, XML
 ###
 ################################################################################
 ################################################################################
@@ -19,7 +21,6 @@ require(data.table)
 require(magrittr)
 require(RCurl)
 require(XML)
-
 require(parallel)
 cl <- makeCluster(detectCores())
 
@@ -38,6 +39,7 @@ extractGameIDs <- function(date) {
 
 ### Obtain a listing of game ID's.
 gameIDs <- parSapplyLB(cl, game.days, extractGameIDs, simplify = F)
+names(gameIDs) <- game.days
 save(gameIDs, file = 'tmp_data/espn_gameIDs.RData')
 
 ### For each game ID, we scrape the player data.

### 2017-07-03 scrape_games.r
Noted file dependencies.

@@ -8,7 +8,9 @@
 ###
 ### Date: March 2017
 ###
-### Inputs: NA
+### Inputs: 'http://www.basketball-reference.com/leagues/NBA_[year]_games-[month].html'
+###
+### Output: 'tmp_data/game_days.RData'
 ###
 ### Dependencies: data.table, magrittr, XML
 ###
@@ -57,3 +59,15 @@ game.days <- unique(data$date)
 game.days <- as.Date(game.days, format = '%A, %b %d, %Y')
 
 save(game.days, file = 'tmp_data/game_days.RData')
+
+
+
+
+
+
+
+
+
+
+
+

### 2017-07-03 scrape_population.r
Now scraping MSA's instead of counties.

@@ -21,36 +21,30 @@ require(RCurl)
 require(XML)
 
 ### Download population estimates by county-year.
-pop <- fread('https://www2.census.gov/programs-surveys/popest/datasets/2010-2016/counties/totals/co-est2016-alldata.csv',
-             select = c('STNAME', 'CTYNAME', paste0('POPESTIMATE', 2010:2016)))
-setnames(pop, tolower(names(pop)))
+## pop <- fread('https://www2.census.gov/programs-surveys/popest/datasets/2010-2016/counties/totals/co-est2016-alldata.csv',
+##              select = c('STNAME', 'CTYNAME', paste0('POPESTIMATE', 2010:2016)))
+## setnames(pop, tolower(names(pop)))
 
 ### Don't forget Washington D.C., which is technically a metro area.
-dc <- fread('https://www2.census.gov/programs-surveys/popest/datasets/2010-2016/metro/totals/csa-est2016-alldata.csv',
+pop <- fread('https://www2.census.gov/programs-surveys/popest/datasets/2010-2016/metro/totals/csa-est2016-alldata.csv',
             select = c('NAME', paste0('POPESTIMATE', 2010:2016)))
 Sys.setlocale('LC_ALL','C') 
-dc <- dc[grep("district", NAME, ignore.case = T)]
-dc[, stname := 'Washington DC']
-setnames(dc, c('ctyname', paste0('popestimate', 2010:2016), 'stname'))
-dc[, ctyname := 'District of Columbia']
-pop <- rbind(pop, dc)
+## dc <- dc[grep("district", NAME, ignore.case = T)]
+## dc[, stname := 'Washington DC']
+setnames(pop, c('locality', paste0('popestimate', 2010:2016)))
+## dc[, locality := 'District of Columbia']
+## pop <- rbind(pop, dc)
 
 ### Reshape our data.
-pop <- melt(pop, id.vars = c('stname', 'ctyname'), variable.name = 'year')
+pop <- melt(pop, id.vars = 'locality', variable.name = 'year')
 pop[, year := gsub('popestimate', '', year) %>% as.integer]
 setnames(pop, 'value', 'population')
-
-setnames(pop, c('stname', 'ctyname'), c('state', 'county'))
 
 ### We use lagged data.
 setnames(pop, 'year', 'season')
 pop[, season := season + 1]
-pop <- pop[state != county]
-pop[, county := paste(county, state, sep = ', ')]
-pop[, state := NULL]
[…]

### 2017-07-03 writing/draft.tex
Updated based on MSA analysis.

@@
 We also found that athletes tend to party with other A-list celebrities,
-particularly musicians.
-This motivated a continuous measure of nightlife, based on a simple count of the number of musicians in the county in which the stadium resides: if the last game was played at most 24-hours ago, use the log of the total number
+particularly musicians. This motivated a continuous measure of nightlife, based on a simple count of the number of musicians in the metropolitan statistical area (MSA) 
+in which the stadium resides: if the last game was played at most 24-hours ago, use the log of the total number
 of musical groups, sound recording studios, and musical publishers in county of the 
@@
 \paragraph{How an unknown spread variable may (not) affect our inference}
-For example, if book-keepers
+For example, if bookkeepers
 are already aware of the effects of night-life on performance, it could
@@
 a true effect. Alternatively,
-if the house book-keeper systematically mis-values
+if the house bookkeeper systematically mis-values
 teams which are more likely to be encountered after playing a city with active nightlife,
@@
 \subsection{BLS data} We also obtained data from
-\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the county-quarter level. We obtain this data for each county corresponding
+\href{https://www.bls.gov/data/}{Bureau of Labor Statistics} which records the number of establishments by business type at the MSA-quarter level.
+We obtain this data for each metropolitan statistical area corresponding
 to an NBA or MLB team from years 2010-2016. 
@@
 are for a particular county-year: we simply calculate the total number of establishments across all three music categories listed above.
-We caveat that since the Toronto Raptors are located in Canada, we don't have BLS data for this team.\footnote{We don't expect this to be a party city.}
+We caveat that since the Toronto Raptors and Toronto Blue Jays are located in Canada,
+we don't have BLS data for these teams.\footnote{We don't expect Toronto, Canada to be a party city.}
 
@@
 have a statistically significant effect on likelihood of meeting the spread.
-In laymens terms, teams under perform book-maker's expectations after playing in cities 
+In laymens terms, teams under perform bookmaker's expectations after playing in cities 
 with active nightlife.
 
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   Party continuous &  & $-$0.199$^{**}$ \\    &  & (0.100) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Logged travel distance & $-$0.017 & $-$0.014 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.146 & $-$0.112 \\    & (0.231) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.003 \\    & (0.043) & (0.043) \\    & & \\   Logged travel distance * east-west & 0.022 & 0.017 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.034 \\    & (0.378) & (0.378) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.675 & $-$6,805.106 \\  Akaike Inf.
+% m0, m1
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   Party continuous &  & $-$0.181$^{**}$ \\    &  & (0.091) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Logged travel distance & $-$0.017 & $-$0.014 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.145 & $-$0.113 \\    & (0.232) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   Logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.028 \\    & (0.378) & (0.378) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.677 & $-$6,805.100 \\  Akaike Inf.
 Crit.
-& 13,619.350 & 13,628.210 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 13,619.350 & 13,628.200 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 the likelihood of meeting the spread, we predict the probability
-a team would win as both a function of the book-makers'
+a team would win as both a function of the bookmakers'
 odds of winning and our party variable. 
@@
 
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.142$^{*}$ &  \\    & (0.083) &  \\    & & \\   Nightlife (no weekend interaction) &  & 0.004 \\    &  & (0.107) \\    & & \\   Bookmaker's odds & 2.600$^{***}$ & 2.604$^{***}$ \\    & (0.162) & (0.163) \\    & & \\   Home-team effect & 0.024 & 0.052 \\    & (0.035) & (0.072) \\    & & \\   Number of rest days & 0.015 & 0.015 \\    & (0.019) & (0.020) \\    & & \\   Logged travel distance & $-$0.001 & $-$0.001 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.042 & 0.0004 \\    & (0.037) & (0.030) \\    & & \\   Constant & $-$1.329$^{***}$ & $-$1.348$^{***}$ \\    & (0.080) & (0.102) \\    & & \\  \hline \\[-1.8ex]  Observations & 25,215 & 24,781 \\  Log Likelihood & $-$17,274.540 & $-$16,977.790 \\  Akaike Inf.
+% m3, m4
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.126$^{*}$ &  \\    & (0.074) &  \\    & & \\   Nightlife (no weekend interaction) &  & $-$0.016 \\    &  & (0.110) \\    & & \\   Bookmaker's odds & 2.638$^{***}$ & 2.583$^{***}$ \\    & (0.163) & (0.165) \\    & & \\   Home-team effect & 0.008 & 0.034 \\    & (0.035) & (0.083) \\    & & \\   Number of rest days & 0.019 & 0.018 \\    & (0.020) & (0.020) \\    & & \\   Logged travel distance & $-$0.002 & $-$0.003 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.044 & 0.003 \\    & (0.038) & (0.030) \\    & & \\   Constant & $-$1.344$^{***}$ & $-$1.324$^{***}$ \\    & (0.081) & (0.113) \\    & & \\  \hline \\[-1.8ex]  Observations & 24,734 & 24,033 \\  Log Likelihood & $-$16,946.080 & $-$16,470.850 \\  Akaike Inf.
 Crit.
-& 34,563.090 & 33,969.580 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 33,906.160 & 32,955.700 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 Angeles or New York, we cannot reject the null hypothesis that player performance is unaffected
-by visitation to cities with active nightlife. The same is true in the MLB.
+by visitation to cities with active nightlife. The analagous placebo test
+for MLB is not relevant, since almost all games are played back to back.
 
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   party placebo & 0.072 \\    & (0.082) \\    & \\   lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   east-west travel direction & $-$0.100 \\    & (0.232) \\    & \\   number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   time of game during day & 0.001 \\    & (0.013) \\    & \\   home team effect & $-$0.009 \\    & (0.043) \\    & \\   logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.070 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.701 \\  Akaike Inf.
+% mp
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   party placebo & 0.072 \\    & (0.082) \\    & \\   lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   east-west travel direction & $-$0.099 \\    & (0.232) \\    & \\   number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   time of game during day & 0.001 \\    & (0.013) \\    & \\   home team effect & $-$0.009 \\    & (0.043) \\    & \\   logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.070 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.702 \\  Akaike Inf.
 Crit.
 & 13,631.400 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Probability of Winning (MLB) \\  \hline \\[-1.8ex]   Party placebo & 0.004 \\    & (0.107) \\    & \\   Bookmaker's odds & 2.604$^{***}$ \\    & (0.163) \\    & \\   Home-team effect & 0.052 \\    & (0.072) \\    & \\   Number of rest days & 0.015 \\    & (0.020) \\    & \\   Logged travel distance & $-$0.001 \\    & (0.005) \\    & \\   Weekend & 0.0004 \\    & (0.030) \\    & \\   Constant & $-$1.348$^{***}$ \\    & (0.102) \\    & \\  \hline \\[-1.8ex]  Observations & 24,781 \\  Log Likelihood & $-$16,977.790 \\  Akaike Inf.
-Crit.
-& 33,969.580 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
 \subsection{Possible confounders or potential issues}
@@
 
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.001$^{*}$ \\    & (0.0005) \\    & \\   Lag change in possessions & 0.00003 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00003 \\    & (0.00002) \\    & \\   Constant & $-$0.008$^{*}$ \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 43,547 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
-Error & 0.104 (df = 43542) \\  F Statistic & 2.981$^{**}$ (df = 4; 43542) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+% m2
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.003$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00003 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00003 \\    & (0.00002) \\    & \\   Constant & $-$0.008$^{*}$ \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 43,547 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
+Error & 0.104 (df = 43542) \\  F Statistic & 2.945$^{**}$ (df = 4; 43542) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 
+% mpa1, tpa, tpo
 \hspace{-100pt}
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.828$^{***}$ &  &  \\    & (0.699) &  &  \\    & & & \\   Continuous party measure &  & 0.475$^{***}$ & 0.009 \\    &  & (0.139) & (0.136) \\    & & & \\   Lag change in possessions &  & 0.267$^{***}$ & 0.285$^{***}$ \\    &  & (0.016) & (0.016) \\    & & & \\   Logged travel distance &  & $-$0.430$^{**}$ & 0.142 \\    &  & (0.191) & (0.187) \\    & & & \\   Number hours since last game &  & 0.012$^{*}$ & 0.017$^{**}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Constant & $-$0.047 & 70.581$^{***}$ & 64.886$^{***}$ \\    & (0.090) & (2.267) & (2.220) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,658 & 6,426 & 6,426 \\  R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Adjusted R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Residual Std.
-Error & 11.159 (df = 15656) & 11.937 (df = 6421) & 11.688 (df = 6421) \\  F Statistic & 16.357$^{***}$ (df = 1; 15656) & 75.841$^{***}$ (df = 4; 6421) & 86.788$^{***}$ (df = 4; 6421) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.828$^{***}$ &  &  \\    & (0.699) &  &  \\    & & & \\   Continuous party measure &  & 1.667$^{***}$ & 0.080 \\    &  & (0.587) & (0.575) \\    & & & \\   Lag change in possessions &  & 0.267$^{***}$ & 0.285$^{***}$ \\    &  & (0.016) & (0.016) \\    & & & \\   Logged travel distance &  & $-$0.447$^{**}$ & 0.143 \\    &  & (0.191) & (0.187) \\    & & & \\   Number hours since last game &  & 0.011 & 0.017$^{**}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Constant & $-$0.047 & 70.710$^{***}$ & 64.859$^{***}$ \\    & (0.090) & (2.271) & (2.223) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,658 & 6,426 & 6,426 \\  R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Adjusted R$^{2}$ & 0.001 & 0.044 & 0.051 \\  Residual Std.
+Error & 11.159 (df = 15656) & 11.941 (df = 6421) & 11.688 (df = 6421) \\  F Statistic & 16.357$^{***}$ (df = 1; 15656) & 74.877$^{***}$ (df = 4; 6421) & 86.792$^{***}$ (df = 4; 6421) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 

### 2017-07-03 poster/poster.tex
Centered 'model' block.
Fixed formatting.

@@
 \documentclass[24pt, a0paper, portrait]{tikzposter}
- \usepackage{algorithm2e} \usepackage{amsmath} \usepackage{amsthm} \usepackage{amsfonts} \usepackage{bbm} \usepackage{color,soul} \usepackage{framed} \usepackage{hyperref} \usepackage{mathtools} \usepackage[utf8]{inputenc}
+
+\usepackage{algorithm2e} 
+\usepackage{amsmath} 
+\usepackage{amsthm} \usepackage{amsfonts} 
+\usepackage{bbm} 
+\usepackage{color,soul} \usepackage{framed} 
+\usepackage{hyperref} 
+\usepackage{mathtools} \usepackage[utf8]{inputenc} 
 \usepackage{subcaption}
- \newtheorem{theorem}{Theorem}[section] \newtheorem{lemma}[theorem]{Lemma} \newtheorem{proposition}[theorem]{Proposition} \newtheorem{corollary}[theorem]{Corollary}  \DeclareMathOperator*{\argmin}{arg\,min} \DeclareMathOperator*{\argmax}{arg\,max} \newcommand{\D}{\mathrm{d}} \SetKwInput{KwInput}{Input} \SetKwInput{KwOutput}{Output}   \title{The effects of night-life on subsequent day sporting performance} \author{Eric Lax, Andreas Santucci} \date{May 19th, 2017} \institute{Institute for Computational and Mathematical Engineering}   \usetheme{Board}   \begin{document}   \maketitle
+
+\newtheorem{theorem}{Theorem}[section] 
+\newtheorem{lemma}[theorem]{Lemma} \newtheorem{proposition}[theorem]{Proposition} \newtheorem{corollary}[theorem]{Corollary}  \DeclareMathOperator*{\argmin}{arg\,min} \DeclareMathOperator*{\argmax}{arg\,max}
+\newcommand{\D}{\mathrm{d}} 
+\SetKwInput{KwInput}{Input} 
+\SetKwInput{KwOutput}{Output}
+
+\title{The effects of night-life on subsequent day sporting performance} 
+\author{Eric Lax, Andreas Santucci} 
+\date{May 19th, 2017} 
+\institute{Institute for Computational and Mathematical Engineering}   
+\usetheme{Board}
+
+\begin{document}
+
+\maketitle
 
@@
 
+\begin{columns}
+
+\column{0.2}
+\column{0.6}
+
 \block{Model}{
@@
 }
+\column{0.2}
+\end{columns}
 

### 2017-07-03 cardinal_analytx_presentation/nightlife_and_sports.tex
Added figure showing why drinking/population don't work, and additional details.

@@
   \frametitle{Who cares?}
-  \begin{block}{\$400 billion spent on sports betting per year in the US.}
+  \begin{block}{\$400 billion spent on sports betting per year in the US.\footnote{NBA Commissioner Adam Silver}}
+
     Apparently, a lot of people.
@@
 
+\begin{frame}   \frametitle{Is the problem hard?}
+  Yes, in general finding arbitrage opportunities is hard.
+
+  Further, it's not obvious that we will find an effect. \end{frame}
+
 \begin{frame}   
@@
   \begin{block}{We can't observe whether players party.}
-    Instead, we consider a quasi-experiment.
+    Instead, we consider a quasi-experiment. Motivating questions:
     \begin{itemize}       \item Do teams play worse the day after visiting a ``party city''?
-      \item How do we even determine whether a city has nightlife?     \end{itemize}
+      \item How do we even determine whether a city has nightlife?    
+      \item What does it mean to ``play worse''? \end{itemize}
   \end{block}
@@
 
+\begin{frame}
+  \centering   \includegraphics[scale=0.55]{../writing/meet_the_spread} \end{frame}
+
 \begin{frame}   \frametitle{Quasi Experiment}
-  \begin{block}{Consider games played back-to-back within 24 hours.}     If odds of beating the spread uncorrelated with next day opponent 
-    $\longrightarrow$ identification.
+  \begin{block}{Consider games played back-to-back within 24 hours.}     
+    If next day opponent is uncorrelated with visitation to a party city $\leadsto$ identification.
 
-  (Under an efficient market, this should be the case)    \end{block}  
+    Weaker, still: if odds of beating the spread uncorrelated \\
+    $\leadsto$ identification.\footnote{Under an efficient market, the latter assumption should hold.}    \end{block}  
 \end{frame}
@@
 \begin{frame}   \frametitle{Bureau of Labor Statistics Data: 2010-2016}
-  We look toward BLS data which describes the number of sound recording
-  studios, musical groups, and musical publishers in each metropolitan statistical area
+  We look toward BLS data which describes the number of business establishments 
+  (of a very particular type) in each metropolitan statistical area
   for each quarter of each year.
@@
 
-  These two locations are excluded from our analysis. \end{frame}
+  These two teams are excluded from our analysis.\footnote{We don't expect Toronto to be a huge party city.} \end{frame}
 
@@
   We add features which describe
-  \begin{itemize}     \item Number of days since the last game (i.e. rest-time),
+  \begin{itemize}     
+    \item Number of days since the last game (i.e. rest-time),
+    \item Lagged changes in possession,
     \item Lagged game location, and
-    \item Travel Distance.   \end{itemize}
+    \item Travel Distance.   
+  \end{itemize}
   We seek to construct a model which accounts for travel-fatigue and jet-lag.
@@
 \begin{frame}   \frametitle{Future Research (General)}
-  \begin{itemize}     \item It's possible there are prolonged effects; recall our magic 24 hour number.
+  \begin{itemize}     \item Possibly prolonged effects; recall our magic 24 hour number.
     \item European soccer leagues.   \end{itemize} \end{frame}
@@
 
+\begin{frame}   \frametitle{Why \# Drinking Establishments or Population won't work}
+  \includegraphics[scale=0.3]{../writing/why_drinks_or_population_dont_work} \end{frame}
+
 \end{document}

### 2017-07-03 cardinal_analytx_presentation/nightlife_and_sports.tex
Added citations to intro. Added plot showing it\'s a coin toss whether we win as a function of the odds.

@@
   \begin{block}{Do pro athletes party to the detriment of their performance?}
+    Anecdotally, yes.\footnote{Cestone ('17), Price ('15), Concepcion ('15), Miller ('15), Campbell ('15).}
+
     It's not unreasonable to think that what happens off the court affects outcome, but!
@@
   \frametitle{Who cares?}
-  \begin{block}{\$400 billion spent on sports betting per year in the US.\footnote{NBA Commissioner Adam Silver}}
+  \begin{block}{\$400 billion spent on sports betting per year in the US.\footnote{NBA Commissioner Adam Silver, 2014.}}
 
@@
     Knowing the causal relation between partying and next-day performance might inform
-    \begin{itemize}       \item How coaches set team-rules, 
-      \item How travel schedules are determined.     \end{itemize}
+    \begin{itemize}       
+      \item How coaches set team-rules, 
+      \item How travel schedules are determined.
+      \item Which players to bench, and when.
+      \end{itemize}
   \end{block}
@@
 
-  Further, it's not obvious that we will find an effect. \end{frame}
+  Further, it's not obvious that we will find an effect.
+
+  \begin{block}{Previous research}
+    Players under-perform proportion of shots made on Sundays.\footnote{Chase ('15): 1,092 NBA players under-perform relative to their own season-long average out of 2,035 who took at least one shot on Sunday.}
+
+    Teams under-perform bookmakers expectations on Sunday afternoons when on the road.\footnote{Ezekowitz, ('14): Only 129 out of 284 games did the away team meet the spread, using 2008-14 data.}   \end{block}
+
+\end{frame}
 
@@
 \begin{frame}
-  \frametitle{How can we measure unexpected changes in performance?}   \begin{block}{Bookmaker spreads}     The expected score differential between two teams.
+  \frametitle{How can we measure unexpected changes in performance?}     \begin{block}{Bookmaker spreads}     
+    The expected score differential between two teams.
 
@@
 
-    A team is expected to meet the spread $\frac{1}{2}$ the time.   \end{block} \end{frame}
+    A team is expected to meet the spread $\frac{1}{2}$ the time.\footnote{Dubner and Levitt, 2006}
+  \end{block}
+  \begin{block}{Bookmaker money-lines}     The odds that a team will beat their opponent.
+
+    After conditioning on odds, successful prediction is like a coin-flip.   \end{block} \end{frame}
 

### 2017-07-03 cardinal_analytx_presentation/nightlife_and_sports.tex
Added previous research section alongside citations.

@@
 
-  \begin{block}{Previous research}
-    Players under-perform proportion of shots made on Sundays.\footnote{Chase ('15): 1,092 NBA players under-perform relative to their own season-long average out of 2,035 who took at least one shot on Sunday.}
+\end{frame}
 
-    Teams under-perform bookmakers expectations on Sunday afternoons when on the road.\footnote{Ezekowitz, ('14): Only 129 out of 284 games did the away team meet the spread, using 2008-14 data.}   \end{block}
+\begin{frame}
+  \frametitle{Previous Research - Player shooting average}     Players under-perform proportion of shots made on Sundays.\footnote{Chase, ('15).}
+    \begin{itemize}       \item 1,092 of 2,035 NBA players who took $\geq 1$ shot on Sundays under-perform relative to their own season-long average; $p=0.001$.
+      \item Effect size is not considered.
+      \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, should that be considered?}     \end{itemize}
+\end{frame}
 
+\begin{frame}   \frametitle{Previous Research - Team's ability to meet spread}
+  Teams under-perform bookmakers expectations on Sunday afternoons when on the road.\footnote{Ezekowitz, ('14).}
+    \begin{itemize}       
+    \item Only 129 out of 284 games did the away team meet the spread, 
+      using 2008-14 data; $p=0.1089$.
+    \item Not quite significant at the 90\% confidence level.     
+    \end{itemize}
   \end{frame}
@@
 \begin{frame}
-  \frametitle{How can we measure unexpected changes in performance?}     \begin{block}{Bookmaker spreads}     
+  \frametitle{How can we measure unexpected changes in performance?}     \begin{block}{Bookmaker spreads (NBA)}     
     The expected score differential between two teams.
@@
   \end{block}
-  \begin{block}{Bookmaker money-lines}     The odds that a team will beat their opponent.
+  \begin{block}{Bookmaker money-lines (MLB)}     The odds that a team will beat their opponent.
 
@@
     Weaker, still: if odds of beating the spread uncorrelated \\
-    $\leadsto$ identification.\footnote{Under an efficient market, the latter assumption should hold.}    \end{block}  
+    $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}    \end{block}  
 \end{frame}

### 2017-07-03 cardinal_analytx_presentation/nightlife_and_sports.tex
Citation for jetlag.

@@
 \begin{frame}   \frametitle{Accounting for Jetlag}
+  Prolonged air travel negatively effects sporting performance.\footnote{Lee and Galvez ('12).}
+
   In order to account for jet-lag, we need to be mindful of the direction of travel. We calculate
@@
   \end{itemize}
-  Finally, to account for jetlag we interact this east-west measure with travel distance. \end{frame}
+  Finally, to account for jetlag we interact this east-west measure with travel distance.\footnote{Lallensack ('17)} \end{frame}
 
@@
 \begin{frame}   \frametitle{Analysis: rebounds}
-  \begin{itemize}     \item For each player and each game, we calculate the number of rebounds obtained per-minute of gamep-lay
+  \begin{itemize}     \item For each player and each game, we calculate the number of rebounds obtained per-minute of game-play
     \item From each observation, we subtract the player's average number of rebounds per minute taken across the entire season.

### 2017-07-04 functions.r
Removed in favor of a package, thank you BMaples!

@@ -1,169 +1,0 @@
-
-### Function: MyGeoCode
-### -------------------
-### Given a listing of teams for a particular sport, this function
-###   uses `ggmap::geocode` to determine a longitude/latitude for each location.
-MyGeoCode <- function(teams, sport) {
-    locs <- lapply(teams, geocode, output = 'more') %>% rbindlist(., fill = T)
-    setnames(locs, gsub('_', '.', names(locs)))
-    locs <- locs[, lapply(.SD, as.character), .SDcols = 1:ncol(locs)]
-    ### We take care to fetch the MSA for each area (i.e. locality)
-    ### Sometimes, we need to make manual adjustments. We do this for NBA and MLB.
-    if (sport == 'nba')  {
-        locs[is.na(locality), locality := 'New York']
-        locs[locality == 'El Segundo', locality := 'Los Angeles']
-    }
-    locs <- data.table(team = teams, locality = locs$locality,
-                       state = locs$administrative.area.level.1, lon = locs$lon, lat = locs$lat)
-    if (sport == 'mlb') {
-        locs[is.na(locality) & state == 'Florida', `:=`(locality = 'Atlanta', state = 'Georgia')]
-        locs[is.na(locality) & state == 'New York', locality := 'New York']
-    }
-    save(locs, file = paste0('tmp_data/', sport, '_team_locations.RData'))
-}
-
-### Function: Clean BLS
-### -------------------
-### Given a string describing what kinds of business patterns to look for, a labeling suffix,
-###   a sport, and an aggregation function, this functional subsets each year of raw BLS
-###   data such that only observations matching the party.regex are retained.
-###   The data are then aggregated using the aggregation function FUN.
-###   Results are saved to disk according to a labeling which uses suffix and sport.
-CleanBLS <- function(party.regex, suffix, sport, FUN, years = 2010:2016, RESCRAPE = T) {
-    ### Subset each year of BLS data to only observations matching regular expression.
-    parLapplyLB(cl, years, SubsetBLS, party.regex = party.regex, sport = sport, suffix = suffix)
-    files <- list.files(path = 'tmp_data/bls',
-                        pattern = paste0(sport, '_', suffix, '\\.csv$'), full.names = T)
-    bls <- lapply(files, fread) %>% rbindlist
-    ### Fetch a listing of team-names from our points-spread/money-lines data.
-    if (sport == 'nba') {
[…]

### 2017-07-04 hangover/R/Bet.r
Init commit.

@@ -1,0 +1,22 @@
+#' Place a bet on yet-unseen data.
+#'
+#' This function compares our prediction with the actual outcome of the game.
+#' We allow the house bookmaker a $10 cut for each bet placed, regardless of outcome.
+#' @param our.prediction A fitted label (0 or 1).
+#' @param actual.outcome A true label (0 or 1).
+#' @param odds Bookkeeper's odds of success, i.e. of actual outcome == 1.
+#' @keywords Bet
+#' @export
+#' @examples
+#' Bet(1, 0, 0.55)
+Bet <- function(our.prediction, actual.outcome, odds) {
+    actual.outcome = as.integer(actual.outcome)
+    if (is.na(our.prediction)) return(NA_integer_)
+    if (our.prediction == actual.outcome) {
+        if (our.prediction > 0)
+            return(100 / odds - 100 - 10)
+        else if (our.prediction == 0)
+            return(100 / (1-odds) - 100 - 10)
+    } else if (our.prediction != actual.outcome)
+        return(-100)
+}

### 2017-07-04 hangover/R/CleanBLS.r

@@ -1,0 +1,55 @@
+#' A function to clean BLS data.
+#'
+#' Given a string describing what kinds of business patterns to look for, a labeling suffix,
+#'  a sport, and an aggregation function, this functional subsets each year of raw BLS
+#'   data such that only observations matching the party.regex are retained.
+#'   The data are then aggregated using the aggregation function FUN.
+#'   Results are saved to disk according to a labeling which uses suffix and sport.
+#' @param party.regex A character vector of patterns to search for in our BLS data.
+#' @param suffix A labeling suffix to be used when saving the output file.
+#' @param sport A string, one of 'nba' or 'mlb'
+#' @param FUN An aggregation function to be used in collapsing BLS data by locality.
+#' @param years A vector of integers describing which years to collect data for.
+#' @param RESCRAPE A boolean indicating whether team-locations should be rescraped.
+#' @keywords BLS, clean
+#' @export
+#' @examples
+#' CleanBLS(c('recording studio', 'musical group'), 'DJ_proxy', 'nba', sum)
+CleanBLS <- function(party.regex, suffix, sport, FUN, years = 2010:2016, RESCRAPE = T) {
+    ### Subset each year of BLS data to only observations matching regular expression.
+    parLapplyLB(cl, years, SubsetBLS, party.regex = party.regex, sport = sport, suffix = suffix)
+    files <- list.files(path = 'tmp_data/bls',
+                        pattern = paste0(sport, '_', suffix, '\\.csv$'), full.names = T)
+    bls <- lapply(files, fread) %>% rbindlist
+    ### Fetch a listing of team-names from our points-spread/money-lines data.
+    if (sport == 'nba') {
+        load(file = 'tmp_data/covers_lines.RData')   # Load lines data to get team-names (overkill)
+        lines[, pct := NULL]
+        teams <- unique(lines$team)
+    } else if (sport == 'mlb') {
+        abbrs <- fread(file = 'mlb_abbrs.csv')
+        setnames(abbrs, c('abbr', 'full', 'nick'))
+        teams <- abbrs$full
+    }
+    ### Geocode the location of each team within the sport, and add this to our BLS data.
+    if (RESCRAPE) MyGeoCode(teams, sport)
+    load(file = paste0('tmp_data/', sport, '_team_locations.RData'))
+    bls[, area.title := gsub(' msa$', '', area.title, ignore.case = T)]
[…]

### 2017-07-04 hangover/R/MyGeoCode.r

@@ -1,0 +1,28 @@
+#' A function to geocode team-locations.
+#'
+#' This function takes as input a listing of team-names for which we
+#' retrieve location data.
+#' @param teams A vector of strings, each a team-name.
+#' @param sport A string; one of 'nba' or 'mlb'.
+#' @keywords geocode, ggmap
+#' @export
+#' @examples
+#' myGeoCode('Los Angeles Lakers', 'nba')
+MyGeoCode <- function(teams, sport) {
+    locs <- lapply(teams, geocode, output = 'more') %>% rbindlist(., fill = T)
+    setnames(locs, gsub('_', '.', names(locs)))
+    locs <- locs[, lapply(.SD, as.character), .SDcols = 1:ncol(locs)]
+    ### We take care to fetch the MSA for each area (i.e. locality)
+    ### Sometimes, we need to make manual adjustments. We do this for NBA and MLB.
+    if (sport == 'nba')  {
+        locs[is.na(locality), locality := 'New York']
+        locs[locality == 'El Segundo', locality := 'Los Angeles']
+    }
+    locs <- data.table(team = teams, locality = locs$locality,
+                       state = locs$administrative.area.level.1, lon = locs$lon, lat = locs$lat)
+    if (sport == 'mlb') {
+        locs[is.na(locality) & state == 'Florida', `:=`(locality = 'Atlanta', state = 'Georgia')]
+        locs[is.na(locality) & state == 'New York', locality := 'New York']
+    }
+    save(locs, file = paste0('tmp_data/', sport, '_team_locations.RData'))
+}

### 2017-07-04 hangover/R/Predict.r

@@ -1,0 +1,20 @@
+#' Label a fitted probability as 0, 1, or NA.
+#'
+#' A labeling function which outputs a value if and only if the prediction
+#' differs sufficiently from the bookmakers expectations.
+#' @param prediction A fitted probability.
+#' @param vegas Bookkeeper's odds.
+#' @param threshold A tuning parametr: we only place a bet if we differ from bookkeeper's expectations by `threshold` amount.
+#' @keywords predict
+#' @export
+#' @examples
+#' Predict(0.6, 0.5, 0.05)
+Predict <- function(prediction, vegas, threshold) {
+    if (is.na(prediction)) return(NA)
+    dif <- prediction - vegas
+    if (abs(dif) > threshold) {
+        if (dif < 0)  return(0)
+        else          return(1)
+    }
+    return(NA)
+}

### 2017-07-04 hangover/R/SubsetBLS.r

@@ -1,0 +1,36 @@
+#' A function to subset BLS data into manageable chunks.
+#'
+#' Subset BLS data into manageable chunks, keeping only observations we care about.
+#' @param year An integer describing which year of BLS data to subset.
+#' @param party.regex A character vector describing patterns to look for in BLS data.
+#' @param sport A string, one of 'nba' or 'mlb'.
+#' @param suffix A string describing how to prefix the output file.
+#' @param raw.path A string describing the input file path. Defaults to 'raw_data/bls/'.
+#' @param save.path A string describing the output file path. Defaults to 'tmp_data/bls/'.
+#' @keywords subset, bls
+#' @export
+#' @examples
+#' SubsetBLS(2010, 'sound recording', 'nba', 'sound_recording_studios')
+SubsetBLS <- function(year, party.regex, sport, suffix,
+                      raw.path = 'raw_data/bls/', save.path = 'tmp_data/bls/') {
+    ### Create a listing of files for all relevant MSA's, and load data.
+    foldr <- list.files(path = raw.path, pattern = paste0(year, '.q1'), full.names = T)
+    files <- list.files(path = foldr,
+                        pattern = 'm(icro)?sa\\.csv',
+                        full.names = T, ignore.case = T)
+    bls <- lapply(files, fread, colClasses = 'character',
+                  select = c('area_fips', 'year', 'qtr', 'industry_code',
+                             'area_title', 'industry_title', 'agglvl_title',
+                             'size_title', 'qtrly_estabs_count')) %>% rbindlist
+    setnames(bls, gsub('_', '.', names(bls)))
+    ### We will take our own total later, so for now we take granular data.
+    bls <- bls[grep("NAICS 6-digit", agglvl.title)]
+    ### Create regex to subset to drinking establshimensts.
+    party.regex <- paste0('(', party.regex, ')', collapse = '|')
+    bls <- bls[grep(party.regex, industry.title, ignore.case = T)]
+    write.csv(x = bls, file = paste0(save.path, year, '_', sport, '_', suffix, '.csv'),
+              row.names = F)
+    rm(bls)
+    gc()
+    return(NULL)
+}

### 2017-07-04 hangover/R/collectData.r

@@ -1,0 +1,53 @@
+
+### Fetch is used as a helper function within CollectData (below).
+#' A helper function used in now defunct collectData.
+#'
+#' Fetches data for a particular team from raw data from gold-sheet.com
+#' @param team A string describing a team name.
+#' @param beg A starting index of relevant data.
+#' @param end An ending index of relevant data.
+#' @param data A dataframe from which to subset.
+#' @keywords Fetch, gold-sheet.
+#' @export
+#' @examples
+#' Fetch('los angeles lakers', 10, 20, data)
+Fetch <- function(team, beg, end, data) {
+    datum <- data[beg:end]
+    ### We have to be careful for data entry errors. Ex: missing delimiter.
+    datum <- gsub("W\\+12 89-98", "W  +12  89-98", datum) %>%               # Separate fields.
+        gsub("([0-9]{3}[OU][[:blank:]]+)(.*)", "\\1", .) %>% # Throw out playoff game (not formatted)
+        strsplit('[[:blank:]]{2,}')
+    datum <- Filter(function(x) !any(grepl("CAN\\.", x)), datum)
+    datum <- do.call(rbind, args = datum) %>%
+        data.table(.) %>% cbind(team)
+}
+
+#' A function to retrieve data from a raw datafile.
+#'
+#' Given a raw datafile from gold-sheet.com, this function parses the data.
+#' @param datafile A string describing input file.
+#' @keywords collect, legacy
+#' @export
+#' @examples
+#' collectData('raw_data/2011.txt')
+collectData <- function(datafile) {
+    data <- readLines(datafile) # 'raw_data/2011.txt')
+    ### Remove blank lines.
+    blanks <- grep("^[[:blank:]]*$", data)
+    data <- data[-blanks]
+    ### Find team names, aka starting indices. 
+    beg <- grep("^\\(SUR", data) - 1     # Look one row before team-summary for team-name.
+    end <- c(beg[-1] - 1, length(data))  # Find limits of data for each team.
+    teams <- data[beg]
+    ### This data frame describes where data start-and-stop for each team.
+    teams <- data.frame(team = teams, beg = beg+2, end = end)
+    data <- mapply(Fetch, team = teams$team, beg = teams$beg, end = teams$end,
+                   MoreArgs = list(data=data), SIMPLIFY = F)
+    data <- rbindlist(data)
+    data[['season']] = gsub('raw_data/([0-9_]{9})\\.txt', '\\1', datafile)
[…]

### 2017-07-04 clean_bls.r
Now using 'hangover' package.

@@ -24,7 +24,8 @@
 ### Set up Workspace
 ##################################################
 
-source('functions.r') # for subsetBLS, CleanBLS.
+require(hangover)
+
 require(data.table)
 require(ggmap)
 require(magrittr)

### 2017-07-04 clean_bls_mlb.r

@@ -18,6 +18,8 @@
 ################################################################################
 ################################################################################
 
+require(hangover)
+
 require(data.table)
 require(ggmap)
 require(magrittr)
@@ -30,9 +32,6 @@ clusterCall(cl, function() {
     require(data.table)
     require(magrittr)
 })
-
-### All the heavy lifting is done in our CleanBLS function...
-source('functions.r')
 
 ### Load raw data, subset based on party.regex, aggregate using FUN, and save using suffix/sport.
 CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),

### 2017-07-04 clean_covers_mlb.r

@@ -21,7 +21,7 @@ require(data.table)
 require(magrittr)
 require(sampling)
 
-source('functions.r')
+require(hangover)
 
 ### A character vector of full team names.
 fulls <- c('arizona diamondbacks', 'atlanta braves', 'baltimore orioles', 'boston redsox',

### 2017-07-04 init_model.rnw
Meet the spread for both NBA and MLB (after conditioning on odds, is just a coinflip.

@@
 require(sampling)
+require(stargazer)
 set.seed(4052017)
@@
 
+### Meet the spread for both NBA and MLB
+pdf('writing/meet_the_spread.pdf')
+par(mfcol = c(2, 1))
+nba.lines[, list(meet.spread = mean(outcome == 'W')), 
+          by = list(season, team)][, hist(meet.spread, 
+                                          main = paste('How often do teams meet the spread? (NBA)', 
+                                                       'By season-team, 2010-11 through 2016-17',
+                                                       sep = '\n'),
+                                          xlab = 'Meet the Spread', breaks = 25)]
+### There should be 'no information' left over after conditioning on odds.
+noinf <- lm(I(team.score > opponent.score) ~ odds, data = mlb.lines)
+preds <- fitted(noinf)
+mlb.lines[, fitted := fitted(noinf)]
+mlb.lines[, resid  := resid(noinf)]
+mlb.lines[, list(avg.residual = mean(resid)), by = round(fitted, 3)] %>% 
+    plot(., xlab = 'Fitted Probability', ylab = 'Average Residual', 
+         main = paste('After conditioning on bookkeeper\'s odds (MLB),',
+                      'successful prediction is a coin-flip', sep = '\n'))
+abline(a = 0, b = 0, col = 'red', lty = 'dashed')
+dev.off()
+
 @

### 2017-07-04 hangover/R/degToCompass.r
Init commit.

@@ -1,0 +1,14 @@
+#' Converts from degrees to a compass direction.
+#'
+#' Source: https://stackoverflow.com/questions/7490660/converting-wind-direction-in-angles-to-text-words
+#' @param x A double representing directional degrees.
+#' @keywords degrees, compass, direction
+#' @export
+#' @examples
+#' degToCompass(300)
+degToCompass <- function(x) {
+    val <- ((x/22.5)+.5)
+    arr <- c("N", "NNE", "NE", "ENE", "E", "ESE", "SE", "SSE",
+             "S", "SSW", "SW", "WSW", "W", "WNW", "NW", "NNW")
+    arr[(val %% 16)]
+}

### 2017-07-04 hangover/R/getDirection.r

@@ -1,0 +1,12 @@
+#' Lookup direction of travel between two locations.
+#'
+#' Given two locations and a (character) matrix describing direction, retrieve direction of travel from last location to current location.
+#' @param last.loc A string or integer describing a row index.
+#' @param cur.loc A string or integer describing a column index.
+#' @param dirs A character matrix describing direction of travel between locations.
+#' @keywords get, direction
+#' @export
+#' @examples
+getDirection <- function(last.loc, cur.loc, dirs)
+    if(!is.na(last.loc) && !is.na(cur.loc))
+        dirs[last.loc, cur.loc]

### 2017-07-04 hangover/R/getDist.r

@@ -1,0 +1,17 @@
+#' Lookup distance between two locations.
+#'
+#' Given two locations and a distance matrix, perform a table-lookup.
+#' @param current A string describing a row name or an integer describing row-index.
+#' @param last A string describing a column name or an integer describing column-index.
+#' @param distances A matrix containing distances between pairwise points.
+#' @keywords get, distance
+#' @export
+#' @examples
+#' teams <- c('los angeles lakers', 'new york knicks')
+#' mat <- matrix(data = c(0, 999, 999, 0), nrow = 2, byrow = T)
+#' rownames(mat) <- teams
+#' colnames(mat) <- teams
+#' getDist('los angeles lakers', 'new york knicks', mat)
+getDist <- function(current, last, distances)
+    if (!is.na(current) && !is.na(last))
+        distances[current, last]

### 2017-07-04 hangover/R/getFullTeamnames.r

@@ -1,0 +1,32 @@
+#' Query full team names for one of NBA or MLB.
+#'
+#' A simple wrapper to return full team names for a particular sport.
+#' @param sport A string, one of 'nba' or 'mlb'
+#' @keywords get, teamnames, full
+#' @export
+#' @examples
+#' getFullTeamnames('nba')
+getFullTeamnames <- function(sport) {
+    if (sport == 'nba')
+        c('Atlanta Hawks',          'Brooklyn Nets',          'Boston Celtics',
+          'Charlotte Bobcats',      'Chicago Bulls',          'Cleveland Cavaliers',
+          'Dallas Mavericks',       'Denver Nuggets',         'Detroit Pistons',
+          'Golden State Warriors',  'Houston Rockets',        'Indiana Pacers',
+          'Los Angeles Clippers',   'Los Angeles Lakers',     'Memphis Grizzlies',
+          'Miami Heat',             'Milwaukee Bucks',        'Minnesota Timberwolves',
+          'New Orleans Hornets',    'New York Knicks',        'Oklahoma City Thunder',
+          'Orlando Magic',          'Philadelphia Sixers',    'Phoenix Suns',
+          'Portland Trail Blazers', 'San Antonio Spurs',      'Sacramento Kings',
+          'Toronto Raptors',        'Utah Jazz',              'Washington Wizards')
+    else if (sport == 'mlb')
+        c('arizona diamondbacks', 'atlanta braves',      'baltimore orioles',
+          'boston redsox',        'chicago cubs',        'chicago whitesox',
+          'cincinati reds',       'cleveland indians',   'colorado rockets',
+          'detroit tigers',       'houston astros',      'kansas city royals',
+          'los angeles angels',   'los angeles dodgers', 'miami marlins',
+          'milwaukee brewers',    'minnesota twins',     'new york mets',
+          'new york yankees',     'oakland athletics',   'philadelphia phillies',
+          'pittsburgh pirates',   'san diego padres',    'seattle mariners',
+          'san francisco giants', 'st louis cardinals',  'tampa bay rays',
+          'texas rangers',        'toronto blue jays',   'washington nationals')
+}

### 2017-07-04 clean_covers.r
Now using hangover package.

@@ -21,6 +21,8 @@
 ### Set up Workspace, Read in Raw Data
 ##################################################
 
+require(hangover)
+
 require(data.table)
 require(geosphere)
 require(ggmap)
@@ -47,14 +49,7 @@ lines[, ou := NULL]
 lines[, `:=`(home = tolower(home), away = tolower(away))]
 
 abbrs <- unique(lines$home) %>% sort
-fulls <- c('Atlanta Hawks',          'Brooklyn Nets',          'Boston Celtics',    'Charlotte Bobcats',
-           'Chicago Bulls',          'Cleveland Cavaliers',    'Dallas Mavericks',  'Denver Nuggets',
-           'Detroit Pistons',        'Golden State Warriors',  'Houston Rockets',   'Indiana Pacers',
-           'Los Angeles Clippers',   'Los Angeles Lakers',     'Memphis Grizzlies', 'Miami Heat',
-           'Milwaukee Bucks',        'Minnesota Timberwolves', 'New Orleans Hornets', 'New York Knicks',
-           'Oklahoma City Thunder',  'Orlando Magic',          'Philadelphia Sixers', 'Phoenix Suns',
-           'Portland Trail Blazers', 'San Antonio Spurs', 'Sacramento Kings',         'Toronto Raptors',
-           'Utah Jazz',              'Washington Wizards')
+fulls <- hangover::getFullTeamnames('nba')
 
 teams <- data.frame(team = fulls %>% tolower, abbr = abbrs %>% tolower, stringsAsFactors = F)
 write.csv(teams, file = 'tmp_data/team_abbreviations.csv', row.names = F)
@@ -133,11 +128,7 @@ lines[, last.game.time := c(NA, lag(hour(date))[1:.N-1]), by = list(team, season
 ## save(dmat, file = 'tmp_data/distance_matrix.RData')
 load(file = 'tmp_data/distance_matrix.RData')
 
-getDist <- function(current, last, distances)
-    if (!is.na(current) && !is.na(last))
-        distances[current, last]
-
-lines[, travel.dist := getDist(location, last.game.loc, dmat),
+lines[, travel.dist := hangover::getDist(location, last.game.loc, dmat),
       by = list(location, last.game.loc)]
 
 ##################################################
@@ -153,23 +144,12 @@ for (i in 1:(nrow(dirs)))
                                   p2 = locs[j, list(lon, lat)])
 diag(dirs) <- NA
 
[…]

### 2017-07-04 clean_covers_mlb.r
Now using hangover package.

@@ -24,14 +24,7 @@ require(sampling)
 require(hangover)
 
 ### A character vector of full team names.
-fulls <- c('arizona diamondbacks', 'atlanta braves', 'baltimore orioles', 'boston redsox',
-           'chicago cubs', 'chicago whitesox', 'cincinati reds', 'cleveland indians',
-           'colorado rockets', 'detroit tigers', 'houston astros', 'kansas city royals',
-           'los angeles angels', 'los angeles dodgers', 'miami marlins', 'milwaukee brewers',
-           'minnesota twins', 'new york mets', 'new york yankees', 'oakland athletics',
-           'philadelphia phillies', 'pittsburgh pirates', 'san diego padres', 'seattle mariners',
-           'san francisco giants', 'st louis cardinals', 'tampa bay rays', 'texas rangers',
-           'toronto blue jays', 'washington nationals')
+fulls <- hangover::getFullTeamnames('mlb')
 
 ##################################################
 ### Load raw money-lines from covers.com.
@@ -160,11 +153,8 @@ lines <- lines[ndays.lgame != 0]
 ## save(dmat, file = 'tmp_data/distance_matrix_mlb.RData')
 load(file = 'tmp_data/distance_matrix_mlb.RData')
 
-getDist <- function(current, last, distances)
-    if (!is.na(current) && !is.na(last))
-        distances[current, last]
-
-lines[, travel.dist := getDist(location, last.game.loc, dmat), by = list(location, last.game.loc)]
+lines[, travel.dist := hangover::getDist(location, last.game.loc, dmat),
+      by = list(location, last.game.loc)]
 
 
 lines[, c('team.score', 'opponent.score') := tstrsplit(score, split = '-')]

### 2017-07-04 legacy/clean_spread.r
Moved legacy code.

@@ -1,0 +1,102 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Clean Spreads
+###
+### Andreas Santucci
+###
+### Date: February 2017
+###
+### Inputs: spreads data from goldensheet.com
+###
+################################################################################
+################################################################################
+
+require(data.table)
+require(magrittr)
+source('functions.r')
+
+##############################
+### Load Data
+##############################
+
+files <- list.files('raw_data', pattern = '[[:digit:]]{4}.txt', full.names = T)[-(1:3)]
+data <- lapply(files, collectData)
+data <- rbindlist(data)
+
+##############################
+### Clean Team Names
+##############################
+
+data[, team := as.character(team)]
+data[, team := tolower(team)]
+data[, opponent := tolower(opponent)]
+
+data[opponent == 'okla. city', opponent := 'oklahoma city']
+data[, opponent := gsub('^la ', 'los angeles ', opponent)]
+
+data[team == 'okla. city thunder', team := 'oklahoma city thunder']
+data[, team := gsub('^[[:blank:]]+', '', team)]
+
+### Relationship between 'new jersey nets' and 'brooklyn nets', collapse into one?
+  # #data[grepl("\\<nets$", team), team := 'brooklyn nets']
+
+### Map team-abbreviations to proper names.
+full.teams <- data[, unique(opponent)] %>%
+    sapply(., grep, x = unique(data$team), value = T) %>%
+    data.frame
+full.teams[['abbr']] = rownames(full.teams)
+setDT(full.teams)
+setnames(full.teams, c('opponent.team', 'opponent'))
+full.teams[, opponent.team := as.character(opponent.team)]
+
+### Merge in propert opponent team names.
+data <- merge(data, full.teams, by = 'opponent', all.x = T)
+setnames(data, c('opponent', 'opponent.team'), c('opponent.team', 'opponent'))
+
+##############################
+### Spreads and Over-Unders
+##############################
+
+data <- data[spread != 'P'] # We lose 14 observations here...
+data[, spread := gsub('\'$', '', spread) %>% as.numeric]
+
+### Over-under.
+data[, ou.out := gsub('^[0-9]+([OUN])$', '\\1', ou, ignore.case = T)]
[…]

### 2017-07-04 legacy/explore.r

@@ -1,0 +1,138 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Exploratory Data Analysis
+###
+### Andreas Santucci
+###
+### Date: March 2017
+###
+### Inputs: 'tmp_data/spreads_wdist.RData', 'tmp_data/standings.RData'
+###
+################################################################################
+################################################################################
+
+##################################################
+### Set up Workspace, load data
+##################################################
+
+require(data.table)
+require(ggmap)
+require(ggplot2)
+require(magrittr)
+require(sp)
+source('functions.r')
+
+load(file = 'tmp_data/spreads_wdist.RData')
+load(file = 'tmp_data/standings.RData')
+
+data[, season := gsub('_20', '_', season)]
+
+standings <- subset(standings, select = c('team', 'season', 'pct'))
+
+data <- merge(data, standings, by = c('season', 'team'), all.x = T)
+
+### Remove playoff games (sketchy)
+data <- data[data[, head(.I, 82), by = list(season, team)]$V1]
+
+##################################################
+### Define a Party Variable
+##################################################
+
+party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
+party.rgx <- paste0('(', party.cities, ')', collapse = "|")
+
+data[, party := 0]
+data[, weekend := grepl("(Friday)|(Saturday)|(Sunday)", day) %>% as.integer]
+
+### We (1) exclude people who already party, (2) check if last game played against party city
+### (3) check last game was played 'away' i.e. at party city and (4) party last night
+data[!grepl(party.rgx, team) & grepl(party.rgx, last.game.loc) & ndays.lgame == 1, party := 1]
+
+data[grepl(party.rgx, team) & grepl(party.rgx, last.game.loc) &
+     team != last.game.loc & ndays.lgame == 1, party := 1]
+
+### We also consider: when the team had the day off the day before playing a party city
+### data[grepl('V', location) & grepl(party.rgx, opponent) & last.game > 1, table(outcome)]
+
+data[, .N, by = list(season, team, party)]
+m <- data[party==1, .N, by = list(season, team)]
+ggplot(m, aes(x = N)) +
[…]

### 2017-07-04 legacy/explore_covers.r

@@ -1,0 +1,115 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Exploring Covers Data
+###
+### Andreas Santucci
+###
+### Date: April 2017
+###
+### Inputs: 
+###
+### Dependencies: 
+###
+################################################################################
+################################################################################
+
+require(data.table)
+require(magrittr)
+require(ggmap)
+require(sp)
+
+load('tmp_data/covers_lines.RData')
+
+##################################################
+### Exploring data
+##################################################
+
+require(ggplot2)
+
+hist(lines[, mean(outcome == 'W'), by = list(season, team)][, V1],
+     breaks = seq(0.3, 0.75, by = 0.025),
+     main = "How often are teams meeting the spread?",
+     xlab = "Proportion of games for which a team-year meets the spread")
+
+tmp <- lines[, list(meet.spread = mean(outcome == 'W'), .N),
+             by = list(last.game.loc, ndays.lgame)][order(ndays.lgame)]
+
+ggplot(tmp %>% na.omit, aes(x = ndays.lgame, y = meet.spread, label = N)) +
+    geom_point() +
+    geom_label() + 
+    geom_hline(yintercept = 0.5) + 
+    facet_wrap(facets = ~ last.game.loc)
+
+lines[party==1, .N, by = list(season, team)][, N] %>%
+    hist(main = "Number of party days by season-team", xlab = "Number of days",
+         ylab = "Frequency")
+
+tmp <- lines[, list(meet.spread = mean(outcome == 'W'), .N), 
+            by = list(pct, party)][order(pct)]
+ggplot(tmp, aes(x = pct, y = meet.spread, color = as.factor(party), label = N)) + 
+    geom_label() +
+    geom_hline(yintercept = 0.5) + 
+    geom_smooth()
+
+m <- lines[,#grepl(party.rgx, game.loc) & !grepl(party.rgx, team),
+          list(meet.spread = mean(outcome == 'W'), pct = unique(pct), .N), by = list(team, season)]
+ggplot(m, aes(x = pct, y = meet.spread)) +
+    geom_point() +
+    geom_smooth() +
+    facet_wrap(facets = ~season) + 
+    labs(title = "In select seasons, lagged performance not predictive of meet the spread",
+         x = "Percentage of games won last season",
[…]

### 2017-07-04 legacy/explore_starters.r

@@ -1,0 +1,214 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Exploring Starters Data (from ESPN)
+###
+### Andreas Santucci
+###
+### Date: April 2017
+###
+### Inputs: 
+###
+### Dependencies: 
+###
+################################################################################
+################################################################################
+
+
+require(data.table)
+require(ggplot2)
+require(lubridate)
+require(magrittr)
+require(sampling)
+set.seed(4052017)
+
+load(file = 'tmp_data/covers_lines.RData')
+load(file = 'tmp_data/nmusician_estabs.RData')
+load(file = 'tmp_data/population_by_county_year.RData')
+
+musicians <- merge(musicians, pop, by = c('county', 'season'), all.x = T)
+musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
+
+lines <- merge(lines, musicians,
+               by.x = c('season', 'last.game.loc'),
+               by.y = c('season', 'team'), all.x = T)
+
+setkey(lines, season, team, date)
+
+HOURS <- 24
+lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, nmusicians, 0)]
+
+##################################################
+### Total Points Allowed
+##################################################
+
+lines[, tpa := team.pts.admitted - mean(team.pts.admitted), by = list(team, season)]
+lines[, tps := team.pts.scored   - mean(team.pts.scored), by = list(team, season)]
+
+lm(tpa ~ party + I(hour(date))  + nhours.lgame, data = lines[nhours.lgame <= HOURS]) %>% summary
+
+##################################################
+### ESPN
+##################################################
+
+espn <- fread('tmp_data/espn_player_data.csv')
+
+### keeping starters makes a big difference in lag.chg.pos effect
+#espn <- espn[espn[, .I[1:5], by = list(team, date)][, V1]]
+espn[, date := as.Date(date)]
+lines[, simple.date := as.character(date) %>% substr(., 1, 10) %>% as.Date]
+data <- merge(lines, espn,
[…]

### 2017-07-04 legacy/model.r

@@ -1,0 +1,42 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Model
+###
+### Andreas Santucci
+###
+### Date: March 2017
+###
+### Inputs: 'tmp_data/spreads_with_standings.RData'
+###
+################################################################################
+################################################################################
+
+require(data.table)
+require(magrittr)
+source('functions.r')
+
+load(file = 'tmp_data/spreads_with_standings.RData')
+
+table(data$party, data$outcome)      # Only win 44% of the games.
+table(data[ndays.lgame == 1, outcome]) # Win 49% of the games they play the day before.
+
+data[, pq := cut(pct, breaks = quantile(pct)), by = season]
+data[, hp := ifelse(pq == '(0.244,0.378]' & party == 1, 1, 0)]
+
+party.teams <- c('charlotte bobcats',   'chicago bulls',      'denver nuggets',
+                 'new orleans hornets', 'philadelphia 76ers', 'houston rockets')
+
+model <- glm(outcome=='W' ~ party + weekend + ndays.lgame + pct + I(travel.dist>1350),
+             family = 'binomial', data = data, subset = season != '2008_09' & team %in% party.teams)
+
+model <- glm(outcome == 'W' ~ party + weekend + ndays.lgame + pct +
+                 travel.dist + team + season + team:season,
+             family = 'binomial', data = data, subset = season != '2008_09')
+
+### logistic
+### probability of meeting the spread ~ # days since last game + # games played in last week +
+###  + day of week + team-ranking + party city
+
+### proportion of games where spread is met if spread is over abs(X)

### 2017-07-04 legacy/eda_report.rnw
Moved legacy code.

--- 2017-03-29 eda_report.rnw
@@
 
-\subsection*{Motivation}
+<<echo=F,results='hide',include=FALSE>>=
+require(data.table)
+require(ggplot2)
+require(magrittr)
+
+load(file = 'tmp_data/spreads_wdist.RData')
+load(file = 'tmp_data/standings.RData')
+
+data[, season := gsub('_20', '_', season)]
+
+standings <- subset(standings, select = c('team', 'season', 'pct'))
+standings[, pct := shift(pct, type = 'lag'), by = team]
+
+data <- merge(data, standings, by = c('season', 'team'), all.x = T)
+data <- data[data[, head(.I, 82), by = list(season, team)]$V1]
+
+### Define a Party Variable
+#party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
+party.cities <- c('boston', 'brooklyn', 'los angeles', 'new york')
+party.rgx <- paste0('(', party.cities, ')', collapse = "|")
+
+data[, party := 0]
+data[, weekend := grepl("(Friday)|(Saturday)|(Sunday)", day) %>% as.integer]
+
+data[!grepl(party.rgx, team) & grepl(party.rgx, last.game.loc) & ndays.lgame == 1, party := 1]
+@
+
+\section{Motivation}
 It's not unreasonable to think that game performance can be affected (in part) by what takes place off the basketball court.
@@
 
+\section{Which teams to include as party cities?}
+
 \paragraph{Defining a party city}
@@
 who travel to a party city and then play a game in another city the day after; each of the thirty NBA teams encounters this situation on average twice per season.
+
+<<echo=-1>>=
+
+tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
+     by = list(last.game.loc, ndays.lgame)][order(last.game.loc, ndays.lgame)]
+
+tmp[ndays.lgame == 1][order(meet.spread)]
+
+@
+
+<<results='hide'>>=
+
+### What about by game location
+ggplot(tmp, aes(x = ndays.lgame, y = meet.spread, label = N)) +
+    geom_point() + 
+    geom_label(nudge_y = 0.2) +
+    geom_hline(yintercept = 0.5) + 
+    facet_wrap(facets = ~ last.game.loc)
+
+@
 
@@
 
-<<echo=F,results='hide',include=FALSE>>=
-require(data.table)
-require(ggplot2)
-require(magrittr)
-
-load(file = 'tmp_data/spreads_wdist.RData')
-load(file = 'tmp_data/standings.RData')
-
-data[, season := gsub('_20', '_', season)]
-
-standings <- subset(standings, select = c('team', 'season', 'pct'))
-standings[, pct := shift(pct, type = 'lag'), by = team]
-
-data <- merge(data, standings, by = c('season', 'team'), all.x = T)
-data <- data[data[, head(.I, 82), by = list(season, team)]$V1]
-
-### Define a Party Variable
-#party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
-party.cities <- c('boston', 'brooklyn', 'los angeles', 'new york')
-party.rgx <- paste0('(', party.cities, ')', collapse = "|")
-
-data[, party := 0]
-data[, weekend := grepl("(Friday)|(Saturday)|(Sunday)", day) %>% as.integer]
-
-data[!grepl(party.rgx, team) & grepl(party.rgx, last.game.loc) & ndays.lgame == 1, party := 1]
-@
-
 <<echo=F,out.height="6in",warning=F>>=
@@
     hist(main = "Number of party days by season-team", xlab = "Number of days", ylab = "Frequency")
-@
-
-\subsection*{Which teams to include as party cities?}
-<<echo=-1>>=
-
-tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
-     by = list(last.game.loc, ndays.lgame)][order(last.game.loc, ndays.lgame)]
-
-tmp[ndays.lgame == 1][order(meet.spread)]
-
-@
-
-<<results='hide'>>=
-
-### What about by game location
-ggplot(tmp, aes(x = ndays.lgame, y = meet.spread, label = N)) +
-    geom_point() + 
-    geom_label(nudge_y = 0.2) +
-    geom_hline(yintercept = 0.5) + 
-    facet_wrap(facets = ~ last.game.loc)
-
 @
@@
 
-summary(model)$coef[c(1:6, sample(7:nrow(coef(summary(model))), size = 6)), ]
+summary(model)
 @

### 2017-07-04 clean_spread.r
Moved.

@@ -1,102 +1,0 @@
-################################################################################
-################################################################################
-################################################################################
-###
-### Title: Clean Spreads
-###
-### Andreas Santucci
-###
-### Date: February 2017
-###
-### Inputs: spreads data from goldensheet.com
-###
-################################################################################
-################################################################################
-
-require(data.table)
-require(magrittr)
-source('functions.r')
-
-##############################
-### Load Data
-##############################
-
-files <- list.files('raw_data', pattern = '[[:digit:]]{4}.txt', full.names = T)[-(1:3)]
-data <- lapply(files, collectData)
-data <- rbindlist(data)
-
-##############################
-### Clean Team Names
-##############################
-
-data[, team := as.character(team)]
-data[, team := tolower(team)]
-data[, opponent := tolower(opponent)]
-
-data[opponent == 'okla. city', opponent := 'oklahoma city']
-data[, opponent := gsub('^la ', 'los angeles ', opponent)]
-
-data[team == 'okla. city thunder', team := 'oklahoma city thunder']
-data[, team := gsub('^[[:blank:]]+', '', team)]
-
-### Relationship between 'new jersey nets' and 'brooklyn nets', collapse into one?
-  # #data[grepl("\\<nets$", team), team := 'brooklyn nets']
-
-### Map team-abbreviations to proper names.
-full.teams <- data[, unique(opponent)] %>%
-    sapply(., grep, x = unique(data$team), value = T) %>%
-    data.frame
-full.teams[['abbr']] = rownames(full.teams)
-setDT(full.teams)
-setnames(full.teams, c('opponent.team', 'opponent'))
-full.teams[, opponent.team := as.character(opponent.team)]
-
-### Merge in propert opponent team names.
-data <- merge(data, full.teams, by = 'opponent', all.x = T)
-setnames(data, c('opponent', 'opponent.team'), c('opponent.team', 'opponent'))
-
-##############################
-### Spreads and Over-Unders
-##############################
-
-data <- data[spread != 'P'] # We lose 14 observations here...
-data[, spread := gsub('\'$', '', spread) %>% as.numeric]
-
-### Over-under.
-data[, ou.out := gsub('^[0-9]+([OUN])$', '\\1', ou, ignore.case = T)]
[…]

### 2017-07-04 explore_covers.r
Moved to legacy.

@@ -1,115 +1,0 @@
-################################################################################
-################################################################################
-################################################################################
-###
-### Title: Exploring Covers Data
-###
-### Andreas Santucci
-###
-### Date: April 2017
-###
-### Inputs: 
-###
-### Dependencies: 
-###
-################################################################################
-################################################################################
-
-require(data.table)
-require(magrittr)
-require(ggmap)
-require(sp)
-
-load('tmp_data/covers_lines.RData')
-
-##################################################
-### Exploring data
-##################################################
-
-require(ggplot2)
-
-hist(lines[, mean(outcome == 'W'), by = list(season, team)][, V1],
-     breaks = seq(0.3, 0.75, by = 0.025),
-     main = "How often are teams meeting the spread?",
-     xlab = "Proportion of games for which a team-year meets the spread")
-
-tmp <- lines[, list(meet.spread = mean(outcome == 'W'), .N),
-             by = list(last.game.loc, ndays.lgame)][order(ndays.lgame)]
-
-ggplot(tmp %>% na.omit, aes(x = ndays.lgame, y = meet.spread, label = N)) +
-    geom_point() +
-    geom_label() + 
-    geom_hline(yintercept = 0.5) + 
-    facet_wrap(facets = ~ last.game.loc)
-
-lines[party==1, .N, by = list(season, team)][, N] %>%
-    hist(main = "Number of party days by season-team", xlab = "Number of days",
-         ylab = "Frequency")
-
-tmp <- lines[, list(meet.spread = mean(outcome == 'W'), .N), 
-            by = list(pct, party)][order(pct)]
-ggplot(tmp, aes(x = pct, y = meet.spread, color = as.factor(party), label = N)) + 
-    geom_label() +
-    geom_hline(yintercept = 0.5) + 
-    geom_smooth()
-
-m <- lines[,#grepl(party.rgx, game.loc) & !grepl(party.rgx, team),
-          list(meet.spread = mean(outcome == 'W'), pct = unique(pct), .N), by = list(team, season)]
-ggplot(m, aes(x = pct, y = meet.spread)) +
-    geom_point() +
-    geom_smooth() +
-    facet_wrap(facets = ~season) + 
-    labs(title = "In select seasons, lagged performance not predictive of meet the spread",
-         x = "Percentage of games won last season",
[…]

### 2017-07-04 explore_starters.r

@@ -1,193 +1,0 @@
-################################################################################
-################################################################################
-################################################################################
-###
-### Title: Exploring Starters Data (from ESPN)
-###
-### Andreas Santucci
-###
-### Date: April 2017
-###
-### Inputs: 
-###
-### Dependencies: 
-###
-################################################################################
-################################################################################
-
-
-require(data.table)
-require(ggplot2)
-require(lubridate)
-require(magrittr)
-require(sampling)
-set.seed(4052017)
-
-load(file = 'tmp_data/covers_lines.RData')
-load(file = 'tmp_data/nmusician_estabs.RData')
-load(file = 'tmp_data/population_by_county_year.RData')
-
-musicians <- merge(musicians, pop, by = c('county', 'season'), all.x = T)
-musicians[, county := gsub(' (county)|(division),.*$', '', county, ignore.case = T)]
-
-lines <- merge(lines, musicians,
-               by.x = c('season', 'last.game.loc'),
-               by.y = c('season', 'team'), all.x = T)
-
-setkey(lines, season, team, date)
-
-HOURS <- 24
-lines[, party := ifelse(nhours.lgame <= HOURS & last.game.loc != team, nmusicians, 0)]
-
-##################################################
-### Total Points Allowed
-##################################################
-
-lines[, tpa := team.pts.admitted - mean(team.pts.admitted), by = list(team, season)]
-lines[, tps := team.pts.scored   - mean(team.pts.scored), by = list(team, season)]
-
-lm(tpa ~ party + I(hour(date))  + nhours.lgame, data = lines[nhours.lgame <= HOURS]) %>% summary
-
-##################################################
-### ESPN
-##################################################
-
-espn <- fread('tmp_data/espn_player_data.csv')
-
-### keeping starters makes a big difference in lag.chg.pos effect
-#espn <- espn[espn[, .I[1:5], by = list(team, date)][, V1]]
-espn[, date := as.Date(date)]
-lines[, simple.date := as.character(date) %>% substr(., 1, 10) %>% as.Date]
-data <- merge(lines, espn,
[…]

### 2017-07-04 explore.r
Moved to legacy.

@@ -1,138 +1,0 @@
-################################################################################
-################################################################################
-################################################################################
-###
-### Title: Exploratory Data Analysis
-###
-### Andreas Santucci
-###
-### Date: March 2017
-###
-### Inputs: 'tmp_data/spreads_wdist.RData', 'tmp_data/standings.RData'
-###
-################################################################################
-################################################################################
-
-##################################################
-### Set up Workspace, load data
-##################################################
-
-require(data.table)
-require(ggmap)
-require(ggplot2)
-require(magrittr)
-require(sp)
-source('functions.r')
-
-load(file = 'tmp_data/spreads_wdist.RData')
-load(file = 'tmp_data/standings.RData')
-
-data[, season := gsub('_20', '_', season)]
-
-standings <- subset(standings, select = c('team', 'season', 'pct'))
-
-data <- merge(data, standings, by = c('season', 'team'), all.x = T)
-
-### Remove playoff games (sketchy)
-data <- data[data[, head(.I, 82), by = list(season, team)]$V1]
-
-##################################################
-### Define a Party Variable
-##################################################
-
-party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
-party.rgx <- paste0('(', party.cities, ')', collapse = "|")
-
-data[, party := 0]
-data[, weekend := grepl("(Friday)|(Saturday)|(Sunday)", day) %>% as.integer]
-
-### We (1) exclude people who already party, (2) check if last game played against party city
-### (3) check last game was played 'away' i.e. at party city and (4) party last night
-data[!grepl(party.rgx, team) & grepl(party.rgx, last.game.loc) & ndays.lgame == 1, party := 1]
-
-data[grepl(party.rgx, team) & grepl(party.rgx, last.game.loc) &
-     team != last.game.loc & ndays.lgame == 1, party := 1]
-
-### We also consider: when the team had the day off the day before playing a party city
-### data[grepl('V', location) & grepl(party.rgx, opponent) & last.game > 1, table(outcome)]
-
-data[, .N, by = list(season, team, party)]
-m <- data[party==1, .N, by = list(season, team)]
-ggplot(m, aes(x = N)) +
[…]

### 2017-07-04 model.r

@@ -1,42 +1,0 @@
-################################################################################
-################################################################################
-################################################################################
-###
-### Title: Model
-###
-### Andreas Santucci
-###
-### Date: March 2017
-###
-### Inputs: 'tmp_data/spreads_with_standings.RData'
-###
-################################################################################
-################################################################################
-
-require(data.table)
-require(magrittr)
-source('functions.r')
-
-load(file = 'tmp_data/spreads_with_standings.RData')
-
-table(data$party, data$outcome)      # Only win 44% of the games.
-table(data[ndays.lgame == 1, outcome]) # Win 49% of the games they play the day before.
-
-data[, pq := cut(pct, breaks = quantile(pct)), by = season]
-data[, hp := ifelse(pq == '(0.244,0.378]' & party == 1, 1, 0)]
-
-party.teams <- c('charlotte bobcats',   'chicago bulls',      'denver nuggets',
-                 'new orleans hornets', 'philadelphia 76ers', 'houston rockets')
-
-model <- glm(outcome=='W' ~ party + weekend + ndays.lgame + pct + I(travel.dist>1350),
-             family = 'binomial', data = data, subset = season != '2008_09' & team %in% party.teams)
-
-model <- glm(outcome == 'W' ~ party + weekend + ndays.lgame + pct +
-                 travel.dist + team + season + team:season,
-             family = 'binomial', data = data, subset = season != '2008_09')
-
-### logistic
-### probability of meeting the spread ~ # days since last game + # games played in last week +
-###  + day of week + team-ranking + party city
-
-### proportion of games where spread is met if spread is over abs(X)

### 2017-07-04 legacy/scrape_play_by_play.r
Moved to legacy.

@@ -1,0 +1,60 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Scrape Play-by-Play data
+###
+### Andreas Santucci
+###
+### Date: April 2017
+###
+### Inputs: 
+###
+### Dependencies: 
+###
+################################################################################
+################################################################################
+
+require(parallel)
+require(XML)
+cl <- makeCluster(detectCores())
+
+load(file = 'tmp_data/espn_gameIDs.RData')
+
+extractGame <- function(gameID, date) {
+    tryCatch(expr = {
+        pg <- paste0('http://scores.espn.com/nba/playbyplay?gameId=', gameID)
+        tb <- readHTMLTable(pg)
+        scores <- tb[[1]]
+        teams  <- paste0(scores[, 1], collapse = '-')
+        ls <- sapply(tb, nrow)
+        tb <- tb[ls > 32]
+        data <- do.call(rbind, tb)
+        data$matchup <- teams
+        data$date <- date
+        write.csv(data,
+                  file = paste0('raw_data/play_by_play/', gameID, '.csv'),
+                  row.names = F)
+    }, error = function(e) return(NULL))
+}
+
+## extractPlays(gameIDs[[1]][1], names(gameIDs)[1])
+## We do have 'flagrant foul'
+
+extractPlays <- function(gameIDs, date, rescrape = F) {
+    if (!rescrape) {
+        scraped <- list.files('raw_data/play_by_play') %>% gsub('\\.csv', '', .)
+        gameIDs <- setdiff(gameIDs, scraped)
+    }
+    lapply(gameIDs, extractGame, date = date) %>% rbindlist
+}
+
+clusterCall(cl, function() {
+    require(XML)
+    require(magrittr)
+    require(data.table)
+})
+clusterExport(cl, 'extractGame')
+clusterMap(cl, fun = extractPlays, gameIDs = gameIDs, date = names(gameIDs),
+           .scheduling = 'dynamic')
+#mapply(scrapeGames, gameIDs = gameIDs, date = game.days)

### 2017-07-04 scrape_play_by_play.r
Moved to legacy.

@@ -1,60 +1,0 @@
-################################################################################
-################################################################################
-################################################################################
-###
-### Title: Scrape Play-by-Play data
-###
-### Andreas Santucci
-###
-### Date: April 2017
-###
-### Inputs: 
-###
-### Dependencies: 
-###
-################################################################################
-################################################################################
-
-require(parallel)
-require(XML)
-cl <- makeCluster(detectCores())
-
-load(file = 'tmp_data/espn_gameIDs.RData')
-
-extractGame <- function(gameID, date) {
-    tryCatch(expr = {
-        pg <- paste0('http://scores.espn.com/nba/playbyplay?gameId=', gameID)
-        tb <- readHTMLTable(pg)
-        scores <- tb[[1]]
-        teams  <- paste0(scores[, 1], collapse = '-')
-        ls <- sapply(tb, nrow)
-        tb <- tb[ls > 32]
-        data <- do.call(rbind, tb)
-        data$matchup <- teams
-        data$date <- date
-        write.csv(data,
-                  file = paste0('raw_data/play_by_play/', gameID, '.csv'),
-                  row.names = F)
-    }, error = function(e) return(NULL))
-}
-
-## extractPlays(gameIDs[[1]][1], names(gameIDs)[1])
-## We do have 'flagrant foul'
-
-extractPlays <- function(gameIDs, date, rescrape = F) {
-    if (!rescrape) {
-        scraped <- list.files('raw_data/play_by_play') %>% gsub('\\.csv', '', .)
-        gameIDs <- setdiff(gameIDs, scraped)
-    }
-    lapply(gameIDs, extractGame, date = date) %>% rbindlist
-}
-
-clusterCall(cl, function() {
-    require(XML)
-    require(magrittr)
-    require(data.table)
-})
-clusterExport(cl, 'extractGame')
-clusterMap(cl, fun = extractPlays, gameIDs = gameIDs, date = names(gameIDs),
-           .scheduling = 'dynamic')
-#mapply(scrapeGames, gameIDs = gameIDs, date = game.days)

### 2017-07-04 clean_bls.r
Made clear the dependancy on hangover package.

@@ -40,9 +40,9 @@ clusterCall(cl, function() {
 })
 
 ### Loads raw data. Subsets based on regex. Aggregate using FUN. Save using suffix and sport.
-CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
-         suffix = 'studios_and_artists', sport = 'nba', FUN = sum)
+hangover::CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
+                   suffix = 'studios_and_artists', sport = 'nba', FUN = sum)
 
 ### We add option 'rescrape = F' to avoid scraping location data a second time.
-CleanBLS(party.regex = c('alcohol', 'drinking'),
-         suffix = 'alcohol_and_drinking', sport = 'nba', FUN = sum, RESCRAPE = F)
+hangover::CleanBLS(party.regex = c('alcohol', 'drinking'),
+                   suffix = 'alcohol_and_drinking', sport = 'nba', FUN = sum, RESCRAPE = F)

### 2017-07-04 clean_bls_mlb.r
Made clear the dependancy on hangover package.

@@ -34,9 +34,9 @@ clusterCall(cl, function() {
 })
 
 ### Load raw data, subset based on party.regex, aggregate using FUN, and save using suffix/sport.
-CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
-         suffix = 'studios_and_artists', sport = 'mlb', FUN = sum)
+hangover::CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
+                   suffix = 'studios_and_artists', sport = 'mlb', FUN = sum)
 
 ### Avoids rescraping data. This output used for a comparison plot in the paper.
-CleanBLS(party.regex = c('alcohol', 'drinking'),
-         suffix = 'alcohol_and_drinking', sport = 'mlb', FUN = sum, RESCRAPE = F)
+hangover::CleanBLS(party.regex = c('alcohol', 'drinking'),
+                   suffix = 'alcohol_and_drinking', sport = 'mlb', FUN = sum, RESCRAPE = F)

### 2017-07-04 hangover/R/calculateDistances.r
Init commit. Functions appear sound; we have reproduced old results.

@@ -1,0 +1,20 @@
+#' Calculate distances between locations. Save to path.
+#'
+#' Given a listing of locations (including longitude and latitude) and a save.path,
+#' this function calculates pairwise distances and converts output to miles.
+#' @param locations A data.table describing location longitude and latitude.
+#' @param save.path A string describing output file location.
+#' @param team.names A character vector describing team-names, for labelling of output matrix.
+#' @param kmToMiles A magic number: the number of kilometers in a mile.
+#' @keywords calculate, distances
+#' @export
+#' @examples
+#'
+calculateDistances <- function(locations, save.path, team.names, kmToMiles = 1.60934) {
+    dmat <- spDists(locations[, c('lon', 'lat')] %>% as.matrix, longlat = T)
+    dmat <- dmat / kmToMiles
+    rownames(dmat) <- team.names
+    colnames(dmat) <- team.names
+    save(dmat, file = save.path)
+    return(NULL)
+}

### 2017-07-04 hangover/R/createPanelDataset.r

@@ -1,0 +1,24 @@
+#' Create a panel-data set from lines data.
+#'
+#' Given a dataframe of game-outcomes, we create a panel dataset.
+#' In particular, for each game we want two observations.
+#' The observations will be different since the lagged variables will be different.
+#' I.e. the number of days since last game and distance traveled will be different for each
+#' of the teams. To do this, we simply 'reverse' each obseration, carefully.
+#' @param dt A data.table containing game-outcomes.
+#' @param sport A string, one of 'nba' or 'mlb'
+#' @keywords create, panel
+#' @export
+#' @examples
+#'
+createPanelDataset <- function(dt, sport) {
+    cp <- copy(dt)
+    setnames(cp, c('team', 'opponent'), c('opponent', 'team'))
+    set(cp, j = 'score', value = strsplit(cp$score, split = '-') %>%
+                             lapply(., rev) %>%
+                             sapply(., paste, collapse = '-'))
+    if (sport == 'nba')      set(cp, j = 'line', value = -1 * cp$line)
+    else if (sport == 'mlb') set(cp, j = 'odds', value =  1 - cp$odds)
+    if (sport == 'nba')      set(cp, j = 'outcome', value = ifelse(cp$outcome == 'L', 'W', 'L'))
+    rbind(dt, cp)
+}

### 2017-07-04 hangover/R/getLocations.r

@@ -1,0 +1,14 @@
+#' Get lon/lat for a listing of teams.
+#'
+#' Geocode team locations to get lat-lon and addresses.
+#' @param teams A character vector of team-names we wish to geocode.
+#' @keywords get, locations
+#' @export
+#' @examples
+#'
+getLocations <- function(teams) {
+    locs <- sapply(teams, geocode, simplify = F) %>% do.call(rbind, .)
+    locs[['team']] <- rownames(locs)
+    setDT(locs)
+    return(locs)
+}

### 2017-07-04 clean_covers.r
Now using hangover package, re-using functions across 'nba' and 'mlb'.

@@ -88,23 +88,7 @@ lines[, location := team]
 ### Creating a panel-data set.
 ##################################################
 
-### We first need to reshape our data, since we want a panel-data set.
-### I.e., for each time we want a listing of games played in a season.
-### This does in fact mean that we have the same game listed twice in our data,
-### but the observations will be different since the lagged variables will be different.
-### I.e. the number of days since last game and distance traveled will be different for each
-### of the teams.
-### To do this, we simply 'reverse' each obseration, carefully.
-lines.dup <- copy(lines)
-lines.dup[, `:=`(team = opponent, opponent = team,
-                 score = strsplit(score, split = '-') %>%
-                     lapply(., rev) %>%
-                     sapply(., paste, collapse = '-'),
-                 line = -1 * line,
-                 location = location,
-                 outcome = ifelse(outcome == 'L', 'W', 'L'))]
-
-lines <- rbind(lines, lines.dup)
+lines <- hangover::createPanelDataset(lines, 'nba')
 
 setkey(lines, team, season, date)
 lines[, nhours.lgame   := c(NA, diff(date)),              by = list(team, season)]
@@ -115,17 +99,9 @@ lines[, last.game.time := c(NA, lag(hour(date))[1:.N-1]), by = list(team, season
 ### Distance Traveled.
 ##################################################
 ### Here, we geocode team locations to get lat-lon, and also addresses.
-## teams <- lines[, unique(team)]
-## locs <- sapply(teams, geocode, simplify = F)
-## locs <- do.call(rbind, locs)
-## locs[['team']] <- rownames(locs)
-## setDT(locs)
+locs <- hangover::getLocations(unique(lines$team))
 
-## dmat <- spDists(locs[, list(lon, lat)] %>% as.matrix, longlat = T) # Returns distance in kilometers.
-## dmat <- dmat / 1.60934  # Convert to Miles.
-## rownames(dmat) <- locs$team
-## colnames(dmat) <- locs$team
-## save(dmat, file = 'tmp_data/distance_matrix.RData')
+hangover::calculateDistances(locs, 'tmp_data/distance_matrix.RData', unique(lines$team))
 load(file = 'tmp_data/distance_matrix.RData')
 
 lines[, travel.dist := hangover::getDist(location, last.game.loc, dmat),
[…]

### 2017-07-04 clean_covers_mlb.r

@@ -115,16 +115,7 @@ data[, start.time := NULL]
 ##################################################
 ### Creating panel data-set.
 ##################################################
-lines.dup <- copy(data)
-lines.dup[, `:=`(team = opponent, opponent = team,
-                 score = strsplit(score, split = '-') %>%
-                     lapply(., rev) %>%
-                     sapply(., paste, collapse = '-'),
-                 odds = 1 - odds,
-                 location = location)]
-
-lines <- rbind(data, lines.dup)
-
+lines <- hangover::createPanelDataset(data, 'mlb')
 lines[, season := year(date)]
 
 setkey(lines, team, season, date)
@@ -138,19 +129,9 @@ lines <- lines[ndays.lgame != 0]
 ### Team Locations and Distance Traveled.
 ##################################################
 ### Here, we geocode team locations to get lat-lon, and also addresses.
-## require(ggmap)
-## require(sp)
-## teams <- lines[, unique(team)]
-## locs <- sapply(teams, geocode, simplify = F)
-## locs <- do.call(rbind, locs)
-## locs[['team']] <- rownames(locs)
-## setDT(locs)
+locs <- hangover::getLocations(unique(lines$team))
 
-## dmat <- spDists(locs[, list(lon, lat)] %>% as.matrix, longlat = T) # Returns distance in kilometers.
-## dmat <- dmat / 1.60934  # Convert to Miles.
-## rownames(dmat) <- locs$team
-## colnames(dmat) <- locs$team
-## save(dmat, file = 'tmp_data/distance_matrix_mlb.RData')
+hangover::calculateDistances(locs, 'tmp_data/distance_matrix_mlb.RData', unique(lines$team))
 load(file = 'tmp_data/distance_matrix_mlb.RData')
 
 lines[, travel.dist := hangover::getDist(location, last.game.loc, dmat),
@@ -190,7 +171,7 @@ save(lines, file = 'tmp_data/covers_lines_mlb.RData')
 ### Causal Model, includes all observations up through 2016.
 m <- glm(I(team.score > opponent.score) ~ party + odds + I(team == location) +
              ndays.lgame + travel.dist + weekend,
-         data = lines, family = 'binomial')
+         data = lines[season < 2017], family = 'binomial')
 summary(m)
 
 ##################################################

### 2017-07-04 legacy/get_locations.r
Moved to legacy.

@@ -1,0 +1,41 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Get Locations (for each team)
+###
+### Andreas Santucci
+###
+### Date: February 2017
+###
+### Inputs: 'tmp_data/spreads.RData'
+###
+################################################################################
+################################################################################
+
+require(data.table)
+require(ggmap)
+require(magrittr)
+require(sp)
+
+load('tmp_data/spreads.RData')
+
+### Here, we geocode team locations to get lat-lon, and also addresses.
+teams <- data[, unique(team)]
+locs <- sapply(teams, geocode, simplify = F)
+locs <- do.call(rbind, locs)
+locs[['team']] <- rownames(locs)
+setDT(locs)
+
+dmat <- spDists(locs[, list(lon, lat)] %>% as.matrix, longlat = T) # Returns distance in kilometers.
+dmat <- dmat / 1.60934  # Convert to Miles.
+rownames(dmat) <- locs$team
+colnames(dmat) <- locs$team
+
+getDist <- function(current, last, distances)
+    if (!is.na(current) && !is.na(last))
+        distances[current, last]
+
+data[, travel.dist := getDist(game.loc, last.game.loc, dmat), by = list(game.loc, last.game.loc)]
+
+save(data, file = 'tmp_data/spreads_wdist.RData')

### 2017-07-04 get_locations.r
Moved to legacy.

@@ -1,41 +1,0 @@
-################################################################################
-################################################################################
-################################################################################
-###
-### Title: Get Locations (for each team)
-###
-### Andreas Santucci
-###
-### Date: February 2017
-###
-### Inputs: 'tmp_data/spreads.RData'
-###
-################################################################################
-################################################################################
-
-require(data.table)
-require(ggmap)
-require(magrittr)
-require(sp)
-
-load('tmp_data/spreads.RData')
-
-### Here, we geocode team locations to get lat-lon, and also addresses.
-teams <- data[, unique(team)]
-locs <- sapply(teams, geocode, simplify = F)
-locs <- do.call(rbind, locs)
-locs[['team']] <- rownames(locs)
-setDT(locs)
-
-dmat <- spDists(locs[, list(lon, lat)] %>% as.matrix, longlat = T) # Returns distance in kilometers.
-dmat <- dmat / 1.60934  # Convert to Miles.
-rownames(dmat) <- locs$team
-colnames(dmat) <- locs$team
-
-getDist <- function(current, last, distances)
-    if (!is.na(current) && !is.na(last))
-        distances[current, last]
-
-data[, travel.dist := getDist(game.loc, last.game.loc, dmat), by = list(game.loc, last.game.loc)]
-
-save(data, file = 'tmp_data/spreads_wdist.RData')

### 2017-07-04 hangover/R/scrapeCoverLinesMLB.r
Moved into re-usable function.

@@ -1,0 +1,55 @@
+
+#' Scrape money-lines data for a particular day.
+#'
+#' Scrape money-lines data for a particular day. At the moment, we only keep
+#' the first money-line we run across.
+#' @param link A url describing where game data are stored.
+#' @keywords scrape, lines, mlb
+#' @export
+#' @examples
+#'
+scrapeLines <- function(link) {
+    tryCatch(expr = {
+        lines <- readHTMLTable(link, which = 1)
+        if (is.null(lines))
+            return(data.frame("no", "lines", "data"))
+        else
+            head(lines, 1)
+    }, error = function(e) return(data.frame("no", "lines", "data")))
+}
+
+### For a particular date,this function fetches a listing of games
+### that occurred, alongside their corresponding outcome/score and line.
+#' Scrapes a listing of games that occured on a given date, alongside outcome/score and money-line.
+#'
+#' Given a date, this function attempts to scrape game outcome information along with a money-line.
+#' @param date A date object.
+#' @keywords scrape, mlb
+#' @export
+#' @examples
+#' scrapeCoverLinesMLB('2010-06-15')
+scrapeCoverLinesMLB <- function(date) {
+    tryCatch(expr = {
+        pg <- paste0('http://www.covers.com/sports/MLB/matchups?selectedDate=', date)
+        lines.links <- htmlParse(pg) %>%
+            xpathSApply(., '//a/@href') %>%
+            grep('linehistory', ., value = T, ignore.case = T)
+        ### Here, we are careful with games that got postponed (i.e. scheduled to play but no outcome).
+        game.ids.with.scores <- htmlParse(pg) %>%
+            xpathSApply(., '//a/@href') %>%
+            grep('boxscore', ., value = T, ignore.case = T) %>%
+            gsub('^.*boxscore([0-9]+)\\.html$', '\\1', .)
+        game.ids.with.scores <- paste0('(', game.ids.with.scores, ')', collapse = '|')
+        lines.links <- Filter(function(x) grepl(game.ids.with.scores, x), lines.links)
+        games.stats <- readHTMLTable(pg)
+        teams <- sapply(games.stats, function(x)
+            x[, 1] %>% as.character  %>% paste(., collapse = '-'))
+        scores <- sapply(games.stats, function(x)
+            x[, 'R'] %>% as.character %>% paste(., collapse = '-'))
+        lines <- lapply(lines.links, scrapeLines) %>% rbindlist
[…]

### 2017-07-04 hangover/R/scrapeCoverLinesNBA.r

@@ -1,0 +1,72 @@
+#' For internal use only, to be used within scrapeCoverLinesNBA.
+#'
+#' For each game played, we scrape the outcome alongside a spread.
+#' @param link A url describing where the game-outcomes are stored online.
+#' @keywords scrape, game
+#' @export
+#' @examples
+#' 
+scrapeGame <- function(link) {
+    tryCatch(expr = {
+        text <- readLines(link, warn = F)
+        ### We manually extract the date-field from HTML text.
+        date <- grep("\\<Date[^(]", text, value = T) %>%
+            gsub('^.*- (.*)<br>', '\\1', .)
+        ### Determine the event-id from the link, such that we may scrape-lines.
+        event.id <- gsub('^.*boxscore([0-9]+).html$', '\\1', link)
+        lines <- paste0('http://www.covers.com/odds/linehistory.aspx?eventId=',
+                        event.id, '&sport=NBA')
+        ### Scrape lines, just grabbing one entry for now.
+        lines <- readHTMLTable(lines, header = T, which = 1)[1, ]  # <-- One entry.
+        data  <- data.frame(date = date, lines)
+        setnames(data, c('date', 'line.ts', 'line', 'ou'))
+        return(data)
+    }, error = function(e) {
+        Sys.sleep(1)
+        cat(paste0('Error when downloading from: ', link, '...trying again\n'))
+        scrapeGame(link)
+    })
+}
+
+#' Scrape game outcomes and spreads for a given date.
+#'
+#' Given a date, this function queries `covers.com` for game-outcomes alongside a spread.
+#' @param date A date object.
+#' @param rescrape A boolean indicating whether to re-scrape raw-data if it's already
+#' been downloaded.
+#' @param save.path A string inidcating path to save output files.
+#' @param ntries A ceiling on the number of attempts to fetch data.
+#' @keywords scrape, covers
+#' @export
+#' @examples
+#' scrapeCoverLinesNBA('2012-11-19')
+scrapeCoverLinesNBA <- function(date, rescrape = F, save.path = 'raw_data/covers/', ntries = 0) {
+    if (ntries > 5) return(NULL)
+    if (!rescrape && file.exists(paste0(save.path, date, '.csv')))
+        return(NULL)
+    cat(paste0("Now trying to download covers for date: ", date, "\n"))
+    ### Bring up a listing of games played on a particular day.
+    tryCatch(expr = {
[…]

### 2017-07-04 scrape_covers.r
Now using reusable functions from R package 'hangover'.

@@ -21,6 +21,7 @@
 ### Set up Workspace
 ##############################
 
+require(hangover)
 require(parallel)
 
 cl <- makeCluster(detectCores())
@@ -37,67 +38,10 @@ game.days <- Filter(function(x) x < Sys.Date() & x > '2010-11-25', game.days)
 game.days <- gsub('-0([0-9])-', '-\\1-', game.days)
 
 ##############################
-### Define Functions
-##############################
-scrapeGame <- function(link) {
-    tryCatch(expr = {
-        text <- readLines(link, warn = F)
-        ### We manually extract the date-field from HTML text.
-        date <- grep("\\<Date[^(]", text, value = T) %>%
-            gsub('^.*- (.*)<br>', '\\1', .)
-        ### Determine the event-id from the link, such that we may scrape-lines.
-        event.id <- gsub('^.*boxscore([0-9]+).html$', '\\1', link)
-        lines <- paste0('http://www.covers.com/odds/linehistory.aspx?eventId=',
-                        event.id, '&sport=NBA')
-        ### Scrape lines, just grabbing one entry for now.
-        lines <- readHTMLTable(lines, header = T, which = 1)[1, ]  # <-- One entry.
-        data  <- data.frame(date = date, lines)
-        setnames(data, c('date', 'line.ts', 'line', 'ou'))
-        return(data)
-    }, error = function(e) {
-        Sys.sleep(1)
-        cat(paste0('Error when downloading from: ', link, '...trying again\n'))
-        scrapeGame(link)
-    })
-}
-
-scrapeCoverLines <- function(date, rescrape = F, save.path = 'raw_data/covers/', ntries = 0) {
-    if (ntries > 5) return(NULL)
-    if (!rescrape && file.exists(paste0(save.path, date, '.csv')))
-        return(NULL)
-    cat(paste0("Now trying to download covers for date: ", date, "\n"))
-    ### Bring up a listing of games played on a particular day.
-    tryCatch(expr = {
-        pg <- paste0('http://www.covers.com/sports/NBA/matchups?selectedDate=', date)
-        doc <- htmlParse(pg)
-        ### Each entry in games.played is a link to a detailed score-report.
-        games.played <- xpathSApply(doc, "//a/@href") %>%
-            grep('boxscore', ., value = T)
-        ### Games.stats contains quarter-by-quarter scores, alongside team-names.
-        games.stats <- readHTMLTable(doc)
[…]

### 2017-07-04 scrape_covers_mlb.r

@@ -21,6 +21,8 @@
 ### Set up workspace
 ##############################
 
+require(hangover)
+
 require(data.table)
 require(magrittr)
 require(parallel)
@@ -29,50 +31,6 @@ load(file = 'tmp_data/game_days_mlb.RData')
 
 game.days <- Filter(function(x) x > as.Date('2011-01-01'),
                     x = game.days) %>% as.character
-
-##############################
-### Define functions
-##############################
-
-### Scrapes lines data for a particular day.
-### At the moment, just returns one betting observation.
-scrapeLines <- function(link) {
-    tryCatch(expr = {
-        lines <- readHTMLTable(link, which = 1)
-        if (is.null(lines))
-            return(data.frame("no", "lines", "data"))
-        else
-            head(lines, 1)
-    }, error = function(e) return(data.frame("no", "lines", "data")))
-}
-
-### For a particular date,this function fetches a listing of games
-### that occurred, alongside their corresponding outcome/score and line.
-scrapeListingOfGames <- function(date) {
-    tryCatch(expr = {
-        pg <- paste0('http://www.covers.com/sports/MLB/matchups?selectedDate=', date)
-        lines.links <- htmlParse(pg) %>%
-            xpathSApply(., '//a/@href') %>%
-            grep('linehistory', ., value = T, ignore.case = T)
-        ### Here, we are careful with games that got postponed (i.e. scheduled to play but no outcome).
-        game.ids.with.scores <- htmlParse(pg) %>%
-            xpathSApply(., '//a/@href') %>%
-            grep('boxscore', ., value = T, ignore.case = T) %>%
-            gsub('^.*boxscore([0-9]+)\\.html$', '\\1', .)
-        game.ids.with.scores <- paste0('(', game.ids.with.scores, ')', collapse = '|')
-        lines.links <- Filter(function(x) grepl(game.ids.with.scores, x), lines.links)
-        games.stats <- readHTMLTable(pg)
-        teams <- sapply(games.stats, function(x)
-            x[, 1] %>% as.character  %>% paste(., collapse = '-'))
-        scores <- sapply(games.stats, function(x)
-            x[, 'R'] %>% as.character %>% paste(., collapse = '-'))
-        lines <- lapply(lines.links, scrapeLines) %>% rbindlist
-        data <- data.frame(date = date, matchup = teams, lines, score = scores)
[…]

### 2017-07-04 init_model.rnw
Added markers to show why drinks and population won't work.

@@
 
+require(hangover)
 require(data.table)
@@
 points(x = ticks, y = tmp$log.population, pch = 18, col = 'orange', type = 'l')
+### Houston has large population, but isn't necessarily known for nightlife.
+segments(x0 = which(tmp$team == 'houston rockets'), 
+         x1 = which(tmp$team == 'houston rockets'),
+         y0 = -1,
+         y1 = tmp[team == 'houston rockets', log.population], col = 'gray70', lty = 'dashed')
+segments(x0 = which(tmp$team == 'houston rockets'), 
+         x1 = length(ticks),
+         y0 = tmp[team == 'houston rockets', log.population],
+         y1 = tmp[team == 'houston rockets', log.population], col = 'gray70', lty = 'dashed')
+abline(h = tmp[team == 'houston rockets', log.ndrinks], col = 'gray70', lty = 'dashed')
+### New orleans is infamous for party, but Population places it at the bottom of the index.
+segments(x0 = which(tmp$team == 'new orleans hornets'),
+         x1 = which(tmp$team == 'new orleans hornets'), 
+         y0 = -1, 
+         y1 = tmp[team == 'new orleans hornets', log.population],col = 'gray70', lty = 'dashed')
+segments(x0 = 0, x1 = which(tmp$team == 'new orleans hornets'),
+         y0 = tmp[team == 'new orleans hornets', log.population],
+         y1 = tmp[team == 'new orleans hornets', log.population], col = 'gray70', lty = 'dashed')
 axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>%

### 2017-07-04 writing/draft.tex
Added a couple anecdotal citations.

@@
 34 percent on Sunday afternoon games in the 2012-13 season.
-Presumably, this was because such games followed a long weekend of partying.\citep{princeofthecity}
+Presumably, this was because such games followed a long weekend of partying.\citep{princeofthecity} It has been suggested that because
+NBA games are played in the evenings, athletes can get away with late-nights.\citep{drodmanbleacherreport}
 
@@
 Finally, there was no methodology used behind the selection of party cities: the authors first define ad-hoc five ``party cities'', and when that did not yield a significant result they estimate fixed-effects for each city.
-The author does carry out a binomial test of how many home-game covers vs.
-away-game covers were met, and concludes with weak statistical evidence that teams
-systematically play worse on Sunday afternoons when on the road, regardless of city, as measured by
-their ability to meet bookmakers' expectations.
+The author does carry out a binomial test, using 284 games, of how many home-game covers vs.
+away-game covers were met; the author concludes with weak statistical evidence that teams  systematically play worse on Sunday afternoons when on the road,
+regardless of city, as measured by their ability to meet bookmakers' expectations.
 
@@
 We also found that athletes tend to party with other A-list celebrities,
-particularly musicians. This motivated a continuous measure of nightlife, based on a simple count of the number of musicians in the metropolitan statistical area (MSA) 
+particularly musicians.\citep{jbutlerdwade} After winning the NBA finals
+in 2016, the Cleveland Cavaliers were found celebrating in nightclubs.\citep{xsclub} 
+This motivated a continuous measure of nightlife, based on a simple count of the number of musicians in the metropolitan statistical area (MSA) 
 in which the stadium resides: if the last game was played at most 24-hours ago, use the log of the total number

### 2017-07-04 cardinal_analytx_presentation/nightlife_and_sports.tex
Added anecdotal citations.

@@
   \begin{block}{Do pro athletes party to the detriment of their performance?}
-    Anecdotally, yes.\footnote{Cestone ('17), Price ('15), Concepcion ('15), Miller ('15), Campbell ('15).}
+    Anecdotally, yes.\footnote{Duffy ('17), Cestone ('17), Duffy ('16), 
+      Price ('15), Concepcion ('15), Miller ('15), Campbell ('15), Ball ('12)}
 
@@
   \begin{itemize}     \item distance traveled between games,
-    \item final direction (bearing) traveling from $a \to b$, and
+    \item final direction (bearing) traveling from $a \leadsto b$, and
     \item direction of travel (in east-west degrees) by taking sine.

### 2017-07-05 cardinal_analytx_presentation/nightlife_and_sports.tex
Added citations and clarifying details.

@@
   \begin{block}{Do pro athletes party to the detriment of their performance?}
-    Anecdotally, yes.\footnote{Duffy ('17), Cestone ('17), Duffy ('16), 
-      Price ('15), Concepcion ('15), Miller ('15), Campbell ('15), Ball ('12)}
+    Anecdotally, they do party.\footnote{Duffy ('17), Cestone ('17), Duffy ('16), 
+      Price ('15), Concepcion ('15), Miller ('15), Campbell ('15), Ball ('12).}
 
-    It's not unreasonable to think that what happens off the court affects outcome, but!
+    \vspace{12pt}It's not unreasonable to think that what happens off the court affects game-outcomes, but!
     \begin{itemize}       \item Can we even measure this?
@@
 
-    Knowing the causal relation between partying and next-day performance might inform
+    \vspace{12pt}Even aside from bettors, knowing the causal relation between partying and next-day performance might inform
     \begin{itemize}       
-      \item How coaches set team-rules, 
-      \item How travel schedules are determined.
-      \item Which players to bench, and when.
+      \item how coaches set team-rules, 
+      \item how travel schedules are determined, as well as
+      \item which players to bench, and when.
       \end{itemize}
@@
 
-\begin{frame}   \frametitle{Is the problem hard?}
-  Yes, in general finding arbitrage opportunities is hard.
+\begin{frame}   
+  \frametitle{Is the problem hard?}
 
-  Further, it's not obvious that we will find an effect.
+  \begin{block}{Finding causal-mechanisms which significantly determine game-outcomes is difficult.}       It's not obvious whether we will find an effect.
+
+      \vspace{12pt}After all, we can't even observe if players party, let alone when.
+
+      Even if we could, can we exploit this in the marketplace?   \end{block}
 
@@
 \begin{frame}
-  \frametitle{Previous Research - Player shooting average}     Players under-perform proportion of shots made on Sundays.\footnote{Chase, ('15).}
-    \begin{itemize}       \item 1,092 of 2,035 NBA players who took $\geq 1$ shot on Sundays under-perform relative to their own season-long average; $p=0.001$.
+  \frametitle{Previous Research - NBA player shooting average}     
+  NBA Players under-perform proportion of shots made on Sundays.\footnote{Chase, ('15).}
+    \begin{itemize}       
+    \item 1,092 of 2,035 NBA players who took $\geq 1$ shot on Sundays under-perform 
+      relative to their own season-long average; $p=0.001$.
       \item Effect size is not considered.
-      \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, should that be considered?}     \end{itemize}
+      \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, is that relevant?}     \end{itemize}
 \end{frame}
 
-\begin{frame}   \frametitle{Previous Research - Team's ability to meet spread}
+\begin{frame}   \frametitle{Prev. Research - Likelihood of meeting the spread in NBA}
   Teams under-perform bookmakers expectations on Sunday afternoons when on the road.\footnote{Ezekowitz, ('14).}
+\vspace{12pt}
     \begin{itemize}       
     \item Only 129 out of 284 games did the away team meet the spread, 
-      using 2008-14 data; $p=0.1089$.
+      using 2006-14 data; $p=0.1089$.
     \item Not quite significant at the 90\% confidence level.     
+    \item If we bet \$100 on each Sunday game, we'd only profit \$1,310 over 8 years.\footnote{$(284 - 129) \times 100 + 129 \times (-110) = 1,310$.} Not exciting.
     \end{itemize}
@@
   \begin{block}{We can't observe whether players party.}
-    Instead, we consider a quasi-experiment. Motivating questions:
+    Instead, we will consider a quasi-experiment. Motivating questions:
     \begin{itemize}       \item Do teams play worse the day after visiting a ``party city''?
@@
   \end{block}
+
+  \vspace{12pt}We define ``play worse'' relative to bookmaker/public expectations.
 \end{frame}
@@
 \begin{frame}
-  \frametitle{How can we measure unexpected changes in performance?}     \begin{block}{Bookmaker spreads (NBA)}     
+  \frametitle{How can we measure unexpected changes in performance?}     \begin{block}{Bookmaker point-spreads (NBA)}     
     The expected score differential between two teams.
 
-    Conditions on observed performance, and future expectations.
+    Conditions on every conceivable element of the game:
+    \begin{itemize}     \item strengths and weaknesses against particular opponents,
+    \item momentum,
+    \item injuries, and
+    \item even possibly a weather forecast.     \end{itemize}
 
-    A team is expected to meet the spread $\frac{1}{2}$ the time.\footnote{Dubner and Levitt, 2006}
+    A team is expected to meet the spread $\frac{1}{2}$ the time.\footnote{Dubner and Levitt, ('06). Losers pay 10\% fee on top of wagers; a balanced book guarantees 5\% gain for the house-bookmaker.}
   \end{block}
@@
 
+\begin{frame}   \frametitle{Identification}
+    If next day opponent is uncorrelated with visitation to a party city,
+    then looking at the \emph{actual outcome} of the game $\leadsto$ identification.
+
+\vspace{15pt}
+
+    Weaker, still: if odds of beating the spread uncorrelated with previous game location
+    $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}     \end{frame}
+
 \begin{frame}   \frametitle{Quasi Experiment}
-  \begin{block}{Consider games played back-to-back within 24 hours.}     
-    If next day opponent is uncorrelated with visitation to a party city $\leadsto$ identification.
+  Consider games played back-to-back within 24 hours.
 
-    Weaker, still: if odds of beating the spread uncorrelated \\
-    $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}    \end{block}  
-\end{frame}
+  \vspace{12pt}
+  Examine whether spread is met as a function of nightlife index for lagged game location.
+
+  \vspace{12pt} (Alcohol and loss of sleep are well correlated with nightlife; both have been shown to have negative effects on sports performance.)\footnote{Shirreffs and Maughan ('06), Barnes ('14), Pilcher and Huffcutt ('96).} \end{frame}
 
 \begin{frame}   \frametitle{Sports: NBA and MLB}
-  \begin{block}{We perform the analysis for both NBA and MLB}     If we yield a causal model, we can realize a betting schema.   \end{block} \end{frame}
+  \begin{block}{We perform the analysis for both NBA and MLB}     
+    Observe a team traveling to a city with active nightlife.
+    \vspace{12pt}
+    \begin{itemize}       \item For NBA teams playing another game in the next 24 hours, are they less likely to meet the points-spread?
+      \item For MLB teams, after conditioning on the odds of winning, are they less likely to beat their opponent?     \end{itemize}
+    \vspace{12pt}\begin{block}{If we yield a causal model, we may realize a betting schema.}     \end{block}
+  \end{block} 
+\end{frame}
 
@@
 
-  We query \href{http://www.basketball-reference.com/leagues/}{basketball-reference.com}
+  \vspace{12pt}We query \href{http://www.basketball-reference.com/leagues/}{basketball-reference.com}
   for a listing of game-days, and then 
   \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{covers.com} for 
-  game outcomes alongside a point-spread.
+  game-outcomes alongside a point-spread.
 
+  \vspace{12pt}  
   \begin{itemize}     \item Missing the first couple months of 2010-11 season.
@@
 
-  We query \href{http://www.baseball-reference.com/leagues/}{baseball-reference.com}
-  for a listing of game days, and again look to Covers.com for game outcomes and money-lines.
+  \vspace{12pt}We query \href{http://www.baseball-reference.com/leagues/}{baseball-reference.com}
+  for a listing of game days, and again look to covers.com for game-outcomes and money-lines.
 
+  \vspace{12pt}
   \begin{itemize}     \item We have on average 2,100 games per season.
-    \item (Missing observations due to broken links?)
+    \item (Missing observations due to broken links, postponed games?)
     \item Total of 12,709 games in our dataset.    \end{itemize} \end{frame}
@@
 
-  All teams are in the US with the exception of 
-  \begin{itemize}     \item Toronto Raptors (NBA)
+  \vspace{12pt}  All teams are in the US with the exception of 
+  \vspace{12pt}\begin{itemize}     \item Toronto Raptors (NBA)
     \item Toronto Blue Jays (MLB)   \end{itemize}
 
-  These two teams are excluded from our analysis.\footnote{We don't expect Toronto to be a huge party city.} \end{frame}
+  \vspace{12pt}These two teams are excluded from our analysis.\footnote{We don't expect Toronto to be a huge party city.} \end{frame}
 
@@
   For NBA, we also collect minutes data, which describes the number of 
-  \begin{itemize}     \item Free-throws
+  \vspace{12pt}
+  \begin{itemize}     
+    \item Free-throws
     \item Field goals
-    \item Three-pointers   \end{itemize}
-  attempted and made by each player in each game. For each season and each team,
-  we have nearly complete minutes data; 85\% of our season-teams' have no more than
-  3 missing games. \end{frame}
+    \item Three-pointers   
+  \end{itemize}
+  \vspace{12pt}attempted and made by each player in each game. For each season and each team,
+  we have nearly complete minutes data; 85\% of our season-teams have no more than 3 missing games. \end{frame}
 
@@
 
-  \begin{itemize}     
+  \vspace{12pt}\begin{itemize}     
     \item Home team effect
@@
 
-  The benefit of using spread data is that it controls for salient factors which affect game outcomes.
+  \vspace{12pt}The benefit of using spread data is that it controls for salient factors which affect game-outcomes.
   The downside is that we don't know exactly what it controls for. \end{frame}
@@
   We add features which describe
+  \vspace{12pt}  
   \begin{itemize}     
     \item Number of days since the last game (i.e. rest-time),
-    \item Lagged changes in possession,
+    \item Lagged changes in possession,\footnote{A crude measure of fatigue incurred from last game.}
     \item Lagged game location, and
@@
   \end{itemize}
-  We seek to construct a model which accounts for travel-fatigue and jet-lag.
+  \vspace{12pt}We seek to construct a model which accounts for travel-fatigue and jet-lag.
   For this, we need a little bit more. \end{frame}
@@
 
-  In order to account for jet-lag, we need to be mindful of the direction of travel. We calculate
-  \begin{itemize}     \item distance traveled between games,
+  \vspace{12pt}In order to account for jet-lag, we need to be mindful of the direction of travel. We calculate
+  \vspace{12pt}\begin{itemize}     \item distance traveled between games,
     \item final direction (bearing) traveling from $a \leadsto b$, and
@@
   \end{itemize}
-  Finally, to account for jetlag we interact this east-west measure with travel distance.\footnote{Lallensack ('17)} \end{frame}
+  \vspace{12pt}Finally, to account for jetlag we interact this east-west measure with travel distance.\footnote{Lallensack ('17)} \end{frame}
 
@@
 \begin{frame}   \frametitle{Discrete Indicator}
-  \begin{block}{Anecdotally, LA and NY appear in news articles frequently.}     This motivates a binary indicator for party when either of these two cities have been visited within 24-hours of another game.
+  \begin{block}{Anecdotally, LA and NY appear in news articles frequently.}     \vspace{12pt}This motivates a binary indicator for party when either of these two cities have been visited within 24-hours of another game.
+\vspace{12pt}
     \[
@@
 
-\begin{frame}   \frametitle{Placebo Test: what if we let players rest?}
+\begin{frame}   \frametitle{Placebo Test (discrete indicator): what if players rest?}
   \centering
   \tiny{
-  \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   party placebo & 0.072 \\    & (0.082) \\    & \\   lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   east-west travel direction & $-$0.099 \\    & (0.232) \\    & \\   number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   time of game during day & 0.001 \\    & (0.013) \\    & \\   home team effect & $-$0.009 \\    & (0.043) \\    & \\   logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.070 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.702 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.072 \\    & (0.082) \\    & \\   Lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   Logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.099 \\    & (0.231) \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   Time of game during day & 0.001 \\    & (0.013) \\    & \\   Home team effect & $-$0.009 \\    & (0.043) \\    & \\   Logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.069 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.701 \\  Akaike Inf.
 Crit.
-& 13,631.400 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-} \end{frame}
+& 13,631.400 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  } \end{frame}
 
@@
   The same analysis can be replicated for MLB. Notable differences:
-  \begin{itemize}     \item Money-lines are used instead of point-spreads. 
+  \vspace{12pt}  
+  \begin{itemize}     
+    \item Money-lines are used instead of point-spreads. 
       So instead of predicting likelihood of meeting the spread,
@@
 
-  \begin{itemize}     \item Avoid using black-box feature
-    \item Pinpoint more precisely what elements of game-play are affected by nightlife visitation.   \end{itemize}
+  \vspace{12pt}
+  \begin{itemize}     
+    \item Avoid using black-box feature
+    \item Pinpoint more precisely what elements of game-play are affected by nightlife visitation.   
+\end{itemize}
 
-  Anecdotally, rebounds are considered a ``hustle'' statistic, in the sense that players
+\vspace{12pt}  Anecdotally, rebounds are considered a ``hustle'' statistic, in the sense that players
   who work harder on the court and expend energy boxing out other players will have higher
@@
 
-  Common knowledge suggests that overall defense of a team is strongly influenced by individual player effort.
+  \vspace{12pt}Common knowledge suggests that overall defense of a team is strongly influenced by individual player effort.
 
-  Is it the case the hangover effects are more prevalent on defensive aspects of gameplay? \end{frame}
+  \vspace{12pt}Is it the case the hangover effects are more prevalent on defensive aspects of gameplay? \end{frame}
 
@@
 
-\begin{frame}   \frametitle{Future Research}
+\begin{frame}   \frametitle{Future Research - NBA}
   Within the NBA
-  \begin{itemize}     \item Are teams more likely to party after a win or a loss?
+  \begin{itemize}     
+    \item Are teams more likely to party after a win or a       loss?
     \item Do teams who aren't going to make the playoffs party more?
@@
   \end{itemize}
-  A preliminary and ad-hoc analysis shows that injury risk increases as a function of party.
-
+  \vspace{12pt}  A preliminary and ad-hoc analysis shows that injury risk increases as a function of party.\footnote{We look for instances where players are put on the bench and stop playing (at least several) games as a function of our nightlife indicator.}
   This might inform coaches which players to bench and when.
 
-  Within the MLB, we can look at statistics such as batting averages or general fielding statistics. \end{frame}
+\end{frame}
 
-\begin{frame}   \frametitle{Future Research (General)}
-  \begin{itemize}     \item Possibly prolonged effects; recall our magic 24 hour number.
+\begin{frame}   \frametitle{Future Research (MLB and General)}
+  \begin{itemize}     
+  \item Within the MLB, we can look at statistics such as batting averages or general fielding statistics. 
+    \item Possibly prolonged effects; recall our magic 24 hour number.
     \item European soccer leagues.   \end{itemize} \end{frame}

### 2017-07-05 writing/draft.tex
Added a couple more anecdotal references.

@@
 the same is \emph{not true} when flying from east to west, in which case a morning flight is reasonable.
+
 The fact that there is some correlation between last game location and next day opponent, even if only in west coast cities, increases the motivation for conditioning on spread, which may help account for systematic biases to the extent that some of these next-day opponents are particularly (not) skilled.
+In fact, using point-spread data we may weaken our assumption: we only need
+likelihood of meeting the spread to be uncorrelated with last game location
+in order to have identification in our variable of interest.
 
@@
 % mp
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   party placebo & 0.072 \\    & (0.082) \\    & \\   lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   east-west travel direction & $-$0.099 \\    & (0.232) \\    & \\   number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   time of game during day & 0.001 \\    & (0.013) \\    & \\   home team effect & $-$0.009 \\    & (0.043) \\    & \\   logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.070 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.702 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.072 \\    & (0.082) \\    & \\   Lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   Logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.099 \\    & (0.231) \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   Time of game during day & 0.001 \\    & (0.013) \\    & \\   Home team effect & $-$0.009 \\    & (0.043) \\    & \\   Logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.069 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.701 \\  Akaike Inf.
 Crit.
 & 13,631.400 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+
 \subsection{Possible confounders or potential issues}

### 2017-07-05 init_model.rnw
Now using a density overlay on histogram of season-teams meet the spread plot.

@@
 
-stargazer(mp, covariate.labels = c('party placebo', 'lag changes in posession', 'logged travel distance', 'east-west travel direction', 'number hours rest time', 'time of game during day', 'home team effect', 'logged travel distance * east-west', 'Constant'),
+stargazer(mp, covariate.labels = c('Party placebo', 'Lag changes in posession', 
+                                   'Logged travel distance', 'East-west travel direction', 
+                                   'Number hours rest time', 'Time of game during day', 
+                                   'Home team effect', 'Logged travel distance * east-west', 
+                                   'Constant'),
           dep.var.labels = 'Meet the Spread (NBA)')
@@
 par(mfcol = c(2, 1))
-nba.lines[, list(meet.spread = mean(outcome == 'W')), 
-          by = list(season, team)][, hist(meet.spread, 
-                                          main = paste('How often do teams meet the spread? (NBA)', 
+tmp <- nba.lines[, list(meet.spread = mean(outcome == 'W')), 
+          by = list(season, team)]
+hist(tmp$meet.spread, prob = T, main = paste('How often do teams meet the spread? (NBA)', 
                                              'By season-team, 2010-11 through 2016-17',
                                              sep = '\n'),
-                                          xlab = 'Meet the Spread', breaks = 25)]
+     xlab = 'Meet the Spread', breaks = 25)
+lines(density(tmp$meet.spread, adjust = 2), col = 'orange', lty = 'dashed')
 ### There should be 'no information' left over after conditioning on odds.
@@
                       'successful prediction is a coin-flip', sep = '\n'))
-abline(a = 0, b = 0, col = 'red', lty = 'dashed')
+abline(a = 0, b = 0, col = 'orange', lty = 'dashed')
 dev.off()

### 2017-07-05 cardinal_analytx_presentation/nightlife_and_sports.tex
Formatting of title-page.

@@
 
+ % \usetheme{Dresden} % \usecolortheme{beaver} % \setbeamercolor{itemize item}{fg=darkred!80!black}
+
 \setbeamertemplate{navigation symbols}{}
 
-\usetikzlibrary{shapes,arrows} \setbeamerfont{author}{size=\Huge} \setbeamerfont{institute}{size=\normalsize\itshape} \setbeamerfont{title}{size=\fontsize{30}{36}\bfseries} \setbeamerfont{subtitle}{size=\Large\normalfont\slshape}  \setbeamertemplate{title page} { \begin{tikzpicture}[remember picture,overlay] \fill[orange]   ([yshift=15pt]current page.west) rectangle (current page.south east); \node[anchor=east]    at ([yshift=-50pt]current page.north east) (author)   {\parbox[t]{.6\paperwidth} {\raggedleft      \usebeamerfont{author}\textcolor{orange} {     \textpdfrender{     TextRenderingMode=FillStroke,     FillColor=orange,     LineWidth=.1ex,     }{\insertauthor}}}};
+\usetikzlibrary{shapes,arrows} \setbeamerfont{author}{size=\Huge} 
+\setbeamerfont{institute}{size=\normalsize\itshape} \setbeamerfont{title}{size=\fontsize{30}{36}\bfseries} \setbeamerfont{subtitle}{size=\Large\normalfont\slshape}  \setbeamertemplate{title page} {
+  \begin{tikzpicture}[remember picture,overlay]     \fill[orange]   ([yshift=15pt]current page.west) rectangle (current page.south     east); \node[anchor=east]    at ([yshift=-50pt]current page.north east) (author)       {\parbox[t]{.6\paperwidth} {\raggedleft              \usebeamerfont{author}\textcolor{orange} {     \textpdfrender{                 TextRenderingMode=FillStroke,     FillColor=orange,     LineWidth=.1ex,               }{\insertauthor}}}};
     \node[anchor=north east]
@@
       \parbox[t]{.6\paperwidth}{\raggedleft\usebeamerfont{subtitle}\textcolor{black}{\insertsubtitle}}};  
-\end{tikzpicture} }  \author{Andreas Santucci} \institute{Stanford Teaching Fellow \\ Computational Math, Data Science} \title{Effects of nightlife on sports} \subtitle{Cardinal Analytx Interview} % \titlegraphic{\includegraphics[width=2cm]{ctanlion}}
+  \end{tikzpicture}
+}
+
+\author{Andreas Santucci} \institute{Stanford Teaching Fellow \\ Computational Math, Data Science} \title{Effects of nightlife on sports} \subtitle{Cardinal Analytx Interview} % \titlegraphic{\includegraphics[width=2cm]{ctanlion}}
  \begin{document}  \begin{frame} \maketitle \end{frame}

### 2017-07-06 cardinal_analytx_presentation/nightlife_and_sports.tex
Changed frametitle's from Camel Case to proper punctuation. Changed ordering of 'Whats our Solution'.

@@
 
-\author{Andreas Santucci} \institute{Stanford Teaching Fellow \\ Computational Math, Data Science} \title{Effects of nightlife on sports} \subtitle{Cardinal Analytx Interview} % \titlegraphic{\includegraphics[width=2cm]{ctanlion}}
+\author{Andreas Santucci} \institute{Stanford Teaching Fellow \\ Computational Math, Data Science} \title{Effects of nightlife on sports} \subtitle{``Hangover Effects''} % \titlegraphic{\includegraphics[width=2cm]{ctanlion}}
  \begin{document}  \begin{frame} \maketitle \end{frame}
@@
     \vspace{12pt}It's not unreasonable to think that what happens off the court affects game-outcomes, but!
-    \begin{itemize}       \item Can we even measure this?
-      \item Are there systematic tendencies?     \end{itemize}
-  \end{block} \end{frame}
+    \begin{itemize}       
+      \item Can we even measure a hangover effect?
+      \item Are there systematic tendencies?
+      \end{itemize}
+  \end{block} 
+\end{frame}
+
+% What is our solution?
+
+\begin{frame}   
+  \frametitle{Whether players actually party is a latent variable}
+  \begin{block}{We can't observe whether players party.}
+    Instead, we will consider a quasi-experiment where players are exposed to nightlife. 
+    Motivating questions:
+    \vspace{12pt} 
+    \begin{itemize}       \item Do teams play worse the day after visiting a ``party city''?
+      \item How do we even determine whether a city has nightlife?    
+      \item What does it mean to ``play worse''? \end{itemize}
+  \end{block}
+
+  \vspace{12pt}We define ``play worse'' relative to bookmaker/public expectations.
+\end{frame}
 
@@
 
-      \vspace{12pt}After all, we can't even observe if players party, let alone when.
+      \vspace{12pt}After all, we can't reliably observe if players party, let alone when.
 
@@
 \begin{frame}
-  \frametitle{Previous Research - NBA player shooting average}     
+  \frametitle{Previous research - NBA player shooting average}     
   NBA Players under-perform proportion of shots made on Sundays.\footnote{Chase, ('15).}
@@
 
-\begin{frame}   \frametitle{Prev. Research - Likelihood of meeting the spread in NBA}
+\begin{frame}   \frametitle{Prev. research - likelihood of meeting the spread in NBA}
   Teams under-perform bookmakers expectations on Sunday afternoons when on the road.\footnote{Ezekowitz, ('14).}
@@
     \item Not quite significant at the 90\% confidence level.     
-    \item If we bet \$100 on each Sunday game, we'd only profit \$1,310 over 8 years.\footnote{$(284 - 129) \times 100 + 129 \times (-110) = 1,310$.} Not exciting.
+    \item If we bet \$100 on each Sunday game, we'd only profit \$1,310.\footnote{$(284 - 129) \times 100 + 129 \times (-110) = 1,310$.} 
     \end{itemize}
-  \end{frame}
-
-\begin{frame}   
-  \frametitle{What is our solution?}
-  \begin{block}{We can't observe whether players party.}
-    Instead, we will consider a quasi-experiment. Motivating questions:
-    \begin{itemize}       \item Do teams play worse the day after visiting a ``party city''?
-      \item How do we even determine whether a city has nightlife?    
-      \item What does it mean to ``play worse''? \end{itemize}
-  \end{block}
-
-  \vspace{12pt}We define ``play worse'' relative to bookmaker/public expectations.
   \end{frame}
@@
 
-    Conditions on every conceivable element of the game:
+    Conditions on (almost) every conceivable element of the game:
     \begin{itemize}     \item strengths and weaknesses against particular opponents,
@@
 
-\begin{frame}   \frametitle{Quasi Experiment}
+\begin{frame}   \frametitle{Our solution: a quasi-experiment}
   Consider games played back-to-back within 24 hours.
@@
 
-\begin{frame}   \frametitle{NBA Data: 2010-11 through 2016-17 season}
+\begin{frame}   \frametitle{NBA data: 2010-11 through 2016-17 season}
   \begin{itemize}     \item Each team plays 82 games per season.
@@
 
-\begin{frame}   \frametitle{MLB Data: 2011 through 2017 season}
+\begin{frame}   \frametitle{MLB data: 2011 through 2017 season}
   \begin{itemize}     \item Each team plays 162 games per season.
@@
 
-\begin{frame}   \frametitle{Bureau of Labor Statistics Data: 2010-2016}
+\begin{frame}   \frametitle{Bureau of Labor Statistics data: 2010-2016}
   We look toward BLS data which describes the number of business establishments 
@@
 \begin{frame}
-  \frametitle{Possible Confounders}
+  \frametitle{Possible confounders}
   What might influence game-outcomes apart from player-ability?
@@
 
-\begin{frame}   \frametitle{Feature Creation}
+\begin{frame}   \frametitle{Feature creation}
   We add features which describe
@@
 
-\begin{frame}   \frametitle{Accounting for Jetlag}
+\begin{frame}   \frametitle{Accounting for jetlag}
   Prolonged air travel negatively effects sporting performance.\footnote{Lee and Galvez ('12).}
@@
 \begin{frame}   
-  \frametitle{Travel Distance Comparison}
+  \frametitle{Travel distance comparison}
   \centering \includegraphics[scale=0.25]{../writing/travel_dist_density_by_party} 
@@
 
-\begin{frame}   \frametitle{Continuous Measure}
+\begin{frame}   \frametitle{Continuous measure}
   \begin{block}{A-list celebrities tend to party with other celebrities.}     Idea: proxy nightlife using number of musicians in previous location
@@
 \begin{frame}   
-  \frametitle{Continuous Measure of Party (NBA)}
+  \frametitle{Continuous measure of party (NBA)}
   \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location} 
@@
 
-\begin{frame}   \frametitle{Placebo Test (discrete indicator): what if players rest?}
+\begin{frame}   \frametitle{Placebo test (discrete indicator): what if players rest?}
   \centering
@@
 \begin{frame}   
-  \frametitle{Continuous Measure of Party (MLB)}
+  \frametitle{Continuous measure of party (MLB)}
   \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location_mlb}   
@@
 
-\begin{frame}   \frametitle{Returning to NBA: Player Performance}
+\begin{frame}   \frametitle{Returning to NBA: player performance}
   We additionally look toward actual performance metrics, instead of relying on point-spread data.
@@
 
-\begin{frame}   \frametitle{Future Research - NBA}
+\begin{frame}   \frametitle{Future research - NBA}
   Within the NBA
@@
 
-\begin{frame}   \frametitle{Future Research (MLB and General)}
+\begin{frame}   \frametitle{Future research (MLB and general)}
   \begin{itemize}     
@@
 
-\begin{frame}   \frametitle{Why \# Drinking Establishments or Population won't work}
+\begin{frame}   \frametitle{Why \# drinking establishments or population won't work}
   \includegraphics[scale=0.3]{../writing/why_drinks_or_population_dont_work} \end{frame}

### 2017-07-06 cardinal_analytx_presentation/nightlife_and_sports.tex
The first couple slides are *starting* to look more polished.

@@
 \begin{frame}   \frametitle{What is the problem?}
-  \begin{block}{Do pro athletes party to the detriment of their performance?}
-    Anecdotally, they do party.\footnote{Duffy ('17), Cestone ('17), Duffy ('16), 
-      Price ('15), Concepcion ('15), Miller ('15), Campbell ('15), Ball ('12).}
-
-    \vspace{12pt}It's not unreasonable to think that what happens off the court affects game-outcomes, but!
+  \begin{block}{Do pro athletes party to the detriment of their performance?}     It's not unreasonable to think that what happens off the court affects game-outcomes, but!
     \begin{itemize}       
-      \item Can we even measure a hangover effect?
+      \item Can we even measure these phenomena?
       \item Are there systematic tendencies?
-      \end{itemize}
-  \end{block} 
+    \end{itemize}       \end{block}
 \end{frame}
@@
 \begin{frame}   
-  \frametitle{Whether players actually party is a latent variable}
-  \begin{block}{We can't observe whether players party.}
-    Instead, we will consider a quasi-experiment where players are exposed to nightlife. 
-    Motivating questions:
-    \vspace{12pt} 
-    \begin{itemize}       \item Do teams play worse the day after visiting a ``party city''?
-      \item How do we even determine whether a city has nightlife?    
-      \item What does it mean to ``play worse''? \end{itemize}
+  \frametitle{Why is this problem hard?}
+  \begin{block}{Partying is a latent variable.}
+    Instead, we proxy for the possibility of athletes partying by looking at 
+    play following games in party cities. Our question becomes...
   \end{block}
+  \vspace{12pt}\begin{block}{Do teams exhibit a decline in performance the day following a game in a city with active nightlife?}     I.e. we are concerned with ``hangover effects''.
+  \end{block}
+  \vspace{12pt}\begin{block}{Part of the question is determining a measure for nightlife. }
+    We colloquially call cities with high levels of nightlife ``party cities''.   \end{block}
+\end{frame}
 
-  \vspace{12pt}We define ``play worse'' relative to bookmaker/public expectations.
-\end{frame}
+  %   \vspace{12pt} 
+  %   \begin{itemize}       \item Do teams play worse the day after visiting a ``party city''?
+  %     \item How do we even determine whether a city has nightlife?    
+  %     \item What does it mean to ``play worse''? \end{itemize}
+
+  % \vspace{12pt}We define ``play worse'' relative to bookmaker/public expectations.
 
@@
 
-    Apparently, a lot of people.
+    Apparently, a lot of people. Aside from bettors...
+  \end{block}
 
-    \vspace{12pt}Even aside from bettors, knowing the causal relation between partying and next-day performance might inform
+  \vspace{12pt}\begin{block}{The causal relationship might inform more optimal strategy.}     For example,
+    % Even aside from bettors, knowing the causal relation between % partying and next-day performance might inform
     \begin{itemize}       
       \item how coaches set team-rules, 
-      \item how travel schedules are determined, as well as
-      \item which players to bench, and when.
+      \item how managers determine travel schedules, as well as
+      \item which players to bench, and when.\footnote{Are players more likely to get injured the next day as a result of partying?}
     \end{itemize}
@@
 
-\begin{frame}   
-  \frametitle{Is the problem hard?}
+% \begin{frame}   
+%   \frametitle{Is the problem hard?}
 
-  \begin{block}{Finding causal-mechanisms which significantly determine game-outcomes is difficult.}       It's not obvious whether we will find an effect.
+%   \begin{block}{Finding causal-mechanisms which significantly determine % game-outcomes is difficult.}       It's not obvious whether we will find an % effect.
 
-      \vspace{12pt}After all, we can't reliably observe if players party, let alone when.
+%       \vspace{12pt}After all, we can't reliably observe if players party, let % alone when.
 
-      Even if we could, can we exploit this in the marketplace?   \end{block}
+%       Even if we could, can we exploit this in the marketplace?   \end{block}
 
-\end{frame}
-
-\begin{frame}
-  \frametitle{Previous research - NBA player shooting average}     
-  NBA Players under-perform proportion of shots made on Sundays.\footnote{Chase, ('15).}
-    \begin{itemize}       
-    \item 1,092 of 2,035 NBA players who took $\geq 1$ shot on Sundays under-perform 
-      relative to their own season-long average; $p=0.001$.
-      \item Effect size is not considered.
-      \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, is that relevant?}     \end{itemize}
-\end{frame}
-
-\begin{frame}   \frametitle{Prev. research - likelihood of meeting the spread in NBA}
-  Teams under-perform bookmakers expectations on Sunday afternoons when on the road.\footnote{Ezekowitz, ('14).}
-\vspace{12pt}
-    \begin{itemize}       
-    \item Only 129 out of 284 games did the away team meet the spread, 
-      using 2006-14 data; $p=0.1089$.
-    \item Not quite significant at the 90\% confidence level.     
-    \item If we bet \$100 on each Sunday game, we'd only profit \$1,310.\footnote{$(284 - 129) \times 100 + 129 \times (-110) = 1,310$.} 
-    \end{itemize}
-  \end{frame}
+% \end{frame}
 
@@
   \end{block}
-  \begin{block}{Bookmaker money-lines (MLB)}     The odds that a team will beat their opponent.
+  \vspace{12pt}\begin{block}{Bookmaker money-lines (MLB)}     The odds that a team will beat their opponent.
 
@@
 \begin{frame}
-  \centering   \includegraphics[scale=0.55]{../writing/meet_the_spread} \end{frame}
+  \centering   \includegraphics[scale=0.55]{../writing/meet_the_spread} 
+\end{frame}
 
-\begin{frame}   \frametitle{Identification}
-    If next day opponent is uncorrelated with visitation to a party city,
+\begin{frame}   \frametitle{Our solution: a quasi-experiment.}
+  \begin{block}{Systematic deviations represent unexplained mechanisms.}     We construct a model which attempts to estimate the causal effect of visiting a ``party city''
+    on subsequent day performance, as measured by the odds of meeting the spread.       \end{block}
+
+  \begin{block}{}     We only consider the hangover effect on games played back-to-back within 24 hours of each other.   \end{block} \end{frame}
+
+\begin{frame}   \frametitle{Identifying variation}
+  \begin{block}{Explain performance as a function of prev. day's nightlife index.}     If next day opponent is uncorrelated with visitation to a party city,
     then looking at the \emph{actual outcome} of the game $\leadsto$ identification.
@@
     Weaker, still: if odds of beating the spread uncorrelated with previous game location
-    $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}     \end{frame}
+    $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}       \end{block}
+\end{frame}
 
-\begin{frame}   \frametitle{Our solution: a quasi-experiment}
-  Consider games played back-to-back within 24 hours.
+\begin{frame}
+  \frametitle{Previous research - NBA player shooting average}
+  \begin{block}{Empirical evidence: players party to the detriment of the game.}
+  NBA Players under-perform proportion of shots made on Sundays.\footnote{Chase, ('15).}
+  \end{block}
+    \begin{itemize}       
+    \item 1,092 of 2,035 NBA players who took $\geq 1$ shot on Sundays under-perform 
+      relative to their own season-long average; $p=0.001$.
+      \item Effect size is not considered.
+      \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, is that relevant?}     \end{itemize}
+\end{frame}
 
-  \vspace{12pt}
-  Examine whether spread is met as a function of nightlife index for lagged game location.
+\begin{frame}   \frametitle{Prev. research - likelihood of meeting the spread in NBA}
+  \begin{block}{Empirical evidence: away teams under-perform on Sundays.}
+  Teams under-perform bookmakers expectations on Sunday afternoons when on the road.\footnote{Ezekowitz, ('14).}
+    \begin{itemize}       
+    \item Only 129 out of 284 games did the away team meet the spread, 
+      using 2006-14 data; $p=0.1089$.
+    \item Not quite significant at the 90\% confidence level.     
+    \item If we bet \$100 on each Sunday game, we'd only profit \$1,310.\footnote{$(284 - 129) \times 100 + 129 \times (-110) = 1,310$.} 
+    \end{itemize}
+  \end{block}
+  \end{frame}
 
-  \vspace{12pt} (Alcohol and loss of sleep are well correlated with nightlife; both have been shown to have negative effects on sports performance.)\footnote{Shirreffs and Maughan ('06), Barnes ('14), Pilcher and Huffcutt ('96).} \end{frame}
+% \begin{frame}   \frametitle{Our solution: a quasi-experiment}
+%   Consider games played back-to-back within 24 hours.
+
+%   \vspace{12pt}
+%   Examine whether spread is met as a function of nightlife index for lagged game % location.
+
+%   \vspace{12pt} (Alcohol and loss of sleep are well correlated with nightlife; % both have been shown to have negative effects on sports % performance.)\footnote{Shirreffs and Maughan ('06), Barnes ('14), Pilcher and % Huffcutt ('96).} \end{frame}
 
@@
   \begin{block}{We perform the analysis for both NBA and MLB}     
-    Observe a team traveling to a city with active nightlife.
+    If we observe a team traveling to a city with active nightlife:
     \vspace{12pt}
@@
       \item For MLB teams, after conditioning on the odds of winning, are they less likely to beat their opponent?     \end{itemize}
-    \vspace{12pt}\begin{block}{If we yield a causal model, we may realize a betting schema.}     \end{block}
+    \vspace{12pt}\begin{block}{If we yield a causal model, we may realize a betting schema.}     
+\end{block}
   \end{block} 
@@
 
-\begin{frame}   \frametitle{Further motivation for conditioning on the spread}
+\begin{frame}   %\frametitle{Further motivation for conditioning on the spread}
   \begin{block}{With some correlation between next-day opponent, this further motivates conditioning

### 2017-07-06 init_model.rnw
Now only betting once per game!

@@
 require(hangover)
+
 require(data.table)
@@
 ##################################################
-### Table for uncorrelated next day opponent
+### Table for (un)-correlated next day opponent
 ##################################################
@@
 dev.off()
+
+## tmp <- nba.lines[, list(meet.spread = mean(outcome == 'W'), .N), by = list(last.game.loc, team)][order(N)]
+## hist(tmp$meet.spread, probability = T)
+## lines(density(tmp$meet.spread, adjust = 1))
 
@@
                                        cor(tmp$log.nmusicians, tmp$log.population) %>% round(., 3)))
+points(x = c(which(tmp$team == 'new orleans hornets'),
+             which(tmp$team == 'houston rockets'),
+             which(tmp$team == 'houston rockets')),
+       y = c(tmp[team == 'new orleans hornets', log.population],
+             tmp[team == 'houston rockets', log.population],
+             tmp[team == 'houston rockets', log.ndrinks]),
+       type = 'p', col = 'red')
 dev.off()
@@
     m <- glm(I(team.score > opponent.score) ~ party + odds,
-             data = mlb.lines[year(date) < yr], family = 'binomial')
+             data = mlb.lines[season < yr], family = 'binomial')
     p <- predict(m, newdata = mlb.lines[year(date) == yr], type = 'response')
-    idx <- which(p < .45)
-    p1 <- mapply(Predict, prediction = p[idx], 
-                 vegas = mlb.lines[year(date) == yr, odds][idx],
-                 MoreArgs=list(threshold = 0.01))
-    b <- mapply(Bet, our.prediction = p1,
-                actual.outcome = mlb.lines[year(date) == yr, team.score > opponent.score][idx],
-                odds = mlb.lines[year(date) == yr, odds][idx], SIMPLIFY = T)
-    cat(paste0(sum(b, na.rm = T), '\n'))
-    ###
-#    pdf(paste0('mlb_bets_', yr, '.pdf'))
-    betting.days <- mlb.lines[year(date) == yr][idx, date]
-    running.prof <- cumsum(ifelse(is.na(b), 0, b))
-    plot(x = 1:length(b), y = running.prof, 
-         main = paste0('Profits over the ', yr, ' season\n',
+    ### Subset data to relevant year we wish to predict on; make predictions.
+    tmp <- mlb.lines[season == yr, .(matchup, date, team, opponent, 
+                                     team.score, opponent.score, odds)]
+    tmp[, fitted := predict(m, newdata = mlb.lines[year(date) == yr], type = 'response')]
+    ### Realize that we can only make one prediction per game. So take the stronger bet.
+    setkey(tmp, date, matchup)
+    tmp <- tmp[, .SD[which.max(abs(fitted - .5))], by = list(matchup, date)]
+    tmp <- mlb.lines[tmp, on = intersect(names(tmp), names(mlb.lines))]
+    ### Translate fitted probabilities into 0-1 predictions if our odds differ by threshold amount.
+    tmp[, prediction := mapply(hangover::Predict, prediction = tmp$fitted, vegas = tmp$odds, 
+                               MoreArgs = list(threshold = 0.02))] # 0.02 was chosen ad-hoc.
+    tmp <- tmp[!is.na(prediction)]
+    ### Translate prediction into a monetary gain or loss based on odds and actual outcome.
+    tmp[, bet := mapply(hangover::Bet, our.prediction = prediction, 
+                        actual.outcome = team.score > opponent.score,
+                        odds = odds)]
+    ### Plot the cumulative sum of profits over the course of the season.
+    tmp[, running.prof := cumsum(bet)]    
+    cat(paste0(sum(tmp$bet), '\n'))
+    plot(x = 1:nrow(tmp), y = tmp$running.prof, 
+         main = paste0('Profit over the ', yr, ' season: $', 
+                       format(round(sum(tmp$bet), digits = -1), big.mark = ','), '\n',
                        'Training on 2011-', yr-1, ' data'),
@@
     abline(a = 0, b = 0, lty = 3)
-    points(x = which.min(running.prof), y = min(running.prof), type = 'p')
-    text(x = which.min(running.prof), y = min(running.prof), 
-         labels = round(min(running.prof)), pos = 4)
-    return(b)
-#    dev.off()
+    points(x = which.min(tmp$running.prof), y = min(tmp$running.prof), type = 'p')
+    text(x = which.min(tmp$running.prof), y = min(tmp$running.prof), 
+         labels = round(min(tmp$running.prof)), pos = 4)
+    return(tmp[, bet])
 })
@@
      xlab = 'Meet the Spread', breaks = 25)
-lines(density(tmp$meet.spread, adjust = 2), col = 'orange', lty = 'dashed')
+lines(density(tmp$meet.spread, adjust = 2), col = 'orange', lty = 'dashed', lwd = 3)
 ### There should be 'no information' left over after conditioning on odds.
@@
 mlb.lines[, resid  := resid(noinf)]
-mlb.lines[, list(avg.residual = mean(resid)), by = round(fitted, 3)] %>% 
-    plot(., xlab = 'Fitted Probability', ylab = 'Average Residual', 
+tmp <- mlb.lines[, .(avg.residual = mean(resid)), by = .(fitted = round(fitted, 3))]
+plot(x = tmp$fitted, y = tmp$avg.residual, 
+     xlab = 'Fitted Probability', ylab = 'Average Residual', 
      main = paste('After conditioning on bookkeeper\'s odds (MLB),',
                   'successful prediction is a coin-flip', sep = '\n'))
-abline(a = 0, b = 0, col = 'orange', lty = 'dashed')
+abline(lm(avg.residual ~ fitted, tmp), col = 'orange', lty = 'dashed', lwd = 3)
 dev.off()
@@
 m <- glm(I(team.score > opponent.score) ~ party + odds,
-         data = mlb.lines[year(date) < 2016], family = 'binomial')
-p <- predict(m, newdata = mlb.lines[year(date) == 2016], type = 'response')
+         data = mlb.lines[season < 2016], family = 'binomial')
+p <- predict(m, newdata = mlb.lines[season == 2016], type = 'response')
 idx <- which(p < .45)
 p1 <- mapply(Predict, prediction = p[idx], 
-             vegas = mlb.lines[year(date) == 2016, odds][idx],
+             vegas = mlb.lines[season == 2016, odds][idx],
              MoreArgs=list(threshold = 0.01))
 b <- mapply(Bet, our.prediction = p1,
-                actual.outcome = mlb.lines[year(date) == 2016, team.score > opponent.score][idx],
-            odds = mlb.lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
+                actual.outcome = mlb.lines[season == 2016, team.score > opponent.score][idx],
+            odds = mlb.lines[season == 2016, odds][idx], SIMPLIFY = T)
 
@@
 ### an optimized betting strategy.
-hist(p- mlb.lines[year(date) == 2016, odds])
+hist(p- mlb.lines[season == 2016, odds])
 
@@
     p1 <- mapply(Predict, prediction = p[idx], 
-                 vegas = mlb.lines[year(date) == 2016, odds][idx],
+                 vegas = mlb.lines[season == 2016, odds][idx],
                  MoreArgs=list(threshold = t))
     b <- mapply(Bet, our.prediction = p1,
-                actual.outcome = mlb.lines[year(date) == 2016, team.score > opponent.score][idx],
-                odds = mlb.lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
+                actual.outcome = mlb.lines[season == 2016, team.score > opponent.score][idx],
+                odds = mlb.lines[season == 2016, odds][idx], SIMPLIFY = T)
     return(c(sum(b, na.rm = T) / sum(!is.na(b)), sum(!is.na(b)))) # Avg profit per game.

### 2017-07-07 init_draft.r
This takes place of 'init_model.rnw', which was getting a bit long for a weave file.

@@ -1,0 +1,581 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Andreas Santucci
+###
+### Andreas Santucci
+###
+### Date: July 2017
+###
+### Inputs: 
+###
+### Dependencies: 
+###
+################################################################################
+################################################################################
+
+
+Sys.setlocale('LC_ALL','C')  # Needed for population data (within area.title variable)
+require(hangover)
+
+require(data.table)
+require(ggplot2)
+require(magrittr)
+require(sampling)
+require(stargazer)
+set.seed(4052017)
+
+### Load data.
+load(file = 'tmp_data/covers_lines.RData')
+load(file = 'tmp_data/population_by_MSA_year.RData')
+espn <- fread('tmp_data/espn_player_data.csv')
+espn[, date := as.Date(date)]
+
+##############################
+### Assemble NBA Data.
+##############################
+
+load(file = 'tmp_data/studios_and_artists.RData')
+musicians <- copy(dt)
+rm(dt)
+setnames(musicians, 'variable', 'nmusicians')
+
+load(file = 'tmp_data/alcohol_and_drinking.RData')
+drinking <- copy(dt)
+rm(dt)
+setnames(drinking, 'variable', 'ndrinks')
+
+### Bring together musicians and drinking establishments.
+bls <- merge(musicians, drinking, on = intersect(names(musicians), names(drinking)))
+bls[, area.title := gsub(' msa$', '', area.title, ignore.case = T)]
+
+### Geocode nba team locations. Prepare data for merge.
+## locs <- MyGeoCode(unique(lines$team), 'nba')
+## save(locs, file = 'tmp_data/nba_team_locations.RData')
+load('tmp_data/nba_team_locations.RData')  # <-- locs.
+states <- data.table(state = state.name, state.abb = state.abb)
+states <- rbind(states, data.frame(state = 'District of Columbia', state.abb = 'DC'))
+locs <- locs[states, on = 'state', nomatch = 0]  # Gets rid of Toronto. That's ok!
+
+### Attach team-location information to lines data.
+lines <- lines[locs, on = 'team']  # Here, we lose Toronto (that's OK).
+
+
+### We can't simply merge. We have to look for the county within the MSA.
+### For each location, we fetch a panel of corresponding musicians data, 2010-2016.
[…]

### 2017-07-07 hangover/R/CleanBLS.r
Code runs OK, but we lost significance for a couple models.

@@ -16,39 +16,42 @@
 #' # not run: CleanBLS(c('recording studio', 'musical group'), 'DJ_proxy', 'nba', sum)
 CleanBLS <- function(party.regex, suffix, sport, FUN, years = 2010:2016) {
     ### Subset each year of BLS data to only observations matching regular expression.
-    lapply(years, hangover::SubsetBLS,
-           party.regex = party.regex, sport = sport, suffix = suffix)
+    lapply(years, hangover::SubsetBLS, party.regex = party.regex, suffix = suffix)
     files <- list.files(path = 'tmp_data/bls',
-                        pattern = paste0(sport, '_', suffix, '\\.csv$'), full.names = T)
+                        pattern = paste0(suffix, '\\.csv$'), full.names = T)
     bls <- lapply(files, fread) %>% rbindlist
-    ### Fetch a listing of team-names from our points-spread/money-lines data.
-    if (sport == 'nba') {
-        load(file = 'tmp_data/covers_lines.RData')   # Load lines data to get team-names (overkill)
-        lines[, pct := NULL]
-        teams <- unique(lines$team)
-    } else if (sport == 'mlb') {
-        abbrs <- fread(file = 'mlb_abbrs.csv')
-        setnames(abbrs, c('abbr', 'full', 'nick'))
-        teams <- abbrs$full
-    }
-    ### Geocode the location of each team within the sport, and add this to our BLS data.
-    locs <- MyGeoCode(teams, sport)
-    bls[, area.title := gsub(' msa$', '', area.title, ignore.case = T)]
-    states <- data.table(state = state.name, state.abb = state.abb)
-    states <- rbind(states, data.frame(state = 'District of Columbia', state.abb = 'DC'))
-    locs <- locs[states, on = 'state', nomatch = 0]
-    ### We can't simply merge. We have to look for the county within the MSA.
-    dt <- lapply(1:nrow(locs), function(i) { 
-        matched <- bls[grepl(locs[i, locality],  area.title, ignore.case = T) &
-                       grepl(locs[i, state.abb], area.title, ignore.case = T)]
-        matched[, c('locality', 'state', 'team') := locs[i, list(locality, state.abb, team)]]
-        return(matched)
-    }) %>% rbindlist
-    dt <- dt[, list(variable = FUN(qtrly.estabs.count %>% as.numeric)),
[…]

### 2017-07-07 hangover/R/MyGeoCode.r

@@ -19,12 +19,13 @@ MyGeoCode <- function(teams, sport) {
         locs[locality == 'El Segundo', locality := 'Los Angeles']
     }
     locs <- data.table(team = teams, locality = locs$locality,
-                       state = locs$administrative.area.level.1, lon = locs$lon, lat = locs$lat)
+                       state = locs$administrative.area.level.1,
+                       lon = locs$lon %>% as.numeric, lat = locs$lat %>% as.numeric)
     if (sport == 'mlb') {
         locs[is.na(locality) & state == 'Florida', `:=`(locality = 'Atlanta', state = 'Georgia')]
         locs[is.na(locality) & state == 'New York', locality := 'New York']
         locs[locality == 'Broomfield' & state == 'Colorado', locality := 'Denver']
         locs[team == 'texas rangers', locality := 'Dallas']
     }
-    save(locs, file = paste0('tmp_data/', sport, '_team_locations.RData'))
+    return(locs)
 }

### 2017-07-07 hangover/R/SubsetBLS.r

@@ -3,7 +3,6 @@
 #' Subset BLS data into manageable chunks, keeping only observations we care about.
 #' @param year An integer describing which year of BLS data to subset.
 #' @param party.regex A character vector describing patterns to look for in BLS data.
-#' @param sport A string, one of 'nba' or 'mlb'.
 #' @param suffix A string describing how to prefix the output file.
 #' @param raw.path A string describing the input file path. Defaults to 'raw_data/bls/'.
 #' @param save.path A string describing the output file path. Defaults to 'tmp_data/bls/'.
@@ -11,7 +10,7 @@
 #' @export
 #' @examples
 #' # not run: SubsetBLS(2010, 'sound recording', 'nba', 'sound_recording_studios')
-SubsetBLS <- function(year, party.regex, sport, suffix,
+SubsetBLS <- function(year, party.regex, suffix,
                       raw.path = 'raw_data/bls/', save.path = 'tmp_data/bls/') {
     ### Create a listing of files for all relevant MSA's, and load data.
     foldr <- list.files(path = raw.path, pattern = paste0(year, '.q1'), full.names = T)
@@ -25,10 +24,10 @@ SubsetBLS <- function(year, party.regex, sport, suffix,
     setnames(bls, gsub('_', '.', names(bls)))
     ### We will take our own total later, so for now we take granular data.
     bls <- bls[grep("NAICS 6-digit", agglvl.title)]
-    ### Create regex to subset to drinking establshimensts.
+    ### Create regex to subset to relevant establshimensts.
     party.regex <- paste0('(', party.regex, ')', collapse = '|')
     bls <- bls[grep(party.regex, industry.title, ignore.case = T)]
-    write.csv(x = bls, file = paste0(save.path, year, '_', sport, '_', suffix, '.csv'),
+    write.csv(x = bls, file = paste0(save.path, year, '_', suffix, '.csv'),
               row.names = F)
     rm(bls)
     gc()

### 2017-07-07 hangover/R/CleanBLS.r
Removed extraneous code.

@@ -24,34 +24,4 @@ CleanBLS <- function(party.regex, suffix, sport, FUN, years = 2010:2016) {
               by = list(area.title, year)]
     dt[, year := as.integer(year)]
     save(dt, file = paste0('tmp_data/', suffix, '.RData'))
-    ## ### Fetch a listing of team-names from our points-spread/money-lines data. (overkill)
-    ## if (sport == 'nba') {
-    ##     load(file = 'tmp_data/covers_lines.RData')
-    ##     lines[, pct := NULL]
-    ##     teams <- unique(lines$team)
-    ## } else if (sport == 'mlb') {
-    ##     abbrs <- fread(file = 'mlb_abbrs.csv')
-    ##     setnames(abbrs, c('abbr', 'full', 'nick'))
-    ##     teams <- abbrs$full
-    ## }
-    ## ### Geocode the location of each team within the sport, and add this to our BLS data.
-    ## locs <- MyGeoCode(teams, sport)
-    ## bls[, area.title := gsub(' msa$', '', area.title, ignore.case = T)]
-    ## states <- data.table(state = state.name, state.abb = state.abb)
-    ## states <- rbind(states, data.frame(state = 'District of Columbia', state.abb = 'DC'))
-    ## locs <- locs[states, on = 'state', nomatch = 0]
-    ## ### We can't simply merge. We have to look for the county within the MSA.
-    ## dt <- lapply(1:nrow(locs), function(i) { 
-    ##     matched <- bls[grepl(locs[i, locality],  area.title, ignore.case = T) &
-    ##                    grepl(locs[i, state.abb], area.title, ignore.case = T)]
-    ##     matched[, c('locality', 'state', 'team') := locs[i, list(locality, state.abb, team)]]
-    ##     return(matched)
-    ## }) %>% rbindlist
-    ## dt <- dt[, list(variable = FUN(qtrly.estabs.count %>% as.numeric)),
-    ##          by = list(locality, state, year)][order(year, variable)]
-    ## setnames(dt, 'year', 'season')
-    ## dt[, season := as.numeric(season)]
-    ## ### NBA season spans new-years, e.g. change 2010-11 season --> 2011.
-    ## if (sport == 'nba') dt[, season := season + 1]
-    ## save(dt, file = paste0('tmp_data/', suffix, '_', sport, '.RData'))
 }

### 2017-07-07 cardinal_analytx_presentation/nightlife_and_sports.tex
Keeping old 'rebounds per minute', since it is not significant with our newly updated scripts. Other than that, everything looks updated!

@@
 \begin{frame}   \frametitle{Identifying variation}
-  \begin{block}{Explain performance as a function of prev. day's nightlife index.}     If next day opponent is uncorrelated with visitation to a party city,
+  \begin{block}{Performance as a function of previous day's nightlife index.}     If next day opponent is uncorrelated with visitation to a party city,
     then looking at the \emph{actual outcome} of the game $\leadsto$ identification.
@@
   \centering
-  \tiny{
-  \begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline   \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\ & & \\   [-1.8ex] Party continuous &  & $-$0.181$^{**}$ \\    &  & (0.091) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Logged travel distance & $-$0.017 & $-$0.014 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.145 & $-$0.113 \\    & (0.232) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   Logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.028 \\    & (0.378) & (0.378) \\ [-1.8ex] & & \\ \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.677 & $-$6,805.100 \\  Akaike Inf.
+  \tiny{ % m0, m1
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.555$^{***}$ &  \\    & p = 0.0002 &  \\    & & \\   Party continuous &  & $-$0.153 \\    &  & p = 0.107 \\    & & \\   Lag changes in posession & 0.001 & 0.001 \\    & p = 0.346 & p = 0.373 \\    & & \\   Logged travel distance & $-$0.014 & $-$0.009 \\    & p = 0.577 & p = 0.733 \\    & & \\   East-west travel direction & $-$0.074 & $-$0.039 \\    & p = 0.754 & p = 0.867 \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & p = 0.218 & p = 0.295 \\    & & \\   Time of game during day & 0.007 & 0.005 \\    & p = 0.603 & p = 0.715 \\    & & \\   Home team effect & 0.002 & 0.004 \\    & p = 0.960 & p = 0.936 \\    & & \\   Logged travel distance * east-west & 0.012 & 0.007 \\    & p = 0.749 & p = 0.846 \\    & & \\   Constant & $-$0.247 & $-$0.229 \\    & p = 0.443 & p = 0.477 \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,587.456 & $-$6,593.324 \\  Akaike Inf.
 Crit.
-& 13,619.350 & 13,628.200 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-}   \end{frame}
+& 13,192.910 & 13,204.650 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
+\end{frame}
 
@@
   \centering
-  \tiny{
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.072 \\    & (0.082) \\    & \\   Lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   Logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.099 \\    & (0.231) \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   Time of game during day & 0.001 \\    & (0.013) \\    & \\   Home team effect & $-$0.009 \\    & (0.043) \\    & \\   Logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.069 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.701 \\  Akaike Inf.
+  \tiny{ % mp
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.056 \\    & p = 0.504 \\    & \\   Lag changes in posession & 0.001 \\    & p = 0.408 \\    & \\   Logged travel distance & $-$0.003 \\    & p = 0.920 \\    & \\   East-west travel direction & $-$0.036 \\    & p = 0.880 \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & p = 0.098 \\    & \\   Time of game during day & 0.003 \\    & p = 0.827 \\    & \\   Home team effect & $-$0.006 \\    & p = 0.896 \\    & \\   Logged travel distance * east-west & 0.006 \\    & p = 0.866 \\    & \\   Constant & $-$0.257 \\    & p = 0.425 \\    & \\  \hline \\[-1.8ex]  Observations & 9,517 \\  Log Likelihood & $-$6,594.404 \\  Akaike Inf.
 Crit.
-& 13,631.400 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  } \end{frame}
+& 13,206.810 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}      }
+\end{frame}
 
@@
   \centering
-  \tiny{
-  \begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.126$^{*}$ &  \\    & (0.074) &  \\    & & \\   Nightlife (no weekend interaction) &  & $-$0.016 \\    &  & (0.110) \\    & & \\   Bookmaker's odds & 2.638$^{***}$ & 2.583$^{***}$ \\    & (0.163) & (0.165) \\    & & \\   Home-team effect & 0.008 & 0.034 \\    & (0.035) & (0.083) \\    & & \\   Number of rest days & 0.019 & 0.018 \\    & (0.020) & (0.020) \\    & & \\   Logged travel distance & $-$0.002 & $-$0.003 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.044 & 0.003 \\    & (0.038) & (0.030) \\    & & \\   Constant & $-$1.344$^{***}$ & $-$1.324$^{***}$ \\    & (0.081) & (0.113) \\    & & \\  \hline \\[-1.8ex]  Observations & 24,734 & 24,033 \\  Log Likelihood & $-$16,946.080 & $-$16,470.850 \\  Akaike Inf.
+  \tiny{ % m3, m4
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.120$^{*}$ &  \\    & p = 0.091 &  \\    & & \\   Nightlife (no weekend interaction) &  & 0.137 \\    &  & p = 0.307 \\    & & \\   Bookmaker's odds & 2.682$^{***}$ & 2.669$^{***}$ \\    & p = 0.000 & p = 0.000 \\    & & \\   Home-team effect & 0.014 & 0.144 \\    & p = 0.674 & p = 0.159 \\    & & \\   Number of rest days & 0.018 & 0.016 \\    & p = 0.399 & p = 0.456 \\    & & \\   Logged travel distance & $-$0.005 & $-$0.003 \\    & p = 0.377 & p = 0.519 \\    & & \\   Weekend & 0.044 & 0.002 \\    & p = 0.246 & p = 0.935 \\    & & \\   Constant & $-$1.360$^{***}$ & $-$1.468$^{***}$ \\    & p = 0.000 & p = 0.000 \\    & & \\  \hline \\[-1.8ex]  Observations & 26,473 & 26,473 \\  Log Likelihood & $-$18,124.390 & $-$18,125.300 \\  Akaike Inf.
 Crit.
-& 33,906.160 & 32,955.700 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-} \end{frame}
+& 36,262.780 & 36,264.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
+\end{frame}
 
@@
   \centering
-  \tiny{
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.003$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00003 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00003 \\    & (0.00002) \\    & \\   Constant & $-$0.008$^{*}$ \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 43,547 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
-Error & 0.104 (df = 43542) \\  F Statistic & 2.945$^{**}$ (df = 4; 43542) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-} \end{frame}
+  \tiny{ % m2
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.004$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00002 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00002 \\    & (0.00002) \\    & \\   Constant & $-$0.007 \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 40,684 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
+Error & 0.104 (df = 40679) \\  F Statistic & 2.696$^{**}$ (df = 4; 40679) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  } \end{frame}
 
@@
   \begin{adjustwidth}{-.75cm}{}
-  \tiny{
-  \begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.828$^{***}$ &  &  \\    & (0.699) &  &  \\    & & & \\   Continuous party measure &  & 1.667$^{***}$ & 0.080 \\    &  & (0.587) & (0.575) \\    & & & \\   Lag change in possessions &  & 0.267$^{***}$ & 0.285$^{***}$ \\    &  & (0.016) & (0.016) \\    & & & \\   Logged travel distance &  & $-$0.447$^{**}$ & 0.143 \\    &  & (0.191) & (0.187) \\    & & & \\   Number hours since last game &  & 0.011 & 0.017$^{**}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Constant & $-$0.047 & 70.710$^{***}$ & 64.859$^{***}$ \\    & (0.090) & (2.271) & (2.223) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,658 & 6,426 & 6,426 \\  R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Adjusted R$^{2}$ & 0.001 & 0.044 & 0.051 \\  Residual Std.
-Error & 11.159 (df = 15656) & 11.941 (df = 6421) & 11.688 (df = 6421) \\  F Statistic & 16.357$^{***}$ (df = 1; 15656) & 74.877$^{***}$ (df = 4; 6421) & 86.792$^{***}$ (df = 4; 6421) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-}
-\end{adjustwidth} \end{frame}
+  \tiny{  % mpa
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & p = 0.0001 &  &  \\    & & & \\   Continuous party measure &  & 1.803$^{***}$ & $-$0.160 \\    &  & p = 0.007 & p = 0.806 \\    & & & \\   Lag change in possessions &  & 0.118$^{***}$ & 0.115$^{***}$ \\    &  & p = 0.000 & p = 0.000 \\    & & & \\   Logged travel distance &  & $-$0.568$^{***}$ & 0.037 \\    &  & p = 0.005 & p = 0.852 \\    & & & \\   Number hours since last game &  & 0.011 & 0.018$^{**}$ \\    &  & p = 0.115 & p = 0.013 \\    & & & \\   Constant & $-$0.046 & 89.567$^{***}$ & 86.136$^{***}$ \\    & p = 0.612 & p = 0.000 & p = 0.000 \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 6,234 & 6,234 \\  R$^{2}$ & 0.001 & 0.023 & 0.021 \\  Adjusted R$^{2}$ & 0.001 & 0.023 & 0.021 \\  Residual Std.
+Error & 11.137 (df = 15137) & 12.071 (df = 6229) & 11.910 (df = 6229) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 37.032$^{***}$ (df = 4; 6229) & 33.756$^{***}$ (df = 4; 6229) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
+\end{adjustwidth} 
+\end{frame}
 

### 2017-07-07 writing/draft.tex
Updated tables; now all but demeaned rebounds contain p-values instead of std. err.

@@
 % m0, m1
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   Party continuous &  & $-$0.181$^{**}$ \\    &  & (0.091) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Logged travel distance & $-$0.017 & $-$0.014 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.145 & $-$0.113 \\    & (0.232) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   Logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.028 \\    & (0.378) & (0.378) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.677 & $-$6,805.100 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.555$^{***}$ &  \\    & p = 0.0002 &  \\    & & \\   Party continuous &  & $-$0.153 \\    &  & p = 0.107 \\    & & \\   Lag changes in posession & 0.001 & 0.001 \\    & p = 0.346 & p = 0.373 \\    & & \\   Logged travel distance & $-$0.014 & $-$0.009 \\    & p = 0.577 & p = 0.733 \\    & & \\   East-west travel direction & $-$0.074 & $-$0.039 \\    & p = 0.754 & p = 0.867 \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & p = 0.218 & p = 0.295 \\    & & \\   Time of game during day & 0.007 & 0.005 \\    & p = 0.603 & p = 0.715 \\    & & \\   Home team effect & 0.002 & 0.004 \\    & p = 0.960 & p = 0.936 \\    & & \\   Logged travel distance * east-west & 0.012 & 0.007 \\    & p = 0.749 & p = 0.846 \\    & & \\   Constant & $-$0.247 & $-$0.229 \\    & p = 0.443 & p = 0.477 \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,587.456 & $-$6,593.324 \\  Akaike Inf.
 Crit.
-& 13,619.350 & 13,628.200 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 13,192.910 & 13,204.650 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 % m3, m4
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.126$^{*}$ &  \\    & (0.074) &  \\    & & \\   Nightlife (no weekend interaction) &  & $-$0.016 \\    &  & (0.110) \\    & & \\   Bookmaker's odds & 2.638$^{***}$ & 2.583$^{***}$ \\    & (0.163) & (0.165) \\    & & \\   Home-team effect & 0.008 & 0.034 \\    & (0.035) & (0.083) \\    & & \\   Number of rest days & 0.019 & 0.018 \\    & (0.020) & (0.020) \\    & & \\   Logged travel distance & $-$0.002 & $-$0.003 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.044 & 0.003 \\    & (0.038) & (0.030) \\    & & \\   Constant & $-$1.344$^{***}$ & $-$1.324$^{***}$ \\    & (0.081) & (0.113) \\    & & \\  \hline \\[-1.8ex]  Observations & 24,734 & 24,033 \\  Log Likelihood & $-$16,946.080 & $-$16,470.850 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.120$^{*}$ &  \\    & p = 0.091 &  \\    & & \\   Nightlife (no weekend interaction) &  & 0.137 \\    &  & p = 0.307 \\    & & \\   Bookmaker's odds & 2.682$^{***}$ & 2.669$^{***}$ \\    & p = 0.000 & p = 0.000 \\    & & \\   Home-team effect & 0.014 & 0.144 \\    & p = 0.674 & p = 0.159 \\    & & \\   Number of rest days & 0.018 & 0.016 \\    & p = 0.399 & p = 0.456 \\    & & \\   Logged travel distance & $-$0.005 & $-$0.003 \\    & p = 0.377 & p = 0.519 \\    & & \\   Weekend & 0.044 & 0.002 \\    & p = 0.246 & p = 0.935 \\    & & \\   Constant & $-$1.360$^{***}$ & $-$1.468$^{***}$ \\    & p = 0.000 & p = 0.000 \\    & & \\  \hline \\[-1.8ex]  Observations & 26,473 & 26,473 \\  Log Likelihood & $-$18,124.390 & $-$18,125.300 \\  Akaike Inf.
 Crit.
-& 33,906.160 & 32,955.700 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 36,262.780 & 36,264.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 % mp
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.072 \\    & (0.082) \\    & \\   Lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   Logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.099 \\    & (0.231) \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   Time of game during day & 0.001 \\    & (0.013) \\    & \\   Home team effect & $-$0.009 \\    & (0.043) \\    & \\   Logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.069 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.701 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.056 \\    & p = 0.504 \\    & \\   Lag changes in posession & 0.001 \\    & p = 0.408 \\    & \\   Logged travel distance & $-$0.003 \\    & p = 0.920 \\    & \\   East-west travel direction & $-$0.036 \\    & p = 0.880 \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & p = 0.098 \\    & \\   Time of game during day & 0.003 \\    & p = 0.827 \\    & \\   Home team effect & $-$0.006 \\    & p = 0.896 \\    & \\   Logged travel distance * east-west & 0.006 \\    & p = 0.866 \\    & \\   Constant & $-$0.257 \\    & p = 0.425 \\    & \\  \hline \\[-1.8ex]  Observations & 9,517 \\  Log Likelihood & $-$6,594.404 \\  Akaike Inf.
 Crit.
-& 13,631.400 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 13,206.810 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 % m2
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.003$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00003 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00003 \\    & (0.00002) \\    & \\   Constant & $-$0.008$^{*}$ \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 43,547 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
-Error & 0.104 (df = 43542) \\  F Statistic & 2.945$^{**}$ (df = 4; 43542) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.004$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00002 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00002 \\    & (0.00002) \\    & \\   Constant & $-$0.007 \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 40,684 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
+Error & 0.104 (df = 40679) \\  F Statistic & 2.696$^{**}$ (df = 4; 40679) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 \paragraph{The hangover correlates with allowing more points}
@@
 \hspace{-100pt}
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.828$^{***}$ &  &  \\    & (0.699) &  &  \\    & & & \\   Continuous party measure &  & 1.667$^{***}$ & 0.080 \\    &  & (0.587) & (0.575) \\    & & & \\   Lag change in possessions &  & 0.267$^{***}$ & 0.285$^{***}$ \\    &  & (0.016) & (0.016) \\    & & & \\   Logged travel distance &  & $-$0.447$^{**}$ & 0.143 \\    &  & (0.191) & (0.187) \\    & & & \\   Number hours since last game &  & 0.011 & 0.017$^{**}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Constant & $-$0.047 & 70.710$^{***}$ & 64.859$^{***}$ \\    & (0.090) & (2.271) & (2.223) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,658 & 6,426 & 6,426 \\  R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Adjusted R$^{2}$ & 0.001 & 0.044 & 0.051 \\  Residual Std.
-Error & 11.159 (df = 15656) & 11.941 (df = 6421) & 11.688 (df = 6421) \\  F Statistic & 16.357$^{***}$ (df = 1; 15656) & 74.877$^{***}$ (df = 4; 6421) & 86.792$^{***}$ (df = 4; 6421) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & p = 0.0001 &  &  \\    & & & \\   Continuous party measure &  & 1.803$^{***}$ & $-$0.160 \\    &  & p = 0.007 & p = 0.806 \\    & & & \\   Lag change in possessions &  & 0.118$^{***}$ & 0.115$^{***}$ \\    &  & p = 0.000 & p = 0.000 \\    & & & \\   Logged travel distance &  & $-$0.568$^{***}$ & 0.037 \\    &  & p = 0.005 & p = 0.852 \\    & & & \\   Number hours since last game &  & 0.011 & 0.018$^{**}$ \\    &  & p = 0.115 & p = 0.013 \\    & & & \\   Constant & $-$0.046 & 89.567$^{***}$ & 86.136$^{***}$ \\    & p = 0.612 & p = 0.000 & p = 0.000 \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 6,234 & 6,234 \\  R$^{2}$ & 0.001 & 0.023 & 0.021 \\  Adjusted R$^{2}$ & 0.001 & 0.023 & 0.021 \\  Residual Std.
+Error & 11.137 (df = 15137) & 12.071 (df = 6229) & 11.910 (df = 6229) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 37.032$^{***}$ (df = 4; 6229) & 33.756$^{***}$ (df = 4; 6229) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 \section{Conclusions and Avenues for Future Research}

### 2017-07-07 clean_bls.r
No longer sport specific!

@@ -10,7 +10,7 @@
 ###
 ### Inputs: 'raw_data/bls/20[xx].q1-q4.by_area'
 ###
-### Output: 'tmp_data/bls/20[xx]_[sport]_[suffix].csv'
+### Output: 'tmp_data/bls/20[xx]_[suffix].csv'
 ###
 ### Dependencies: data.table, ggmap, magrittr, parallel
 ###
@@ -30,10 +30,9 @@ require(data.table)
 require(ggmap)
 require(magrittr)
 
-### Loads raw data. Subsets based on regex. Aggregate using FUN. Save using suffix and sport.
+### Loads raw data. Subsets based on regex. Aggregate using FUN. Save using suffix.
 hangover::CleanBLS(party.regex = c('sound recording', 'music publisher', 'musical group'),
-                   suffix = 'studios_and_artists', sport = 'nba', FUN = sum)
+                   suffix = 'studios_and_artists', FUN = sum)
 
-### We add option 'rescrape = F' to avoid scraping location data a second time.
 hangover::CleanBLS(party.regex = c('alcohol', 'drinking'),
-                   suffix = 'alcohol_and_drinking', sport = 'nba', FUN = sum)
+                   suffix = 'alcohol_and_drinking', FUN = sum)

### 2017-07-07 clean_covers.r
Commented out old code.

@@ -99,7 +99,9 @@ lines[, last.game.time := c(NA, lag(hour(date))[1:.N-1]), by = list(team, season
 ### Distance Traveled.
 ##################################################
 ### Here, we geocode team locations to get lat-lon, and also addresses.
-locs <- hangover::getLocations(unique(lines$team))
+locs <- hangover::MyGeoCode(unique(lines$team), 'nba')
+## load(file = 'tmp_data/nba_team_locations.RData')
+locs[, `:=`(lon = as.numeric(lon), lat = as.numeric(lat))]
 
 hangover::calculateDistances(locs, 'tmp_data/distance_matrix.RData', unique(lines$team))
 load(file = 'tmp_data/distance_matrix.RData')
@@ -110,8 +112,8 @@ lines[, travel.dist := hangover::getDist(location, last.game.loc, dmat),
 ##################################################
 ### Direction Traveled.
 ##################################################
-load(file = 'tmp_data/nba_team_locations.RData')
-locs[, `:=`(lon = as.numeric(lon), lat = as.numeric(lat))]
+## load(file = 'tmp_data/nba_team_locations.RData')
+## locs[, `:=`(lon = as.numeric(lon), lat = as.numeric(lat))]
 ### Get directions.
 dirs <- matrix(nrow = nrow(dmat), ncol = ncol(dmat), dimnames = dimnames(dmat))
 for (i in 1:(nrow(dirs)))

### 2017-07-07 clean_covers_mlb.r
Commented out old code.

@@ -52,6 +52,7 @@ data[, c('matchup', 'team', 'opponent') := NULL]
 ### Rename variables.
 setnames(data, gsub('\\.full', '', tolower(names(data))))
 setcolorder(data, c('date', 'team', 'opponent', 'line', 'score'))
+data[, matchup := paste(team, opponent, sep = '-')]
 
 ##################################################
 ### Merge in Game-Information
@@ -76,6 +77,7 @@ write.csv(abbrs, file = 'mlb_abbrs.csv', row.names = F)
 game.info[, c('team.nick', 'opponent.nick') := tstrsplit(matchup, split = ' Vs\\. ')]
 game.info[, team.nick := gsub('(.*) sox', '\\1sox', team.nick, ignore.case = T) %>% tolower]
 game.info[, opponent.nick := gsub('(.*) sox', '\\1sox', opponent.nick, ignore.case = T) %>% tolower]
+game.info[, matchup := NULL]
 
 ### Set up column names and types for the merge.
 setnames(game.info, 'location', 'stadium')
@@ -87,8 +89,8 @@ data <- data[year(date) > 2010]
 ##################################################
 ### Variable creation and column classes.
 ##################################################
-
 ### Create a numeric money-line variable.
+data <- data[!is.na(line) & !(line %in% c('OFF', 'lines', 'Line'))]
 data[, line := gsub(' \\(Open\\)$', '', line) %>% as.numeric]
 data <- data[!is.na(line) & !is.na(matchup)]
 
@@ -106,6 +108,7 @@ data[, odds := ifelse(line < 0, 100 / (-line + 100), line / (line + 100))]
 data[, weekday := weekdays(as.Date(date))]
 data[, duration := gsub('Game Duration: ', '', duration)]
 data[, stadium  := gsub('Venue: ', '', stadium)]
+cat("The following line of code will generate a warning; it's related to attendance. Don't worry.\n")
 data[, attendance := gsub('Attendance: ', '', attendance) %>% gsub(',', '', . ) %>% as.numeric]
 data[, start.time := gsub('Start Time: ', '', start.time)]
 data[, start.time := gsub(' +(ET)|(Local)$', '', start.time)]
@@ -129,7 +132,8 @@ lines <- lines[ndays.lgame != 0]
 ### Team Locations and Distance Traveled.
 ##################################################
 ### Here, we geocode team locations to get lat-lon, and also addresses.
[…]

### 2017-07-07 init_draft.r
Now printing out p-values; also using hangover::LookupLocation.

@@ -62,17 +62,8 @@ locs <- locs[states, on = 'state', nomatch = 0]  # Gets rid of Toronto. That's o
 lines <- lines[locs, on = 'team']  # Here, we lose Toronto (that's OK).
 
 
-### We can't simply merge. We have to look for the county within the MSA.
-### For each location, we fetch a panel of corresponding musicians data, 2010-2016.
-LookupLocation <- function(loc, st, data, team = NULL) {
-    sb <- data[grepl(loc, area.title, ignore.case = T) & grepl(st,  area.title)]
-    sb[, `:=`(locality = loc, state = st)]
-    if (!is.null(team)) sb[, team := team]
-    return(sb)
-}
-
 ### Attach nba team-names to BLS data.
-demog <- mapply(LookupLocation, loc = locs$locality, st = locs$state.abb, team = locs$team,
+demog <- mapply(hangover::LookupLocation, loc = locs$locality, st = locs$state.abb, team = locs$team,
                 MoreArgs = list(data = bls), SIMPLIFY = F) %>% rbindlist
 demog <- merge(demog, pop, by = c('area.title', 'year'), all.x = T)
 setnames(demog, 'state', 'state.abb')
@@ -158,7 +149,7 @@ stargazer(m0, m1, covariate.labels = c('Party discrete', 'Party continuous',
                                        'Home team effect', 
                                        'Logged travel distance * east-west', 
                                        'Constant'),
-          dep.var.labels = 'Meet the Spread')
+          dep.var.labels = 'Meet the Spread', report = 'vc*p')
 
 ##################################################
 ### Dot plot of Party-Continuous by Team Location
@@ -214,7 +205,7 @@ stargazer(mp, covariate.labels = c('Party placebo', 'Lag changes in posession',
                                    'Number hours rest time', 'Time of game during day', 
                                    'Home team effect', 'Logged travel distance * east-west', 
                                    'Constant'),
-          dep.var.labels = 'Meet the Spread (NBA)')
+          dep.var.labels = 'Meet the Spread (NBA)', report = 'vc*p')
 
 ##################################################
 ### Table for (un)-correlated next day opponent
@@ -349,7 +340,7 @@ stargazer(m2, dep.var.labels = 'Demeanead Rebounds Per Minute',
                                'Lag change in possessions',
                                'Logged travel distance',
                                'Number hours since last game',
-                               'Constant'))
+                               'Constant'), report = 'vc*p')
 
[…]

### 2017-07-07 init_model.rnw
Now retiring this script, as it was getting too long. Code was ported over to init_draft.r.

@@
 
+Sys.setlocale('LC_ALL','C')  # Needed for population data (within area.title variable)
+
 require(hangover)
@@
 load(file = 'tmp_data/population_by_MSA_year.RData')
-load(file = 'tmp_data/studios_and_artists_nba.RData')
+load(file = 'tmp_data/studios_and_artists.RData')
 
@@
 
-### This is simply ad-hoc.
-musicians[, population := 0.0]
-for (i in 1:nrow(musicians))
-    set(musicians, i, j = 'population', 
-        value = pop[grepl(musicians[i, locality], locality) &
-                    grepl(musicians[i, state],    locality) &
-                    musicians[i, season] == season, mean(population)])
+load(file = 'tmp_data/alcohol_and_drinking.RData')
+drinking <- copy(dt)
+rm(dt)
+setnames(drinking, 'variable', 'ndrinks')
+
+musicians <- merge(musicians, drinking, on = intersect(names(musicians), names(drinking)))
+
+### Geocode team locations. Prepare data for merge.
+locs <- MyGeoCode(unique(lines$team), 'nba')
+states <- data.table(state = state.name, state.abb = state.abb)
+states <- rbind(states, data.frame(state = 'District of Columbia', state.abb = 'DC'))
+locs <- locs[states, on = 'state', nomatch = 0]
+
+### We can't simply merge. We have to look for the county within the MSA.
+musicians[, area.title := gsub(' msa$', '', area.title, ignore.case = T)]
+musicians <- lapply(1:nrow(locs), function(i) { 
+    matched <- musicians[grepl(locs[i, locality],  area.title, ignore.case = T) &
+                         grepl(locs[i, state.abb], area.title, ignore.case = T)]
+    matched[, c('locality', 'state') := locs[i, list(locality, state.abb)]]
+    return(matched)
+}) %>% rbindlist %>% unique
+
+demog <- merge(musicians, pop, by = c('area.title', 'year'), all.x = T)
+setnames(demog, 'state', 'state.abb')
+
+lines <- lines[locs, on = 'team']  # Here, we lose Toronto (that's OK).
+
+setnames(demog, 'year', 'season')
+demog[, season := season + 1]
+
+nba.lines <- lines[demog, on = c('locality', 'state.abb', 'season'), allow.cartesian = T]
+
+nba.lines <- nba.lines[last.game.loc != 'toronto raptors' &
+                       location      != 'toronto raptors']
+
+### Exclude playoff games.
+setkey(nba.lines, season, team, date)
+nba.lines <- nba.lines[, head(.SD, n=82), by=list(season,team)]
 
@@
 <<echo=F>>=
-nba.lines <- lines
-nba.lines[, list(date = as.Date(date), team, opponent, line, score)]
+#nba.lines <- lines
+#nba.lines[, list(date = as.Date(date), team, opponent, line, score)]
 @
@@
 <<echo=F>>=
-### Merge lines data with BLS data.
-nba.lines <- merge(nba.lines, musicians,
-               by.x = c('season', 'last.game.loc'),
-               by.y = c('season', 'team'), all.x = T)
 
-### Exclude playoff games.
-setkey(nba.lines, season, team, date)
-nba.lines <- nba.lines[, head(.SD, n=82), by=list(season,team)]
+## nba.lines <- nba.lines[locs, on = 'team']
+
+## ## setnames(nba.lines, 'state.abb', 'state')
+## setnames(musicians, 'year', 'season')
+## musicians[, season := as.integer(season)]
+## merge(nba.lines, musicians, by = c('locality', 'state', 'season'))
+## nba.lines[musicians, on = c('locality', 'state', 'season')]
+## ### Merge lines data with BLS data.
+## nba.lines <- merge(nba.lines, musicians,
+##                by.x = c('season', 'last.game.loc'),
+##                by.y = c('season', 'team'), all.x = T)
 
@@
 <<echo=F>>= 
-musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)),
-          by = team][order(average.nmusicians)][c(1:5, 25:27)]
+#musicians[, list(average.nmusicians = mean(nmusicians) %>% round(., digits = 1)),
+#          by = team][order(average.nmusicians)][c(1:5, 25:27)]
 @
@@
 nba.lines[, party := party / max(party, na.rm = T)]
-
 @
@@
 ##################################################
-tmp <- musicians[, list(party.continuous = mean(log(nmusicians+1))), 
-                 by = team][order(party.continuous)][!is.na(party.continuous)]
+## tmp <- musicians[, list(party.continuous = mean(log(nmusicians+1))), 
+##                  by = team][order(party.continuous)][!is.na(party.continuous)]
+tmp <- nba.lines[, list(party.continuous = mean(log(nmusicians+1))),
+                 by = team][order(party.continuous)]
 
@@
 
+# mp
 stargazer(mp, covariate.labels = c('Party placebo', 'Lag changes in posession', 
@@
 ##################################################
-load(file = 'tmp_data/alcohol_and_drinking_nba.RData')
-drinking <- copy(dt)
-setnames(drinking, 'variable', 'ndrinks')
-rm(dt)
+## load(file = 'tmp_data/alcohol_and_drinking.RData')
+## drinking <- copy(dt)
+## setnames(drinking, 'variable', 'ndrinks')
+## rm(dt)
 
-tmpm <- nba.lines[, list(log.nmusicians = mean(log(nmusicians+1), na.rm = T),
-                         log.population = mean(log(population+1), na.rm = T)), 
-                  by = team][order(log.nmusicians)] %>% na.omit
-tmpd <- drinking[, list(log.ndrinks = mean(log(ndrinks+1))), 
-                  by = team][order(log.ndrinks)] %>% na.omit
-tmp <- merge(tmpm, tmpd, by = 'team')
-setorder(tmp, log.nmusicians)
+## tmpm <- nba.lines[, list(log.nmusicians = mean(log(nmusicians+1), na.rm = T),
+##                          log.population = mean(log(population+1), na.rm = T)), 
+##                   by = team][order(log.nmusicians)] %>% na.omit
+## tmpd <- drinking[, list(log.ndrinks = mean(log(ndrinks+1))), 
+##                   by = team][order(log.ndrinks)] %>% na.omit
+## tmp <- merge(tmpm, tmpd, by = 'team')
+## setorder(tmp, log.nmusicians)
+
+tmp <- nba.lines[, list(log.nmusicians = mean(log(nmusicians+1)),
+                        log.population = mean(log(population+1)),
+                        log.ndrinks    = mean(log(ndrinks+1))), by = team][order(log.nmusicians)]
 
@@
      ylab = 'Log sum of Variable',
-     xlab = '', axes = F, pch = 1, ylim = c(0, max(tmp$log.population)), type = 'l')
+     xlab = '', axes = F, pch = 1, ylim = c(0, max(tmp$log.population,na.rm = T)), type = 'l')
 points(x = ticks, y = tmp$log.ndrinks, pch = 16, col = 'purple', type = 'l')
@@
 
-HOURS <- 36
-m <- glm(outcome == 'W' ~ party, 
-         data = nba.lines[season < 2016 & last.game.loc != team])
+## HOURS <- 36
+## m <- glm(outcome == 'W' ~ party, 
+##          data = nba.lines[season < 2016 & last.game.loc != team])
 
-### We're only interested in making predictions when teams party.
-p <- predict(m, newdata = nba.lines[season == 2016 & nhours.lgame <= HOURS & 
-                                last.game.loc != team], type = 'response')
+## ### We're only interested in making predictions when teams party.
+## p <- predict(m, newdata = nba.lines[season == 2016 & nhours.lgame <= HOURS & 
+##                                 last.game.loc != team], type = 'response')
 
-idx <- which(p < 0.476)
-out <- table(nba.lines[season == 2016 & nhours.lgame <= HOURS & last.game.loc != team][idx, outcome])
+## idx <- which(p < 0.476)
+## out <- table(nba.lines[season == 2016 & nhours.lgame <= HOURS & last.game.loc != team][idx, outcome])
 
-(out['L'] * 100 - 110*out['W']) %>% as.numeric
+## (out['L'] * 100 - 110*out['W']) %>% as.numeric
 
-table(round(p), nba.lines[season == 2016 & nhours.lgame <= HOURS & 
-                      last.game.loc != team, outcome])
+## table(round(p), nba.lines[season == 2016 & nhours.lgame <= HOURS & 
+##                       last.game.loc != team, outcome])
 
@@
 <<echo=1>>=
-source('functions.r')
-m <- glm(outcome == 'W' ~ party,
-         data = nba.lines[season < 2016 & last.game.loc != team])
+## ## source('functions.r')
+## m <- glm(outcome == 'W' ~ party,
+##          data = nba.lines[season < 2016 & last.game.loc != team])
 
-p <- predict(m, newdata = nba.lines[season == 2016 & nhours.lgame <= HOURS & 
-                                last.game.loc != team], type = 'response')
+## p <- predict(m, newdata = nba.lines[season == 2016 & nhours.lgame <= HOURS & 
+##                                 last.game.loc != team], type = 'response')
 
-table(round(p), nba.lines[season == 2016 & nhours.lgame <= HOURS & 
-                      last.game.loc != team, outcome])
+## table(round(p), nba.lines[season == 2016 & nhours.lgame <= HOURS & 
+##                       last.game.loc != team, outcome])
 
@@
 nba.lines[, demeaned.pts.admitted := team.pts.admitted - mean(team.pts.admitted), by = list(season, team)]
+nba.lines[, party.discrete := grepl("(los angeles)|(new york)|(brooklyn)", last.game.loc) &
+                nhours.lgame <= HOURS &
+                last.game.loc != team]
+
 nba.lines[, mean(demeaned.pts.admitted), by = list(last.game.loc, party.discrete)][order(V1)]
@@
 load(file = 'tmp_data/covers_lines_mlb.RData')
-mlb.lines <- lines
+locs <- MyGeoCode(unique(lines$team), 'mlb')
+states <- data.table(state = state.name, state.abb = state.abb)
+states <- rbind(states, data.frame(state = 'District of Columbia', state.abb = 'DC'))
+locs <- locs[states, on = 'state', nomatch = 0]
+lines <- lines[locs, on = 'team']  # Here, we lose Toronto (that's OK).
+mlb.lines <- lines[demog, on = c('locality', 'state.abb', 'season'), allow.cartesian = T]
 
@@
 
-pdf('writing/next_day_opponent_mlb.pdf', width = 14, height = 14)
-ggplot(tmp, aes(x = location, y = last.game.loc, fill = log(N))) + 
-    geom_tile() +
-    facet_wrap(~lst.game + cur.game) + 
-    labs(x = 'Current Game Location', y = 'Last Game Location', 
-         title = 'Distribution of Next Day Opponent - MLB') +
-    theme_bw() +
-    theme(axis.text.x = element_text(angle = 315, vjust = 0.5),
-          plot.title = element_text(hjust=.5))
-dev.off()
+## pdf('writing/next_day_opponent_mlb.pdf', width = 14, height = 14)
+## ggplot(tmp, aes(x = location, y = last.game.loc, fill = log(N))) + 
+##     geom_tile() +
+##     facet_wrap(~lst.game + cur.game) + 
+##     labs(x = 'Current Game Location', y = 'Last Game Location', 
+##          title = 'Distribution of Next Day Opponent - MLB') +
+##     theme_bw() +
+##     theme(axis.text.x = element_text(angle = 315, vjust = 0.5),
+##           plot.title = element_text(hjust=.5))
+## dev.off()
 
@@
 ##################################################
-load(file = 'tmp_data/studios_and_artists_mlb.RData')
-setnames(dt, 'variable', 'nmusicians')
+## load(file = 'tmp_data/studios_and_artists_mlb.RData')
+## setnames(dt, 'variable', 'nmusicians')
 
-tmp <- dt[, list(party.continuous = mean(log(nmusicians+1))),
+tmp <- mlb.lines[, list(party.continuous = mean(log(nmusicians+1))),
                  by = team][order(party.continuous)][!is.na(party.continuous)]
 
-pdf('writing/Party_by_Team_Location_mlb.pdf', width = 14, height = 8.5)
-ticks = 1:nrow(tmp)
-plot(x = ticks, 
-     y = tmp$party.continuous, 
-     main = 'Continuous measure of party by Team Location: MLB',
-     ylab = 'Log sum of Musicians and Sound Recording Studios',
-     xlab = '',
-     axes = F)
-axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>% 
-                               gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
-axis(side = 2)
-dev.off()
+## pdf('writing/Party_by_Team_Location_mlb.pdf', width = 14, height = 8.5)
+## ticks = 1:nrow(tmp)
+## plot(x = ticks, 
+##      y = tmp$party.continuous, 
+##      main = 'Continuous measure of party by Team Location: MLB',
+##      ylab = 'Log sum of Musicians and Sound Recording Studios',
+##      xlab = '',
+##      axes = F)
+## axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>% 
+##                                gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
+## axis(side = 2)
+## dev.off()
 
@@
                                      team.score, opponent.score, odds)]
-    tmp[, fitted := predict(m, newdata = mlb.lines[year(date) == yr], type = 'response')]
+    tmp[, fitted := predict(m, newdata = mlb.lines[season == yr], type = 'response')]
     ### Realize that we can only make one prediction per game. So take the stronger bet.
@@
 ### There should be 'no information' left over after conditioning on odds.
+mlb.lines <- na.omit(mlb.lines)
 noinf <- lm(I(team.score > opponent.score) ~ odds, data = mlb.lines)
@@
 
+### Musicians by area.
+## load(file = 'tmp_data/studios_and_artists.RData')
+## tmp <- demog[, list(log.nmusicians = log(nmusicians + 1),
+##                       log.ndrinks    = log(ndrinks    + 1),
+##                       log.population = log(population + 1)), by = locality][order(log.nmusicians)]
+## points(demog[, log(nmusicians + 1)])
+
 @

### 2017-07-07 hangover/R/LookupLocation.r
Init commit for a simple helper function.

@@ -1,0 +1,17 @@
+#' Subset data to observations satisfying regex containing location and state.
+#'
+#' We can't simply merge. We have to look for the county within the MSA.
+#' For each location, we fetch a panel of corresponding musicians data, 2010-2016.
+#' @param loc A string describing location to search for.
+#' @param st A string describing the state to search for.
+#' @param data A data.frame containing the observations we wish to extract.
+#' @param team A string to be used as a feature describing team.
+#' @keywords lookup, location
+#' @export
+#' @examples
+LookupLocation <- function(loc, st, data, team = NULL) {
+    sb <- data[grepl(loc, area.title, ignore.case = T) & grepl(st,  area.title)]
+    sb[, `:=`(locality = loc, state = st)]
+    if (!is.null(team)) sb[, team := team]
+    return(sb)
+}

### 2017-07-08 cardinal_analytx_presentation/nightlife_and_sports.tex
Regained significance; added a seasonal trend.

@@
   \tiny{ % m0, m1
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.555$^{***}$ &  \\    & p = 0.0002 &  \\    & & \\   Party continuous &  & $-$0.153 \\    &  & p = 0.107 \\    & & \\   Lag changes in posession & 0.001 & 0.001 \\    & p = 0.346 & p = 0.373 \\    & & \\   Logged travel distance & $-$0.014 & $-$0.009 \\    & p = 0.577 & p = 0.733 \\    & & \\   East-west travel direction & $-$0.074 & $-$0.039 \\    & p = 0.754 & p = 0.867 \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & p = 0.218 & p = 0.295 \\    & & \\   Time of game during day & 0.007 & 0.005 \\    & p = 0.603 & p = 0.715 \\    & & \\   Home team effect & 0.002 & 0.004 \\    & p = 0.960 & p = 0.936 \\    & & \\   Logged travel distance * east-west & 0.012 & 0.007 \\    & p = 0.749 & p = 0.846 \\    & & \\   Constant & $-$0.247 & $-$0.229 \\    & p = 0.443 & p = 0.477 \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,587.456 & $-$6,593.324 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.557$^{***}$ &  \\    & (0.149) &  \\    & & \\   Party continuous &  & $-$0.158$^{*}$ \\    &  & (0.095) \\    & & \\   Lag changes in posession & 0.189 & 0.181 \\    & (0.140) & (0.140) \\    & & \\   Logged travel distance & $-$0.015 & $-$0.010 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.072 & $-$0.037 \\    & (0.234) & (0.234) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.006 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.002 & 0.004 \\    & (0.044) & (0.044) \\    & & \\   Logged travel distance * east-west & 0.011 & 0.007 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.923 & $-$0.875 \\    & (0.695) & (0.695) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,586.320 & $-$6,592.159 \\  Akaike Inf.
 Crit.
-& 13,192.910 & 13,204.650 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
+& 13,200.640 & 13,212.320 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+}   
 \end{frame}
@@
   \tiny{ % mp
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.056 \\    & p = 0.504 \\    & \\   Lag changes in posession & 0.001 \\    & p = 0.408 \\    & \\   Logged travel distance & $-$0.003 \\    & p = 0.920 \\    & \\   East-west travel direction & $-$0.036 \\    & p = 0.880 \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & p = 0.098 \\    & \\   Time of game during day & 0.003 \\    & p = 0.827 \\    & \\   Home team effect & $-$0.006 \\    & p = 0.896 \\    & \\   Logged travel distance * east-west & 0.006 \\    & p = 0.866 \\    & \\   Constant & $-$0.257 \\    & p = 0.425 \\    & \\  \hline \\[-1.8ex]  Observations & 9,517 \\  Log Likelihood & $-$6,594.404 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.053 \\    & (0.084) \\    & \\   Lag changes in posession & 0.170 \\    & (0.140) \\    & \\   Logged travel distance & $-$0.004 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.034 \\    & (0.235) \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   Time of game during day & 0.002 \\    & (0.013) \\    & \\   Home team effect & $-$0.006 \\    & (0.044) \\    & \\   Logged travel distance * east-west & 0.006 \\    & (0.036) \\    & \\   Constant & $-$0.865 \\    & (0.695) \\    & \\  \hline \\[-1.8ex]  Observations & 9,517 \\  Log Likelihood & $-$6,593.344 \\  Akaike Inf.
 Crit.
-& 13,206.810 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}      }
+& 13,214.690 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+      }
 \end{frame}
@@
   \tiny{ % m3, m4
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.120$^{*}$ &  \\    & p = 0.091 &  \\    & & \\   Nightlife (no weekend interaction) &  & 0.137 \\    &  & p = 0.307 \\    & & \\   Bookmaker's odds & 2.682$^{***}$ & 2.669$^{***}$ \\    & p = 0.000 & p = 0.000 \\    & & \\   Home-team effect & 0.014 & 0.144 \\    & p = 0.674 & p = 0.159 \\    & & \\   Number of rest days & 0.018 & 0.016 \\    & p = 0.399 & p = 0.456 \\    & & \\   Logged travel distance & $-$0.005 & $-$0.003 \\    & p = 0.377 & p = 0.519 \\    & & \\   Weekend & 0.044 & 0.002 \\    & p = 0.246 & p = 0.935 \\    & & \\   Constant & $-$1.360$^{***}$ & $-$1.468$^{***}$ \\    & p = 0.000 & p = 0.000 \\    & & \\  \hline \\[-1.8ex]  Observations & 26,473 & 26,473 \\  Log Likelihood & $-$18,124.390 & $-$18,125.300 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.120$^{*}$ &  \\    & (0.071) &  \\    & & \\   Nightlife (no weekend interaction) &  & 0.137 \\    &  & (0.134) \\    & & \\   Bookmaker's odds & 2.682$^{***}$ & 2.669$^{***}$ \\    & (0.158) & (0.158) \\    & & \\   Home-team effect & 0.014 & 0.144 \\    & (0.034) & (0.102) \\    & & \\   Number of rest days & 0.018 & 0.016 \\    & (0.021) & (0.021) \\    & & \\   Logged travel distance & $-$0.005 & $-$0.003 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.044 & 0.002 \\    & (0.038) & (0.029) \\    & & \\   Constant & $-$1.363$^{***}$ & $-$1.468$^{***}$ \\    & (0.084) & (0.122) \\    & & \\  \hline \\[-1.8ex]  Observations & 26,473 & 26,473 \\  Log Likelihood & $-$18,124.350 & $-$18,125.300 \\  Akaike Inf.
 Crit.
-& 36,262.780 & 36,264.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
+& 36,272.700 & 36,264.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  }
 \end{frame}
@@
   \begin{adjustwidth}{-.75cm}{}
-  \tiny{  % mpa
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & p = 0.0001 &  &  \\    & & & \\   Continuous party measure &  & 1.803$^{***}$ & $-$0.160 \\    &  & p = 0.007 & p = 0.806 \\    & & & \\   Lag change in possessions &  & 0.118$^{***}$ & 0.115$^{***}$ \\    &  & p = 0.000 & p = 0.000 \\    & & & \\   Logged travel distance &  & $-$0.568$^{***}$ & 0.037 \\    &  & p = 0.005 & p = 0.852 \\    & & & \\   Number hours since last game &  & 0.011 & 0.018$^{**}$ \\    &  & p = 0.115 & p = 0.013 \\    & & & \\   Constant & $-$0.046 & 89.567$^{***}$ & 86.136$^{***}$ \\    & p = 0.612 & p = 0.000 & p = 0.000 \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 6,234 & 6,234 \\  R$^{2}$ & 0.001 & 0.023 & 0.021 \\  Adjusted R$^{2}$ & 0.001 & 0.023 & 0.021 \\  Residual Std.
-Error & 11.137 (df = 15137) & 12.071 (df = 6229) & 11.910 (df = 6229) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 37.032$^{***}$ (df = 4; 6229) & 33.756$^{***}$ (df = 4; 6229) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
+  \tiny{  % mpa, tpa, tps
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std.
+Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}      }
 \end{adjustwidth} 

### 2017-07-08 writing/draft.tex
Regained significance; added a seasonal trend.

--- 2017-07-05 writing/draft.tex
@@
 % m0, m1
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.513$^{***}$ &  \\    & (0.145) &  \\    & & \\   Party continuous &  & $-$0.181$^{**}$ \\    &  & (0.091) \\    & & \\   Lag changes in posession & 0.0001 & 0.00004 \\    & (0.002) & (0.002) \\    & & \\   Logged travel distance & $-$0.017 & $-$0.014 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.145 & $-$0.113 \\    & (0.232) & (0.231) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.005 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.001 & 0.004 \\    & (0.043) & (0.043) \\    & & \\   Logged travel distance * east-west & 0.022 & 0.018 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.048 & $-$0.028 \\    & (0.378) & (0.378) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,823 & 9,823 \\  Log Likelihood & $-$6,800.677 & $-$6,805.100 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.557$^{***}$ &  \\    & (0.149) &  \\    & & \\   Party continuous &  & $-$0.158$^{*}$ \\    &  & (0.095) \\    & & \\   Lag changes in posession & 0.189 & 0.181 \\    & (0.140) & (0.140) \\    & & \\   Logged travel distance & $-$0.015 & $-$0.010 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.072 & $-$0.037 \\    & (0.234) & (0.234) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.006 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.002 & 0.004 \\    & (0.044) & (0.044) \\    & & \\   Logged travel distance * east-west & 0.011 & 0.007 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.923 & $-$0.875 \\    & (0.695) & (0.695) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,586.320 & $-$6,592.159 \\  Akaike Inf.
 Crit.
-& 13,619.350 & 13,628.200 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 13,200.640 & 13,212.320 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 % m3, m4
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.126$^{*}$ &  \\    & (0.074) &  \\    & & \\   Nightlife (no weekend interaction) &  & $-$0.016 \\    &  & (0.110) \\    & & \\   Bookmaker's odds & 2.638$^{***}$ & 2.583$^{***}$ \\    & (0.163) & (0.165) \\    & & \\   Home-team effect & 0.008 & 0.034 \\    & (0.035) & (0.083) \\    & & \\   Number of rest days & 0.019 & 0.018 \\    & (0.020) & (0.020) \\    & & \\   Logged travel distance & $-$0.002 & $-$0.003 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.044 & 0.003 \\    & (0.038) & (0.030) \\    & & \\   Constant & $-$1.344$^{***}$ & $-$1.324$^{***}$ \\    & (0.081) & (0.113) \\    & & \\  \hline \\[-1.8ex]  Observations & 24,734 & 24,033 \\  Log Likelihood & $-$16,946.080 & $-$16,470.850 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.120$^{*}$ &  \\    & (0.071) &  \\    & & \\   Nightlife (no weekend interaction) &  & 0.137 \\    &  & (0.134) \\    & & \\   Bookmaker's odds & 2.682$^{***}$ & 2.669$^{***}$ \\    & (0.158) & (0.158) \\    & & \\   Home-team effect & 0.014 & 0.144 \\    & (0.034) & (0.102) \\    & & \\   Number of rest days & 0.018 & 0.016 \\    & (0.021) & (0.021) \\    & & \\   Logged travel distance & $-$0.005 & $-$0.003 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.044 & 0.002 \\    & (0.038) & (0.029) \\    & & \\   Constant & $-$1.363$^{***}$ & $-$1.468$^{***}$ \\    & (0.084) & (0.122) \\    & & \\  \hline \\[-1.8ex]  Observations & 26,473 & 26,473 \\  Log Likelihood & $-$18,124.350 & $-$18,125.300 \\  Akaike Inf.
 Crit.
-& 33,906.160 & 32,955.700 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+& 36,272.700 & 36,264.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
@@
 % mp
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.072 \\    & (0.082) \\    & \\   Lag changes in posession & 0.00001 \\    & (0.002) \\    & \\   Logged travel distance & $-$0.005 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.099 \\    & (0.231) \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   Time of game during day & 0.001 \\    & (0.013) \\    & \\   Home team effect & $-$0.009 \\    & (0.043) \\    & \\   Logged travel distance * east-west & 0.015 \\    & (0.036) \\    & \\   Constant & $-$0.069 \\    & (0.378) \\    & \\  \hline \\[-1.8ex]  Observations & 9,823 \\  Log Likelihood & $-$6,806.701 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.053 \\    & (0.084) \\    & \\   Lag changes in posession & 0.170 \\    & (0.140) \\    & \\   Logged travel distance & $-$0.004 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.034 \\    & (0.235) \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   Time of game during day & 0.002 \\    & (0.013) \\    & \\   Home team effect & $-$0.006 \\    & (0.044) \\    & \\   Logged travel distance * east-west & 0.006 \\    & (0.036) \\    & \\   Constant & $-$0.865 \\    & (0.695) \\    & \\  \hline \\[-1.8ex]  Observations & 9,517 \\  Log Likelihood & $-$6,593.344 \\  Akaike Inf.
 Crit.
-& 13,631.400 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
+& 13,214.690 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 \subsection{Possible confounders or potential issues}
@@
 
-\paragraph{Rebounds: the ``hustle'' statistic}
-In particular, we see that the average number of rebounds obtained per minute
-of game-play does decrease as a function of being exposed to a city with active night-life. Rebounds are often considered a ``hustle-statistic'' in the sense 
-that anecdotally, players who work harder on the court and expend energy boxing out
-other players will have higher rebound rates. If this common knowledge is accepted,
-our analysis suggests that players are working less hard following travel to a 
-city with active nightlife.
+% \paragraph{Rebounds: the ``hustle'' statistic}
+% In particular, we see that the average number of rebounds obtained per minute
+% of game-play does decrease as a function of being exposed to a city with active % night-life. Rebounds are often considered a ``hustle-statistic'' in the sense 
+% that anecdotally, players who work harder on the court and expend energy boxing % out other players will have higher rebound rates.
+If this common knowledge is % accepted, our analysis might suggest that players are working less hard following travel to a
+% city with active nightlife.
 
-The method was to first calculate for each player and each game the number of rebounds obtained
-per minute of game-play. Then, we subtract from each observation the player's average number of rebounds per minute taken across the entire season.
-This is an ad-hoc approach that cannot be used for betting, since at game-time we are relying on future data for use in our average statistic which we must calculate and sweep out from each observation.
+% The method was to first calculate for each player and each game the number of % rebounds obtained
+% per minute of game-play. Then, we subtract from each observation the player's % average number of rebounds per minute taken across the entire season.
+% This is an ad-hoc approach that cannot be used for betting, since at game-time % we are relying on future data for use in our average statistic which we must % calculate and sweep out from each observation.
 We then run a linear regression
-explaining demeaned rebounds per minute as a function of our nightlife variable of interest
-as well as our usual features accounting for player-fatigue.
+% explaining demeaned rebounds per minute as a function of our nightlife variable % of interest
+% as well as our usual features accounting for player-fatigue.
 
-% m2
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.003$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00003 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00003 \\    & (0.00002) \\    & \\   Constant & $-$0.008$^{*}$ \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 43,547 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
-Error & 0.104 (df = 43542) \\  F Statistic & 2.945$^{**}$ (df = 4; 43542) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+% % m2
+% \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & % \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & % Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & % $-$0.004$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00002 % \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    % & \\   Number hours since last game & 0.00002 \\    & (0.00002) \\    & \\   % Constant & $-$0.007 \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations % & 40,684 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
+% Error & 0.104 (df = 40679) \\  F Statistic & 2.696$^{**}$ (df = 4; 40679) \\  % \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; % $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 
 \paragraph{The hangover correlates with allowing more points}
-Similarly, common knowledge also suggests that the overall defense of a team
+Common knowledge suggests that the overall defense of a team
 is effected strongly by individual player effort. It seems likely that a hangover effect
@@
 It's also interesting to see that after visiting LA or NY, for any given
-team their next day performance is strongly correlated with allowing the other team to
-score two and a half points more. 
-It's even more interesting to note that points spreads are on a similar order of magnitude,  
-i.e. this could often be the difference between meeting the spread or not.
+team their next day performance is strongly correlated with allowing the other team to score two and a half points more. 
+It's even more interesting to note that points spreads are on a similar order of magnitude, i.e. this could often be the difference between meeting the spread or not.
 
 % mpa1, tpa, tpo
-\hspace{-100pt}
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.828$^{***}$ &  &  \\    & (0.699) &  &  \\    & & & \\   Continuous party measure &  & 1.667$^{***}$ & 0.080 \\    &  & (0.587) & (0.575) \\    & & & \\   Lag change in possessions &  & 0.267$^{***}$ & 0.285$^{***}$ \\    &  & (0.016) & (0.016) \\    & & & \\   Logged travel distance &  & $-$0.447$^{**}$ & 0.143 \\    &  & (0.191) & (0.187) \\    & & & \\   Number hours since last game &  & 0.011 & 0.017$^{**}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Constant & $-$0.047 & 70.710$^{***}$ & 64.859$^{***}$ \\    & (0.090) & (2.271) & (2.223) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,658 & 6,426 & 6,426 \\  R$^{2}$ & 0.001 & 0.045 & 0.051 \\  Adjusted R$^{2}$ & 0.001 & 0.044 & 0.051 \\  Residual Std.
-Error & 11.159 (df = 15656) & 11.941 (df = 6421) & 11.688 (df = 6421) \\  F Statistic & 16.357$^{***}$ (df = 1; 15656) & 74.877$^{***}$ (df = 4; 6421) & 86.792$^{***}$ (df = 4; 6421) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
+\hspace{-125pt}
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std.
+Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 \section{Conclusions and Avenues for Future Research}
@@
 
-\begin{figure}   
-  \centering
+\begin{figure}[h]
   \hspace{-55pt}
@@
   \includegraphics[scale=0.5]{Party_by_Team_Location_mlb}
-  \caption{Continuous measure of nightlife by team location for MLB.
-What's interesting here is that because New York Yankees are in the Bronx, they show up as having few musicians and sound recording studios.
-We might wonder if this is a bad thing, since with NBA New York city was deemed a major party city.
-It turns out that when we define a discrete indicator of party based on New York or Los Angeles, it is not at all significant, and in fact the sign is incorrect.
-When we use our continuous
- measure of nightlife, its significant, and it helps inform our profitable betting model.
-This is good news for our proxy, although begs the question why New York should be treated so differently by MLB players when compared with NBA players.
-One point worth noting is that we haven't yet made the important distinction that in the MLB, the Los Angeles Angels are in Orange County, for example, which is much less bustling than the NBA's  Los Angeles Lakers who are located in downtown Los Angeles.}
+  \caption{Continuous measure of nightlife by team location for MLB.}
 \end{figure}
-
+  % What's   % interesting here is that because New York Yankees are in the Bronx, they show up % as having few musicians and sound recording studios.
+We might wonder if this is % a bad thing, since with NBA New York city was deemed a major party city.
+It % turns out that when we define a discrete indicator of party based on New York or % Los Angeles, it is not at all significant, and in fact the sign is incorrect.
+% When we use our continuous
+%  measure of nightlife, its significant, and it helps inform our profitable % betting model.
+% This is good news for our proxy, although begs the question why New York should % be treated so differently by MLB players when compared with NBA players.
+One % point worth noting is that we haven't yet made the important distinction that in % the MLB, the Los Angeles Angels are in Orange County, for example, which is much % less bustling than the NBA's  Los Angeles Lakers who are located in downtown Los % Angeles.
 \begin{figure}   
@@
     and even in spite of the extremal observations being ranked consistently across each statistic, 
-    i.e. Charlotte Bobcats toward the bottom and Los Angeles more or less at the top,  
+    i.e. Memphis toward the bottom and Los Angeles more or less at the top,  
     the problem is that population and number of drinking establishments have sinusoidal 
@@
     that do not properly reflect variation in nightlife between cities. For example, both 
-    incorrectly rank Detroit Michigan as having a high nightlife index. Another example, 
-    population incorrectly places New Orleans, Louisiana as having a low nightlife index, 
-    when in reality it is famous for its vibrant city life.}
+    incorrectly rank San Antonio, Texas as having one of the higher nightlife indices. Another example, 
+    population incorrectly places New Orleans, Louisiana as having a low     nightlife index, when in reality it is famous for its vibrant city life.}
 \end{sidewaysfigure}

### 2017-07-08 clean_espn.r

@@ -94,7 +94,7 @@ data[, (paste0('demeaned.', cols)) :=
 ##################################################
 ### Remove players who only show up < 10 times per season...
 ##################################################
-insufficient.obs <- data[, .N, keyby = list(season, player)][N < 10]
+insufficient.obs <- data[, .N, keyby = list(season, player)][N < 15]
 data <- data[!insufficient.obs, on = c('season', 'player')]
 
 data[, position := gregexpr(pattern = '[A-Z]{1,2}$', player) %>%
@@ -111,4 +111,3 @@ data[, (paste0('lag.', cols)) := shift(.SD),
      by = list(season, player), .SDcols = cols]
 
 write.csv(data, file = 'tmp_data/espn_player_data.csv', row.names = F)
-

### 2017-07-08 init_draft.r

@@ -33,7 +33,7 @@ load(file = 'tmp_data/mlb.RData')  # mlb.lines
 
 ### Discrete indicator.
 m0 <- glm(outcome == 'W' ~ party.discrete + lag.chg.pos + log(travel.dist+1)*ew + 
-             nhours.lgame + I(hour(date)) + I(team==location),
+             nhours.lgame + I(hour(date)) + I(team==location) + as.factor(season),
          data = nba.lines[season < 2017 & !is.na(party)], family = 'binomial',
          na.action = 'na.exclude')
 
@@ -44,25 +44,26 @@ m1 <- update(m0, . ~ party + . - party.discrete)
 mp <- update(m0, . ~ party.placebo + . - party.discrete)
 
 ### Player specific model - demeaned rebounds per minute.
-m2 <- lm(demeaned.reb.per.min ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame,
-   data = espn[last.game.loc != team & season < 2017] %>% na.omit)
+## m2 <- lm(demeaned.reb.per.min ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame +
+##              as.factor(season),
+##    data = espn[last.game.loc != team] %>% na.omit)
 
 ### Total points admitted and scored.
-tpa <- update(m2, team.pts.admitted ~ .)
+tpa <- lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame,
+   data = espn[last.game.loc != team] %>% na.omit)
 tps <- update(tpa, team.pts.scored ~ .)
 
 ### Points  allowed as afunction of last game location (by party)
 nba.lines[, demeaned.pts.admitted := team.pts.admitted - mean(team.pts.admitted),
           by = list(season, team)]
 mpa1 <- lm(demeaned.pts.admitted ~ party.discrete, data = nba.lines)
-mpa2 <- lm(demeaned.pts.admitted ~ party,          data = nba.lines)
 
 ##################################################
 ### MLB Modeling
 ##################################################
 
 m3 <- glm(I(team.score > opponent.score) ~ party + odds + I(team == location) + 
-             ndays.lgame + log(travel.dist+1) + weekend,
+             ndays.lgame + log(travel.dist+1) + weekend + as.factor(season),
          data = mlb.lines[season < 2017], family = 'binomial')
 summary(m3)
 m4 <- update(m3, . ~ party.now + . - party)
[…]

### 2017-07-08 init_draft.r
Combined plot of party-measure by location.

@@ -160,39 +160,34 @@ stargazer(m3, m4, covariate.labels = c('Continuous measure of nightlife',
 ##################################################
 ### Dot plot of Party-Continuous by Team Location
 ##################################################
-tmp <- nba.lines[, list(party.continuous = mean(log(nmusicians+1))),
-                 by = team][order(party.continuous)]
+
+data <- rbind(nba.lines[, list(avg.party = mean(log(nmusicians+1), na.rm = T),
+                               sport = 'nba'), by = .(team, locality, state.abb)],
+              mlb.lines[, list(avg.party = mean(log(nmusicians+1), na.rm = T),
+                               sport = 'mlb'), by = .(team, locality, state.abb)])
+setkey(data, avg.party)
+data[, idx := .GRP, by = locality]
 
 pdf('writing/Party_by_Team_Location.pdf', width = 14, height = 8.5)
-ticks = 1:nrow(tmp)
-plot(x = ticks, 
-     y = tmp$party.continuous, 
-     main = 'Continuous measure of party by Team Location: NBA',
+plot(x = data[sport == 'nba', jitter(idx)],
+     y = data[sport == 'nba', avg.party],
+     main = 'Continuous measure of party by Location',
      ylab = 'Log sum of Musicians and Sound Recording Studios',
      xlab = '',
-     axes = F)
-axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>% 
-                               gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
+     axes = F, pch = 24, col = 'orange', lwd = 5)
+axis(side = 1, at = data[sport == 'nba', idx],
+     labels = gsub('([a-z]{4,}) ', '\\1\n', data[sport == 'nba', locality]) %>% 
+         gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
 axis(side = 2)
+points(x = data[sport == 'mlb', jitter(idx)],
+       y = data[sport == 'mlb', avg.party], pch = 25, col = 'purple', lwd = 5)
+axis(side = 1, at = data[sport == 'mlb', idx],
+     labels = gsub('([a-z]{4,}) ', '\\1\n', data[sport == 'mlb', locality]) %>% 
+         gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
[…]

### 2017-07-08 cardinal_analytx_presentation/nightlife_and_sports.tex
Removed demeaned.rebounds.per.minute, added section on interpreting logistic regression results.

@@
 \begin{frame}   
-  \frametitle{Continuous measure of party (NBA)}
+  \frametitle{Continuous measure of party}
   \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location} 
@@
 }   
+\end{frame}
+
+\begin{frame}   \frametitle{Interpreting our results}
+  Let's consider the simplest model, where $Y_i | X \sim \textrm{Bern}(p)$, i.e.
+  $\Pr(Y_i = y | X) = p^{y} (1-p)^{1-y}$. Using a generalized linear model,
+  we fit a linear predictor to a transformation of the expected value of our
+  variable of interest. I.e.
+  \[
+    \textrm{logit} \left( \mathbb E[Y_i | X]\right) = \textrm{logit} \left(p\right) = \ln \left( \frac{p}{1-p}\right) = X \beta.
+  \]
+  So, we interpret our regression coefficient as follows: visiting LA or NY within
+  24 hours of another game corresponds to being $e^{\beta_1} \approx e^{-0.55} \approx \frac{3}{5}$
+  as likely to meet the spread, all other factors held constant. 
 \end{frame}
@@
 
-\begin{frame}   
-  \frametitle{Continuous measure of party (MLB)}
-  \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location_mlb}   
-\end{frame}
+% \begin{frame}   
+%   \frametitle{Continuous measure of party (MLB)}
+%   \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location_mlb}   
+% \end{frame}
 
@@
 
-\begin{frame}   \frametitle{Analysis: rebounds}
-  \begin{itemize}     \item For each player and each game, we calculate the number of rebounds obtained per-minute of game-play
-    \item From each observation, we subtract the player's average number of rebounds per minute taken across the entire season.
-    \item We then run a linear regression explaining demeaned rebounds as a function of 
-      nightlife visitation and other features describing player fatigue.   \end{itemize} \end{frame}
+% \begin{frame}   \frametitle{Analysis: rebounds}
+%   \begin{itemize}     \item For each player and each game, we calculate the % number of rebounds obtained per-minute of game-play
+%     \item From each observation, we subtract the player's average number of % rebounds per minute taken across the entire season.
+%     \item We then run a linear regression explaining demeaned rebounds as a % function of 
+%       nightlife visitation and other features describing player fatigue.   % \end{itemize} \end{frame}
 
-\begin{frame}   \frametitle{Results: rebounds}
-  \centering
-  \tiny{ % m2
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & $-$0.004$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00002 \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    & \\   Number hours since last game & 0.00002 \\    & (0.00002) \\    & \\   Constant & $-$0.007 \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations & 40,684 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
-Error & 0.104 (df = 40679) \\  F Statistic & 2.696$^{**}$ (df = 4; 40679) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  } \end{frame}
+% \begin{frame}   \frametitle{Results: rebounds}
+%   \centering
+%   \tiny{ % m2
+% \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & % \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & % Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & % $-$0.004$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00002 % \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    % & \\   Number hours since last game & 0.00002 \\    & (0.00002) \\    & \\   % Constant & $-$0.007 \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations % & 40,684 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
+% Error & 0.104 (df = 40679) \\  F Statistic & 2.696$^{**}$ (df = 4; 40679) \\  % \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; % $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  } \end{frame}
 

### 2017-07-08 cardinal_analytx_presentation/nightlife_and_sports.tex
Changed where 'prev research' slides placed. Updated tables and formatting.

@@
 
+\begin{frame}
+  \frametitle{Previous research - NBA player shooting average}
+  \begin{block}{Empirical evidence: players party to the detriment of the game.}
+  NBA Players under-perform proportion of shots made on Sundays.\footnote{Chase, ('15).}
+  \end{block}
+    \begin{itemize}       
+    \item 1,092 of 2,035 NBA players who took $\geq 1$ shot on Sundays under-perform 
+      relative to their own season-long average; $p=0.001$.
+      \item Effect size is not considered.
+      \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, is that relevant?}     \end{itemize}
+\end{frame}
+
+\begin{frame}   \frametitle{Prev. research - likelihood of meeting the spread in NBA}
+  \begin{block}{Empirical evidence: away teams under-perform on Sundays.}
+  Teams under-perform bookmakers expectations on Sunday afternoons when on the road.\footnote{Ezekowitz, ('14).}
+    \begin{itemize}       
+    \item Only 129 out of 284 games did the away team meet the spread, 
+      using 2006-14 data; $p=0.1089$.
+    \item Not quite significant at the 90\% confidence level.     
+    \item If we bet \$100 on each Sunday game, we'd only profit \$1,310.\footnote{$(284 - 129) \times 100 + 129 \times (-110) = 1,310$.} 
+    \end{itemize}
+  \end{block}
+  \end{frame}
+
 % \begin{frame}   
@@
     \item injuries, and
-    \item even possibly a weather forecast.     \end{itemize}
+    \item even possibly a weather forecast (for outdoor sports).     \end{itemize}
 
@@
     $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}       \end{block}
-\end{frame}
-
-\begin{frame}
-  \frametitle{Previous research - NBA player shooting average}
-  \begin{block}{Empirical evidence: players party to the detriment of the game.}
-  NBA Players under-perform proportion of shots made on Sundays.\footnote{Chase, ('15).}
-  \end{block}
-    \begin{itemize}       
-    \item 1,092 of 2,035 NBA players who took $\geq 1$ shot on Sundays under-perform 
-      relative to their own season-long average; $p=0.001$.
-      \item Effect size is not considered.
-      \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, is that relevant?}     \end{itemize}
-\end{frame}
-
-\begin{frame}   \frametitle{Prev. research - likelihood of meeting the spread in NBA}
-  \begin{block}{Empirical evidence: away teams under-perform on Sundays.}
-  Teams under-perform bookmakers expectations on Sunday afternoons when on the road.\footnote{Ezekowitz, ('14).}
-    \begin{itemize}       
-    \item Only 129 out of 284 games did the away team meet the spread, 
-      using 2006-14 data; $p=0.1089$.
-    \item Not quite significant at the 90\% confidence level.     
-    \item If we bet \$100 on each Sunday game, we'd only profit \$1,310.\footnote{$(284 - 129) \times 100 + 129 \times (-110) = 1,310$.} 
-    \end{itemize}
-  \end{block}
 \end{frame}
@@
 
-\begin{frame}   \frametitle{Partying as a latent variable}
+\begin{frame}   \frametitle{Recall: partying is a latent variable}
   \begin{block}{We can't observe when players go for a night on the town.}
@@
   \frametitle{Continuous measure of party}
-  \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location} 
+  \centering \includegraphics[scale=0.375]{../writing/Party_by_Team_Location} 
 \end{frame}
@@
   \tiny{ % m0, m1
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.557$^{***}$ &  \\    & (0.149) &  \\    & & \\   Party continuous &  & $-$0.158$^{*}$ \\    &  & (0.095) \\    & & \\   Lag changes in posession & 0.189 & 0.181 \\    & (0.140) & (0.140) \\    & & \\   Logged travel distance & $-$0.015 & $-$0.010 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.072 & $-$0.037 \\    & (0.234) & (0.234) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.006 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.002 & 0.004 \\    & (0.044) & (0.044) \\    & & \\   Logged travel distance * east-west & 0.011 & 0.007 \\    & (0.036) & (0.036) \\    & & \\   Constant & $-$0.923 & $-$0.875 \\    & (0.695) & (0.695) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,586.320 & $-$6,592.159 \\  Akaike Inf.
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.557$^{***}$ &  \\    & (0.149) &  \\    & & \\   Party continuous &  & $-$0.158$^{*}$ \\    &  & (0.095) \\    & & \\   Lag changes in posession & 0.189 & 0.181 \\    & (0.140) & (0.140) \\    & & \\   Logged travel distance & $-$0.015 & $-$0.010 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.072 & $-$0.037 \\    & (0.234) & (0.234) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.006 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.002 & 0.004 \\    & (0.044) & (0.044) \\    & & \\   Logged travel distance * east-west & 0.011 & 0.007 \\    & (0.036) & (0.036) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,586.320 & $-$6,592.159 \\  Akaike Inf.
 Crit.
-& 13,200.640 & 13,212.320 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-}   
+& 13,200.640 & 13,212.320 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  }
 \end{frame}
@@
   \[
-    \textrm{logit} \left( \mathbb E[Y_i | X]\right) = \textrm{logit} \left(p\right) = \ln \left( \frac{p}{1-p}\right) = X \beta.
+    \textrm{logit} \left( \mathbb E[Y_i | X_i]\right) = \textrm{logit} \left(p\right) = \ln \left( \frac{p}{1-p}\right) = X_i^T \beta.
   \]
@@
   \tiny{  % mpa, tpa, tps
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std.
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std.
 Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  }
@@
 \begin{frame}   \frametitle{Conclusions}
-  \begin{itemize}     \item NBA and MLB players under-perform bookmakers expectations after visiting LA and NY, and in general we see that performance decreases as a function of (our proxy for) nightlife.
-    \item The decrease in likelihood can (in part) be explained by a drop in defensive performance.
-    \item With a causal model established, we realize a profitable betting scheme.   \end{itemize} \end{frame}
+  \begin{itemize}     \item NBA and MLB players under-perform bookmakers expectations after visiting LA and NY.
+    \item In general we see that performance decreases as a function of (our proxy for) nightlife.
+\item The decrease in likelihood can (in part) be explained by a drop in defensive performance in the NBA.
+    \item With a causal model established, we realize a profitable betting scheme in MLB.   \end{itemize} \end{frame}
 

### 2017-07-08 init_draft.r
Stargazer table formatting; added a required dependency (ggplot2).

@@ -11,7 +11,7 @@
 ### Inputs: 'tmp_data/nba.RData'
 ###         'tmp_data/mlb.RData'
 ###
-### Dependencies: data.table, magrittr, stargazer
+### Dependencies: data.table, ggplot2, magrittr, stargazer
 ###
 ################################################################################
 ################################################################################
@@ -21,6 +21,7 @@
 ##################################################
 
 require(data.table)
+require(ggplot2)
 require(magrittr)
 require(stargazer)
 
@@ -121,7 +122,7 @@ stargazer(m0, m1, covariate.labels = c('Party discrete', 'Party continuous',
                                        'Home team effect', 
                                        'Logged travel distance * east-west', 
                                        'Constant'),
-          omit = 'as.factor\\(season\\)',
+          omit = '(as.factor)|(Constant)',
           dep.var.labels = 'Meet the Spread')
 
 stargazer(mp, covariate.labels = c('Party placebo', 'Lag changes in posession', 
@@ -140,7 +141,8 @@ stargazer(mp, covariate.labels = c('Party placebo', 'Lag changes in posession',
 ##                                'Number hours since last game',
 ##                                'Constant'))
 
-stargazer(mpa1, tpa, tps, dep.var.labels = c('Points Admitted by Team', 'Team Points Admitted', 'Team Points Scored'),
+stargazer(mpa1, tpa, tps,
+          dep.var.labels = c('Team Points Admitted', 'Team Points Admitted', 'Team Points Scored'),
           covariate.labels = c('Discrete party indicator',
                                'Continuous party measure',
                                'Lag change in possessions',
@@ -168,22 +170,23 @@ data <- rbind(nba.lines[, list(avg.party = mean(log(nmusicians+1), na.rm = T),
 setkey(data, avg.party)
 data[, idx := .GRP, by = locality]
 
-pdf('writing/Party_by_Team_Location.pdf', width = 14, height = 8.5)
+pdf('writing/Party_by_Team_Location.pdf', width = 13, height = 8.5)
 plot(x = data[sport == 'nba', jitter(idx)],
      y = data[sport == 'nba', avg.party],
-     main = 'Continuous measure of party by Location',
[…]

### 2017-07-08 writing/draft.tex
Using combined plot of party-measure by location.
Added a 'meet the spread' figure to the 'logic for using betting metrics' section. Updated tables.

@@
   \caption{We examine the relationship between last game location and next day opponent in the NBA, sorted by distance from New York City. 
-Starting from the top left, notice that the most likely scenario across all seasons is that a game will be played at home, and the last game was also at home.
+Top left: notice that the most likely scenario across all seasons is that a game will be played at home, and the last game was also at home.
 This makes sense; 1/2 of all games are played at home, and 1/2 are played away, so clearly each team is most likely to be found playing a home game for any given game.
@@
 whether any players are injured, the momentum of success each team is carrying
-in the season, etc. To that end, we utilize spread data.\citep{anderson} The benefit of
-using spread data is that it controls for a lot, but the downside is
+in the season, etc. To that end, we utilize spread data.\citep{anderson} The benefit of using spread data is that it controls for a lot, but the downside is
 we don't know exactly what it controls for.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[scale=0.65]{meet_the_spread}
+  \label{fig: meet the spread}
+  \caption{Empirically, in NBA any given team is apt to meet the spread
+half the time across a particular season; in MLB, after conditioning on bookmakers' odds of
+success, successful prediction of game outcome is like a coin-flip. 
+In general, systematic deviations represent
+unexplained mechanisms. }
+\end{figure}
 
@@
   \includegraphics[scale=0.6,angle=270]{Party_by_Team_Location}
-  \caption{We plot our continuous measure of nightlife for each city within the NBA. We present the analagous plot for MLB in the appendix.\ref{fig: party by team location mlb}}
+  \caption{We plot our continuous measure of nightlife for each city within the NBA and MLB. We add a small amount of random noise to our x-coordinate such that cities
+with both an NBA and MLB team are visible.}
 \end{figure}
@@
 \hspace{-125pt}
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Points Admitted by Team & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std.
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std.
 Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+
 \section{Conclusions and Avenues for Future Research}
@@
 \section{Appendix}
-
+% \begin{figure}[h]
+%   \hspace{-55pt}
+%   \label{fig: party by team location mlb}
+%   \includegraphics[scale=0.5]{Party_by_Team_Location_mlb}
+%   \caption{Continuous measure of nightlife by team location for MLB.}
+% \end{figure}
+%   % What's   % interesting here is that because New York Yankees are in the % Bronx, they show up % as having few musicians and sound recording studios.
+We % might wonder if this is % a bad thing, since with NBA New York city was deemed a % major party city.
+It % turns out that when we define a discrete indicator of % party based on New York or % Los Angeles, it is not at all significant, and in % fact the sign is incorrect.
+% When we use our continuous
+% %  measure of nightlife, its significant, and it helps inform our profitable % % betting model.
+% % This is good news for our proxy, although begs the question why New York % should % be treated so differently by MLB players when compared with NBA % players.
+One % point worth noting is that we haven't yet made the important % distinction that in % the MLB, the Los Angeles Angels are in Orange County, for % example, which is much % less bustling than the NBA's  Los Angeles Lakers who % are located in downtown Los % Angeles.
 \begin{figure}[h]
-  \hspace{-55pt}
-  \label{fig: party by team location mlb}
-  \includegraphics[scale=0.5]{Party_by_Team_Location_mlb}
-  \caption{Continuous measure of nightlife by team location for MLB.}
-\end{figure}
-  % What's   % interesting here is that because New York Yankees are in the Bronx, they show up % as having few musicians and sound recording studios.
-We might wonder if this is % a bad thing, since with NBA New York city was deemed a major party city.
-It % turns out that when we define a discrete indicator of party based on New York or % Los Angeles, it is not at all significant, and in fact the sign is incorrect.
-% When we use our continuous
-%  measure of nightlife, its significant, and it helps inform our profitable % betting model.
-% This is good news for our proxy, although begs the question why New York should % be treated so differently by MLB players when compared with NBA players.
-One % point worth noting is that we haven't yet made the important distinction that in % the MLB, the Los Angeles Angels are in Orange County, for example, which is much % less bustling than the NBA's  Los Angeles Lakers who are located in downtown Los % Angeles.
-\begin{figure}   
   \centering
@@
   \label{fig: next day opponent mlb}
-  \includegraphics[scale=0.5]{next_day_opponent_mlb} 
+  \includegraphics[scale=0.45]{next_day_opponent_mlb} 
   \caption{For the MLB, there are some differences when compared with the     analogous

### 2017-07-11 writing/draft.tex
Updated mean points allowed regression table to include possible confounders for discrete indicator model. Update county --> MSA.

@@
 statistically significant negative effect on a team's likelihood of meeting bookmakers' expectations
-for both NBA and MLB.
-Within the NBA, we analyze specific team performance metrics such as number of points allowed and number of rebounds obtained in a game, and find that both are negatively effected by previous day visitation to a city with active nightlife.
+for both NBA and MLB. Within the NBA, we analyze specific team performance metrics such as number of points allowed, and find that it is 
+negatively effected by previous day visitation to a city with active nightlife.
 Having established a causal model, we realize a profitable betting scheme for MLB.
@@
 last game location. We plot a heat-map displaying the relationship between last game location
-and next game location, and further break this down according to whether the game(s)
-were played at home or away.
+and next game location, and further break this down according to whether the game(s) were played at home or away.
 
@@
 
-We plot a heatmap which displays the logged count of co-occurences between last game
-location and current game location, split by home-and-away games.
+The figure displays the a count of co-occurences between last game location and current game location, faceted by home and away games.
 Our takeaway from this figure is that indeed there is some correlation between last game location
@@
 in which the stadium resides: if the last game was played at most 24-hours ago, use the log of the total number
-of musical groups, sound recording studios, and musical publishers in county of the 
+of musical groups, sound recording studios, and musical publishers in the MSA of the 
 last game location, else emit zero.  We also re-scale this feature to lie within the
@@
 The reason we utilize bookmaker spreads is because this measure already incorporates
-up-to-date information on how well the team as been performing in said season and also
+up-to-date information on how well the team has been performing in said season and also
 their expected (dis)-advantage relative to a particular opponent. Without a feature
@@
 
-\paragraph{MLB} We also obtain data for seasons 2010 through 2017; each season runs from 
+\paragraph{MLB} We also obtain data for seasons 2011 through 2017; each season runs from 
 April through October. There are also exactly 30 teams in the MLB, each plays 162 games per season.
@@
 number of sound recording studios, musical groups, and music publishers there 
-are for a particular county-year: we simply calculate the total number of establishments across all three music categories listed above.
+are for a particular MSA and quarter: we simply calculate the total number of establishments across all three music categories listed above.
 We caveat that since the Toronto Raptors and Toronto Blue Jays are located in Canada,
@@
 the spread as a function of whether the team partied as well as controlling for other confounders
-listed above. Our basic model is as follows accounts for the rest-time since
-last game, jet-lag, home-team effect, player fatigue incurred during the last game,
-and the party effect.
+listed above. Our basic model accounts for the rest-time since
+last game, jet-lag, home-team effect, player fatigue incurred during the last game, and the party effect.
 
@@
     game, we place a bet of \$100 if the expected value of the bet is positive after accounting for the bookmaker's cut. E.g. in the early 2016 season,
-    at its worst we are out of pocket \$750 dollars. By season end, we have     recouped all of
+    at its worst we are out of pocket \$89 dollars. By season end, we have     recouped all of
     our initial investment plus an additional \$11.5k. We are profitable in each season, and mention that 2015 is the ``worst'', but we still come out ahead.}
@@
 \hspace{-125pt}
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std.
-Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.649$^{***}$ &  &  \\    & (0.718) &  &  \\    & & & \\   Continuous party measure &  & 1.833$^{***}$ & 0.447$^{*}$ \\    &  & (0.246) & (0.245) \\    & & & \\   Lag change in possessions & 1.058$^{*}$ & 0.022$^{***}$ & 0.035$^{***}$ \\    & (0.615) & (0.007) & (0.007) \\    & & & \\   Logged travel distance & $-$0.045 & $-$0.028 & $-$0.213$^{***}$ \\    & (0.038) & (0.074) & (0.074) \\    & & & \\   Number hours since last game & 0.009$^{**}$ & 0.015$^{***}$ & 0.018$^{***}$ \\    & (0.004) & (0.003) & (0.003) \\    & & & \\   Home team effect & $-$2.997$^{***}$ & $-$3.412$^{***}$ & 2.211$^{***}$ \\    & (0.218) & (0.116) & (0.116) \\    & & & \\   Constant & $-$3.429 & 101.941$^{***}$ & 100.112$^{***}$ \\    & (2.926) & (0.479) & (0.477) \\    & & & \\  \hline \\[-1.8ex]  Observations & 14,843 & 44,689 & 44,689 \\  R$^{2}$ & 0.018 & 0.021 & 0.010 \\  Adjusted R$^{2}$ & 0.017 & 0.021 & 0.010 \\  Residual Std.
+Error & 11.070 (df = 14831) & 12.002 (df = 44683) & 11.967 (df = 44683) \\  F Statistic & 24.021$^{***}$ (df = 11; 14831) & 196.086$^{***}$ (df = 5; 44683) & 92.809$^{***}$ (df = 5; 44683) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 \section{Conclusions and Avenues for Future Research}

### 2017-07-11 init_draft.r
Now using scale_fill_gradient.

@@ -50,14 +50,16 @@ mp <- update(m0, . ~ party.placebo + . - party.discrete)
 ##    data = espn[last.game.loc != team] %>% na.omit)
 
 ### Total points admitted and scored.
-tpa <- lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame,
+tpa <- lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame +
+         I(team == location),
    data = espn[last.game.loc != team] %>% na.omit)
 tps <- update(tpa, team.pts.scored ~ .)
 
 ### Points  allowed as afunction of last game location (by party)
 nba.lines[, demeaned.pts.admitted := team.pts.admitted - mean(team.pts.admitted),
           by = list(season, team)]
-mpa1 <- lm(demeaned.pts.admitted ~ party.discrete, data = nba.lines)
+mpa1 <- lm(demeaned.pts.admitted ~ party.discrete + lag.chg.pos + log(travel.dist+1) +
+         nhours.lgame + I(team == location) + as.factor(season), data = nba.lines)
 
 ##################################################
 ### MLB Modeling
@@ -148,7 +150,9 @@ stargazer(mpa1, tpa, tps,
                                'Lag change in possessions',
                                'Logged travel distance',
                                'Number hours since last game',
-                               'Constant'))
+                               'Home team effect',
+                               'Constant'),
+          omit = 'as.factor')
 
 stargazer(m3, m4, covariate.labels = c('Continuous measure of nightlife',
                                        'Nightlife (no weekend interaction)', 
@@ -232,9 +236,12 @@ setorder(tmp, 'distance to NYC')
 tmp[, last.game.loc := factor(last.game.loc, levels = dmat[order(distance, decreasing = T), last.game.loc])]
 tmp[, location      := factor(location,      levels = dmat[order(distance, decreasing = T), last.game.loc])]
 
+breaks <- 2^(0:6)
+
 pdf('writing/next_day_opponent.pdf', width = 14, height = 14)
-ggplot(tmp, aes(x = location, y = last.game.loc, fill = log(N))) +
+ggplot(tmp, aes(x = location, y = last.game.loc, fill = N)) +
     geom_tile() +
+    scale_fill_gradient(name = 'Count', trans = 'log', breaks = breaks) +
[…]

### 2017-07-11 cardinal_analytx_presentation/nightlife_and_sports.tex
Restructuring. Still a long way to go.

@@
 
-% What is our solution?
-
 \begin{frame}   
-  \frametitle{Why is this problem hard?}
-  \begin{block}{Partying is a latent variable.}
-    Instead, we proxy for the possibility of athletes partying by looking at 
-    play following games in party cities. Our question becomes...
-  \end{block}
-  \vspace{12pt}\begin{block}{Do teams exhibit a decline in performance the day following a game in a city with active nightlife?}     I.e. we are concerned with ``hangover effects''.
-  \end{block}
-  \vspace{12pt}\begin{block}{Part of the question is determining a measure for nightlife. }
-    We colloquially call cities with high levels of nightlife ``party cities''.   \end{block}
-\end{frame}
-
-  %   \vspace{12pt} 
-  %   \begin{itemize}       \item Do teams play worse the day after visiting a ``party city''?
-  %     \item How do we even determine whether a city has nightlife?    
-  %     \item What does it mean to ``play worse''? \end{itemize}
-
-  % \vspace{12pt}We define ``play worse'' relative to bookmaker/public expectations.
-
-\begin{frame}   
-  \frametitle{Who cares?}
+  \frametitle{Why is this important?}
   \begin{block}{\$400 billion spent on sports betting per year in the US.\footnote{NBA Commissioner Adam Silver, 2014.}}
@@
 \begin{frame}   
+  \frametitle{Why is this problem hard?}
+  \begin{block}{Partying is a latent variable.}
+    Instead, we proxy for the possibility of athletes partying by looking at 
+    play following games in party cities.
+  \end{block}
+  \vspace{12pt}\begin{block}{Part of the question is determining a measure for nightlife. }
+    We colloquially call cities with high levels of nightlife ``party cities''.   \end{block}
+\end{frame}
+
+\begin{frame}   \frametitle{What's the specific question, exactly?}
+  \begin{block}{Do teams/athletes perform worse the day after playing in a party-oriented city?\footnote{  We will define ``perform worse'' relative to bookmaker expectations, which are in part driven by
+  bettors' expectations.}}
+\vspace{12pt}     I.e. we are concerned with ``hangover effects''.
+  \end{block}
+\end{frame}
+
+\begin{frame}
+  \frametitle{How can we measure unexpected changes in performance?}     
+  \begin{block}{Bookmaker point-spreads (NBA)}     
+    The expected score differential between two teams.
+
+    Conditions on (almost) every conceivable element of the game:
+    \begin{itemize}     
+      \item strengths and weaknesses against particular opponents,
+      \item momentum,
+      \item injuries, and
+      \item even possibly a weather forecast (for outdoor sports).     
+    \end{itemize}
+
+    A team is expected to meet the spread $\frac{1}{2}$ the time.\footnote{Dubner and Levitt, ('06). Losers pay 10\% fee on top of wagers; a balanced book guarantees 5\% gain for the house-bookmaker.}
+  \end{block}
+  \vspace{12pt}
+  \begin{block}{Bookmaker money-lines (MLB)}     
+    The odds that a team will beat their opponent.
+
+    After conditioning on odds, successful prediction is like a coin-flip. 
+  \end{block} 
+\end{frame}
+
+\begin{frame}
+  \centering   \includegraphics[scale=0.55]{../writing/meet_the_spread} 
+\end{frame}
+
+\begin{frame}   \frametitle{Our solution: a quasi-experiment.}
+  \begin{block}{Systematic deviations represent unexplained mechanisms.}     We construct a model which attempts to estimate the causal effect of visiting a ``party city''
+    on subsequent day performance, as measured by the odds of meeting the spread.       \end{block}
+
+  \begin{block}{}     We only consider the hangover effect on games played back-to-back within 24 hours of each other.   
+  \end{block} 
+\end{frame}
+
+\begin{frame}   \frametitle{Identifying variation}
+  \begin{block}{Performance as a function of previous day's nightlife index.}     If next day opponent is uncorrelated with visitation to a party city,
+    then looking at the \emph{actual outcome} of the game $\leadsto$ identification.
+
+    \vspace{15pt}
+
+    Weaker, still: if odds of beating the spread uncorrelated with previous game location
+    $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}       \end{block}
+\end{frame}
+
+  %   \vspace{12pt} 
+  %   \begin{itemize}       \item Do teams play worse the day after visiting a ``party city''?
+  %     \item How do we even determine whether a city has nightlife?    
+  %     \item What does it mean to ``play worse''? \end{itemize}
+
+  % \vspace{12pt}We define ``play worse'' relative to bookmaker/public expectations.
+
+\begin{frame}   \frametitle{Anecdotal evidence: players do party}
+  \begin{block}{Open attitude toward partying}     Matt Barnes recently complained, ``there's no nightlife in Utah'' when asked
+    about playing the Jazz in the 2017 playoffs.\footnote{Cestone and Seiter ('17).}   \end{block}
+
+  \vspace{12pt}
+  \begin{block}{Blaming the party scene for poor performance}
+    J.R. Smith's career-to-date success rate of shots taken was 42\%, 
+    but on Sunday afternoon games he only managed 34\% in 2012-13.\footnote{Concepcion ('15).}
+
+    \vspace{12pt}
+    After being traded from NY to Cavaliers, performance increased.
+    When asked why: ``There’s nothing you expect (in Cleveland) but basketball. 
+    There’s nothing, there’s no going out, there’s no late nights.''\footnote{Price, Levy ('15).}
+  \end{block} \end{frame}
+
+\begin{frame}
   \frametitle{Previous research - NBA player shooting average}
@@
     \item 1,092 of 2,035 NBA players who took $\geq 1$ shot on Sundays       under-perform 
-      relative to their own season-long average; $p=0.001$.
-      \item Effect size is not considered.
-      \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, is that relevant?}     \end{itemize}
+      relative to their own career-long average;\footnote{Analysis limited to all players who entered the league after 1974.} $p=0.001$.
+    \end{itemize}
+    \begin{block}{Problems}       \begin{itemize}           \item Effect size is not considered.
+          \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, is that relevant?}            \end{itemize}     \end{block}
 \end{frame}
@@
       using 2006-14 data; $p=0.1089$.
+    \end{itemize}
+    \vspace{12pt}
+    \begin{block}{Only consider Sunday games $\leadsto$ underwhelming results:}       \begin{itemize}
           \item Not quite significant at the 90\% confidence level.     
-    \item If we bet \$100 on each Sunday game, we'd only profit \$1,310.\footnote{$(284 - 129) \times 100 + 129 \times (-110) = 1,310$.} 
-    \end{itemize}
+          \item If we bet \$100 on each Sunday game, we'd only profit \$1,310 over 8 years.\footnote{$(284 - 129) \times 100 + 129 \times (-110) = 1,310$.} 
+      \end{itemize}     \end{block}
   \end{block}
@@
 
-\begin{frame}
-  \frametitle{How can we measure unexpected changes in performance?}     \begin{block}{Bookmaker point-spreads (NBA)}     
-    The expected score differential between two teams.
-
-    Conditions on (almost) every conceivable element of the game:
-    \begin{itemize}     \item strengths and weaknesses against particular opponents,
-    \item momentum,
-    \item injuries, and
-    \item even possibly a weather forecast (for outdoor sports).     \end{itemize}
-
-    A team is expected to meet the spread $\frac{1}{2}$ the time.\footnote{Dubner and Levitt, ('06). Losers pay 10\% fee on top of wagers; a balanced book guarantees 5\% gain for the house-bookmaker.}
-  \end{block}
-  \vspace{12pt}\begin{block}{Bookmaker money-lines (MLB)}     The odds that a team will beat their opponent.
-
-    After conditioning on odds, successful prediction is like a coin-flip. \end{block} \end{frame}
-
-\begin{frame}
-  \centering   \includegraphics[scale=0.55]{../writing/meet_the_spread} 
-\end{frame}
-
-\begin{frame}   \frametitle{Our solution: a quasi-experiment.}
-  \begin{block}{Systematic deviations represent unexplained mechanisms.}     We construct a model which attempts to estimate the causal effect of visiting a ``party city''
-    on subsequent day performance, as measured by the odds of meeting the spread.       \end{block}
-
-  \begin{block}{}     We only consider the hangover effect on games played back-to-back within 24 hours of each other.   \end{block} \end{frame}
-
-\begin{frame}   \frametitle{Identifying variation}
-  \begin{block}{Performance as a function of previous day's nightlife index.}     If next day opponent is uncorrelated with visitation to a party city,
-    then looking at the \emph{actual outcome} of the game $\leadsto$ identification.
-
-    \vspace{15pt}
-
-    Weaker, still: if odds of beating the spread uncorrelated with previous game location
-    $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}       \end{block}
-\end{frame}
-
 % \begin{frame}   \frametitle{Our solution: a quasi-experiment}
@@
 
-\begin{frame}   \frametitle{Sports: NBA and MLB}
+\begin{frame}   \frametitle{Returning to our question within the NBA and MLB}
   \begin{block}{We perform the analysis for both NBA and MLB}     
@@
     \begin{itemize}       \item For NBA teams playing another game in the next 24 hours, are they less likely to meet the points-spread?
-      \item For MLB teams, after conditioning on the odds of winning, are they less likely to beat their opponent?     \end{itemize}
+      \item For MLB teams, after conditioning on the odds of winning, are they less likely to beat their next opponent?     \end{itemize}
     \vspace{12pt}\begin{block}{If we yield a causal model, we may realize a betting schema.}
@@
 \end{frame}
+
+\begin{frame}   \frametitle{Overzealous assumption: next day opponent uncorrelated}
+  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent} \end{frame}
+
+\begin{frame}   \frametitle{A closer look at NBA games on the road}
+  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent_on_tour} 
+\end{frame}
+
+\begin{frame}
+  \frametitle{Weakening our assumptions}
+  \begin{block}{With some correlation between next-day opponent, this further motivates using bookmaker expectations as the response.}
+
+    \vspace{12pt}
+    Helps to account for systematically running into teams the day after visitation to 
+    a party oriented city who are particularly (not)-skilled.
+  \end{block}
+\end{frame}
+
+\begin{frame}   \frametitle{Public scrutiny makes data collection difficult}
+
+  \begin{block}{Whether parties take place is a latent variable}     Due to intense public scrutiny, there is a reluctance to share nightlife experiences
+    of players on social media.
+
+  \end{block} 
+  \vspace{12pt}
+  \begin{block}{We can't observe when players go for a night on the town.}
+
+    Instead, we develop two approaches to proxy for partying.
+
+    A discrete indicator and a continuous measure.
+  \end{block} \end{frame}
+
+\begin{frame}   \frametitle{Discrete indicator}
+  \begin{block}{Anecdotally, LA and NY appear in news articles frequently.}     \vspace{12pt}This motivates a binary indicator for party when either of these two cities have been visited within 24-hours of another game.
+\vspace{12pt}
+    \[
+      \texttt{party} \coloneqq \mathbbm 1 \left \{\substack{\textrm{last game played in LA or NY} \\ \textrm{at most 24-hours ago}}\right \}
+    \]   \end{block} \end{frame}
+
+\begin{frame}   
+  \frametitle{Travel distance comparison}
+  \centering \includegraphics[scale=0.25]{../writing/travel_dist_density_by_party} 
+\end{frame}
+
+\begin{frame}   \frametitle{Continuous measure}
+  \begin{block}{A-list celebrities tend to party with other celebrities.}     Idea: proxy nightlife using number of musicians in previous location
+
+    \[
+      \texttt{party} \coloneqq \begin{cases}       \log \left(\# \textrm{musical groups}\right) &\substack{\textrm{ if last game played} \\ \textrm{at most 24-hours ago,}} \\
+      0 &\textrm{ otherwise.}
+    \end{cases}
+    \]
+    \newline
+    Specifically, we consider the sum total of
+    \[
+      \textrm{Sound recording studios } + \textrm{ musical groups } + \textrm{ musical publishers}.
+    \]   \end{block} \end{frame}
+
+\begin{frame}   
+  \frametitle{Continuous measure of party}
+  \centering \includegraphics[scale=0.375]{../writing/Party_by_Team_Location} 
+\end{frame}
+
+  % \begin{itemize}     \item If bookkeepers aware of hangover effects, our variable of interest may not appear significant, in spite of there being a true effect.
+
+  %   \item If bookkeepers systematically mis-value teams which are more likely to be encountered after visiting a city with active nightlife, we could see a spurious effect.
+
+  %     (Given size of betting market, this is unlikely)
+  %   \end{itemize}
 
@@
   \vspace{12pt}Finally, to account for jetlag we interact this east-west measure with travel distance.\footnote{Lallensack ('17)} \end{frame}
-
-\begin{frame}   \frametitle{Next day opponent: NBA}
-  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent} \end{frame}
-
-\begin{frame}   \frametitle{A closer look at NBA games on the road}
-  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent_on_tour} \end{frame}
-
-\begin{frame}   %\frametitle{Further motivation for conditioning on the spread}
-  \begin{block}{With some correlation between next-day opponent, this further motivates conditioning
-  on bookmakers' expectations.}
-  \begin{itemize}     \item If bookkeepers aware of hangover effects, our variable of interest may not appear significant, in spite of there being a true effect.
-
-    \item If bookkeepers systematically mis-value teams which are more likely to be encountered after visiting a city with active nightlife, we could see a spurious effect.
-
-      (Given size of betting market, this is unlikely)
-    \end{itemize}   \end{block}
-\end{frame}
-
-\begin{frame}   \frametitle{Recall: partying is a latent variable}
-  \begin{block}{We can't observe when players go for a night on the town.}
-    Instead, we develop two approaches to proxy for partying.
-
-    A discrete indicator and a continuous measure.
-  \end{block} \end{frame}
-
-\begin{frame}   \frametitle{Discrete Indicator}
-  \begin{block}{Anecdotally, LA and NY appear in news articles frequently.}     \vspace{12pt}This motivates a binary indicator for party when either of these two cities have been visited within 24-hours of another game.
-\vspace{12pt}
-    \[
-      \texttt{party} \coloneqq \mathbbm 1 \left \{\substack{\textrm{last game played in LA or NY} \\ \textrm{at most 24-hours ago}}\right \}
-    \]   \end{block} \end{frame}
-
-\begin{frame}   
-  \frametitle{Travel distance comparison}
-  \centering \includegraphics[scale=0.25]{../writing/travel_dist_density_by_party} 
-\end{frame}
-
-\begin{frame}   \frametitle{Continuous measure}
-  \begin{block}{A-list celebrities tend to party with other celebrities.}     Idea: proxy nightlife using number of musicians in previous location
-
-    \[
-      \texttt{party} \coloneqq \begin{cases}       \log \left(\# \textrm{musical groups}\right) &\substack{\textrm{ if last game played} \\ \textrm{at most 24-hours ago,}} \\
-      0 &\textrm{ otherwise.}
-    \end{cases}
-    \]
-    \newline
-    Specifically, we consider the sum total of
-    \[
-      \textrm{Sound recording studios } + \textrm{ musical groups } + \textrm{ musical publishers}.
-    \]   \end{block} \end{frame}
-
-\begin{frame}   
-  \frametitle{Continuous measure of party}
-  \centering \includegraphics[scale=0.375]{../writing/Party_by_Team_Location} 
-\end{frame}
 
@@
   \begin{adjustwidth}{-.75cm}{}
-  \tiny{  % mpa, tpa, tps
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.817$^{***}$ &  &  \\    & (0.712) &  &  \\    & & & \\   Continuous party measure &  & 1.942$^{***}$ & 0.377 \\    &  & (0.248) & (0.246) \\    & & & \\   Lag change in possessions &  & 0.022$^{***}$ & 0.035$^{***}$ \\    &  & (0.007) & (0.007) \\    & & & \\   Logged travel distance &  & $-$0.411$^{***}$ & 0.035 \\    &  & (0.074) & (0.073) \\    & & & \\   Number hours since last game &  & 0.012$^{***}$ & 0.020$^{***}$ \\    &  & (0.003) & (0.003) \\    & & & \\   Constant & $-$0.046 & 102.824$^{***}$ & 99.540$^{***}$ \\    & (0.091) & (0.483) & (0.478) \\    & & & \\  \hline \\[-1.8ex]  Observations & 15,139 & 44,689 & 44,689 \\  R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Adjusted R$^{2}$ & 0.001 & 0.002 & 0.002 \\  Residual Std.
-Error & 11.137 (df = 15137) & 12.117 (df = 44684) & 12.015 (df = 44684) \\  F Statistic & 15.670$^{***}$ (df = 1; 15137) & 27.968$^{***}$ (df = 4; 44684) & 24.325$^{***}$ (df = 4; 44684) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  }
+  \tiny{  % mpa1, tpa, tps
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.649$^{***}$ &  &  \\    & (0.718) &  &  \\    & & & \\   Continuous party measure &  & 1.833$^{***}$ & 0.447$^{*}$ \\    &  & (0.246) & (0.245) \\    & & & \\   Lag change in possessions & 1.058$^{*}$ & 0.022$^{***}$ & 0.035$^{***}$ \\    & (0.615) & (0.007) & (0.007) \\    & & & \\   Logged travel distance & $-$0.045 & $-$0.028 & $-$0.213$^{***}$ \\    & (0.038) & (0.074) & (0.074) \\    & & & \\   Number hours since last game & 0.009$^{**}$ & 0.015$^{***}$ & 0.018$^{***}$ \\    & (0.004) & (0.003) & (0.003) \\    & & & \\   Home team effect & $-$2.997$^{***}$ & $-$3.412$^{***}$ & 2.211$^{***}$ \\    & (0.218) & (0.116) & (0.116) \\    & & & \\   Constant & $-$3.429 & 101.941$^{***}$ & 100.112$^{***}$ \\    & (2.926) & (0.479) & (0.477) \\    & & & \\  \hline \\[-1.8ex]  Observations & 14,843 & 44,689 & 44,689 \\  R$^{2}$ & 0.018 & 0.021 & 0.010 \\  Adjusted R$^{2}$ & 0.017 & 0.021 & 0.010 \\  Residual Std.
+Error & 11.070 (df = 14831) & 12.002 (df = 44683) & 11.967 (df = 44683) \\  F Statistic & 24.021$^{***}$ (df = 11; 14831) & 196.086$^{***}$ (df = 5; 44683) & 92.809$^{***}$ (df = 5; 44683) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
 \end{adjustwidth} 

### 2017-07-12 cardinal_analytx_presentation/nightlife_and_sports.tex
Updated tables. Moved slides around.

@@
   \vspace{12pt}\begin{block}{Part of the question is determining a measure for nightlife. }
-    We colloquially call cities with high levels of nightlife ``party cities''.   \end{block}
+    We dub cities with high levels of nightlife ``party cities''.  
+  \vspace{12pt}
+  \begin{block}{Another is how to measure a dip in performance?}
+    We need to consider player ability, team momentum, etc.
+  \end{block} \end{block}
 \end{frame}
@@
 \begin{frame}   \frametitle{What's the specific question, exactly?}
-  \begin{block}{Do teams/athletes perform worse the day after playing in a party-oriented city?\footnote{  We will define ``perform worse'' relative to bookmaker expectations, which are in part driven by
-  bettors' expectations.}}
+  \begin{block}{Do teams/athletes perform worse the day after playing in a party-oriented city?}
 \vspace{12pt}     I.e. we are concerned with ``hangover effects''.
+
+\vspace{12pt}
+
+% E.g. a team visits Las Vegas for the day. How is their next game?
   \end{block}
@@
     then looking at the \emph{actual outcome} of the game $\leadsto$ identification.
+  \end{block}
+\end{frame}
 
-    \vspace{15pt}
+\begin{frame}   \frametitle{Overzealous assumption: next day opponent uncorrelated}
+  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent} 
+\end{frame}
 
-    Weaker, still: if odds of beating the spread uncorrelated with previous game location
-    $\leadsto$ identification.\footnote{Under an efficient market, this assumption should hold.}       \end{block}
+\begin{frame}   \frametitle{A closer look at NBA games on the road}
+  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent_on_tour} 
+\end{frame}
+
+\begin{frame}
+  \frametitle{Weakening our assumptions}
+  \begin{block}{With some correlation between next-day opponent, this further motivates using bookmaker expectations as the response.}
+
+    Helps to account for systematically running into teams the day after visitation to 
+    a party oriented city who are particularly (not)-skilled.
+  \end{block}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Implications and pitfalls of using a black-box feature}
+  \begin{block}{If bookkeepers aware of hangover effects...}     then our variable of interest may not appear significant, in spite of there being a true effect.   \end{block}
+  \vspace{12pt}
+  \begin{block}{If bookkeepers systematically mis-value teams...}
+     which are more likely to be encountered after visiting a city with active nightlife, we could see a spurious effect.
+
+      (Given size of betting market, this is unlikely)
+    \end{block}   
 \end{frame}
@@
     There’s nothing, there’s no going out, there’s no late nights.''\footnote{Price, Levy ('15).}
-  \end{block} \end{frame}
+  \end{block} 
+\end{frame}
 
@@
 
-\begin{frame}   \frametitle{Overzealous assumption: next day opponent uncorrelated}
-  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent} \end{frame}
-
-\begin{frame}   \frametitle{A closer look at NBA games on the road}
-  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent_on_tour} 
-\end{frame}
-
-\begin{frame}
-  \frametitle{Weakening our assumptions}
-  \begin{block}{With some correlation between next-day opponent, this further motivates using bookmaker expectations as the response.}
-
-    \vspace{12pt}
-    Helps to account for systematically running into teams the day after visitation to 
-    a party oriented city who are particularly (not)-skilled.
-  \end{block}
-\end{frame}
-
 \begin{frame}   \frametitle{Public scrutiny makes data collection difficult}
@@
       \texttt{party} \coloneqq \mathbbm 1 \left \{\substack{\textrm{last game played in LA or NY} \\ \textrm{at most 24-hours ago}}\right \}
-    \]   \end{block} \end{frame}
+    \]   \end{block} 
+\end{frame}
 
@@
 \end{frame}
-
-  % \begin{itemize}     \item If bookkeepers aware of hangover effects, our variable of interest may not appear significant, in spite of there being a true effect.
-
-  %   \item If bookkeepers systematically mis-value teams which are more likely to be encountered after visiting a city with active nightlife, we could see a spurious effect.
-
-  %     (Given size of betting market, this is unlikely)
-  %   \end{itemize}
 
@@
 
-\vspace{12pt}  Anecdotally, rebounds are considered a ``hustle'' statistic, in the sense that players
+  \vspace{12pt}  
+  \begin{block}{Is it the case the hangover effects are more prevalent on defensive aspects of gameplay?}     Anecdotally, rebounds are considered a ``hustle'' statistic, in the sense that players
     who work harder on the court and expend energy boxing out other players will have higher
-  rebound rates. \end{frame}
-
-% \begin{frame}   \frametitle{Analysis: rebounds}
-%   \begin{itemize}     \item For each player and each game, we calculate the % number of rebounds obtained per-minute of game-play
-%     \item From each observation, we subtract the player's average number of % rebounds per minute taken across the entire season.
-%     \item We then run a linear regression explaining demeaned rebounds as a % function of 
-%       nightlife visitation and other features describing player fatigue.   % \end{itemize} \end{frame}
-
-% \begin{frame}   \frametitle{Results: rebounds}
-%   \centering
-%   \tiny{ % m2
-% \begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & % \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & % Demeanead Rebounds Per Minute \\  \hline \\[-1.8ex]   Continuous party measure & % $-$0.004$^{*}$ \\    & (0.002) \\    & \\   Lag change in possessions & 0.00002 % \\    & (0.0001) \\    & \\   Logged travel distance & 0.001 \\    & (0.001) \\    % & \\   Number hours since last game & 0.00002 \\    & (0.00002) \\    & \\   % Constant & $-$0.007 \\    & (0.004) \\    & \\  \hline \\[-1.8ex]  Observations % & 40,684 \\  R$^{2}$ & 0.0003 \\  Adjusted R$^{2}$ & 0.0002 \\  Residual Std.
-% Error & 0.104 (df = 40679) \\  F Statistic & 2.696$^{**}$ (df = 4; 40679) \\  % \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; % $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  } \end{frame}
-
-\begin{frame}   \frametitle{Hangover correlates with allowing more points}
-  We examine points allowed and points scored (on the team level).
-
-  \vspace{12pt}Common knowledge suggests that overall defense of a team is strongly influenced by individual player effort.
-
-  \vspace{12pt}Is it the case the hangover effects are more prevalent on defensive aspects of gameplay? \end{frame}
+    rebound rates.    \end{block}
+\end{frame}
 
 \begin{frame}   \frametitle{Defense in the NBA}
-  \begin{adjustwidth}{-.75cm}{}
   \tiny{  % mpa1, tpa, tps
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.649$^{***}$ &  &  \\    & (0.718) &  &  \\    & & & \\   Continuous party measure &  & 1.833$^{***}$ & 0.447$^{*}$ \\    &  & (0.246) & (0.245) \\    & & & \\   Lag change in possessions & 1.058$^{*}$ & 0.022$^{***}$ & 0.035$^{***}$ \\    & (0.615) & (0.007) & (0.007) \\    & & & \\   Logged travel distance & $-$0.045 & $-$0.028 & $-$0.213$^{***}$ \\    & (0.038) & (0.074) & (0.074) \\    & & & \\   Number hours since last game & 0.009$^{**}$ & 0.015$^{***}$ & 0.018$^{***}$ \\    & (0.004) & (0.003) & (0.003) \\    & & & \\   Home team effect & $-$2.997$^{***}$ & $-$3.412$^{***}$ & 2.211$^{***}$ \\    & (0.218) & (0.116) & (0.116) \\    & & & \\   Constant & $-$3.429 & 101.941$^{***}$ & 100.112$^{***}$ \\    & (2.926) & (0.479) & (0.477) \\    & & & \\  \hline \\[-1.8ex]  Observations & 14,843 & 44,689 & 44,689 \\  R$^{2}$ & 0.018 & 0.021 & 0.010 \\  Adjusted R$^{2}$ & 0.017 & 0.021 & 0.010 \\  Residual Std.
-Error & 11.070 (df = 14831) & 12.002 (df = 44683) & 11.967 (df = 44683) \\  F Statistic & 24.021$^{***}$ (df = 11; 14831) & 196.086$^{***}$ (df = 5; 44683) & 92.809$^{***}$ (df = 5; 44683) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
-\end{adjustwidth} 
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & \multicolumn{2}{c}{Team Points Admitted} & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.970$^{***}$ &  &  \\    & (0.835) &  &  \\    & & & \\   Continuous party measure &  & 1.661$^{***}$ & $-$0.139 \\    &  & (0.644) & (0.637) \\    & & & \\   Lag change in possessions & 6.694$^{***}$ & 6.751$^{***}$ & 5.858$^{***}$ \\    & (0.973) & (0.973) & (0.963) \\    & & & \\   Logged travel distance & $-$0.114 & $-$0.132 & $-$0.285 \\    & (0.198) & (0.198) & (0.196) \\    & & & \\   Number hours since last game & 0.011$^{*}$ & 0.015$^{**}$ & 0.016$^{**}$ \\    & (0.006) & (0.007) & (0.007) \\    & & & \\   Home team effect & $-$3.269$^{***}$ & $-$3.269$^{***}$ & 2.417$^{***}$ \\    & (0.303) & (0.303) & (0.300) \\    & & & \\   Constant & 68.114$^{***}$ & 67.686$^{***}$ & 71.348$^{***}$ \\    & (4.805) & (4.810) & (4.758) \\    & & & \\  \hline \\[-1.8ex]  Observations & 6,234 & 6,234 & 6,234 \\  R$^{2}$ & 0.091 & 0.090 & 0.084 \\  Adjusted R$^{2}$ & 0.087 & 0.086 & 0.079 \\  Residual Std.
+Error (df = 6201) & 11.670 & 11.675 & 11.549 \\  F Statistic (df = 32; 6201) & 19.453$^{***}$ & 19.247$^{***}$ & 17.720$^{***}$ \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
+\end{frame}
+
+\begin{frame}
+  \centering
+  \includegraphics[scale=0.25]{../writing/distribution_point_spreads_nba}
 \end{frame}

### 2017-07-12 init_draft.r
Now plotting distribution of point-spreads. Also, updated models and as a result achieved more impactful results!

@@ -45,21 +45,22 @@ m1 <- update(m0, . ~ party + . - party.discrete)
 mp <- update(m0, . ~ party.placebo + . - party.discrete)
 
 ### Player specific model - demeaned rebounds per minute.
-## m2 <- lm(demeaned.reb.per.min ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame +
-##              as.factor(season),
-##    data = espn[last.game.loc != team] %>% na.omit)
+## m2 <- lm(reb.per.min ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame +
+##              as.factor(season) + player*as.factor(season),
+##          data = espn[last.game.loc != team] %>% na.omit)
+
+## require(lfe)
+## m2 <- felm(reb.per.min ~ party.discrete + lag.chg.pos + log(travel.dist+1) + nhours.lgame +
+##              as.factor(season) + player,
+##          data = espn[last.game.loc != team] %>% na.omit)
 
 ### Total points admitted and scored.
 tpa <- lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame +
-         I(team == location),
-   data = espn[last.game.loc != team] %>% na.omit)
-tps <- update(tpa, team.pts.scored ~ .)
+              I(team == location) + team,
+          data = nba.lines[last.game.loc != team] %>% na.omit)
+mpa1 <- update(tpa, . ~ party.discrete + . - party)
+tps <- update(tpa, team.pts.scored ~ . + team)
 
-### Points  allowed as afunction of last game location (by party)
-nba.lines[, demeaned.pts.admitted := team.pts.admitted - mean(team.pts.admitted),
-          by = list(season, team)]
-mpa1 <- lm(demeaned.pts.admitted ~ party.discrete + lag.chg.pos + log(travel.dist+1) +
-         nhours.lgame + I(team == location) + as.factor(season), data = nba.lines)
 
 ##################################################
 ### MLB Modeling
@@ -144,7 +145,7 @@ stargazer(mp, covariate.labels = c('Party placebo', 'Lag changes in posession',
 ##                                'Constant'))
 
 stargazer(mpa1, tpa, tps,
-          dep.var.labels = c('Team Points Admitted', 'Team Points Admitted', 'Team Points Scored'),
+          dep.var.labels = c('Team Points Admitted', 'Team Points Scored'),
           covariate.labels = c('Discrete party indicator',
[…]

### 2017-07-12 writing/draft.tex
Updated table on Team Points Allowed/Scored.

@@
   \centering
-  \includegraphics[scale=0.55]{travel_dist_density_by_party}
+  \includegraphics[scale=0.425]{travel_dist_density_by_party}
   \label{fig: density by party}
@@
     whether the last game was in a (non) party city. We see that in general,
-  travel distance follows a similar distribution regardless of last game location, i.e. 
-  it's not as though players encounter significantly
-  more travel fatigue after playing in a party city.
+    when traveling from a non party-oriented city to next game location, the distribution
+    of travel-distance has fatter tails. This makes sense: as we saw in our previous
+    heat-map, next day opponent geographically correlated with current game location.
+    Since LA and NY are on the coast, their next day opponent is likely to be nearby; compare
+    this to inland cities, for example, who may have to travel farther to reach their 
+    nearest neighbor.
 We mention that the same     plot is not interesting to look at within the MLB, since games are played in     series, and therefore most travel-distances between games are zero miles.}
@@
 % mpa1, tpa, tpo
-\hspace{-125pt}
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & Team Points Admitted & Team Points Admitted & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.649$^{***}$ &  &  \\    & (0.718) &  &  \\    & & & \\   Continuous party measure &  & 1.833$^{***}$ & 0.447$^{*}$ \\    &  & (0.246) & (0.245) \\    & & & \\   Lag change in possessions & 1.058$^{*}$ & 0.022$^{***}$ & 0.035$^{***}$ \\    & (0.615) & (0.007) & (0.007) \\    & & & \\   Logged travel distance & $-$0.045 & $-$0.028 & $-$0.213$^{***}$ \\    & (0.038) & (0.074) & (0.074) \\    & & & \\   Number hours since last game & 0.009$^{**}$ & 0.015$^{***}$ & 0.018$^{***}$ \\    & (0.004) & (0.003) & (0.003) \\    & & & \\   Home team effect & $-$2.997$^{***}$ & $-$3.412$^{***}$ & 2.211$^{***}$ \\    & (0.218) & (0.116) & (0.116) \\    & & & \\   Constant & $-$3.429 & 101.941$^{***}$ & 100.112$^{***}$ \\    & (2.926) & (0.479) & (0.477) \\    & & & \\  \hline \\[-1.8ex]  Observations & 14,843 & 44,689 & 44,689 \\  R$^{2}$ & 0.018 & 0.021 & 0.010 \\  Adjusted R$^{2}$ & 0.017 & 0.021 & 0.010 \\  Residual Std.
-Error & 11.070 (df = 14831) & 12.002 (df = 44683) & 11.967 (df = 44683) \\  F Statistic & 24.021$^{***}$ (df = 11; 14831) & 196.086$^{***}$ (df = 5; 44683) & 92.809$^{***}$ (df = 5; 44683) \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+\hspace{-35pt}
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & \multicolumn{2}{c}{Team Points Admitted} & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.970$^{***}$ &  &  \\    & (0.835) &  &  \\    & & & \\   Continuous party measure &  & 1.661$^{***}$ & $-$0.139 \\    &  & (0.644) & (0.637) \\    & & & \\   Lag change in possessions & 6.694$^{***}$ & 6.751$^{***}$ & 5.858$^{***}$ \\    & (0.973) & (0.973) & (0.963) \\    & & & \\   Logged travel distance & $-$0.114 & $-$0.132 & $-$0.285 \\    & (0.198) & (0.198) & (0.196) \\    & & & \\   Number hours since last game & 0.011$^{*}$ & 0.015$^{**}$ & 0.016$^{**}$ \\    & (0.006) & (0.007) & (0.007) \\    & & & \\   Home team effect & $-$3.269$^{***}$ & $-$3.269$^{***}$ & 2.417$^{***}$ \\    & (0.303) & (0.303) & (0.300) \\    & & & \\   Constant & 68.114$^{***}$ & 67.686$^{***}$ & 71.348$^{***}$ \\    & (4.805) & (4.810) & (4.758) \\    & & & \\  \hline \\[-1.8ex]  Observations & 6,234 & 6,234 & 6,234 \\  R$^{2}$ & 0.091 & 0.090 & 0.084 \\  Adjusted R$^{2}$ & 0.087 & 0.086 & 0.079 \\  Residual Std.
+Error (df = 6201) & 11.670 & 11.675 & 11.549 \\  F Statistic (df = 32; 6201) & 19.453$^{***}$ & 19.247$^{***}$ & 17.720$^{***}$ \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
 \section{Conclusions and Avenues for Future Research}

### 2017-07-12 init_draft.r
Changed rounding for residual plot. Added a baseline zero-earnings line for MLB betting.

@@ -105,7 +105,7 @@ b <- lapply(2012:2016, function(yr) {
                        format(round(sum(tmp$bet), digits = -1), big.mark = ','), '\n',
                        'Training on 2011-', yr-1, ' data'),
          xlab = 'Game Number', ylab = 'Profit', type = 'l')
-    abline(a = 0, b = 0, lty = 3)
+    abline(a = 0, b = 0, lty = 'dashed', col = 'orange', lwd = 3)
     points(x = which.min(tmp$running.prof), y = min(tmp$running.prof), type = 'p')
     text(x = which.min(tmp$running.prof), y = min(tmp$running.prof), 
          labels = round(min(tmp$running.prof)), pos = 4)
@@ -375,7 +375,7 @@ mlb.lines <- na.omit(mlb.lines)
 noinf <- glm(I(team.score > opponent.score) ~ odds, data = mlb.lines, family = 'binomial')
 mlb.lines[, fitted := fitted(noinf)]
 mlb.lines[, resid  := resid(noinf)]
-tmp <- mlb.lines[, .(avg.residual = mean(resid)), by = .(fitted = round(fitted, 2))]
+tmp <- mlb.lines[, .(avg.residual = mean(resid)), by = .(fitted = round(fitted, 3))]
 plot(tmp$fitted, tmp$avg.residual, 
      xlab = 'Fitted Probability', ylab = 'Average Residual', 
      main = paste('After conditioning on bookkeeper\'s odds (MLB),',

### 2017-07-12 writing/draft.tex
Added note on how further research might investigate how the hangover effect tapers as a function of time.

@@
 In both sports, it's possible there are prolonged effects of partying, and applying some sort of kernel on the number of hours played since the last game could be interesting; recall that our decision to only look at back-to-back games
-within 24 hours of each other was only motivated by intuition and not guided by data. 
-Future work may consider similar effects in European soccer leagues.
+within 24 hours of each other was only motivated by intuition and not guided by data; How does the effect taper off as a function of time? 
+Future work may consider similar effects in NHL or European soccer leagues.
 

### 2017-07-12 cardinal_analytx_presentation/nightlife_and_sports.tex
Added a footnote (size of betting market). Added NHL as candidate sport to investigate.

@@
     Helps to account for systematically running into teams the day after visitation to 
-    a party oriented city who are particularly (not)-skilled.
+    a party oriented city who are particularly (not)-skilled.\footnote{Given the size of the market, this is reasonable.}
   \end{block}
@@
 \begin{frame}   \frametitle{Minutes data for NBA}
-  For NBA, we also collect minutes data, which describes the number of 
+  For NBA, we also collect minutes data from ESPN, which describes the number of 
   \vspace{12pt}
@@
 
-\begin{frame}   \frametitle{Defense in the NBA}
+\begin{frame}   \frametitle{Defense in the NBA is affected by our measure for party}
   \tiny{  % mpa1, tpa, tps
@@
     \item Possibly prolonged effects; recall our magic 24 hour number.
-    \item European soccer leagues.   \end{itemize} \end{frame}
+    \item NHL or European soccer leagues.   \end{itemize} \end{frame}
 

### 2017-07-14 cardinal_analytx_presentation/nightlife_and_sports.tex
Added commentary to conclusion slide.

@@
     \item Within the MLB, we can look at statistics such as batting averages or general fielding statistics. 
-    \item Possibly prolonged effects; recall our magic 24 hour number.
-    \item NHL or European soccer leagues.   \end{itemize} \end{frame}
+    \vspace{12pt}
+    \item Possibly prolonged effects: recall our magic 24 hour number.
+      How does the hangover effect taper off as a function of time?
+    \vspace{12pt}
+    \item NHL or European soccer leagues.   
+  \end{itemize} 
+\end{frame}
 
 \begin{frame}   \frametitle{Conclusions}
-  \begin{itemize}     \item NBA and MLB players under-perform bookmakers expectations after visiting LA and NY.
+  \begin{itemize}
+    \vspace{12pt}
+    \item NBA and MLB players under-perform bookmakers expectations after visiting LA and NY.
+    \vspace{12pt}
     \item In general we see that performance decreases as a function of (our proxy for) nightlife.     
+    \vspace{12pt}
       \item The decrease in likelihood can (in part) be explained by a drop in defensive performance in the NBA.
-    \item With a causal model established, we realize a profitable betting scheme in MLB.   \end{itemize} \end{frame}
+    \vspace{12pt}
+    \item With a causal model established, we realize a profitable betting scheme in MLB.   
+  \end{itemize} 
+\end{frame}
 

### 2017-07-14 cardinal_analytx_presentation/nightlife_and_sports.tex
Condensed 'Data' section down to one slide!

@@
  \begin{document}  \begin{frame} \maketitle \end{frame}
+
+\begin{frame}   \frametitle{Anecdotal evidence: athletes party}
+  \begin{block}{Open attitude toward partying}     Matt Barnes recently complained, ``there's no nightlife in Utah'' when asked
+    about playing the Jazz in the 2017 playoffs.\footnote{Cestone and Seiter ('17).}   \end{block}
+
+  \vspace{12pt}
+  \begin{block}{Blaming the party scene for poor performance}
+    J.R. Smith's career-to-date success rate of shots taken was 42\%, 
+    but on Sunday afternoon games he only managed 34\% in 2012-13.\footnote{Concepcion ('15).}
+
+    \vspace{12pt}
+    After being traded from NY to Cavaliers, performance increased.
+    When asked why: ``There’s nothing you expect (in Cleveland) but basketball. 
+    There’s nothing, there’s no going out, there’s no late nights.''\footnote{Price, Levy ('15).}
+  \end{block} 
+\end{frame}
 
@@
 
-\begin{frame}   \frametitle{Anecdotal evidence: players do party}
-  \begin{block}{Open attitude toward partying}     Matt Barnes recently complained, ``there's no nightlife in Utah'' when asked
-    about playing the Jazz in the 2017 playoffs.\footnote{Cestone and Seiter ('17).}   \end{block}
-
-  \vspace{12pt}
-  \begin{block}{Blaming the party scene for poor performance}
-    J.R. Smith's career-to-date success rate of shots taken was 42\%, 
-    but on Sunday afternoon games he only managed 34\% in 2012-13.\footnote{Concepcion ('15).}
-
-    \vspace{12pt}
-    After being traded from NY to Cavaliers, performance increased.
-    When asked why: ``There’s nothing you expect (in Cleveland) but basketball. 
-    There’s nothing, there’s no going out, there’s no late nights.''\footnote{Price, Levy ('15).}
-  \end{block} 
-\end{frame}
-
 \begin{frame}
@@
 
-\begin{frame}   \frametitle{NBA data: 2010-11 through 2016-17 season}
-  \begin{itemize}     \item Each team plays 82 games per season.
-    \item 30 teams in the NBA.
-    \item Total of $82 \times 30 / 2 = 1,230$ games per season.  
-  \end{itemize}
-
-  \vspace{12pt}We query \href{http://www.basketball-reference.com/leagues/}{basketball-reference.com}
-  for a listing of game-days, and then 
-  \href{http://www.covers.com/sports/NBA/matchups?selectedDate=2011-1-01}{covers.com} for 
-  game-outcomes alongside a point-spread.
-
-  \vspace{12pt}  
-  \begin{itemize}     \item Missing the first couple months of 2010-11 season.
-    \item At most two missing games in each of remaining seasons.
+\begin{frame}   \frametitle{Data: 2010-11 through 2016-17}
+  \begin{block}{NBA Data}     \begin{itemize}       \item Total of $82 \times 30 / 2 = 1,230$ games per season.
+      % \item Missing the first couple months of 2010-11 season.
       \item Total of 7,829 games in our dataset.     \end{itemize}
-\end{frame}
-
-\begin{frame}   \frametitle{MLB data: 2011 through 2017 season}
-  \begin{itemize}     \item Each team plays 162 games per season.
-    \item 30 teams in the MLB.
-    \item Total of $162 \times 30 / 2 = 2,430$ games per season.   \end{itemize}
-
-  \vspace{12pt}We query \href{http://www.baseball-reference.com/leagues/}{baseball-reference.com}
-  for a listing of game days, and again look to covers.com for game-outcomes and money-lines.
-
-  \vspace{12pt}
-  \begin{itemize}     \item We have on average 2,100 games per season.
-    \item (Missing observations due to broken links, postponed games?)
-    \item Total of 12,709 games in our dataset.    \end{itemize} \end{frame}
-
-\begin{frame}   \frametitle{Bureau of Labor Statistics data: 2010-2016}
-  We look toward BLS data which describes the number of business establishments 
-  (of a very particular type) in each metropolitan statistical area
-  for each quarter of each year.
-
-  \vspace{12pt}  All teams are in the US with the exception of 
-  \vspace{12pt}\begin{itemize}     \item Toronto Raptors (NBA)
-    \item Toronto Blue Jays (MLB)   \end{itemize}
-
-  \vspace{12pt}These two teams are excluded from our analysis.\footnote{We don't expect Toronto to be a huge party city.} \end{frame}
-
-\begin{frame}   \frametitle{Minutes data for NBA}
-  For NBA, we also collect minutes data from ESPN, which describes the number of 
-  \vspace{12pt}
-  \begin{itemize}     
-    \item Free-throws
-    \item Field goals
-    \item Three-pointers   
-  \end{itemize}
-  \vspace{12pt}attempted and made by each player in each game. For each season and each team,
-  we have nearly complete minutes data; 85\% of our season-teams have no more than 3 missing games. \end{frame}
+  \end{block} 
+  \begin{block}{MLB Data}     \begin{itemize}       \item Total of $162 \times 30 / 2 = 2,430$ games per season.
+      % \item We have on average 2,100 games per season.
+      \item Total of 12,709 games in our dataset.     \end{itemize}   \end{block}
+  \begin{block}{BLS Data}
+    \begin{itemize}       \item \# establishments by MSA-quarter-year.
+      \item Excludes Toronto from our analysis.     \end{itemize}   \end{block}
+  \begin{block}{ESPN Data}     \begin{itemize}       \item Describes \# shots taken for each player in each game.
+      \item 85\% of our season-teams have no more than 3 missing games.     \end{itemize}   \end{block} \end{frame}
 

### 2017-07-15 init_draft.r
Now coloring histogram of point-spreads.

@@ -386,11 +386,11 @@ dev.off()
 ##################################################
 ### Distribution of point spreads.
 ##################################################
-
+h <- hist(nba.lines$line, prob = T, breaks = 32)
+cuts <- cut(h$breaks, breaks = c(-Inf, -3, 3, Inf))
 pdf('writing/distribution_point_spreads_nba.pdf', width = 16, height = 12)
-hist(nba.lines$line, prob = T, breaks = 20,
+plot(h, col = c('white', 'orange', 'white')[cuts],
      main = paste('Distribution of point-spreads in the NBA',
                   '2010-11 through 2016-17', sep = '\n'),
      xlab = 'Point-spread')
-lines(density(nba.lines$line, adjust = 2), lty = 'dashed', lwd = 3, col = 'orange')
 dev.off()

### 2017-07-15 cardinal_analytx_presentation/nightlife_and_sports.tex
Minor changes. Added a footnote.

@@
     \end{itemize}
-    \begin{block}{Problems}       \begin{itemize}           \item Effect size is not considered.
+    \begin{block}{Problems}       \begin{itemize}           \item Effect size is not considered.\footnote{What if a player scores 50\% of shots taken, but only 49\% on Sundays?}
           \item Density of support is questionable.\footnote{What if we miss the only shot we took on Sunday, is that relevant?}            \end{itemize}     \end{block}
@@
       \item Excludes Toronto from our analysis.     \end{itemize}   \end{block}
-  \begin{block}{ESPN Data}     \begin{itemize}       \item Describes \# shots taken for each player in each game.
+  \begin{block}{ESPN Data (NBA)}     \begin{itemize}       \item Describes \# shots taken for each player in each game.
       \item 85\% of our season-teams have no more than 3 missing games.     \end{itemize}   \end{block} \end{frame}
@@
   \begin{itemize}     
-    \item Number of days since the last game (i.e. rest-time),
+    \item Number of days since the last game (i.e. rest-time),\footnote{Note: if we are missing games in our dataset, this metric is affected.}
     \item Lagged changes in possession,\footnote{A crude measure of fatigue incurred from last game.}

### 2017-07-15 init_presentation.r
Using larger text for figures, as per Charlie's suggestion.

@@ -1,0 +1,414 @@
+################################################################################
+################################################################################
+################################################################################
+###
+### Title: Andreas Santucci
+###
+### Andreas Santucci
+###
+### Date: July 2017
+###
+### Inputs: 'tmp_data/nba.RData'
+###         'tmp_data/mlb.RData'
+###
+### Dependencies: data.table, ggplot2, magrittr, stargazer
+###
+################################################################################
+################################################################################
+
+##################################################
+### Set up workspace.
+##################################################
+
+require(data.table)
+require(ggplot2)
+require(magrittr)
+require(stargazer)
+
+load(file = 'tmp_data/nba.RData')  # nba.lines, espn.
+load(file = 'tmp_data/mlb.RData')  # mlb.lines
+
+##################################################
+### NBA Modeling.
+##################################################
+
+### Discrete indicator.
+m0 <- glm(outcome == 'W' ~ party.discrete + lag.chg.pos + log(travel.dist+1)*ew + 
+             nhours.lgame + I(hour(date)) + I(team==location) + as.factor(season),
+         data = nba.lines[season < 2017 & !is.na(party)], family = 'binomial',
+         na.action = 'na.exclude')
+
+### Continuous measure.
+m1 <- update(m0, . ~ party + . - party.discrete)
+
+### Placebo model.
+mp <- update(m0, . ~ party.placebo + . - party.discrete)
+
+### Player specific model - demeaned rebounds per minute.
+## m2 <- lm(reb.per.min ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame +
+##              as.factor(season) + player*as.factor(season),
+##          data = espn[last.game.loc != team] %>% na.omit)
+
+## require(lfe)
+## m2 <- felm(reb.per.min ~ party.discrete + lag.chg.pos + log(travel.dist+1) + nhours.lgame +
+##              as.factor(season) + player,
+##          data = espn[last.game.loc != team] %>% na.omit)
+
+### Total points admitted and scored.
+tpa <- lm(team.pts.admitted ~ party + lag.chg.pos + log(travel.dist+1) + nhours.lgame +
+              I(team == location) + team,
+          data = nba.lines[last.game.loc != team] %>% na.omit)
+mpa1 <- update(tpa, . ~ party.discrete + . - party)
+tps <- update(tpa, team.pts.scored ~ . + team)
+
+
[…]

### 2017-07-15 init_presentation.r
Changed output filepath for figures.

@@ -75,7 +75,7 @@ m4 <- update(m3, . ~ party.now + . - party)
 ##################################################
 ### MLB Betting.
 ##################################################
-pdf('writing/mlb_bets_by_season.pdf', width = 18, height = 13)
+pdf('cardinal_analytx_presentation/mlb_bets_by_season.pdf', width = 18, height = 13)
 par(mfrow = c(3,2))
 b <- lapply(2012:2016, function(yr) {
     m <- glm(I(team.score > opponent.score) ~ party + odds,
@@ -177,7 +177,7 @@ data <- rbind(nba.lines[, list(avg.party = mean(log(nmusicians+1), na.rm = T),
 setkey(data, avg.party)
 data[, idx := .GRP, by = locality]
 
-pdf('writing/Party_by_Team_Location.pdf', width = 18, height = 12)
+pdf('cardinal_analytx_presentation/Party_by_Team_Location.pdf', width = 18, height = 12)
 plot(x = data[sport == 'nba', jitter(idx)],
      y = data[sport == 'nba', avg.party],
      main = 'Continuous measure of party by location',
@@ -207,7 +207,7 @@ dev.off()
 ##################################################
 
 
-pdf('writing/travel_dist_density_by_party.pdf', width = 15, height = 13)
+pdf('cardinal_analytx_presentation/travel_dist_density_by_party.pdf', width = 15, height = 13)
 plot(density(nba.lines[(party.discrete) & !is.na(travel.dist), travel.dist]), 
      xlim = c(-100, 18e2), lty = 1,
      main = 'Density of travel distance for (non) party cities: NBA',
@@ -246,7 +246,7 @@ tmp[, location      := factor(location,      levels = dmat[order(distance, decre
 
 breaks <- 2^(0:6)
 
-pdf('writing/next_day_opponent.pdf', width = 14, height = 14)
+pdf('cardinal_analytx_presentation/next_day_opponent.pdf', width = 14, height = 14)
 ggplot(tmp, aes(x = location, y = last.game.loc, fill = N)) +
     geom_tile() +
     scale_fill_gradient(name = 'Count', trans = 'log', breaks = breaks) +
@@ -258,7 +258,7 @@ ggplot(tmp, aes(x = location, y = last.game.loc, fill = N)) +
           plot.title = element_text(hjust=.5))
 dev.off()
 
-pdf('writing/next_day_opponent_on_tour.pdf', width = 16, height = 12)
[…]

### 2017-07-15 cardinal_analytx_presentation/nightlife_and_sports.tex
Changed where figures are read from.

@@
 \begin{frame}
-  \centering   \includegraphics[scale=0.55]{../writing/meet_the_spread} 
+  \centering   \includegraphics[scale=0.55]{meet_the_spread} 
 \end{frame}
@@
 
-\begin{frame}   \frametitle{Overzealous assumption: next day opponent uncorrelated}
-  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent} 
-\end{frame}
+% \begin{frame}   \frametitle{Overzealous assumption: next day opponent % uncorrelated}
+%   \centering \includegraphics[scale=0.25]{next_day_opponent} 
+% \end{frame}
 
 \begin{frame}   \frametitle{A closer look at NBA games on the road}
-  \centering \includegraphics[scale=0.25]{../writing/next_day_opponent_on_tour} 
+  \centering \includegraphics[scale=0.25]{next_day_opponent_on_tour} 
 \end{frame}
@@
   \frametitle{Travel distance comparison}
-  \centering \includegraphics[scale=0.25]{../writing/travel_dist_density_by_party} 
+  \centering \includegraphics[scale=0.25]{travel_dist_density_by_party} 
 \end{frame}
@@
 \begin{frame}   
-  \frametitle{Continuous measure of party}
-  \centering \includegraphics[scale=0.375]{../writing/Party_by_Team_Location} 
+  % \frametitle{Continuous measure of party}
+  \hspace*{-25pt}
+  \includegraphics[scale=0.275]{Party_by_Team_Location} 
 \end{frame}
@@
 %   \frametitle{Continuous measure of party (MLB)}
-%   \centering \includegraphics[scale=0.3]{../writing/Party_by_Team_Location_mlb}   
+%   \centering \includegraphics[scale=0.3]{Party_by_Team_Location_mlb}   
 % \end{frame}
@@
   \centering
-  \includegraphics[scale=0.5]{../writing/mlb_bets_by_season} \end{frame}
+  \includegraphics[scale=0.25]{mlb_bets_by_season} \end{frame}
 
@@
   \centering
-  \includegraphics[scale=0.25]{../writing/distribution_point_spreads_nba}
+  \includegraphics[scale=0.25]{distribution_point_spreads_nba}
 \end{frame}
@@
 
-\begin{frame}   \frametitle{Why \# drinking establishments or population won't work}
-  \includegraphics[scale=0.3]{../writing/why_drinks_or_population_dont_work} \end{frame}
+\begin{frame}   
+  % \frametitle{Why \# drinking establishments or population won't work}
+  \includegraphics[scale=0.25]{why_drinks_or_population_dont_work} \end{frame}
 

### 2017-07-15 init_presentation.r
Now only keeping party variable in each regression table.

@@ -127,7 +127,7 @@ stargazer(m0, m1, covariate.labels = c('Party discrete', 'Party continuous',
                                        'Home team effect', 
                                        'Logged travel distance * east-west', 
                                        'Constant'),
-          omit = '(as.factor)|(Constant)',
+          keep = 'party',
           dep.var.labels = 'Meet the Spread')
 
 stargazer(mp, covariate.labels = c('Party placebo', 'Lag changes in posession', 
@@ -135,7 +135,7 @@ stargazer(mp, covariate.labels = c('Party placebo', 'Lag changes in posession',
                                    'Number hours rest time', 'Time of game during day', 
                                    'Home team effect', 'Logged travel distance * east-west', 
                                    'Constant'),
-          omit = 'as.factor',
+          keep = 'party',
           dep.var.labels = 'Meet the Spread (NBA)')
 
 
@@ -155,14 +155,14 @@ stargazer(mpa1, tpa, tps,
                                'Number hours since last game',
                                'Home team effect',
                                'Constant'),
-          omit = '(as.factor)|(team[^ ])')
+          keep = 'party')
 
 stargazer(m3, m4, covariate.labels = c('Continuous measure of nightlife',
                                        'Nightlife (no weekend interaction)', 
                                        'Bookmaker\'s odds', 'Home-team effect',
                                        'Number of rest days', 'Logged travel distance',
                                        'Weekend', 'Constant'),
-          omit = 'as.factor',
+          keep = 1:2,
           dep.var.labels = 'Probability of Winning')
 
 

### 2017-07-15 cardinal_analytx_presentation/nightlife_and_sports.tex
Now including only the party variable(s) of interest in regression tables. Listing what we control for in a footnote.

@@
 
-\begin{frame}   \frametitle{Results: NBA}
+\begin{frame}   \frametitle{Results: NBA\footnote{\tiny{We also account for: \texttt{Lag changes in posession, logged travel distance, east-west travel direction, number hours rest time, time of game during day, home team effect.}}}}
   \vspace{-2pt}
   \centering 
-  \tiny{ % m0, m1
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.557$^{***}$ &  \\    & (0.149) &  \\    & & \\   Party continuous &  & $-$0.158$^{*}$ \\    &  & (0.095) \\    & & \\   Lag changes in posession & 0.189 & 0.181 \\    & (0.140) & (0.140) \\    & & \\   Logged travel distance & $-$0.015 & $-$0.010 \\    & (0.026) & (0.026) \\    & & \\   East-west travel direction & $-$0.072 & $-$0.037 \\    & (0.234) & (0.234) \\    & & \\   Number hours rest time & 0.001 & 0.001 \\    & (0.001) & (0.001) \\    & & \\   Time of game during day & 0.006 & 0.004 \\    & (0.013) & (0.013) \\    & & \\   Home team effect & 0.002 & 0.004 \\    & (0.044) & (0.044) \\    & & \\   Logged travel distance * east-west & 0.011 & 0.007 \\    & (0.036) & (0.036) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,586.320 & $-$6,592.159 \\  Akaike Inf.
+  % m0, m1
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Meet the Spread} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Party discrete & $-$0.557$^{***}$ &  \\    & (0.149) &  \\    & & \\   Party continuous &  & $-$0.158$^{*}$ \\    &  & (0.095) \\    & & \\  \hline \\[-1.8ex]  Observations & 9,517 & 9,517 \\  Log Likelihood & $-$6,586.320 & $-$6,592.159 \\  Akaike Inf.
 Crit.
-& 13,200.640 & 13,212.320 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  }
-\end{frame}
+& 13,200.640 & 13,212.320 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{frame}
 
@@
 
-\begin{frame}   \frametitle{Placebo test (discrete indicator): what if players rest?}
+\begin{frame}   \frametitle{Placebo test (discrete indicator): what if players rest?\footnote{\tiny{We also account for: \texttt{Lag changes in posession, logged travel distance, east-west travel direction, number hours rest time, time of game during day, home team effect.}}}}
   \centering
-  \tiny{ % mp
-\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.053 \\    & (0.084) \\    & \\   Lag changes in posession & 0.170 \\    & (0.140) \\    & \\   Logged travel distance & $-$0.004 \\    & (0.026) \\    & \\   East-west travel direction & $-$0.034 \\    & (0.235) \\    & \\   Number hours rest time & 0.001$^{*}$ \\    & (0.001) \\    & \\   Time of game during day & 0.002 \\    & (0.013) \\    & \\   Home team effect & $-$0.006 \\    & (0.044) \\    & \\   Logged travel distance * east-west & 0.006 \\    & (0.036) \\    & \\   Constant & $-$0.865 \\    & (0.695) \\    & \\  \hline \\[-1.8ex]  Observations & 9,517 \\  Log Likelihood & $-$6,593.344 \\  Akaike Inf.
+% mp
+\begin{tabular}{@{\extracolsep{5pt}}lc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{1}{c}{\textit{Dependent variable:}} \\  \cline{2-2}  \\[-1.8ex] & Meet the Spread (NBA) \\  \hline \\[-1.8ex]   Party placebo & 0.053 \\    & (0.084) \\    & \\  \hline \\[-1.8ex]  Observations & 9,517 \\  Log Likelihood & $-$6,593.344 \\  Akaike Inf.
 Crit.
-& 13,214.690 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
-}
-\end{frame}
+& 13,214.690 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{frame}
 
@@
 
-\begin{frame}   \frametitle{Results: MLB}
+\begin{frame}   \frametitle{Results: MLB\footnote{\tiny{We also account for \texttt{bookmaker's odds, home-team effect, number of rest days, logged travel distance, weekend indicator.}}}}
   \centering
-  \tiny{ % m3, m4
-\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.120$^{*}$ &  \\    & (0.071) &  \\    & & \\   Nightlife (no weekend interaction) &  & 0.137 \\    &  & (0.134) \\    & & \\   Bookmaker's odds & 2.682$^{***}$ & 2.669$^{***}$ \\    & (0.158) & (0.158) \\    & & \\   Home-team effect & 0.014 & 0.144 \\    & (0.034) & (0.102) \\    & & \\   Number of rest days & 0.018 & 0.016 \\    & (0.021) & (0.021) \\    & & \\   Logged travel distance & $-$0.005 & $-$0.003 \\    & (0.005) & (0.005) \\    & & \\   Weekend & 0.044 & 0.002 \\    & (0.038) & (0.029) \\    & & \\   Constant & $-$1.363$^{***}$ & $-$1.468$^{***}$ \\    & (0.084) & (0.122) \\    & & \\  \hline \\[-1.8ex]  Observations & 26,473 & 26,473 \\  Log Likelihood & $-$18,124.350 & $-$18,125.300 \\  Akaike Inf.
+% m3, m4
+\begin{tabular}{@{\extracolsep{5pt}}lcc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{2}{c}{\textit{Dependent variable:}} \\  \cline{2-3}  \\[-1.8ex] & \multicolumn{2}{c}{Probability of Winning} \\  \\[-1.8ex] & (1) & (2)\\  \hline \\[-1.8ex]   Continuous measure of nightlife & $-$0.120$^{*}$ &  \\    & (0.071) &  \\    & & \\   Nightlife (no weekend interaction) &  & 0.138 \\    &  & (0.135) \\    & & \\  \hline \\[-1.8ex]  Observations & 26,473 & 26,473 \\  Log Likelihood & $-$18,124.350 & $-$18,125.260 \\  Akaike Inf.
 Crit.
-& 36,272.700 & 36,264.600 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  }
-\end{frame}
+& 36,272.700 & 36,274.510 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{frame}
 
@@
 
-\begin{frame}   \frametitle{Defense in the NBA is affected by our measure for party}
-  \tiny{  % mpa1, tpa, tps
-\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & \multicolumn{2}{c}{Team Points Admitted} & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.970$^{***}$ &  &  \\    & (0.835) &  &  \\    & & & \\   Continuous party measure &  & 1.661$^{***}$ & $-$0.139 \\    &  & (0.644) & (0.637) \\    & & & \\   Lag change in possessions & 6.694$^{***}$ & 6.751$^{***}$ & 5.858$^{***}$ \\    & (0.973) & (0.973) & (0.963) \\    & & & \\   Logged travel distance & $-$0.114 & $-$0.132 & $-$0.285 \\    & (0.198) & (0.198) & (0.196) \\    & & & \\   Number hours since last game & 0.011$^{*}$ & 0.015$^{**}$ & 0.016$^{**}$ \\    & (0.006) & (0.007) & (0.007) \\    & & & \\   Home team effect & $-$3.269$^{***}$ & $-$3.269$^{***}$ & 2.417$^{***}$ \\    & (0.303) & (0.303) & (0.300) \\    & & & \\   Constant & 68.114$^{***}$ & 67.686$^{***}$ & 71.348$^{***}$ \\    & (4.805) & (4.810) & (4.758) \\    & & & \\  \hline \\[-1.8ex]  Observations & 6,234 & 6,234 & 6,234 \\  R$^{2}$ & 0.091 & 0.090 & 0.084 \\  Adjusted R$^{2}$ & 0.087 & 0.086 & 0.079 \\  Residual Std.
-Error (df = 6201) & 11.670 & 11.675 & 11.549 \\  F Statistic (df = 32; 6201) & 19.453$^{***}$ & 19.247$^{***}$ & 17.720$^{***}$ \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}    }
-\end{frame}
+\begin{frame}   \frametitle{Defense in the NBA is affected by our measure for party\footnote{\tiny{We also account for \texttt{Lag changes in posession, logged travel distance, number of hours since last game, home team effect.}}}}
+% mpa1, tpa, tps
+\scriptsize{
+\begin{tabular}{@{\extracolsep{5pt}}lccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]   & \multicolumn{3}{c}{\textit{Dependent variable:}} \\  \cline{2-4}  \\[-1.8ex] & \multicolumn{2}{c}{Team Points Admitted} & Team Points Scored \\  \\[-1.8ex] & (1) & (2) & (3)\\  \hline \\[-1.8ex]   Discrete party indicator & 2.970$^{***}$ &  &  \\    & (0.835) &  &  \\    & & & \\   Continuous party measure &  & 1.661$^{***}$ & $-$0.139 \\    &  & (0.644) & (0.637) \\    & & & \\  \hline \\[-1.8ex]  Observations & 6,234 & 6,234 & 6,234 \\  R$^{2}$ & 0.091 & 0.090 & 0.084 \\  Adjusted R$^{2}$ & 0.087 & 0.086 & 0.079 \\  Residual Std.
+Error (df = 6201) & 11.670 & 11.675 & 11.549 \\  F Statistic (df = 32; 6201) & 19.453$^{***}$ & 19.247$^{***}$ & 17.720$^{***}$ \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}
+} \end{frame}
 
@@
   \includegraphics[scale=0.25]{distribution_point_spreads_nba}
-\end{frame}
-
-\begin{frame}   \frametitle{Future research - NBA}
-  Within the NBA
-  \begin{itemize}     
-    \item Are teams more likely to party after a win or a       loss?
-    \item Do teams who aren't going to make the playoffs party more?
-    \item Are there demographic differences which influence propensity to party? E.g. are younger, unmarried players more susceptible to our treatment effect?
-    \item We could also examine how shots taken vs. shots scored are affected for star players.
-  \end{itemize}
-  \vspace{12pt}  A preliminary and ad-hoc analysis shows that injury risk increases as a function of party.\footnote{We look for instances where players are put on the bench and stop playing (at least several) games as a function of our nightlife indicator.}
-  This might inform coaches which players to bench and when.
-
-\end{frame}
-
-\begin{frame}   \frametitle{Future research (MLB and general)}
-  \begin{itemize}     
-    \item Within the MLB, we can look at statistics such as batting averages or general fielding statistics. 
-    \vspace{12pt}
-    \item Possibly prolonged effects: recall our magic 24 hour number.
-      How does the hangover effect taper off as a function of time?
-    \vspace{12pt}
-    \item NHL or European soccer leagues.   
-  \end{itemize} 
 \end{frame}
@@
 
+\begin{frame}   \frametitle{Future research - NBA}
+  Within the NBA
+  \begin{itemize}     
+    \item Are teams more likely to party after a win or a       loss?
+    \item Do teams who aren't going to make the playoffs party more?
+    \item Are there demographic differences which influence propensity to party? E.g. are younger, unmarried players more susceptible to our treatment effect?
+    \item We could also examine how shots taken vs. shots scored are affected for star players.
+  \end{itemize}
+  \vspace{12pt}  A preliminary and ad-hoc analysis shows that injury risk increases as a function of party.\footnote{We look for instances where players are put on the bench and stop playing (at least several) games as a function of our nightlife indicator.}
+  This might inform coaches which players to bench and when.
+\end{frame}
+
+\begin{frame}   \frametitle{Future research (MLB and general)}
+  \begin{itemize}     
+    \item Within the MLB, we can look at statistics such as batting averages or general fielding statistics. 
+    \vspace{12pt}
+    \item Possibly prolonged effects: recall our magic 24 hour number.
+      How does the hangover effect taper off as a function of time?
+    \vspace{12pt}
+    \item NHL or European soccer leagues.   
+  \end{itemize} 
+\end{frame}
+
 \begin{frame}   

### 2017-07-17 cardinal_analytx_presentation/nightlife_and_sports.tex
Updated interpretation slide. Also included mention of team-fixed effects for poitns scored.

@@
   \begin{itemize}     
-    \item Number of days since the last game (i.e. rest-time),\footnote{Note: if we are missing games in our dataset, this metric is affected.}
-    \item Lagged changes in possession,\footnote{A crude measure of fatigue incurred from last game.}
+    \item Number of hours since the last game (i.e. rest-time),\footnote{Note: if we are missing games in our dataset, this metric is affected.}
+    \item Lagged changes in possession,\footnote{A crude measure of fatigue incurred from last game.
+To back this out: $\Delta \textrm{possessions} \approx \textrm{Shots made} + \textrm{turnovers} + (\textrm{shots missed} - \textrm{offensive rebounds})$.
+}
     \item Lagged game location, and
@@
   Let's consider the simplest model, where $Y_i | X \sim \textrm{Bern}(p)$, i.e.
-  $\Pr(Y_i = y | X) = p^{y} (1-p)^{1-y}$. Using a generalized linear model,
-  we fit a linear predictor to a transformation of the expected value of our
+  $\Pr(Y_i = y | X) = p^{y} (1-p)^{1-y}$. Using a GLM,
+  we fit our predictor to a transformation of the expected value of our
   variable of interest. I.e.
@@
   \]
-  So, we interpret our regression coefficient as follows: visiting LA or NY within
-  24 hours of another game corresponds to being $e^{\beta_1} \approx e^{-0.55} \approx \frac{3}{5}$
-  as likely to meet the spread, all other factors held constant. 
+  \begin{block}{Laymen's interpretation}   Visiting LA or NY within
+  24 hours of another game corresponds to being $e^{\beta_1} \approx e^{-0.55}   \approx \frac{60}{100}$
+  as likely to meet the spread, all other factors held constant.        \end{block}
+
+  \vspace{12pt}Our continuous 
+  measure suggests a team is $\frac{85}{100}$ as likely to meet the spread, given a 
+  one unit increase in last night's party index.
 \end{frame}
@@
 
-\begin{frame}   \frametitle{Defense in the NBA is affected by our measure for party\footnote{\tiny{We also account for \texttt{Lag changes in posession, logged travel distance, number of hours since last game, home team effect.}}}}
+\begin{frame}   \frametitle{Defense in the NBA is affected by our measure for party\footnote{\tiny{We also account for \texttt{Lag changes in posession, logged travel distance, number of hours since last game, home team effect}.
+In addition, we account for team-fixed effects.}}}
 % mpa1, tpa, tps

### 2017-07-17 init_presentation.r
Added legend title for travel density. Estimated SD of points-spread.

@@ -214,7 +214,8 @@ plot(density(nba.lines[(party.discrete) & !is.na(travel.dist), travel.dist]),
      xlab = 'Travel Distance (Miles)',
      cex.main = 3, cex.axis = 2, cex.lab = 2, ylab = '')
 lines(density(nba.lines[!(party.discrete) & !is.na(travel.dist), travel.dist]), lty = 2)
-legend(x = 12e2, y = 0.0015, legend = c('LA/NY', 'non-party cities'), lty = 1:2, cex = 2)
+legend(x = 12e2, y = 0.0015, legend = c('LA/NY', 'non-party cities'), lty = 1:2, cex = 2,
+       title = 'Last game location')
 mtext(text = 'Density', side = 2, line = 2.4, las = 3, cex = 2)
 dev.off()
 
@@ -412,3 +413,5 @@ plot(h, col = c('white', 'orange', 'white')[cuts],
      cex.main = 2.5, cex.axis = 2, cex.lab = 2)
 mtext(text = 'Frequency', side = 2, line = 2.7, las = 3, cex = 2)
 dev.off()
+
+sd(nba.lines$line)

### 2017-07-17 init_presentation.r
Changed input file paths for figures.

@@ -75,7 +75,7 @@ m4 <- update(m3, . ~ party.now + . - party)
 ##################################################
 ### MLB Betting.
 ##################################################
-pdf('cardinal_analytx_presentation/mlb_bets_by_season.pdf', width = 18, height = 13)
+pdf('cardinal_analytx_presentation/figures/mlb_bets_by_season.pdf', width = 18, height = 13)
 par(mfrow = c(3,2))
 b <- lapply(2012:2016, function(yr) {
     m <- glm(I(team.score > opponent.score) ~ party + odds,
@@ -177,7 +177,7 @@ data <- rbind(nba.lines[, list(avg.party = mean(log(nmusicians+1), na.rm = T),
 setkey(data, avg.party)
 data[, idx := .GRP, by = locality]
 
-pdf('cardinal_analytx_presentation/Party_by_Team_Location.pdf', width = 18, height = 12)
+pdf('cardinal_analytx_presentation/figures/Party_by_Team_Location.pdf', width = 18, height = 12)
 plot(x = data[sport == 'nba', jitter(idx)],
      y = data[sport == 'nba', avg.party],
      main = 'Continuous measure of party by location',
@@ -207,7 +207,7 @@ dev.off()
 ##################################################
 
 
-pdf('cardinal_analytx_presentation/travel_dist_density_by_party.pdf', width = 15, height = 13)
+pdf('cardinal_analytx_presentation/figures/travel_dist_density_by_party.pdf', width = 15, height = 13)
 plot(density(nba.lines[(party.discrete) & !is.na(travel.dist), travel.dist]), 
      xlim = c(-100, 18e2), lty = 1,
      main = 'Density of travel distance for (non) party cities: NBA',
@@ -247,7 +247,7 @@ tmp[, location      := factor(location,      levels = dmat[order(distance, decre
 
 breaks <- 2^(0:6)
 
-pdf('cardinal_analytx_presentation/next_day_opponent.pdf', width = 14, height = 14)
+pdf('cardinal_analytx_presentation/figures/next_day_opponent.pdf', width = 14, height = 14)
 ggplot(tmp, aes(x = location, y = last.game.loc, fill = N)) +
     geom_tile() +
     scale_fill_gradient(name = 'Count', trans = 'log', breaks = breaks) +
@@ -259,7 +259,7 @@ ggplot(tmp, aes(x = location, y = last.game.loc, fill = N)) +
[…]

### 2017-07-17 cardinal_analytx_presentation/nightlife_and_sports.tex
Changed input file paths for figures.

@@
 \begin{frame}
-  \centering   \includegraphics[scale=0.55]{meet_the_spread} 
+  \centering   \includegraphics[scale=0.55]{figures/meet_the_spread} 
 \end{frame}
@@
 \begin{frame}   \frametitle{A closer look at NBA games on the road}
-  \centering \includegraphics[scale=0.25]{next_day_opponent_on_tour} 
+  \centering \includegraphics[scale=0.25]{figures/next_day_opponent_on_tour} 
 \end{frame}
@@
   \frametitle{Travel distance comparison}
-  \centering \includegraphics[scale=0.25]{travel_dist_density_by_party} 
+  \centering \includegraphics[scale=0.25]{figures/travel_dist_density_by_party} 
 \end{frame}
@@
   \hspace*{-25pt}
-  \includegraphics[scale=0.275]{Party_by_Team_Location} 
+  \includegraphics[scale=0.275]{figures/Party_by_Team_Location} 
 \end{frame}
@@
   \centering
-  \includegraphics[scale=0.25]{mlb_bets_by_season} \end{frame}
+  \includegraphics[scale=0.25]{figures/mlb_bets_by_season} \end{frame}
 
@@
   \centering
-  \includegraphics[scale=0.25]{distribution_point_spreads_nba}
+  \includegraphics[scale=0.25]{figures/distribution_point_spreads_nba}
 \end{frame}
@@
   % \frametitle{Why \# drinking establishments or population won't work}
-  \includegraphics[scale=0.25]{why_drinks_or_population_dont_work} \end{frame}
+  \includegraphics[scale=0.25]{figures/why_drinks_or_population_dont_work} \end{frame}
 
